### 2011-09-26 misc/papers/icassp12_tied/icassp12.tex
Adding paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@509 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\documentclass{article}
\usepackage{spconf,amsmath,amssymb,graphicx,ngerman,bm}
\selectlanguage{USenglish}

% Example definitions.
% --------------------
\def \x{{\mathbf x}}
\def \v{{\mathbf v}}
\def \w{{\mathbf w}}
\def \M{{\mathbf M}}
\def \m{{\bm \mu}}
\def \k{{\mathbf \Sigma}}
\def \nv{{\mathcal N}}
\def \L{{\cal L}}

% Title.
% ------
\title{REVISITING SEMI-CONTINUOUS HIDDEN MARKOV MODELS}

%
% Single address.
% ---------------
\name{
K.~Riedhammer$^1$,
T.~Bocklet$^1$,
A.~Ghoshal$^2$,
D.~Povey$^3$
}
\address{
$^1$Pattern Recognition Lab, University of Erlangen-Nuremberg, \sc{Germany}\\
$^2$Spoken Language Systems, Saarland University, \sc{Germany}\\
$^3$Microsoft Research, Redmond, WA, {\sc USA}\\
{\small\tt korbinian.riedhammer@informatik.uni-erlangen.de}
}

\begin{document}
\ninept

\maketitle

\begin{abstract}
In the past decade, semi-continuous hidden Markov models (HMMs) have not attracted 
much attention in the speech recognition community. The strongly growing amount
of training data and continuous improvements of the initialization and training
procedures led to the impression that continuous HMMs are the best choice of
acoustic model.
%
However, recent work on tasks involving cross-lingual speech recognition and
languages with little training data faced the same the old problem -- millions
of parameters require a large amount of (unavailable) training data -- thus people are
investigating methods of reducing the number of parameters while maintaining or
extending the modeling capabilities of continuous models.
%
In this work, we compare continuous, semi-continuous, two-level tree based 
multiple-codebook semi-continuous with full covariance matrices and subspace Gaussian
mixture models.
%
Experiments using on the RM and WSJ corpora show that a semi-continuous system
can still yield competitive results while using fewer Gaussian components.
\end{abstract}

\begin{keywords}
automatic speech recognition, acoustic modeling
\end{keywords}

\section{Introduction}
\label{sec:intro}
In the past years, semi-continuous hidden Markov models (SC-HMMs) \cite{huang1989shm}
have not attracted much attention in the automatic speech recognition (ASR) community. 
Most major frameworks (e.g.~CMU {\sc Sphinx} or HTK) more or less retired
semi-continuous models in favor of continuous models, which performed 
considerably better with the growing amount of training data available.
%
However, topics like cross-lingual model initialization and model training
for languages with little training data became more and more popular recently,
along with efforts on how to reduce the number of parameters of a continuous
system while maintaining or even extending the modeling capabilities and thus
the robustness and performance.
%
Typical techniques to reduce the number of parameters for continuous
systems include generalizations of the phonetic context modeling to reduce the 
number of states and state-tying to reduce the number of Gaussians.

It is well known that context dependent states significantly improve the
recognition performance (e.g.~from monophone to triphone). However for a
continuous system, increasing the number of states not only requires more
Gaussians to be estimated, but also implies that the training data for each
state is reduced.
%
This is the probably strongest advantage of semi-continuous models: While the
codebook Gaussians are always estimated on the whole data set, the only state
specific variables are the weights of the Gaussians, i.e.~increasing the number
of states only slightly increases the number of parameters, and the Gaussian
means and variances are still estimated using all the data; the reduced number
of Gaussian parameters for a large data set also allows to robustly estimate
full covariances.
%
This allows to reliably estimate arbitrary large phonetic context 
(``polyphones''), that showed good performance for the recognition of 
spontaneous  speech \cite{schukattalamazzini1994srf,schukattalamazzini1995as} 
while having a rather small number of parameters.
%
Further advantages of the shared codebook Gaussians are that they need to be 
evaluated only once per frame, and that the codebook can be initialized, trained
and adapted in a solely acoustic way.

The idea of semi-continuous models can be extended to multiple codebooks; by
assigning certain groups of states to specific codebooks, one can find a
hybrid between a continuous and semi-continuous models that combine the strength
of state-specific Gaussians with reliable parameter estimation \cite{prasad2004t2b}.

Extending the idea of sharing parameters, subspace Gaussian mixture models
\cite{povey2011sgm} assign state specific weights and means, while the latter
are derived from the means of a shared codebook (universal background model, UBM)
using a state, phone or speaker specific transformation, thus limiting the state
Gaussians to a subspace of the UBM.

In this article, we experiment with classic SC-HMMs and two-level tree based 
multiple-codebook SC-HMMs which are, in terms of parameter sharing, somewhere in 
between continuous and subspace Gaussian mixture models; 
we compare the above four types of acoustic models on a small and medium
sized corpus of read English while keeping acoustic frontend, training and 
decoding unchanged. The software is part of the {\sc Kaldi} speech recognition
toolkit \cite{povey2011tks} and is freely available for download, along with the
example scripts to reproduce the presented experiments.

\section{Acoustic Modeling}
\label{sec:am}
In this section, we summarize the different forms of acoustic model we use: 
the continuous, semi-continuous, multiple-codebook semi-continuous, and 
subspace forms of Gaussian Mixture Models. 
We also describe the phonetic decisoin-tree building process, which is 
necessary background for how we build the multiple-codebook systems.

\subsection{Tree-Building}
The acoustic-phonetic decision tree provides the link between phones
in context and emission probability density functions (pdfs). 
%
The statistics for the tree clustering consist of the 
sufficient statistics needed to estimate a Gaussian for each seen combination 
of (phonetic context, HMM-state).  These statistics are based on a Viterbi 
alignment of a previously built system.  The roots
of the decision can be configured; for the RM experiments, these correspond to
the phones in the system, and on Wall Street Journal, they
correspond to the ``real phones'', i.e. grouping different stress and
position-marked versions of the same phone together.

The splitting procedure can ask questions not only about the context phones,
but also the central phone and the HMM state; the phonetic questions are 
derived from an automatic clustering procedure.  The splitting procedure is
greedy and optimizes the likelihood of the data given a single Gaussian in each
tree leaf; this is subject to a fixed variance floor to avoid problems caused
by singleton counts and numerical roundoff.  We typically split up to a 
specified number of leaves and then cluster the resulting leaves as long as the
likelihood loss due to combining them is less than the likelihood improvement
on the final split of the tree-building process; we typically restrict this 
clustering to avoid combining leaves across different subtrees (e.g., from 
different monophones).  The post-clustering process typically reduces the number 
of leaves by around 20\%.

\subsection{Continuous Models}
For continuous models, every leaf of the tree is assigned an separate
Gaussian mixture model. We initialize the mixtures with a single component
each, and subsequently allocate more components by splitting components every 
iteration until a certain total number of components is reached (in this case
about 9000). We allocate the Gaussians proportional to a small power (e.g.~0.2) 
of the state's occupation count.
Formally, the emission probability of tree leaf $j$ is computed as
\begin{equation}
p(\x | j) = \sum_{i=1}^{N_j} c_{ji} \nv(\x; \m_{ji}, \k_{ji}) 
\end{equation}
where $N_j$ is the number of Gaussians assigned to $s_j$, and the $\m_{ji}$ and
$\k_{ji}$ are the means and covariance matrices of the mixtures.

\subsection{Semi-continuous Models}
The idea of semi-continuous models is to have a large number of Gaussians that
are shared by every tree leaf using individual weights, thus the emission 
probability of tree leaf $j$ can be computed as
\begin{equation}
p(\x | j) = \sum_{i=1}^{N} c_{ji} \nv(\x; \m_i, \k_i) 
\end{equation}
where $c_{ji}$ is the weight of mixture $i$ for leaf $j$. As the means
and covariances are shared, increasing the number of states only implies
a small increase in number of parameters even for full-covariance Gaussians.
Furthermore, the Gaussians need to be evaluated only once for each $\x$.

Another advantage is the initialization and use of the codebook. It can be
initialized and adapted in a fully unsupervised manner using expectation
maximization (EM), maximum a-posteriori (MAP), maximum likelihood linear
regression (MLLR) and similar algorithms.

For better performance, we initialize the codebook using the tree statistics
collected on a prior phone alignment. For each tree leaf, we include
the respective Gaussian as a component in the mixture.
The components are merged and (if necessary) split to eliminate low count 
Gaussians and to match the desired number of components. In a final step, 
5 EM iterations are used to improve the goodness of fit to the acoustic 
training data.

\subsection{Two-level Tree based Multiple-Codebook Semi-continuous Models}
A style of system that lies between the semi-continuous and continuous types of
systems is one based on a two-level tree; this is described in~\cite{prasad2004t2b}
as a state-clustered tied-mixture (SCTM) system. The way we implement this is
to first build the decision tree used for phonetic clustering to a relatively
coarse level (e.g.~100); and to then split more finely (e.g.~2500), but for
each new leaf, remember which leaf of the original tree it corresponds to.
Each leaf of the first tree corresponds to a codebook of Gaussians, i.e.~the 
Gaussian parameters are tied at this level.  For this type of system we do not
do any post-clustering.
The leaves of the second tree contain the weights, and these leaves correspond 
to the actual context-dependent HMM states.

The way we formalize this is that the context-dependent states $j$ each have 
a corresponding codebook indexed $k$, and the function $m(j) \rightarrow k$ 
maps from the leaf index to the codebook index. The likelihood function is now
%
\begin{equation}
p(\x | j) = \sum_{i=1}^{N_{m(j)}} c_{ji} \, \nv\left(\x; \m_{m(j), i}, \k_{m(j), i}\right)
\end{equation}
%
where $\m_{m(j), i}$ is the $i$-th mean vector of the codebook
associated with $j$.

Similar to the traditional semi-continuous codebook initialization, we start
from the tree statistics and distribute the Gaussians to the respective codebooks
using the tree mapping to obtain preliminary codebooks.
The target size $N_k$ of codebook $k$ is determined with respect to a power
of the occupancies of the respective leaves as
\begin{equation}
N_k = N_0 + \frac
  { \left(  \sum_{l \in \{m(l) = k\}} \text{occ}(l)  \right)^q }
  { \sum_r \left(  \sum_{t \in \{m(t) = r\}} \text{occ}(l)  \right)^q } 
  \left( N - K \cdot N_0 \right)
\end{equation}
where $N_0$ is a minimum number of Gaussians per codebooks (e.g.\,3), $N$ is
the total number of Gaussians, $K$ the number of codebooks, $\text{occ}(j)$ 
is the occupancy of tree leaf $j$, and $q$ controls the influence of
the occupancy regarding one codebook. A typical value of $q=0.2$ leads to
rather homogeneous codebook sizes while still attributing more Gaussians
to states with larger occupancies. For the two-level architecture, a $q=1$
yielded best results.
The target sizes of the codebooks are again enforced by either splitting or
merging the components.

\subsection{Subspace Gaussian Mixture Models}
The idea of subspace Gaussian mixture models (SGMM) is, similar to 
semi-continuous models, to reduce the number of parameters by selecting the
Gaussians from a subspace spanned by a universal background model (codebook)
and state specific transformations. In principle, the SGMM emission pdfs can
be computed as
\begin{eqnarray}
p(\x | j) & = & \sum_{i=1}^{N} c_{ji} \nv(\x; \m_{ji}, \k_i) \\
\m_{ji}    & = & \M_i \v_j \\
c_{ji}     & = & \frac{\exp \w_i^T \v_j}{\sum_l^N \exp \w_l^T \v_j}
\end{eqnarray}
where the covariance matrices $\k_i$ are shared between all leafs $j$. The
weights $w_{ji}$ and means $\m_{ji}$ are derived from $\v_j$ together with
$\M_i$ and $\w_i$. The term ``subspace'' indicates that the parameters of the
mixtures are limited to ta sub-space of the entire space of parameters of
the underlying codebook.
%
A detailed description and derivation of the accumulation and update formulas
can be found in \cite{povey2011sgm}. Note that the experiments in this article
are without adaptation.

\section{Smoothing Techniques for Semi-continuous Models}
%
\subsection{Intra-Iteration Smoothing}
Although an increasing number of tree leaves does not imply a huge increase in
parameters, the sufficient statistics for the individual leaves will be
sparser, especially for little training data.
The {\em intra-iteration} smoothing is motivated by the fact that tree leaves
that share the root are similar, thus, if the sufficient statistics of the
weights of a leaf $j$ are very small, they should be interpolated with the 
statistics of closely related leaf nodes.
To do so, we first propagate all sufficient statistics up the tree so that the
statistics of any node is the sum of its children's statistics.
Second, the statistics of each node and leaf are interpolated top-down with 
their parent's using an interpolation weight $\rho$ in
\begin{equation}
\hat\gamma_{ji} \leftarrow 
  \underbrace{\left(\gamma_{ji} + \frac{\rho \, \gamma_{p(j),i}}{\left( \sum_k \gamma_{p(j),k} \right) + \epsilon}\right)}_{\mathrel{\mathop{:}}= \bar\gamma_{ji}}
  \cdot 
  \underbrace{\frac{\sum_k \gamma_{jk}}{\sum_k \bar\gamma_{jk}}}_\text{normalization}
\end{equation}
where $\gamma_{ji}$ is the accumulator of weight $i$ in tree leaf $j$
and $p(j)$ is the mapping to the parent node of $j$.
%
For two-level tree based models, the propagation and interpolation cannot
be applied across different codebooks. 
%
A similar yet different interpolation scheme without normalization was used in 
\cite{schukattalamazzini1994srf,schukattalamazzini1995as} along with a different
tree structure.

%This is similar to the {\em I} step of the {\em APIS} algorithm 
%\cite{schukattalamazzini1994srf,schukattalamazzini1995as} but smoothes the
%statistics with respect to their counts instead of replacing them.
%This modification is necessary as the tree in \cite{schukattalamazzini1995as}
%is constructed in a way, that the tree splits (nodes) represent acoustic
%models (the {\em polyphones}), i.e., the first node level would be the
%monophones, and any phone with more context would be a child node, with only
%the phones with the largest context being a leaf. For the training, any data contributing
%to a node $j$ should also contribute to its parent $p(j)$, and if a child
%has insufficient statistics, these can be interpolated with the parent's. 
%As we attach models only to the leaves, this motivation does not (necessarily)
%hold, thus requiring the above modification of the interpolation scheme.

\subsection{Inter-Iteration Smoothing}
Another typical problem that arises when training semi-continuous models is
that the weights tend to converge on a local optimum over the iterations. 
This is due to the fact that the codebook parameters change rather slowly 
as the statistics are collected from all respective tree leaves, but the 
leaf specific weights converge much faster.
%
To compensate for this, we smooth the newly estimated {\em parameters} 
$\Theta^{(i+1)}$ with the ones from the prior iteration using an interpolation
weight $\varrho$
\begin{equation}
\hat\Theta^{(i+1)} \leftarrow (1 - \varrho) \, \Theta^{(i+1)} + \varrho \, \Theta^{(i)} % \quad .
\end{equation}
which leads to an expontially decreasing impact of the initial parameter set.

%----%<------------------------------------------------------------------------

\section{Systems Description}
The experiments presented in this article are computed on the DARPA Resource 
Management Continuous Speech Corpus (RM) \cite{price1993rm} and the Wall Street
Journal Corpus \cite{garofalo2007wsj}.

For both data sets, we first train an initial monophone continuous system
with about 1000 diagonal Gaussians (122 states) on a small data subset; with 
the final alignments, we train an initial triphone continuous system with 
about 10000 Gaussians (1600 states).  The alignments of this triphone system
are used as a basis for the actual system training.

Note that this baseline may be computed on any other (reduced) data set
and has the sole purpose of providing better-than-linear initial alignments
to speed up the convergence of the models.
%
The details of the model training can also be found in the {\sc Kaldi} recipes.

On the acoustic frontend, we extract 13 Mel-frequency cepstral coefficients,
compute deltas and delta-deltas, and apply cepstral mean normalization. For
the decoding, we use a regular 3-gram estimated using the IRSTLM toolkit 
\cite{federico2008iao}.

\subsection{Resource Management}
For the RM data set, we train on the speaker independent training and development 
set (about 4000 utterances) and test on the six DARPA test runs Mar and Oct '87,
Feb and Oct '89, Feb '91 and Sep '92 (about 1500 utterances total).  The parameters 
were tuned to the test set.
%
\begin{itemize}
\item
  {\em cont}: continuous triphone system using 9011 diagonal covariance
  Gaussians in 1480 tree leaves
\item
  {\em semi}: semi-continuous triphone system using 768 full covariance
  Gaussians in 2500 tree leaves, but no smoothing (explanations below).
\item
  {\em 2lvl}: two-level tree based semi-continuous triphone system using
  3072 full covariance Gaussians in 208 codebooks ($4/14.7/141$ min/average/max
  components), 2500 tree leaves, $\rho = 35$ and $\varrho = 0.2$.
\item
  {\em sgmm}: subspace Gaussian mixture model triphone system using a 400 
  full-covariance Gaussian background model and 2016 tree leaves.
\end{itemize}

\subsection{WSJ}
% TODO fix the WSJ train/dev/eval description
For the WSJ data set, we trained on the si284 training set, tuned on dev
and tested on eval92,93.
% TODO Korbinian: There are more expts ongoing, they should be done by 
%                 tomorrow! I will also fill in the details below then.
\begin{itemize}
\item
  {\em cont}: continuous triphone system using 10000 diagonal covariance
  Gaussians in 1576 tree leaves.
\item
  {\em semi}: semi-continuous triphone system using 768 full covariance
  Gaussians in 2500 tree leaves, but no smoothing (explanations below).
\item
  {\em 2lvl}: two-level tree based semi-continuous triphone system using
  4096 full covariance Gaussians in 208 codebooks ( min/average/max
  components),  tree leaves, $\rho = $ and $\varrho = $.
\item
  {\em sgmm}: subspace Gaussian mixture model triphone system using a 400 
  full-covariance Gaussian background model and 2361 tree leaves.
\end{itemize}

%----%<------------------------------------------------------------------------

\section{Results}

\subsection{RM}

\begin{table}[tb]
\begin{center}
%\footnotesize
%\begin{tabular}{|l||r|r|r|r|r|r||c|}
\begin{tabular}{|l||c|c|c|c|c|c||c|}
\hline
~               & ma87 & oc87 & fe89 & oc89 & fe91 & se92 & {\em avg}  \\ \hline\hline
%{\em mono/cont} &  6.24 &  8.72 &  7.85 &  9.80 &  8.21 & 13.60 & 9.50 \\ \hline
%{\em tri1/cont} &  0.96 &  2.76 &  2.69 &  3.61 &  3.30 &  6.33 & 3.64 \\ \hline\hline
{\em cont} &  1.08 &  2.48 &  2.69 &  3.46 &  2.66 &  5.90 & 3.38 \\ \hline
{\em semi} &  1.80 &  3.19 &  4.72 &  4.62 &  4.15 &  6.88 & 4.66 \\ \hline 
{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & 2.90 \\ \hline
{\em sgmm} &  0.48 &  2.20 &  2.62 &  2.50 &  1.93 &  5.12 & 2.78 \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:res_rm}
Detailed recognition results in \% WER for the six DARPA test sets using 
different acoustic models on the RM data set.
}
\end{table}

The results on the different test sets on the RM data are displayed in 
Tab.~\ref{tab:res_rm}. The classic semi-continuous system shows the worst 
performance of the four. Possible explanations are an insufficient number
of Gaussians paired overfitting of the weights to a little number of 
components. Unfortunately, this could not be eased with the interpolations
and larger numbers of Gaussians would unacceptably increase the computational 
load due to the full covariance matrices.

The two-level tree based multiple-codebook system performance lies within the 
continuous and SGMM system, which goes along with the modeling capabilities 
and the rather little training data. 
% TODO Korbinian: I have some numbere on diag vs. full; maybe not as a table
Interestingly, the full covariances make the difference -- the same system with
diagonal covariances performs significantly worse, but approximates the 
continuous performance when using the same amount of Gaussians.

\subsection{WSJ}

% TODO Korbinian: More expts are ongoing; these include the full training set
%                 and cmvn, and hopefully a good param combination for 2lvl...
\begin{table}[tb]
\begin{center}
\begin{tabular}{|l||c|c|c|c|c|c||c|}
\hline
~          & eval92 & eval93 & {\em avg} \\ \hline\hline
{\em cont} &  13.17 &  18.61 &     15.23 \\ \hline
{\em semi} &   \\ \hline 
% tri2-2lvl-208-3072-4000-0-0-0-0 !! no cmvn !!
{\em 2lvl} &  13.95 &  21.61 &     16.85 \\ \hline
{\em sgmm} &  10.76 &  17.82 &     13.44 \\ \hline
\end{tabular}
\end{center}
\caption{\label{tab:res_wsj}
Detailed recognition results in \% WER for the '92 and '93 test sets using 
different acoustic models on the WSJ data.
}
\end{table}

% TODO Korbinian: This needs some more text once the above table is filled
The results on the different test sets on the WSJ data are displayed in
Tab.~\ref{tab:res_rm}.

\section{Summary}
In this article, we evaluated semi-continuous and two-level tree based
semi-continuous acoustic models of the {\sc Kaldi} toolkit in the context of
both continuous and SGMM models.
%
While the first could not produce convincing results, the two-level tree 
based multiple-codebook architecture shows promising performance, especially for 
little training data.

Although the current performance is below the state-of-the-art, the
rather simple theory and computational complexity, paired with the possibility
of solely acoustic adaptation make two-level tree based semi-continuous acoustic
models an attractive alternative to low-resource applications -- both in terms
of computational power and training data.
%
The take home message from the experiments is that the choice of acoustic model
should be made based on a resource constraint (number of Gaussians, available
compute power), and that both continuous and semi-continuous models can do the job.

For future work, effect of adaptation and especially discriminative training
with respect to the codebook architecture need to be explored.
Furthermore, the SGMM framework can be extended to two-level architecture to 
reduce computational effort in both training and decoding.

%\section{REFERENCES}
%\label{sec:ref}

\bibliographystyle{IEEEbib}
\bibliography{refs-eig,refs}

\end{document}

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Adding icassp12 lattice paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@510 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

% Template for SLT-2006 paper; to be used with:
%          spconf.sty  - ICASSP/ICIP LaTeX style file, and
%          IEEEbib.bst - IEEE bibliography style file.
% --------------------------------------------------------------------------
\documentclass{article}
\usepackage{spconf,amsmath,amssymb,bm,epsfig}
\usepackage{color}
\usepackage{multirow}

\definecolor{MidGray}{rgb}{0.3, 0.3, 0.3}
\definecolor{Pink}{rgb}{1.0, 0.4, 0.4}
\definecolor{Red}{rgb}{0.9, 0.1, 0.1}

% Example definitions.
% --------------------

%%%%%%
% Begin Dan's math definitions.
%%%%%%
\def\Re{{\mathbb{R}}}
\def\x{{\mathbf x}}
\def\a{{\mathbf a}}
\def\U{{\mathbf U}}
\def\HCLG{{\mathit{HCLG}}}
\def\bDelta{{\mathbf \Delta}}
\def\V{{\mathbf V}}
\def\u{{\mathbf u}}
\def\s{{\mathbf s}}
\def\t{{\mathbf t}}
\def\G{{\mathbf G}}
\def\F{{\mathbf F}}
\def\r{{\mathbf r}}
\def\semicolonsep{\,;\,}
\def\tr{{\operatorname{tr}\,}}
\def\z{{\mathbf z}}
\def\y{{\mathbf y}}
\def\Y{{\mathbf Y}}
\def\Z{{\mathbf Z}}
\def\gammatilde{{\gamma}}  %% This was \tilde{\gamma} in the tutorial, mapping to \gamma2 to preserve the distinction in the latex source in case later I want to resurrect that.
\def\R{{\mathbf R}}
\def\D{{\mathbf D}}
\def\half{ {\textstyle {\frac{1}{2}}} }
\def\L{{\mathbf L}}
\def\P{{\mathbf P}}
\def\Q{{\mathbf Q}}
\def\B{{\mathbf B}}
\def\S{{\mathbf S}}
\def\T{{\mathbf T}}
\def\K{{\mathbf K}}
\def\H{{\mathbf H}}
\def\g{{\mathbf g}}
\def\I{{\mathbf I}}
\def\btheta{{\mathbf\theta}}
\def\calN{{\cal N}}
\def\calQ{{\cal Q}}
\def\diag{{\mathrm{diag}}}
\def\pre{{\mathrm{pre}}}
\def\inv{{\mathrm{inv}}}
\def\calT{{\cal T}}
\def\bmu{{\bm \mu}} % mathbf does not work for lowercase greek letters!  But mathbf looks better than bm in most other cases so I only use bm here.
\def\bSigma{{\mathbf \Sigma}}
\def\M{{\mathbf M}}
\def\N{{\mathbf N}}
\def\bv{{\mathbf v}}
\def\w{{\mathbf w}}
\def\W{{\mathbf W}}
\def\A{{\mathbf A}}
\def\o{{\mathbf o}}
\def\b{{\mathbf b}}
\def\p{{\mathbf p}}
%%\newcommand\refeqn[1]{(\ref{#1})}

%%%%%%
% End Dan's math definitions.
%%%%%%

% Title.
% ------
\title{Generating exact lattices in a WFST framework}

\makeatletter
\def\name#1{\gdef\@name{#1\\}}
\makeatother
\name{ Daniel Povey$^1$, Mirko Hannemann$^2$,  \\
  {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^6$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
  {Petr Motl\'{i}\v{c}ek}$^7$, {Yanmin Qian}$^8$, {Ngoc Thang Vu}$^9$, {Korbinian Reidhammer}$^10$, {Karel Vesel\'{y}}$^2$
    \thanks{Thanks here}}
%\makeatletter
%\def\name#1{\gdef\@name{#1\\}}
%\makeatother

%%% TODO: fix thanks.
%\thanks{Arnab Ghoshal was supported by the European Community's Seventh Framework 
% Programme under grant agreement no. 213850 (SCALE); BUT researchers were partially 
%supported by Czech MPO project No. FR-TI1/034.}}

\address{$^1$ Microsoft Research, Redmond, WA, {\normalsize \tt dpovey@microsoft.com} \\
         $^2$ Brno University of Technology, Czech Republic, {\normalsize \tt \{karafiat,schwarzp\}@fit.vutbr.cz} \\
         $^3$ Saarland University, {\normalsize \tt arnab.ghoshal@lsv.uni-saarland.de} }

\begin{document}

\ninept  
%
\maketitle
%
\pagestyle{plain} % was set to empty in spconf.sty... this makes page number appear.

\begin{abstract}
We describe a lattice generation method that is exact, i.e. it satisfies all the 
natural properties we would want from a lattice of alternative transcriptions of
an utterance.  This method is also quite efficient; it does not 
introduce substantial overhead above one-best decoding.  Our method is
most directly applicable when using fully-expanded WFST decoders, i.e. decoders that use a
WFST-based decoding graph that is expanded down to the HMM-state level.  It outputs
lattices that include state-level alignments as well as word labels.
The general idea is to create a state-level lattice during decoding, and to do
a special form of determinization that retains only the best-scoring path
for each word sequence.
\end{abstract}

\begin{keywords}
  Speech Recognition, Lattice Generation
\end{keywords}

\section{Introduction}

The word ``lattice'' is used in the speech recognition literature to mean some
kind of compact representation of the most likely transcriptions of an utterance,
in the form of a graph structure and normally including score and alignment
information in addition to the word labels.  See 
for example~\cite{efficient_general,ney_word_graph,odell_thesis,saon2005anatomy}.

[more history + context here]

\section{Decoding with WFSTs}

The graph creation process we use in our toolkit, Kaldi~\cite{kaldi_paper},
is very close to the standard recipe described in~\cite{wfst},
where the Weighted Finite State Transducer (WFST) decoding graph is
\begin{equation}
  \HCLG = \min(\det(H \circ C \circ L \circ G)),
\end{equation}
where $\circ$ is WFST composisition (note: view $\HCLG$ as a single symbol).  
For concreteness we will speak of ``costs'' rather
than weights, where a cost is a floating point number that typically represents a negated
log-probability.  A WFST has a set of states with one distinguished
start state\footnote{This is the formulation that corresponds best with the toolkit we use.},
each state has a final-cost (or $\infty$ for non-final states);
and there is a set of arcs, where each arc $a$ has a previous-state $p[a]$, a next-state $n[a]$, a
weight $w[a]$ (just think of this as a cost for now), an input label $i[a]$ and an output
label $o[a]$.  In $\HCLG$, the input labels are the identifiers of context-dependent
HMM states, and the output labels represent words.  For both the input and output
symbols, the special label $\epsilon$ may appear meaning ``no label is present.''

\begin{figure}
\begin{center}
\includegraphics[height=2.1in,angle=270]{figures/acceptor_utterance.eps}
  \caption{Acceptor $U$ describing the acoustic scores of an utterance}
\label{fig:acceptor}
\end{center}
\end{figure}

Imagine we have an utterance, of length $T$, and want to ``decode'' it (i.e. find
the most likely word sequence, and the corresponding alignment).  A WFST
interpretation of the decoding problem is as follows.  We construct an
acceptor, or WFSA, as in Fig.~\ref{fig:acceptor} (an acceptor is represented as a
WFST with identical input and output symbols).  It has $T{+}1$ states,
and an arc for each combination of (time, context-dependent HMM state).  The
costs on these arcs are the negated and scaled acoustic log-likelihoods.
Call this acceptor $U$.  Define
\begin{equation}
   S \equiv U \circ \HCLG,
\end{equation}
which is the {\em search graph} for this utterance.  It has approximately $T+1$ times
more states than $\HCLG$ itself.  The decoding problem is equivalent to finding
the best path through $S$.  The input symbol sequence for this best path represents
the state-level alignment, and the output symbol sequence is the corresponding
sentence.  In practice we do not do a full search of $S$, but use beam pruning.
Let $B$ be the searched subset of $S$, containing a subset of the states and arcs
of $S$ obtained by some heuristic pruning procedure.  Then in practice, 
when we do Viterbi decoding with beam-pruning, we are finding the best path through $B$.

Since the beam pruning is a part of any practical search procedure and cannot
easily be avoided, we will define the desired outcome of lattice generation in terms
of the visited subset $B$ of the search graph $S$.

\section{Defining lattices and the lattice generation problem}

There is no generally accepted single definition of a lattice.  In~\cite{efficient_general}
and~\cite{sak2010fly}, it is defined as a labeled, weighted, directed acyclic graph
(i.e. a WFSA, with word labels).  In~\cite{ney_word_graph}, time information
is also included.  In the HTK lattice format~\cite{htkbook}, phone-level time alignments 
are also supported (along with separate language model, acoustic and pronunciation-probability 
scores), and in~\cite{saon2005anatomy}, state-level alignments are also produced.
In our work here we will be producing state-level alignments; in fact, the input-symbols
on our graph, which we call {\em transition-ids}, are slightly more fine-grained
than acoustic states and contain sufficient information to reconstruct the phone
sequence.  Hence, our lattices contain at least as much information as any prior work
we cite.

There is, as far as we know, no generally accepted problem statement
for lattice generation, but all the the authors we cited seem to
be concerned with the accuracy of the information in the lattice (e.g. that the
scores and alignments are correct) and the completeness of such information (e.g.
that no high-scoring word-sequences are missing).  The simplest
way to formalize these concerns is to express them in terms of a lattice 
pruning beam $\alpha > 0$ (interpret $\alpha$ as a natural logarithm).
\begin{itemize}
  \item The lattice should have a path for every word sequence within $\alpha$ of the best-scoring one.
  \item The scores and alignments in the lattice should be accurate.
  \item The lattice should not contain duplicate paths with the same word sequence.
\end{itemize}
Actually, this definition is not completely formalized-- we have left
the second condition (accuracy of scores and alignments) a little vague.
Let the lattice be $L$.  The way we would like to state this requirement is:
\begin{itemize}
  \item For every path in $L$, the score and alignment corresponds to 
   the best-scoring path in in $B$ for the corresponding word
   sequence\footnote{Or one of the best-scoring paths, in case of a tie}.
\end{itemize}
The way we actually have to state the requirement in order to get an efficient procedure is:
\begin{itemize}
  \item For every word-sequence in $B$ within $\alpha$ of the best one, the score and alignment
    for the corresponding path in $L$ is accurate.
  \item All scores and alignments in $L$ correspond to actual paths through $B$ (but not always
   necessarily the best ones).
\end{itemize}
The issue is that we want to be able to prune $B$ before generating a lattice from it,
but doing so could cause paths not within $\alpha$ of the best one to be lost, so we have
to weaken the condition.  In fact this weakening is of very little consequence, since regardless
of pruning, any word-sequence not within $\alpha$ of the best one could be omitted altogether, 
which is the same as being assigned a cost of $\infty$).

We note that by ``word-sequence'' we mean a sequence of whatever symbols are on the
output of $\HCLG$.  In our experiments these symbols represent words, but not including
silence, which we represent via alternative paths in $L$.

\section{Previous lattice generation methods}

Lattice generation algorithms tend to be closely linked to a particular type of decoder,
but are often justified by the same kinds of ideas.
A common assumption underlying lattice generation methods is the {\em word-pair assumption}
of~\cite{ney_word_graph}.  This is the notion that the time boundary between a pair of words
is not affected by the identity of any earlier words.  In a decoder in which there is
a different copy of the lexical tree for each preceding word, this assumption is automatically
satisfied, and if the word-pair assumption holds, in order to generate an accurate lattice it is
sufficient to store a single Viterbi back-pointer at the word level; the entire set of
such back-pointers contains enough information to generate the lattice.  Authors who have used
this type of lattice generation method~\cite{ney_word_graph,odell_thesis} have generally
not been able to evaluate how correct the word-pair assumption is in practice, but it seems
unlikely to cause problems.  Such methods are not applicable for us anyway, as we
use a WFST based decoder in which each copy of the lexical tree does not have a unique
one-word history.

The lattice generation method described in~\cite{efficient_general} 
is applicable to decoders that use WFSTs~\cite{wfst}
expanded down to the $C$ level (i.e. $CLG$), so the input symbols represent
context-dependent phones.  In their WFST based decoding networks, states did
not have a unique one-word history, but they were able to satisfy a similar 
condition at the phone level.  Their method was
to store a single Viterbi back-pointer at the phone level; use this to 
create a phone-level latice; prune the resulting
lattice; project it to leave only word labels; and then convert this into a word-level
lattice via epsilon removal and determinization.
Note that the form of pruning referred to here is not the same as beam pruning as
it takes account of both the forward and backward parts of the cost.
The paper also reported experiments with a different method that did not requir any
phone-pair assumption; these experiments showed that the more
efficient method that relied on the phone-pair assumption 
had a minimal impact on the lattice oracle error rate.  However, the 
experiments did not evaluate how much impact the assumption had on the accuracy of the scores, 
and this information could be important in some applications.

The lattice generation algorithm that was described in~\cite{saon2005anatomy} (although not
in very great detail) is applicable to WFSTs expanded down to the $H$ level (i.e. $\HCLG$),
so the input symbols represent context-dependent states.  It keeps both score and 
state-level alignment information.  In some sense this algorithm also relies
on the word-pair assumption, but since the copies of the lexical tree in the decoding
graph do not have unique word histories, the resulting algorithm has to be quite
different.  Viterbi back-pointers at the word level are used, but in each state
is stored not just the best back-pointer but the $N$ best back-pointers for the $N$ top-scoring distinct
preceding.  Therefore, this algorithm has more in common with the sentence N-best
algorithm than with the Viterbi algorithm.  By limiting $N$ to be quite small (e.g. $N{=}5$)
the algorithm was made efficient, but this is at the cost of possibly losing word sequences
that would be within the lattice-generation beam.  In addition, this algorithm was quite
complex to implement efficiently.

\section{Overview of our algorithm}

\subsection{Version without alignments}

In order to explain our algorithm in the easiest way, we will first explain how it would
be if we did not keep the alignment information, and were storing only a single cost
(i.e. the total acoustic plus language-model cost).  This is just for didactic purposes;
in fact, we have not implemented this simple version.
In this case our algorithm would be quite similar
to~\cite{efficient_general}, except working at the state level rather than the phone level.
We actually store forward rather than backward pointers: for each active state on each
frame, we create a forward link record for each active arc out of that state; this points
to the record for the destination state of the arc on the next frame (or on 
the current frame, for $\epsilon$-input
arcs).   As in~\cite{efficient_general}, at the end of the utterance
we prune the resulting graph to discard any paths that are not within the beam $\alpha$ of the best cost. 
Let the pruned graph be $P$, i.e.
\begin{equation}
  P = \mathrm{prune}(B, \alpha),
\end{equation}
where $B$ is the un-pruned state-level lattice.
We project on the output labels (keeping only the word labels), and remove $\epsilon$
labels and determinize.  In fact, we use a determinization algorithm that does $\epsilon$
removal itself.

As in~\cite{efficient_general}, to save memory we actually do the pruning
periodically rather than waiting for the end of the file (we do it every 25 frames).
Our method is equivalent to their method of linking all currently active states to a ``dummy''
final state and then pruning in the normal way.  However, we implement it in such a way
that the pruning algorithm does not always have to go back to the beginning of the utterance.
For each still-active state, we store the cost difference between the best path including that
state, and the best overall path.  This quantity does not always change between different
iterations of calling the pruning algorithm, and when we detect that these quantities are 
unchanged for a particular frame, the pruning algorithm can stop going backward in time.

After the determinization phase, we prune again using the beam $\alpha$.  This is needed because
the determinization process can introduce a lot of unlikely arcs.  In fact, for particular
utterances the determinization process can cause the lattice to expand enough to
exhaust memory.  To deal with this, we currently just detect when determinization
has produced more than a pre-set maximum number of states, then we prune with a tighter
beam and try again.  In future we may try more sophisticated methods such as a
determinization algorithm that does pruning itself.

This ``simple'' version of the algorithm produces an acyclic, deterministic
WFSA with words as labels.  This is sufficient for applications such as language-model
rescoring.

\subsection{Keeping separate graph and acoustic costs}

A fairly trivial extension of the algorithm described above is to store separately
the acoustic costs and the costs arising from $\HCLG$.  This enables us to do things
like generating output from the lattice with different acoustic scaling factors.
We refer to these two costs as the graph cost and the acoustic cost, since the cost
in $\HCLG$ is not just the language model cost but also contains components
arising from transition probabilities and pronunciation probabilities.  We implement
this by using a semiring that contains two real numbers (one for the graph and one
for the acoustic costs); it keeps track of the two costs separately, but its
$\oplus$ operation returns whichever pair has the lowest sum of costs (graph plus acoustic).

Formally, if each weight is a pair $(a,b)$, then $(a,b) \otimes (c,d) = (a{+}c, b{+}d)$,
and $(a,b) \oplus (c,d)$ is equal to $(a,b)$ if $a{+}b < c{+}d$ or if $a{+}b = c{f}+d$ and $a{-}b < c{-}d$,
and otherwise is equal to $(c,d)$.  This is equivalent to the lexicographic semiring 
of~\cite{roark2011lexicographic}, on the pair $((a{+}b),(a{-}b))$.

\subsection{Keeping state-level alignments}

It is useful for various purposes, e.g. discriminative training and certain kinds of
acoustic rescoring, to keep the state-level alignments in the lattices.  We will now
explain how we can make the alignments ``piggyback'' on top of the computation
defined above, by encoding them in a special semiring.

First, let us define $Q = \inv(P)$, i.e. $Q$ is the inverted, pruned state-level lattice,
so that the input symbols are the words and the output symbols are the p.d.f. labels.
We want to process $Q$ in such a way that we keep only the best path through it
for each word sequence, and get the corresponding alignment.  This is possible
by defining an appropriate semiring and then doing normal determinization.  Firstly, 
let us ignore the fact that we are keeping track of separate graph and acoustic costs, 
to avoid complicating the present discussion.

We will define a semiring in which symbol sequences are encoded into the weights.
Let a weight be a pair $(c, s)$, where $c$ is a cost and $s$ is a sequence of symbols.
We define the $\otimes$ operation as $(c, s) \otimes (c', s') = (c+c', (s,s'))$, where
$(s,s')$ is $s$ and $s'$ appended together.  We define the $\oplus$ operation so that
it returns whichever pair has the smallest weight: that is, $(c,s) \oplus (c',s')$ 
equals $(c,s)$ if $c < c'$, and $(c',s')$ if $c > c'$.  If the weights are identical,
we cannot arbitrarily return the first pair because this would not satisfy the semiring
axioms.  In this case, we return the pair with the shorter string part, and if
the lengths are the same, whichever string appears first in dictionary order.  These
details do not have any practical effect on the behavior of the final algorithm in any
normal situation, but are required to satisfy the semiring axioms.

Suppose we have a WFST $Q$ representing a state-level lattice, with words
as its input symbols and p.d.f.'s as its output symbols.  Let $E$ be an encoding of $Q$ with the 
same number of states and arcs; $E$ is an acceptor, with its symbols
equal to the input symbol on the corresponding arc of $Q$, and
the weights on the arcs of $E$ containing both the the weight and the output symbol
on the corresponding arcs of $Q$.  Let $D = \mathrm{det}(\mathrm{rmeps}(E))$.

In the case we are interested in, determinization
will always succeed because $E$ is acyclic (as long as the original decoding
graph $\HCLG$ has no $\epsilon$-input cycles).  Because $D$ is deterministic
and $\epsilon$-free, it has only one path for each word sequence. 
Determinization preserves equivalence,
and equivalence is defined in such a way that the $\oplus$-sum of the
weights of all the paths through $E$ with a particular word-sequence, must be the same
as the weight of the corresponding path through $D$ with that word-sequence.
It is clear from the definition of $\oplus$ that this path through
$D$ has the cost and alignment of the lowest-cost path through $E$ that has the
same word-sequence on it.

\subsection{Summary of our algorithm}

We now summarize the lattice generation algorithm.  During decoding, we create a data-structure
corresponding to a full state-level lattice.  That is, for every arc of $\HCLG$ we traverse
on every frame, we create a separate arc in the state-level lattice.  These arcs
contain the acoustic and graph costs separately.  We prune the state-level graph using
a beam $\alpha$; we do this periodically (every 25 frames) but this is equivalent to
doing it just once at the end (similar to~\cite{efficient_general}).  Let the pruned
state-level lattice that we get after pruning it at the end of the utterance be $P$.

Let $Q = \inv(P)$, and let $E$ be an encoded version of $Q$ as described above (with the
state labels as part of the weights).  The final lattice we output is
\begin{equation}
   L = \mathrm{prune}(\mathrm{det}(\mathrm{rmeps}(E)), \alpha) .
\end{equation}
The determinization and epsilon removal are done together by a single algorithm
that we will describe below.

The resulting lattice $L$ is a deterministic, acyclic weighted acceptor with the 
words as the labels, and the graph and acoustic costs and the alignments
encoded into the weights.  Of course, the costs and alignments are not in any
sense ``synchronized'' with the words.

\section{Details of our implementation}

\subsection{Epsilon removal and determinization}

We implemented $\epsilon$ removal and determinization as a single algorithm 
because $\epsilon$-removal using the traditional approach would greatly
increase the size of the state-level lattice (this is mentioned 
in~\cite{efficient_general}).  Rather than using completely generic
code as is usually done in OpenFst~\cite{openfst} which our code is based on,
we implemented it in a fairly specialized way for this case.  The issue
is that the determinization process often has to append a single symbol to
a string of symbols (for the state alignments), and the easiest way to
do this in ``generic'' code would involve copying the whole sequence each time.
Instead we use a data structure that enables this to be done in linear
time (it involves a hash table).

We will briefly describe another unique aspect of our
algorithm.  Determinization algorithms involve weighted subsets of states,
of the form
\begin{equation}
   S = \{ (s_1, w_1), (s_2, w_2), \ldots \} .
\end{equation}
Let this weighted subset, as it appears in a conventional determinization
algorithm with epsilon removal, be the {\em canonical representation} of a state.
A typical determinization algorithm would maintain a map from this representation
to a state index.  We define a {\em minimial representation} of a state
to be like the canonical representation, but only keeping states that
are either final, or have non-$\epsilon$ arcs out of them.  We maintain
a map from the minimal representation to the state index.  We can
show that this algorithm is still correct (it will tend to give output
with fewer states, i.e. more minimal).  As an optimization for speed, we also define
the {\em initial representation} to be the same type of subset, but prior
to following through the $\epsilon$ arcs, i.e. it only contains the states
that we reached by following non-$\epsilon$ arcs from a previous determinized
state.  We maintain a separate map from the initial representation to
the output state index; think of this as a ``lookaside buffer'' that helps us
avoid the expense of following $\epsilon$ arcs.  Note that formally
proving the correctness of this algorithm would be a big project, but we
are able to test correctness in a randomized way; this is easy because we
can test WFSA equivalence and we can test whether a WFSA is deterministic.

We should note that although the determinization algorithm
is guaranteed to terminate (since the state-level lattice is acyclic), it can
sometimes exhaust memory, i.e. in practice it can ``blow up'' for particular 
inputs.  Our observation is that in cases where it significantly expands
the input, this tends to be reversed by pruning after determinization.
We currently make our software robust to this phenomenon by 
pre-specifyin a maximum number of states for a lattice, and if the
determinization exceeds this threshold, we prune with a tighter beam
and try again.  In the future we may implement a form of determinization that
includes an efficient form of pruning.

%% The algorithm for $\epsilon$-removal and determinization is also optimized
%% for our particular case.  We will first explain what a generic $\epsilon$-removal
%% and determinization algorithm would look like.  We will explain this using
%% generic weights.  Let us use the notation
%% \begin{equation}
%%    S = \{ (s_1, w_1), (s_2, w_2), \ldots \}
%% \end{equation}
%% for a weighted subset of states in the original FST.  Each subset $S$ corresponds
%% to a state in the determinized FST, i.e. we maintain a map $m(S) \rightarrow s$
%% where $s$ is a state number in the output (in practice there are tolerances
%% involved in this map when the weights contain floating-point numbers).  We need
%% to normalize these subsets to remove any ``common part'' of the weights.  We
%% need to define an associative and commutative ``common divisor'' operation 
%% on weights (call this $\boxplus$; it can be the same as $\oplus$ for many semirings),
%% that gives us the normalizer for a pair of weights.
%% We must be able to  left-divide each individual weight by the result of this operation, i.e.
%% if $c = a \boxplus b$, then we need to be able to solve $a = c a'$ and $b = c b'$ for $a'$ and $b'$.
%% Here, $a'$ and $b'$ become the ``normalized weights''.  Also $\otimes$-multiplication
%% must be left-distributive over this operation, which ensures the right kind of
%% ``uniqueness'' of the normalization operation.  In the particular case of interest
%% $w$ is a pair $(w', s)$ of a ``base-weight'' and a string $s$, and the 
%% operation $(a,s) \boxplus (b,t)$ returns $(a\oplus b, u)$ where $u$ is the longest
%% common prefix of $s$ and $t$ and $a\oplus b$ essentially returns the minimum cost.

%% For the determinization algorithm, we define
%% $\mathrm{normalizer}(S)$ to be the $\boxplus$-sum of the weights in the subset $S$,
%% and $\mathrm{normalize}(S)$ to be $S$ with each weight left-divided by 
%% $\mathrm{normalizer}(S)$.  We also need to define an operation that ``follows through''
%% $\epsilon$ transitions.  Call this $\mathrm{follow-eps}(S)$

%% The key characteristic of our method is that unlike all previously reported 
%% methods that we are aware of (with the sole exception of the inefficient method used 
%% as a baseline in~\cite{efficient_general}), our method is exact even if the word-pair 
%% assumption is invalid.  In addition, we are able to generate a lattice with full
%% state-level alignment information, which would not be possible using the approaches
%% used in~\cite{efficient_general}.  Our algorithm also does not need to store $N$
%% back-pointers like the lattice-generation algorithm previously reported in~\cite{saon2005anatomy}
%% for $H$-level WFSTs (i.e. fully-expanded WFSTs).

\section{Experimental results}
\label{sec:exp}

We report experimental results on the Wall Street Journal database of
read speech.  
Our system is a standard mixture-of-Gaussians system trained on the SI-284
training data; we test on the November 1992 evaluation data.  
For these experiments we generate lattices with the bigram language model
supplied with the WSJ database, and rescore them with the trigram language model.

We report experiments on CallHome English and Switchboard.

Our Callhome English setup is as described in~\cite{ubmconf,ubmpaper}.  
We used PLP features with ceptral mean and variance normalization.  We tested 
with the trigram LM built as described in~\cite{ubmpaper}.

%% [dan's results, not included, used arnab's.]
%%svatava:er: showwer_nist test_arnab_nosmooth_floor02_vecs_rerun/*/*/*0.0.ctm.filt.dtl
%%3/4:48.7% 3/8:48.3% 4/4:47.9% 4/8:47.5% 5/4:47.3% 5/8:47.2% 6/4:47.2% 6/8:46.9% 7/4:47.0% 7/8:46.8% 8/4:46.6% 8/8:46.5%
%%svatava:er: showwer_nist test_symmetric_rerun/*/*/*0.0.ctm.filt.dtl
%%3/4:47.3% 3/8:47.0% 4/4:46.5% 4/8:46.4% 5/4:46.3% 5/8:46.1% 6/4:45.8% 6/8:45.8% 7/4:46.9% 7/8:45.0% 8/4:45.0% 8/8:45.1%

%%3/4:47.3% 3/8:47.1% 4/4:46.6% 4/8:46.3% 5/4:46.3% 5/8:46.1% 6/4:45.8% 6/8:45.9%

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|rcccccc|}           \hline 
  GMM:            &  \multicolumn{6}{c|}{    52.5                          }   \\ \hline  % from csl paper
                  &   \multicolumn{6}{c|}{ \#Substates    }                \\
                  &  2700  &  4k    &  6k  & 9k   &  12k   &  16k    \\  \hline
% in /homes/eva/q/qghoshal/work/a3.0
% showwer_nist 05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
  SGMM:             & 48.8   & 48.2 & 48.0  & 47.7  &{\bf 47.4}   &  47.5  \\
% in /homes/eva/q/qghoshal/work/a3.1
% showwer_nist t05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
  +spk-vecs:        & 47.6   & 47.0  & 46.4  & 46.4  & 46.1  &{\bf  45.9} \\
% showwer_nist 05_test_sgmm+spkvecs+wts/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
  +symmetric,exact: & 46.3    & 45.6 &  45.2  & 44.8  & 44.5 &{\bf 44.4}  \\ 
% ./bin/showwer_nist
%05_test_sgmm+spkvecs+wts-inexact/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
%3/8:46.5% 4/8:45.6% 5/8:45.0% 6/8:44.6% 7/8:44.4%
+symmetric,inexact & 46.5    & 45.6  &  45.0  & 44.6  & {\bf 44.4}&           \\ \hline
  \end{tabular}
   \caption{CallHome English: WERs without CMLLR adaptation }
% \vspace*{-0.2in}
 \label{tab:callhome:nocmllr}
\end{table}

Table~\ref{tab:callhome:nocmllr} shows experiments without CMLLR adaptation; the only
normalization is cepstral mean and variance normalization.  Using the symmetric
model reduced WER from 45.9\% to 44.4\%, a 1.5\% absolute improvement.   The inexact
update gave the same improvement as the exact update.

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|l|cccccc|}           \hline 
  GMM:            &  \multicolumn{6}{c|}{    49.7                             }   \\ 
  +SAT:           &  \multicolumn{6}{c|}{    46.0                             }   \\ \hline 
                  &   \multicolumn{6}{c|}{ \#Substates    }                \\
                  &  2700  &  4k    &  6k   & 9k     &  12k       &  16k    \\  \cline{2-7}
% in /homes/eva/q/qghoshal/work/a3.0
% showwer_nist 05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
  SGMM+spk-vecs:  & 46.5    & 45.5  & 45.2   & 45.4  & 44.8        & {\bf 44.7} \\
 +symmetric,exact & 44.9    & 44.4  &  44.1  & 43.2  & {\bf 42.8}  & 42.9      \\ 
%./bin/showwer_nist
%06_test_cmllr+spkvecs+wts-inexact/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
+symmetric,inexact& 45.2    & 44.1  &  43.5  &  43.4  & {\bf 43.3} &           \\ \hline
  \end{tabular}
   \caption{CallHome English: WERs with CMLLR adaptation}
% \vspace*{-0.2in}
 \label{tab:callhome:cmllr}
\end{table}

Table~\ref{tab:callhome:cmllr} shows experiments with CMLLR adaptation.  The exact
update gives 1.9\% absolute improvement and the inexact update gives 1.4\% absolute
improvement.  Note that these are the same models as the previous table, tested with
CMLLR, and we attribute the difference between the exact and inexact models on this 
setup to statistical noise; further experiments will have to tell us whether, in general,
there is a difference between the exact and inexact updates.

%%Note that Speaker Adaptive Training (SAT) gives a large improvement for the
%%baseline GMM-based model, but we do not do this for our model: as reported
%%in~\cite{ubmpaper}, we did not get an improvement from SAT with the SGMM; we
%%are unsure why SAT gave an unusually large gain for the baseline model.

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|l|ccccccc|}           \hline 
 \multirow{2}{*}{GMM}
            &     \multicolumn{7}{c|}{\#Gauss per state}     \\
            &   20  &  26    &   32  &  34    &   36    &   38   & 40   \\  \cline{2-8}
 -          &       & 36.8   & 36.6  & 36.4   &  36.4   &  36.4  & 36.4  \\  
 CMLLR      &       & 34.8   & 34.5  & 34.4   & 34.3    & 34.5   & 34.3  \\
 STC        &       & 35.4   & 35.3  &        &  35.2   &        &      \\
    +CMLLR  &       & 33.1   & 32.9  &        &  32.9   &        &      \\ \hline
 \multirow{2}{*}{SGMM} 
            &    \multicolumn{7}{c|}{\#Substates}                      \\
            & 30k   &  40k  &   50k  &  75k   &  100k  &  150k  & 200k \\  \cline{2-8}
 unadapted  &  35.7 & 35.7  &  35.1  & 34.7   & 34.3   &  33.9  & 33.7  \\ 
  CMLLR     &       &       &  32.2  &        &        &        &       \\
 +spk-vecs  &  32.0 & 31.7  &  31.4  & 31.2   & 30.8   &        &       \\ 
\hspace{-0.08in}
 +symmetric\hspace{-0.1in} &  31.9& 31.7   & 31.3   & 31.0   & 30.6   &        &       \\ \hline
  \end{tabular}
   \caption{Switchboard: WERs, with VTLN}
% \vspace*{-0.2in}
 \label{tab:switchboard:vtln}
\end{table}

\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|l|cccccc|}           \hline 
 \multirow{2}{*}{GMM}
            &     \multicolumn{6}{c|}{\#Gauss per state}     \\
            &       &        &       &        &   36    &         \\  \cline{2-7}
 -          &       &        &       &        &  39.2   &        \\  
 CMLLR      &       &        &       &        &  37.0   &        \\
 STC        &       &        &       &        &  38.0   &        \\
 +CMLLR     &       &        &       &        &  35.2   &         \\ \hline
 \multirow{2}{*}{SGMM} 
            &    \multicolumn{6}{c|}{\#Substates}                      \\
            & 30k   &  40k  &   50k  &  75k   &  100k  &         \\  \cline{2-7}
 unadapted  &  37.9 & 37.5  &  37.1  & 36.6   &  36.3  &          \\ \hline
CMLLR+spk-vecs&33.9 & 33.5  &  33.4  &        &        &          \\
 +symmetric & 33.8  & 33.0  &  33.2  &        &        &          \\ \hline
  \end{tabular}
   \caption{Switchboard: WERs, no VTLN}
% \vspace*{-0.2in}
 \label{tab:switchboard:novtln}
\end{table}

%% \begin{table}[h]
%%  \renewcommand\arraycolsep{-0.1in}
%%   \begin{tabular}{|l|ccccccc|}           \hline 
%%             &     \multicolumn{7}{c|}{\#Gauss per state}     \\
%%             &  16   &   26   &  32   &   36   & 40       &      &    \\  \cline{2-8}
%%   GMM:      &  37.8 & 36.8   & 36.6  & 36.4   &  36.4    &      &     \\  
%% +STC:       &       &        & 35.6  &        &          &       &     \\ \hline
%%             &    \multicolumn{7}{c|}{\#Substates}                      \\
%%             & 30k   &  40k  &   50k  &  75k   &  100k  &  150k  & 200k \\  \cline{2-8}
%%   SGMM:     &  35.7 & 35.7  &  35.1  & 34.7   & 34.3   &  33.9  & 33.7  \\ \hline
%%   \end{tabular}
%%    \caption{Switchboard: WERs, unadapted (except for VTLN)}
%% % \vspace*{-0.2in}
%%  \label{tab:switchboard:unadapted}
%% \end{table}

Next we discuss Switchboard experiments.  Our Switchboard system was trained on
278 hours of data from Switchboard I and II, and CallHome English.  Models were
tested on the Hub5 Eval01 test set (just over 6 hours long).  We used PLP
features with cepstral mean and variance normalization, and Vocal Tract Length
Normalization (VTLN).  The bigram language model used during decoding was taken
from the AMI RT'07 system described in~\cite{NIST:2007}; we used a recognition
lexicon of 50K words.  Our baseline GMM models were built with
HTK~\cite{htkbook}.  Tables~\ref{tab:switchboard:vtln}
and~\ref{tab:switchboard:novtln} show results with VTLN, and without VTLN,
respectively.  We did the baseline experiments with Constrained MLLR (CMLLR; a.k.a. fMLLR), 
and Semi-tied Covariance (STC; a.k.a. MLLT).  With the SGMMs, 
we used the exact update for the $\u_i$ quantities in the symmetric case.  
In both cases the symmetric extension to the model gives a much smaller
improvement than on the CallHome setup.  We are not sure of the reason for this.
Note that we do not show Speaker Adapted Training (SAT) results for the GMM baseline
because we did not see an improvement (we tried SAT after STC which would anyway
reduce the gains).

%%Table~\ref{tab:switchboard:unadapted}
%%shows experiments without adaptation except for VTLN (i.e. no ``speaker vectors'' or
%%CMLLR).  The GMM line with ``+STC'' includes a global semi-tied covariance transform (STC),
%%also known as a Maximum Likelihood Linear Transform (MLLT).
%%The SGMM is 1.9\% absolute (5.3\% relative) better than the best GMM baseline.

%% \begin{table}[h]
%%  \renewcommand\arraycolsep{-0.1in}
%%   \begin{tabular}{|l|ccccc|}           \hline 
%%                 &     \multicolumn{5}{c|}{\#Gauss per state}     \\
%%                 &  16   &   26   &  32   &   36   &   40       \\  \cline{2-6}
%%       GMM:      &       & 34.8   & 34.5  & 34.3   & 34.3       \\  \hline
%%      +STC:      &       &        & 34.1  &        &           \\  \hline
%%                 &    \multicolumn{5}{c|}{\#Substates}                      \\
%%                 & 30k   &  40k  &   50k  &  75k   &  100k   \\  \cline{2-6}
%%  SGMM+spk-vecs: &  33.0 & 32.5  &  32.0  & 31.7   & 31.4     \\ 
%%  +symmetric:    &  32.9 & 32.3  &  31.9  & 31.7   & 31.3     \\ \hline
%%   \end{tabular}
%%    \caption{Switchboard: WERs, adapted (all with VTLN+CMLLR)}
%% % \vspace*{-0.2in}
%%  \label{tab:switchboard:adapted}
%% \end{table}

%%Table~\ref{tab:switchboard:adapted} shows adapted results on Switchboard; these all include
%%CMLLR, and the SGMM systems also use speaker vectors.  Comparing the best GMM number with the best SGMM
%%number (with the symmetric model), we get 2.8\% absolute improvement (8.2\% relative).  Based on
%%other experiments on a slightly different setup, we believe that Speaker Adapative Training would
%%improve the baseline by about 0.3\% absolute.

%% The unexpected thing in Table~\ref{tab:switchboard:adapted} is that the symmetric modification gave 
%% us almost no improvement.  Although it is common to get slightly different results on different
%% datasets, we are not accustomed to seeing such large differences.

Table~\ref{tab:likes} compares
the average acoustic likelihood in the three passes of decoding, which reveals the effect of the
symmetric modification on the likelihood.  The ``baseline'' rows are the SGMM with speaker vectors 
and CMLLR, but without the symmetric modification.  As expected, the likelihood before adaptation
is slightly worse (because there is a mismatch between the models, which were adaptively trained,
and the data), but it gets better after adaptation.  In both cases the likelihood improvement per
frame, after adaptation, was about 0.1 (in natural-logarithm units).  This makes it 
hard to interpret the differences in results between the CallHome and Switchboard setups, because
the effect on the likelihoods is so similar.  We intend to do further experiments on other
data-sets to find which results are more typical.

% see ~/er/RESULTS.icassp11 on BUT systems.
\begin{table}[h]
 \renewcommand\arraycolsep{-0.1in}
  \begin{tabular}{|cr|ccc|}           \hline 
           &              &   \multicolumn{3}{c|}{ Decoding pass }           \\
           &              &   1         &     2       &          3            \\  
           &              & {\footnotesize (no-adapt)}  &{\footnotesize  +spk-vecs}  & {\footnotesize +CMLLR}         \\  \hline
  Call     &{\footnotesize SGMM+spk-vecs}&  -65.44     &   -63.62    &  -62.56              \\ 
  Home     & {\footnotesize +symmetric}    &  -65.57     &   -63.50    &  -62.45            \\ \hline
  Switch-  & {\footnotesize SGMM+spk-vecs}&  -60.07     &  -57.78     &  -58.86             \\
  board    & {\footnotesize +symmetric}    &  -60.17     &  -57.68     &  -56.76            \\ \hline
  \end{tabular}
   \caption{Acoustic likelihoods on the three test-time decoding passes}
% \vspace*{-0.2in}
 \label{tab:likes}
\end{table}

%% Baseline:
%% merlin:er: grep Per test_arnab*rerun/8/8/decode_fixed_s10.0_p0.0/*.log  | perl -e ' $n=0;while(<>) { @A=split(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/\
%% $tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -65.441097461703 over 655612 frames
%% pass 2: -63.6198843852782 over 655612 frames
%% pass 3: -62.5628536429138 over 655612 frames

%% Experiment:
%% merlin:er: grep Per test_symmetric_rerun/8/8/decode_fixed_s10.0_p0.0/*.log  | perl -e ' $n=0;while(<>) { @A=split(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{\
%% $n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -65.5657124102808 over 655612 frames
%% pass 2: -63.5015519090499 over 655612 frames
%% pass 3: -62.4537370786898 over 655612 frames
%%  Martin's setup:
%% cd /mnt/matylda3/karafiat/SGMM/exp/ctstrain04.v0.xwrd.sn.4.vtln

%% Baseline:
%% merlin:ctstrain04.v0.xwrd.sn.4.vtln: dir=output_sgmm_full_400gz_arnabcmllr39_icassp/hmm7.8.s50000; for x in $dir/asr_out*.log.gz; do gunzip -c $x ; done | grep Per  | perl -e ' $n=0;while(<>) { @A=split\
%% (`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -60.0675595633454 over 2228237 frames
%% pass 2: -57.7841323235589 over 2228237 frames
%% pass 3: -56.8589936376831 over 2228237 frames

%% Experiment:
%% merlin:ctstrain04.v0.xwrd.sn.4.vtln: dir=output_sgmm_full_400gz_arnabcmllr39_sym_icassp/hmm7.8.s50000; for x in $dir/asr_out*.log.gz; do gunzip -c $x ; done | grep Per  | perl -e ' $n=0;while(<>) { @A=s\
%% plit(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
%% pass 1: -60.1726961158082 over 2228237 frames
%% pass 2: -57.681438551532 over 2228237 frames
%% pass 3: -56.7608517686853 over 2228237 frames

%\vspace*{-0.075in}
\section{Conclusions}
\label{sec:conc}
%\vspace*{-0.05in}

We have described a modification to the Subspace Gaussian Mixture Model which we
call the Symmetric SGMM.  This is a very natural extension which removes an
asymmetry in the way the Gaussian mixture weights were previously computed.  The
extra computation is minimal but the memory used for the acoustic model is nearly
doubled.  Our experimental results were inconsistent: on one setup we got a large
improvement of 1.5\% absolute, and on another setup it was much smaller.

We would also like to report our progress on releasing open-source software that
supports the SGMM modeling approach.  An official announcement, with additional
co-authors, will follow within the next year.  We are developing an open-source
(Apache-licensed) C++ speech recognition toolkit that uses the OpenFst
library~\cite{openfst}.  Most aspects of the toolkit are not related directly to
SGMMs, but SGMMs will be one of the acoustic models the toolkit natively
supports.  Most likely the toolkit will already have been released by the time this 
is published.

%\vspace*{-0.1in}
\bibliographystyle{IEEEbib}
\bibliography{refs}

\end{document}

steps/decode_tri3a_latgen.sh
acwt=0.0625
beam=[15.0-7.0]
max_active=15000
lat_beam=[10.0-1.0]
max_arcs=50000 (75000 for lat_beam 10.0)
model=exp/tri3a/final.mdl
graph=exp/graph_tri3a_bg_left/HCLG.fst

scp=data/eval_nov92.scp

333 utterances
87 utterances contain OOV
42 utterances contain annotated noise (actually some more, but listed as OOV utterances here)

5641 tokens (some scripts count 5700 with 59 noise annotations)
107 OOV tokens
1.88% OOV token rate

253204 frames (2532 seconds)

19976 words in decoding vocabulary and LM (not counting <s>,</s>,<unk>)

average lattice link density per frame:
have a wide range over files: for example beam=15.0, lat_beam=9.0:
ranging from 1.5 to 167

running on blades, using 64bit version with double float and lattice-simple-decoder

decoding with bigram LM:
decoding WER:      11.51%
ins/del/sub:       118/57/474   
utterances wrong:  227

experiment: changing lattice beam:

decoding beam:     15.0         15.0       15.0       15.0       15.0         15.0        15.0        15.0        15.0         15.0        15.0
lattice beam:      10.0          9.0        8.0        7.0        6.0          5.0         4.0         3.0         2.0          1.0         0.0
avg. link density: 23.56        16.0       10.82       7.26       4.91         3.38        2.38        1.77        1.4          1.17        1.0
realtime factor:    3.61         3.17       2.59       2.41       2.34         2.98 ?      3.21 ?      1.89        3.12 ?       1.78        --

oracle WER:         2.62%        2.78%      2.85%      3.08%      3.35%        3.86%       4.52%       5.46%       6.98%        8.62%      11.51%
ins/del/sub:       22/0/126     25/0/132   29/0/132   31/1/142   38/5/146     51/11/156   57/15/183   64/20/224   76/32/286    89/40/357   118/57/474
oracle utt wrong:  93           96         96         101        105          116         129         147         168          191         227

rescoring WER(15):  9.59%        9.59%      9.59%      9.59%      9.56%        9.59%       9.61%       9.70%       9.80%       10.21%      11.51%
ins/del/sub:       117/36/388                                    116/36/387   117/35/389  115/35/392  111/38/398  108/39/406   113/42/421  118/57/474
rescore utt wrong: 205                                           204          204         204         202         205          212         227

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Removed some unwanted stuff from lattice paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@511 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 
-We report experiments on CallHome English and Switchboard.
-
-Our Callhome English setup is as described in~\cite{ubmconf,ubmpaper}.  
-We used PLP features with ceptral mean and variance normalization.  We tested 
-with the trigram LM built as described in~\cite{ubmpaper}.
-
-%% [dan's results, not included, used arnab's.]
-%%svatava:er: showwer_nist test_arnab_nosmooth_floor02_vecs_rerun/*/*/*0.0.ctm.filt.dtl
-%%3/4:48.7% 3/8:48.3% 4/4:47.9% 4/8:47.5% 5/4:47.3% 5/8:47.2% 6/4:47.2% 6/8:46.9% 7/4:47.0% 7/8:46.8% 8/4:46.6% 8/8:46.5%
-%%svatava:er: showwer_nist test_symmetric_rerun/*/*/*0.0.ctm.filt.dtl
-%%3/4:47.3% 3/8:47.0% 4/4:46.5% 4/8:46.4% 5/4:46.3% 5/8:46.1% 6/4:45.8% 6/8:45.8% 7/4:46.9% 7/8:45.0% 8/4:45.0% 8/8:45.1%
-
-%%3/4:47.3% 3/8:47.1% 4/4:46.6% 4/8:46.3% 5/4:46.3% 5/8:46.1% 6/4:45.8% 6/8:45.9%
-
-\begin{table}[h]
- \renewcommand\arraycolsep{-0.1in}
-  \begin{tabular}{|rcccccc|}           \hline 
-  GMM:            &  \multicolumn{6}{c|}{    52.5                          }   \\ \hline  % from csl paper
-                  &   \multicolumn{6}{c|}{ \#Substates    }                \\
-                  &  2700  &  4k    &  6k  & 9k   &  12k   &  16k    \\  \hline
-% in /homes/eva/q/qghoshal/work/a3.0
-% showwer_nist 05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
-  SGMM:             & 48.8   & 48.2 & 48.0  & 47.7  &{\bf 47.4}   &  47.5  \\
-% in /homes/eva/q/qghoshal/work/a3.1
-% showwer_nist t05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
-  +spk-vecs:        & 47.6   & 47.0  & 46.4  & 46.4  & 46.1  &{\bf  45.9} \\
-% showwer_nist 05_test_sgmm+spkvecs+wts/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
-  +symmetric,exact: & 46.3    & 45.6 &  45.2  & 44.8  & 44.5 &{\bf 44.4}  \\ 
-% ./bin/showwer_nist
-%05_test_sgmm+spkvecs+wts-inexact/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
-%3/8:46.5% 4/8:45.6% 5/8:45.0% 6/8:44.6% 7/8:44.4%
-+symmetric,inexact & 46.5    & 45.6  &  45.0  & 44.6  & {\bf 44.4}&           \\ \hline
-  \end{tabular}
-   \caption{CallHome English: WERs without CMLLR adaptation }
-% \vspace*{-0.2in}
- \label{tab:callhome:nocmllr}
-\end{table}
-
-Table~\ref{tab:callhome:nocmllr} shows experiments without CMLLR adaptation; the only
-normalization is cepstral mean and variance normalization.  Using the symmetric
-model reduced WER from 45.9\% to 44.4\%, a 1.5\% absolute improvement.   The inexact
-update gave the same improvement as the exact update.
-
-\begin{table}[h]
- \renewcommand\arraycolsep{-0.1in}
-  \begin{tabular}{|l|cccccc|}           \hline 
-  GMM:            &  \multicolumn{6}{c|}{    49.7                             }   \\ 
-  +SAT:           &  \multicolumn{6}{c|}{    46.0                             }   \\ \hline 
-                  &   \multicolumn{6}{c|}{ \#Substates    }                \\
-                  &  2700  &  4k    &  6k   & 9k     &  12k       &  16k    \\  \cline{2-7}
-% in /homes/eva/q/qghoshal/work/a3.0
-% showwer_nist 05_test_sgmm_danupdate/*/*/decode_s10.0_p0.0_3gram.ctm.filt.dtl
-  SGMM+spk-vecs:  & 46.5    & 45.5  & 45.2   & 45.4  & 44.8        & {\bf 44.7} \\
- +symmetric,exact & 44.9    & 44.4  &  44.1  & 43.2  & {\bf 42.8}  & 42.9      \\ 
-%./bin/showwer_nist
-%06_test_cmllr+spkvecs+wts-inexact/*/*/decode_s10.0_p0.0_perspk_3gram.ctm.filt.dtl
-+symmetric,inexact& 45.2    & 44.1  &  43.5  &  43.4  & {\bf 43.3} &           \\ \hline
-  \end{tabular}
-   \caption{CallHome English: WERs with CMLLR adaptation}
-% \vspace*{-0.2in}
- \label{tab:callhome:cmllr}
-\end{table}
-
-Table~\ref{tab:callhome:cmllr} shows experiments with CMLLR adaptation.  The exact
-update gives 1.9\% absolute improvement and the inexact update gives 1.4\% absolute
-improvement.  Note that these are the same models as the previous table, tested with
-CMLLR, and we attribute the difference between the exact and inexact models on this 
-setup to statistical noise; further experiments will have to tell us whether, in general,
-there is a difference between the exact and inexact updates.
-
-%%Note that Speaker Adaptive Training (SAT) gives a large improvement for the
-%%baseline GMM-based model, but we do not do this for our model: as reported
-%%in~\cite{ubmpaper}, we did not get an improvement from SAT with the SGMM; we
-%%are unsure why SAT gave an unusually large gain for the baseline model.
-
-\begin{table}[h]
- \renewcommand\arraycolsep{-0.1in}
-  \begin{tabular}{|l|ccccccc|}           \hline 
- \multirow{2}{*}{GMM}
-            &     \multicolumn{7}{c|}{\#Gauss per state}     \\
-            &   20  &  26    &   32  &  34    &   36    &   38   & 40   \\  \cline{2-8}
- -          &       & 36.8   & 36.6  & 36.4   &  36.4   &  36.4  & 36.4  \\  
- CMLLR      &       & 34.8   & 34.5  & 34.4   & 34.3    & 34.5   & 34.3  \\
- STC        &       & 35.4   & 35.3  &        &  35.2   &        &      \\
-    +CMLLR  &       & 33.1   & 32.9  &        &  32.9   &        &      \\ \hline
- \multirow{2}{*}{SGMM} 
-            &    \multicolumn{7}{c|}{\#Substates}                      \\
-            & 30k   &  40k  &   50k  &  75k   &  100k  &  150k  & 200k \\  \cline{2-8}
- unadapted  &  35.7 & 35.7  &  35.1  & 34.7   & 34.3   &  33.9  & 33.7  \\ 
-  CMLLR     &       &       &  32.2  &        &        &        &       \\
- +spk-vecs  &  32.0 & 31.7  &  31.4  & 31.2   & 30.8   &        &       \\ 
-\hspace{-0.08in}
- +symmetric\hspace{-0.1in} &  31.9& 31.7   & 31.3   & 31.0   & 30.6   &        &       \\ \hline
-  \end{tabular}
-   \caption{Switchboard: WERs, with VTLN}
-% \vspace*{-0.2in}
- \label{tab:switchboard:vtln}
-\end{table}
-
-\begin{table}[h]
- \renewcommand\arraycolsep{-0.1in}
-  \begin{tabular}{|l|cccccc|}           \hline 
- \multirow{2}{*}{GMM}
-            &     \multicolumn{6}{c|}{\#Gauss per state}     \\
-            &       &        &       &        &   36    &         \\  \cline{2-7}
- -          &       &        &       &        &  39.2   &        \\  
- CMLLR      &       &        &       &        &  37.0   &        \\
- STC        &       &        &       &        &  38.0   &        \\
- +CMLLR     &       &        &       &        &  35.2   &         \\ \hline
- \multirow{2}{*}{SGMM} 
-            &    \multicolumn{6}{c|}{\#Substates}                      \\
-            & 30k   &  40k  &   50k  &  75k   &  100k  &         \\  \cline{2-7}
- unadapted  &  37.9 & 37.5  &  37.1  & 36.6   &  36.3  &          \\ \hline
-CMLLR+spk-vecs&33.9 & 33.5  &  33.4  &        &        &          \\
- +symmetric & 33.8  & 33.0  &  33.2  &        &        &          \\ \hline
-  \end{tabular}
-   \caption{Switchboard: WERs, no VTLN}
-% \vspace*{-0.2in}
- \label{tab:switchboard:novtln}
-\end{table}
-
-%% \begin{table}[h]
-%%  \renewcommand\arraycolsep{-0.1in}
-%%   \begin{tabular}{|l|ccccccc|}           \hline 
-%%             &     \multicolumn{7}{c|}{\#Gauss per state}     \\
-%%             &  16   &   26   &  32   &   36   & 40       &      &    \\  \cline{2-8}
-%%   GMM:      &  37.8 & 36.8   & 36.6  & 36.4   &  36.4    &      &     \\  
-%% +STC:       &       &        & 35.6  &        &          &       &     \\ \hline
-%%             &    \multicolumn{7}{c|}{\#Substates}                      \\
-%%             & 30k   &  40k  &   50k  &  75k   &  100k  &  150k  & 200k \\  \cline{2-8}
-%%   SGMM:     &  35.7 & 35.7  &  35.1  & 34.7   & 34.3   &  33.9  & 33.7  \\ \hline
-%%   \end{tabular}
-%%    \caption{Switchboard: WERs, unadapted (except for VTLN)}
-%% % \vspace*{-0.2in}
-%%  \label{tab:switchboard:unadapted}
-%% \end{table}
-
-Next we discuss Switchboard experiments.  Our Switchboard system was trained on
-278 hours of data from Switchboard I and II, and CallHome English.  Models were
-tested on the Hub5 Eval01 test set (just over 6 hours long).  We used PLP
-features with cepstral mean and variance normalization, and Vocal Tract Length
-Normalization (VTLN).  The bigram language model used during decoding was taken
-from the AMI RT'07 system described in~\cite{NIST:2007}; we used a recognition
-lexicon of 50K words.  Our baseline GMM models were built with
-HTK~\cite{htkbook}.  Tables~\ref{tab:switchboard:vtln}
-and~\ref{tab:switchboard:novtln} show results with VTLN, and without VTLN,
-respectively.  We did the baseline experiments with Constrained MLLR (CMLLR; a.k.a. fMLLR), 
-and Semi-tied Covariance (STC; a.k.a. MLLT).  With the SGMMs, 
-we used the exact update for the $\u_i$ quantities in the symmetric case.  
-In both cases the symmetric extension to the model gives a much smaller
-improvement than on the CallHome setup.  We are not sure of the reason for this.
-Note that we do not show Speaker Adapted Training (SAT) results for the GMM baseline
-because we did not see an improvement (we tried SAT after STC which would anyway
-reduce the gains).
-
-%%Table~\ref{tab:switchboard:unadapted}
-%%shows experiments without adaptation except for VTLN (i.e. no ``speaker vectors'' or
-%%CMLLR).  The GMM line with ``+STC'' includes a global semi-tied covariance transform (STC),
-%%also known as a Maximum Likelihood Linear Transform (MLLT).
-%%The SGMM is 1.9\% absolute (5.3\% relative) better than the best GMM baseline.
-
-%% \begin{table}[h]
-%%  \renewcommand\arraycolsep{-0.1in}
-%%   \begin{tabular}{|l|ccccc|}           \hline 
-%%                 &     \multicolumn{5}{c|}{\#Gauss per state}     \\
-%%                 &  16   &   26   &  32   &   36   &   40       \\  \cline{2-6}
-%%       GMM:      &       & 34.8   & 34.5  & 34.3   & 34.3       \\  \hline
-%%      +STC:      &       &        & 34.1  &        &           \\  \hline
-%%                 &    \multicolumn{5}{c|}{\#Substates}                      \\
-%%                 & 30k   &  40k  &   50k  &  75k   &  100k   \\  \cline{2-6}
-%%  SGMM+spk-vecs: &  33.0 & 32.5  &  32.0  & 31.7   & 31.4     \\ 
-%%  +symmetric:    &  32.9 & 32.3  &  31.9  & 31.7   & 31.3     \\ \hline
-%%   \end{tabular}
-%%    \caption{Switchboard: WERs, adapted (all with VTLN+CMLLR)}
-%% % \vspace*{-0.2in}
-%%  \label{tab:switchboard:adapted}
-%% \end{table}
-
-%%Table~\ref{tab:switchboard:adapted} shows adapted results on Switchboard; these all include
-%%CMLLR, and the SGMM systems also use speaker vectors.  Comparing the best GMM number with the best SGMM
-%%number (with the symmetric model), we get 2.8\% absolute improvement (8.2\% relative).  Based on
-%%other experiments on a slightly different setup, we believe that Speaker Adapative Training would
-%%improve the baseline by about 0.3\% absolute.
-
-%% The unexpected thing in Table~\ref{tab:switchboard:adapted} is that the symmetric modification gave 
-%% us almost no improvement.  Although it is common to get slightly different results on different
-%% datasets, we are not accustomed to seeing such large differences.
-
-Table~\ref{tab:likes} compares
-the average acoustic likelihood in the three passes of decoding, which reveals the effect of the
-symmetric modification on the likelihood.  The ``baseline'' rows are the SGMM with speaker vectors 
-and CMLLR, but without the symmetric modification.  As expected, the likelihood before adaptation
-is slightly worse (because there is a mismatch between the models, which were adaptively trained,
-and the data), but it gets better after adaptation.  In both cases the likelihood improvement per
-frame, after adaptation, was about 0.1 (in natural-logarithm units).  This makes it 
-hard to interpret the differences in results between the CallHome and Switchboard setups, because
-the effect on the likelihoods is so similar.  We intend to do further experiments on other
-data-sets to find which results are more typical.
-
-% see ~/er/RESULTS.icassp11 on BUT systems.
-\begin{table}[h]
- \renewcommand\arraycolsep{-0.1in}
-  \begin{tabular}{|cr|ccc|}           \hline 
-           &              &   \multicolumn{3}{c|}{ Decoding pass }           \\
-           &              &   1         &     2       &          3            \\  
-           &              & {\footnotesize (no-adapt)}  &{\footnotesize  +spk-vecs}  & {\footnotesize +CMLLR}         \\  \hline
-  Call     &{\footnotesize SGMM+spk-vecs}&  -65.44     &   -63.62    &  -62.56              \\ 
-  Home     & {\footnotesize +symmetric}    &  -65.57     &   -63.50    &  -62.45            \\ \hline
-  Switch-  & {\footnotesize SGMM+spk-vecs}&  -60.07     &  -57.78     &  -58.86             \\
-  board    & {\footnotesize +symmetric}    &  -60.17     &  -57.68     &  -56.76            \\ \hline
-  \end{tabular}
-   \caption{Acoustic likelihoods on the three test-time decoding passes}
-% \vspace*{-0.2in}
- \label{tab:likes}
-\end{table}
-
-%% Baseline:
-%% merlin:er: grep Per test_arnab*rerun/8/8/decode_fixed_s10.0_p0.0/*.log  | perl -e ' $n=0;while(<>) { @A=split(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/\
-%% $tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
-%% pass 1: -65.441097461703 over 655612 frames
-%% pass 2: -63.6198843852782 over 655612 frames
-%% pass 3: -62.5628536429138 over 655612 frames
-
-%% Experiment:
-%% merlin:er: grep Per test_symmetric_rerun/8/8/decode_fixed_s10.0_p0.0/*.log  | perl -e ' $n=0;while(<>) { @A=split(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{\
-%% $n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
-%% pass 1: -65.5657124102808 over 655612 frames
-%% pass 2: -63.5015519090499 over 655612 frames
-%% pass 3: -62.4537370786898 over 655612 frames
-%%  Martin's setup:
-%% cd /mnt/matylda3/karafiat/SGMM/exp/ctstrain04.v0.xwrd.sn.4.vtln
-
-%% Baseline:
-%% merlin:ctstrain04.v0.xwrd.sn.4.vtln: dir=output_sgmm_full_400gz_arnabcmllr39_icassp/hmm7.8.s50000; for x in $dir/asr_out*.log.gz; do gunzip -c $x ; done | grep Per  | perl -e ' $n=0;while(<>) { @A=split\
-%% (`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
-%% pass 1: -60.0675595633454 over 2228237 frames
-%% pass 2: -57.7841323235589 over 2228237 frames
-%% pass 3: -56.8589936376831 over 2228237 frames
-
-%% Experiment:
-%% merlin:ctstrain04.v0.xwrd.sn.4.vtln: dir=output_sgmm_full_400gz_arnabcmllr39_sym_icassp/hmm7.8.s50000; for x in $dir/asr_out*.log.gz; do gunzip -c $x ; done | grep Per  | perl -e ' $n=0;while(<>) { @A=s\
-%% plit(`` ``,$_); $n=($n%3)+1; $t{$n} += $A[4]; $tt{$n} += $A[6]; } for($n=1;$n<=3;$n++) { $f=$t{$n}/$tt{$n}; print ``pass $n: $f over $tt{$n} frames\n''; }'
-%% pass 1: -60.1726961158082 over 2228237 frames
-%% pass 2: -57.681438551532 over 2228237 frames
-%% pass 3: -56.7608517686853 over 2228237 frames
-
 %\vspace*{-0.075in}
@@
 
-We have described a modification to the Subspace Gaussian Mixture Model which we
-call the Symmetric SGMM.  This is a very natural extension which removes an
-asymmetry in the way the Gaussian mixture weights were previously computed.  The
-extra computation is minimal but the memory used for the acoustic model is nearly
-doubled.  Our experimental results were inconsistent: on one setup we got a large
-improvement of 1.5\% absolute, and on another setup it was much smaller.
-
-We would also like to report our progress on releasing open-source software that
-supports the SGMM modeling approach.  An official announcement, with additional
-co-authors, will follow within the next year.  We are developing an open-source
-(Apache-licensed) C++ speech recognition toolkit that uses the OpenFst
-library~\cite{openfst}.  Most aspects of the toolkit are not related directly to
-SGMMs, but SGMMs will be one of the acoustic models the toolkit natively
-supports.  Most likely the toolkit will already have been released by the time this 
-is published.
+Conclusions here
 

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Improved affiliations on paper
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@512 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 \name{ Daniel Povey$^1$, Mirko Hannemann$^2$,  \\
-  {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^6$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
-  {Petr Motl\'{i}\v{c}ek}$^7$, {Yanmin Qian}$^8$, {Ngoc Thang Vu}$^9$, {Korbinian Reidhammer}$^10$, {Karel Vesel\'{y}}$^2$
+  {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
+  {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Reidhammer}$^9$, {Karel Vesel\'{y}}$^2$
     \thanks{Thanks here}}
@@
 \address{$^1$ Microsoft Research, Redmond, WA, {\normalsize \tt dpovey@microsoft.com} \\
-         $^2$ Brno University of Technology, Czech Republic, {\normalsize \tt \{karafiat,schwarzp\}@fit.vutbr.cz} \\
-         $^3$ Saarland University, {\normalsize \tt arnab.ghoshal@lsv.uni-saarland.de} }
+         $^2$ Brno University of Technology, Czech Republic, {\normalsize \tt ihannema@fit.vutbr.cz} \\
+         $^3$ CRIM, Montreal, Canada \ \  $^4$ SRI International, Menlo Park, CA, USA \\
+         $^5$ University of Edinburgh, U.K. \ \  $^6$  IDIAP, Martigny, Switzerland  \\
+         $^6$ Shanghai Jiao Tong University, China \ \ $^7$ Karlsruhe Institute of Technology, Germany \\
+         $^9$ Friedrich Alexander University of Erlangen-Nurnberg, Germany}
 

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Minor modifications to paper
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@513 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 introduce substantial overhead above one-best decoding.  Our method is
-most directly applicable when using fully-expanded WFST decoders, i.e. decoders that use a
-WFST-based decoding graph that is expanded down to the HMM-state level.  It outputs
+most directly applicable when using WFST decoders where the WFST is
+expanded down to the HMM-state level.  It outputs
 lattices that include state-level alignments as well as word labels.
@@
 Let $B$ be the searched subset of $S$, containing a subset of the states and arcs
-of $S$ obtained by some heuristic pruning procedure.  Then in practice, 
-when we do Viterbi decoding with beam-pruning, we are finding the best path through $B$.
+of $S$ obtained by some heuristic pruning procedure.  
+When we do Viterbi decoding with beam-pruning, we are finding the best path through $B$.
 
@@
 expanded down to the $C$ level (i.e. $CLG$), so the input symbols represent
-context-dependent phones.  In their WFST based decoding networks, states did
-not have a unique one-word history, but they were able to satisfy a similar 
-condition at the phone level.  Their method was
+context-dependent phones.  In WFST based decoding networks, states normally do
+not have a unique one-word history, but the authors of~\cite{efficient_general}
+were able to satisfy a similar condition at the phone level.  Their method was
 to store a single Viterbi back-pointer at the phone level; use this to 
@@
 it takes account of both the forward and backward parts of the cost.
-The paper also reported experiments with a different method that did not requir any
+The paper also reported experiments with a different method that did not require any
 phone-pair assumption; these experiments showed that the more
 efficient method that relied on the phone-pair assumption
-had a minimal impact on the lattice oracle error rate.  However, the 
+had almost the same the lattice oracle error rate as the other method.  However, the 
 experiments did not evaluate how much impact the assumption had on the accuracy of the scores, 
@@
 
-The lattice generation algorithm that was described in~\cite{saon2005anatomy} (although not
-in very great detail) is applicable to WFSTs expanded down to the $H$ level (i.e. $\HCLG$),
+The lattice generation algorithm that was described in~\cite{saon2005anatomy}
+is applicable to WFSTs expanded down to the $H$ level (i.e. $\HCLG$),
 so the input symbols represent context-dependent states.  It keeps both score and 
@@
 graph do not have unique word histories, the resulting algorithm has to be quite
-different.  Viterbi back-pointers at the word level are used, but in each state
-is stored not just the best back-pointer but the $N$ best back-pointers for the $N$ top-scoring distinct
+different.  Viterbi back-pointers at the word level are used, but the algorithm keeps track
+of not just a single back-pointer in each state, but the $N$ best back-pointers for the $N$ top-scoring distinct
 preceding.  Therefore, this algorithm has more in common with the sentence N-best
@@
 (i.e. the total acoustic plus language-model cost).  This is just for didactic purposes;
-in fact, we have not implemented this simple version.
+we have not implemented this simple version.
 In this case our algorithm would be quite similar
-to~\cite{efficient_general}, except working at the state level rather than the phone level.
+to~\cite{efficient_general}, except at the state level rather than the phone level.
 We actually store forward rather than backward pointers: for each active state on each
@@
 where $B$ is the un-pruned state-level lattice.
-We project on the output labels (keeping only the word labels), and remove $\epsilon$
-labels and determinize.  In fact, we use a determinization algorithm that does $\epsilon$
+We project on the output labels (i.e. we keep only the word labels), then remove $\epsilon$
+arcs and determinize.  In fact, we use a determinization algorithm that does $\epsilon$
 removal itself.
@@
 First, let us define $Q = \inv(P)$, i.e. $Q$ is the inverted, pruned state-level lattice,
-so that the input symbols are the words and the output symbols are the p.d.f. labels.
+where the input symbols are the words and the output symbols are the p.d.f. labels.
 We want to process $Q$ in such a way that we keep only the best path through it
@@
 axioms.  In this case, we return the pair with the shorter string part, and if
-the lengths are the same, whichever string appears first in dictionary order.  These
-details do not have any practical effect on the behavior of the final algorithm in any
-normal situation, but are required to satisfy the semiring axioms.
+the lengths are the same, whichever string appears first in dictionary order.
 
-Suppose we have a WFST $Q$ representing a state-level lattice, with words
-as its input symbols and p.d.f.'s as its output symbols.  Let $E$ be an encoding of $Q$ with the 
-same number of states and arcs; $E$ is an acceptor, with its symbols
-equal to the input symbol on the corresponding arc of $Q$, and
-the weights on the arcs of $E$ containing both the the weight and the output symbol
+Let $E$ be an encoding of the inverted state-level lattice $Q$ 
+as described above, with the same number of states and arcs; $E$ is an acceptor, with its symbols
+equal to the input symbol (word) on the corresponding arc of $Q$, and
+the weights on the arcs of $E$ containing both the the weight and the output symbol (p.d.f.)
 on the corresponding arcs of $Q$.  Let $D = \mathrm{det}(\mathrm{rmeps}(E))$.
@@
 
-\section{Details of our implementation}
-
-\subsection{Epsilon removal and determinization}
+\section{Details of our $\epsilon$ removal and determinization algorithm}
 
@@
 We currently make our software robust to this phenomenon by 
-pre-specifyin a maximum number of states for a lattice, and if the
+pre-specifying a maximum number of states for a lattice, and if the
 determinization exceeds this threshold, we prune with a tighter beam

### 2011-09-26 misc/papers/icassp12_tied/icassp12.tex
updates to semicont paper
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@514 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 T.~Bocklet$^1$,
-A.~Ghoshal$^2$,
-D.~Povey$^3$
+A.~Ghoshal$^{2,3}$,
+D.~Povey$^4$
 }
@@
 $^2$Spoken Language Systems, Saarland University, \sc{Germany}\\
-$^3$Microsoft Research, Redmond, WA, {\sc USA}\\
+$^3$Centre for Speech Technology Research, University of Edinburgh, UK\\
+$^4$Microsoft Research, Redmond, WA, {\sc USA}\\
 {\small\tt korbinian.riedhammer@informatik.uni-erlangen.de}
@@
 \begin{abstract}
-In the past decade, semi-continuous hidden Markov models (HMMs) have not attracted 
-much attention in the speech recognition community. The strongly growing amount
-of training data and continuous improvements of the initialization and training
-procedures led to the impression that continuous HMMs are the best choice of
-acoustic model.
+In the past decade, semi-continuous hidden Markov models (SC-HMMs) have not 
+attracted much attention in the speech recognition community. Growing amounts 
+of training data and increasing sophistication of model estimation led to the 
+impression that continuous HMMs are the best choice of acoustic model.
 %
-However, recent work on tasks involving cross-lingual speech recognition and
-languages with little training data faced the same the old problem -- millions
-of parameters require a large amount of (unavailable) training data -- thus people are
-investigating methods of reducing the number of parameters while maintaining or
+However, recent work on recognition of under-resourced languages faces the same the old problem of estimating a large number of parameters from limited amounts 
+of labeled speech.
+This has led to a renewed interest in methods of reducing the number of parameters while maintaining or 
 extending the modeling capabilities of continuous models.
 %
-In this work, we compare continuous, semi-continuous, two-level tree based 
+In this work, we compare continuous, semi-continuous, %two-level tree based 
 multiple-codebook semi-continuous with full covariance matrices and subspace Gaussian
@@
 semi-continuous models in favor of continuous models, which performed 
-considerably better with the growing amount of training data available.
+considerably better with the growing amount of available training data.
 %
@@
 %
-This is the probably strongest advantage of semi-continuous models: While the
-codebook Gaussians are always estimated on the whole data set, the only state
-specific variables are the weights of the Gaussians, i.e.~increasing the number
-of states only slightly increases the number of parameters, and the Gaussian
-means and variances are still estimated using all the data; the reduced number
-of Gaussian parameters for a large data set also allows to robustly estimate
-full covariances.
+This is where semi-continuous models hold an advanatge: the only state specific 
+variables are the weights of the Gaussians, while the codebook Gaussians 
+themselves are always estimated on the whole data set. In other words, increasing the number of states leads to only modest increase in the total number of 
+parameters. The Gaussian means and variances are estimated using all of the available data, making it possible to robustly estimate even full covariance matrices.
 %
-This allows to reliably estimate arbitrary large phonetic context 
-(``polyphones''), that showed good performance for the recognition of 
-spontaneous  speech \cite{schukattalamazzini1994srf,schukattalamazzini1995as} 
-while having a rather small number of parameters.
+This also allows to reliably estimate arbitrarily large phonetic context 
+(``polyphones''), which has been shown to achieve good performance for the 
+recognition of spontaneous  speech \cite{schukattalamazzini1994srf,schukattalamazzini1995as} 
+while having a relatively small number of parameters.
 %
@@
 evaluated only once per frame, and that the codebook can be initialized, trained
-and adapted in a solely acoustic way.
+and adapted using even untranscribed speech.
 
@@
 For continuous models, every leaf of the tree is assigned an separate
-Gaussian mixture model. We initialize the mixtures with a single component
-each, and subsequently allocate more components by splitting components every 
-iteration until a certain total number of components is reached (in this case
-about 9000). We allocate the Gaussians proportional to a small power (e.g.~0.2) 
-of the state's occupation count.
-Formally, the emission probability of tree leaf $j$ is computed as
+Gaussian mixture model. Formally, the emission probability of tree leaf $j$ is 
+computed as
 \begin{equation}
@@
 $\k_{ji}$ are the means and covariance matrices of the mixtures.
+
+We initialize the mixtures with a single component each, and subsequently 
+allocate more components by splitting components at every iteration until a 
+pre-determined number of components is reached (in this case 9000). We allocate 
+the Gaussians proportional to a small power (e.g.~0.2) of the state's 
+occupation count.
 
@@
 \section{Summary}
-In this article, we evaluated semi-continuous and two-level tree based
-semi-continuous acoustic models of the {\sc Kaldi} toolkit in the context of
-both continuous and SGMM models.
+In this article, we compared two types of semi-continuous hidden Markov models, 
+one using a single codebook and the other using multiple codebooks based 
+on a two-level phonetic decision tree, with both continuous HMMs and SGMMs.
 %
 While the first could not produce convincing results, the two-level tree 
-based multiple-codebook architecture shows promising performance, especially for 
-little training data.
+based multiple-codebook architecture shows promising performance, especially 
+with limited training data.
 

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Modified the introduction of the paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@515 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 \makeatother
-\name{ Daniel Povey$^1$, Mirko Hannemann$^2$,  \\
+\name{ Daniel Povey$^1$, Mirko Hannemann$^{1,2}$,  \\
   {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
   {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Reidhammer}$^9$, {Karel Vesel\'{y}}$^2$
-    \thanks{Thanks here}}
+    \thanks{Thanks here.. remember Sanjeev.}}
 %\makeatletter
@@
 
-The word ``lattice'' is used in the speech recognition literature to mean some
-kind of compact representation of the most likely transcriptions of an utterance,
-in the form of a graph structure and normally including score and alignment
-information in addition to the word labels.  See 
-for example~\cite{efficient_general,ney_word_graph,odell_thesis,saon2005anatomy}.
+In Section~\ref{sec:wfst} we give a Weighted Finite State Transducer
+(WFST) interpretation of the speech-recognition decoding problem, in order
+to introduce notation for the rest of the paper.  In Section~\ref{sec:lattices}
+we define the lattice generation problem, and review previous work.
+In Section~\ref{sec:overview} we give an overview of our method,
+and in Section~\ref{sec:details} we summarize some aspects of a determinization
+algorithm that we use in our method.  In Section~\ref{sec:exp} we give
+experimental results, and in Section~\ref{sec:conc} we conclude.
 
-[more history + context here]
-
-\section{Decoding with WFSTs}
+\section{WFSTs and the decoding problem}
+\label{sec:wfst}
 
@@
 
-\section{Defining lattices and the lattice generation problem}
+\section{The lattice generation problem, and previous work}
+\label{sec:lattices}
+
+\subsection{Lattices, and the lattice generation problem}
 
@@
 
-\section{Previous lattice generation methods}
+\subsection{Previous lattice generation methods}
 
@@
 \section{Overview of our algorithm}
+\label{sec:overview}
 
@@
 \section{Details of our $\epsilon$ removal and determinization algorithm}
+\label{sec:details}
 

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Commit further changes to latgen paper, mostly cleanup.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@517 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 each state has a final-cost (or $\infty$ for non-final states);
-and there is a set of arcs, where each arc $a$ has a previous-state $p[a]$, a next-state $n[a]$, a
-weight $w[a]$ (just think of this as a cost for now), an input label $i[a]$ and an output
-label $o[a]$.  In $\HCLG$, the input labels are the identifiers of context-dependent
+and there is a set of arcs, where each arc has a weight,
+weight (just think of this as a cost for now), an input label and an output
+label.  In $\HCLG$, the input labels are the identifiers of context-dependent
 HMM states, and the output labels represent words.  For both the input and output
@@
 way to formalize these concerns is to express them in terms of a lattice 
-pruning beam $\alpha > 0$ (interpret $\alpha$ as a natural logarithm).
+pruning beam $\alpha > 0$ (interpret this as a log likelihood difference).
 \begin{itemize}
@@
 words as the labels, and the graph and acoustic costs and the alignments
-encoded into the weights.  Of course, the costs and alignments are not in any
-sense ``synchronized'' with the words.
+encoded into the weights.  The costs and alignments are not ``synchronized'' 
+with the words.
 
-\section{Details of our $\epsilon$ removal and determinization algorithm}
+\section{Details of our determinization algorithm}
 \label{sec:details}

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
Commit further changes to latgen paper-- shortening some parts
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@518 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 increase the size of the state-level lattice (this is mentioned 
-in~\cite{efficient_general}).  Rather than using completely generic
-code as is usually done in OpenFst~\cite{openfst} which our code is based on,
-we implemented it in a fairly specialized way for this case.  The issue
+in~\cite{efficient_general}).  Our algorithm uses data-structures
+specialized for the particular type of weight we are using.  The issue
 is that the determinization process often has to append a single symbol to
@@
 the output state index; think of this as a ``lookaside buffer'' that helps us
-avoid the expense of following $\epsilon$ arcs.  Note that formally
-proving the correctness of this algorithm would be a big project, but we
-are able to test correctness in a randomized way; this is easy because we
-can test WFSA equivalence and we can test whether a WFSA is deterministic.
-
-We should note that although the determinization algorithm
-is guaranteed to terminate (since the state-level lattice is acyclic), it can
-sometimes exhaust memory, i.e. in practice it can ``blow up'' for particular 
-inputs.  Our observation is that in cases where it significantly expands
-the input, this tends to be reversed by pruning after determinization.
-We currently make our software robust to this phenomenon by 
-pre-specifying a maximum number of states for a lattice, and if the
-determinization exceeds this threshold, we prune with a tighter beam
-and try again.  In the future we may implement a form of determinization that
-includes an efficient form of pruning.
+avoid the expense of following $\epsilon$ arcs.
 

### 2011-09-26 misc/papers/icassp12_lat/paper.tex
More edits to paper.. saved some space.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@519 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 \makeatother
-\name{ Daniel Povey$^1$, Mirko Hannemann$^{1,2}$,  \\
-  {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
-  {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Reidhammer}$^9$, {Karel Vesel\'{y}}$^2$
-    \thanks{Thanks here.. remember Sanjeev.}}
-%\makeatletter
-%\def\name#1{\gdef\@name{#1\\}}
-%\makeatother
+
+\name{ \em Daniel Povey$^1$, Mirko Hannemann$^{1,2}$,  \\
+  \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
+  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Reidhammer}$^9$, {Karel Vesel\'{y}}$^2$
+    \thanks{Thanks here.. remember Sanjeev.} }
 
@@
 natural properties we would want from a lattice of alternative transcriptions of
-an utterance.  This method is also quite efficient; it does not 
-introduce substantial overhead above one-best decoding.  Our method is
+an utterance.  This method does not introduce substantial overhead above one-best decoding.  Our method is
 most directly applicable when using WFST decoders where the WFST is
@@
 to introduce notation for the rest of the paper.  In Section~\ref{sec:lattices}
-we define the lattice generation problem, and review previous work.
+we define the lattice generation problem, and in Section~\ref{sec:previous}
+we review previous work.
 In Section~\ref{sec:overview} we give an overview of our method,
@@
 
-\section{The lattice generation problem, and previous work}
+\section{The lattice generation problem}
 \label{sec:lattices}
-
-\subsection{Lattices, and the lattice generation problem}
 
@@
 than acoustic states and contain sufficient information to reconstruct the phone
-sequence.  Hence, our lattices contain at least as much information as any prior work
-we cite.
+sequence.
 
@@
    the best-scoring path in in $B$ for the corresponding word
-   sequence\footnote{Or one of the best-scoring paths, in case of a tie}.
+   sequence\footnote{Or one of the best-scoring paths, in case of a tie.}.
 \end{itemize}
@@
 but doing so could cause paths not within $\alpha$ of the best one to be lost, so we have
-to weaken the condition.  In fact this weakening is of very little consequence, since regardless
+to weaken the condition.  This is no great loss, since regardless
 of pruning, any word-sequence not within $\alpha$ of the best one could be omitted altogether, 
 which is the same as being assigned a cost of $\infty$).
-
-We note that by ``word-sequence'' we mean a sequence of whatever symbols are on the
+By ``word-sequence'' we mean a sequence of whatever symbols are on the
 output of $\HCLG$.  In our experiments these symbols represent words, but not including
@@
 
-\subsection{Previous lattice generation methods}
+\section{Previous lattice generation methods}
+\label{sec:previous}
 
-Lattice generation algorithms tend to be closely linked to a particular type of decoder,
+Lattice generation algorithms tend to be closely linked to particular types of decoder,
 but are often justified by the same kinds of ideas.
@@
 is not affected by the identity of any earlier words.  In a decoder in which there is
-a different copy of the lexical tree for each preceding word, this assumption is automatically
-satisfied, and if the word-pair assumption holds, in order to generate an accurate lattice it is
+a different copy of the lexical tree for each preceding word, 
+assuming the word-pair assumption holds, in order to generate an accurate lattice it is
 sufficient to store a single Viterbi back-pointer at the word level; the entire set of
@@
 create a phone-level latice; prune the resulting
-lattice; project it to leave only word labels; and then convert this into a word-level
-lattice via epsilon removal and determinization.
+lattice; project it to leave only word labels; and then remove $\epsilon$
+symbols and determinize.
 Note that the form of pruning referred to here is not the same as beam pruning as
@@
 efficient method that relied on the phone-pair assumption
-had almost the same the lattice oracle error rate as the other method.  However, the 
+had almost the same lattice oracle error rate as their more efficient method.  However, the 
 experiments did not evaluate how much impact the assumption had on the accuracy of the scores, 
@@
 is applicable to WFSTs expanded down to the $H$ level (i.e. $\HCLG$),
-so the input symbols represent context-dependent states.  It keeps both score and 
+so the input symbols represent context-dependent states.  It keeps both scores and 
 state-level alignment information.  In some sense this algorithm also relies
@@
 algorithm than with the Viterbi algorithm.  By limiting $N$ to be quite small (e.g. $N{=}5$)
-the algorithm was made efficient, but this is at the cost of possibly losing word sequences
-that would be within the lattice-generation beam.  In addition, this algorithm was quite
-complex to implement efficiently.
+the algorithm was made efficient, but at the cost of losing word sequences
+that would be within the lattice-generation beam.
 
@@
 arising from transition probabilities and pronunciation probabilities.  We implement
-this by using a semiring that contains two real numbers (one for the graph and one
-for the acoustic costs); it keeps track of the two costs separately, but its
+this by using a semiring that contains two real numbers, one for the graph and one
+for the acoustic costs; it keeps track of the two costs separately, but its
 $\oplus$ operation returns whichever pair has the lowest sum of costs (graph plus acoustic).
@@
 Formally, if each weight is a pair $(a,b)$, then $(a,b) \otimes (c,d) = (a{+}c, b{+}d)$,
-and $(a,b) \oplus (c,d)$ is equal to $(a,b)$ if $a{+}b < c{+}d$ or if $a{+}b = c{f}+d$ and $a{-}b < c{-}d$,
+and $(a,b) \oplus (c,d)$ is equal to $(a,b)$ if $a{+}b < c{+}d$ or if $a{+}b = c{+}d$ and $a{-}b < c{-}d$,
 and otherwise is equal to $(c,d)$.  This is equivalent to the lexicographic semiring 
@@
 for each word sequence, and get the corresponding alignment.  This is possible
-by defining an appropriate semiring and then doing normal determinization.  Firstly, 
-let us ignore the fact that we are keeping track of separate graph and acoustic costs, 
+by defining an appropriate semiring and then doing normal determinization.  We shall
+ignore the fact that we are keeping track of separate graph and acoustic costs, 
 to avoid complicating the present discussion.
@@
 $(s,s')$ is $s$ and $s'$ appended together.  We define the $\oplus$ operation so that
-it returns whichever pair has the smallest weight: that is, $(c,s) \oplus (c',s')$ 
-equals $(c,s)$ if $c < c'$, and $(c',s')$ if $c > c'$.  If the weights are identical,
+it returns whichever pair has the smallest cost: that is, $(c,s) \oplus (c',s')$ 
+equals $(c,s)$ if $c < c'$, and $(c',s')$ if $c > c'$.  If the costs are identical,
 we cannot arbitrarily return the first pair because this would not satisfy the semiring
@@
 equal to the input symbol (word) on the corresponding arc of $Q$, and
-the weights on the arcs of $E$ containing both the the weight and the output symbol (p.d.f.)
-on the corresponding arcs of $Q$.  Let $D = \mathrm{det}(\mathrm{rmeps}(E))$.
-
-In the case we are interested in, determinization
+the weights on the arcs of $E$ containing both the the weight and the output symbol (p.d.f.),
+if any, on the corresponding arcs of $Q$.  Let $D = \mathrm{det}(\mathrm{rmeps}(E))$.
+Determinization
 will always succeed because $E$ is acyclic (as long as the original decoding
@@
 is that the determinization process often has to append a single symbol to
-a string of symbols (for the state alignments), and the easiest way to
+a string of symbols, and the easiest way to
 do this in ``generic'' code would involve copying the whole sequence each time.
@@
 
-Conclusions here
+Conclusions here [mention does not rely on word-pair assumption]
 

### 2011-09-27 misc/papers/icassp12_tied/icassp12.tex
Further rewrites, proper WSJ expts still missing; please check if you can contrib to TODOs
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@521 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 %
-However, recent work on recognition of under-resourced languages faces the same the old problem of estimating a large number of parameters from limited amounts 
-of labeled speech.
-This has led to a renewed interest in methods of reducing the number of parameters while maintaining or 
-extending the modeling capabilities of continuous models.
+However, recent work on recognition of under-resourced languages faces the same
+old problem of estimating a large number of parameters from limited amounts 
+of transcribed speech.
+This has led to a renewed interest in methods of reducing the number of parameters 
+while maintaining or extending the modeling capabilities of continuous models.
 %
-In this work, we compare continuous, semi-continuous, %two-level tree based 
-multiple-codebook semi-continuous with full covariance matrices and subspace Gaussian
-mixture models.
+In this work, we compare continuous, classic and multiple-codebook semi-continuous,
+with full covariance matrices, and subspace Gaussian mixture models.
 %
@@
 systems include generalizations of the phonetic context modeling to reduce the 
-number of states and state-tying to reduce the number of Gaussians.
+number of states or state-tying to reduce the number of Gaussians.
 
 It is well known that context dependent states significantly improve the
-recognition performance (e.g.~from monophone to triphone). However for a
-continuous system, increasing the number of states not only requires more
-Gaussians to be estimated, but also implies that the training data for each
-state is reduced.
+recognition performance (e.g.,~from monophone to triphone). However, 
+increasing the number of states of a continuous system not only requires more 
+Gaussians to be estimated, but also implies that the training data for each state 
+is reduced.
 %
-This is where semi-continuous models hold an advanatge: the only state specific 
+This is where semi-continuous models hold an advantage: the only state specific 
 variables are the weights of the Gaussians, while the codebook Gaussians 
-themselves are always estimated on the whole data set. In other words, increasing the number of states leads to only modest increase in the total number of 
-parameters. The Gaussian means and variances are estimated using all of the available data, making it possible to robustly estimate even full covariance matrices.
+themselves are always estimated on the whole data set. 
+In other words, increasing the number of states leads to only modest increase 
+in the total number of parameters. The Gaussian means and variances are estimated 
+using all of the available data, making it possible to robustly estimate even full 
+covariance matrices.
 %
@@
 %
-Further advantages of the shared codebook Gaussians are that they need to be 
+Further advantages of using codebook Gaussians are that they need to be 
 evaluated only once per frame, and that the codebook can be initialized, trained
@@
 
-The idea of semi-continuous models can be extended to multiple codebooks; by
+Semi-continuous models can be extended to multiple codebooks; by
 assigning certain groups of states to specific codebooks, one can find a
@@
 
+% TODO Fix the number of hours for RM and WSJ SI-284
 In this article, we experiment with classic SC-HMMs and two-level tree based 
@@
 between continuous and subspace Gaussian mixture models; 
-we compare the above four types of acoustic models on a small and medium
-sized corpus of read English while keeping acoustic frontend, training and 
-decoding unchanged. The software is part of the {\sc Kaldi} speech recognition
-toolkit \cite{povey2011tks} and is freely available for download, along with the
+we compare the above four types of acoustic models on a small (Resource Management, RM, 5 hours)
+and medium sized corpus (Wall Street Journal, WSJ, 17 hours) of read English while keeping acoustic 
+frontend, training and decoding unchanged. 
+The software is part of the {\sc Kaldi} speech recognition toolkit 
+\cite{povey2011tks} and is freely available for download, along with the
 example scripts to reproduce the presented experiments.
@@
 subspace forms of Gaussian Mixture Models. 
-We also describe the phonetic decisoin-tree building process, which is 
+We also describe the phonetic decision-tree building process, which is 
 necessary background for how we build the multiple-codebook systems.
 
-\subsection{Tree-Building}
+\subsection{Acoustic-Phonetic Decision Tree Building}
 The acoustic-phonetic decision tree provides the link between phones
@@
 The splitting procedure can ask questions not only about the context phones,
-but also the central phone and the HMM state; the phonetic questions are 
+but also about the central phone and the HMM state; the phonetic questions are 
 derived from an automatic clustering procedure.  The splitting procedure is
@@
 \end{equation}
-where $N_j$ is the number of Gaussians assigned to $s_j$, and the $\m_{ji}$ and
+where $N_j$ is the number of Gaussians assigned to $j$, and the $\m_{ji}$ and
 $\k_{ji}$ are the means and covariance matrices of the mixtures.
@@
 maximization (EM), maximum a-posteriori (MAP), maximum likelihood linear
-regression (MLLR) and similar algorithms.
+regression (MLLR) and similar algorithms on untranscribed audio data.
 
@@
 Gaussian parameters are tied at this level.  For this type of system we do not
-do any post-clustering.
+apply any post-clustering.
 The leaves of the second tree contain the weights, and these leaves correspond 
@@
 N_k = N_0 + \frac
-  { \left(  \sum_{l \in \{m(l) = k\}} \text{occ}(l)  \right)^q }
-  { \sum_r \left(  \sum_{t \in \{m(t) = r\}} \text{occ}(l)  \right)^q } 
+%% power rule, maybe some other day...
+%  { \left(  \sum_{l \in \{m(l) = k\}} \text{occ}(l)  \right)^q }
+%  { \sum_r \left(  \sum_{t \in \{m(t) = r\}} \text{occ}(l)  \right)^q } 
+  { \sum_{l \in \{m(l) = k\}} \text{occ}(l) }
+  { \sum_r \sum_{t \in \{m(t) = r\}} \text{occ}(l) } 
   \left( N - K \cdot N_0 \right)
 \end{equation}
-where $N_0$ is a minimum number of Gaussians per codebooks (e.g.\,3), $N$ is
-the total number of Gaussians, $K$ the number of codebooks, $\text{occ}(j)$ 
-is the occupancy of tree leaf $j$, and $q$ controls the influence of
-the occupancy regarding one codebook. A typical value of $q=0.2$ leads to
-rather homogeneous codebook sizes while still attributing more Gaussians
-to states with larger occupancies. For the two-level architecture, a $q=1$
-yielded best results.
+where $N_0$ is a minimum number of Gaussians per codebooks (e.g., 3), $N$ is
+the total number of Gaussians, $K$ the number of codebooks, and $\text{occ}(j)$ 
+is the occupancy of tree leaf $j$. 
+%
+%% Korbinian: No power rule for now, seems to harm results.
+%, and $q$ controls the influence of the occupancy regarding one codebook. 
+%A typical value of $q=0.2$ leads to rather homogeneous codebook sizes while 
+%still attributing more Gaussians to states with larger occupancies. 
+%For the two-level architecture, a $q=1$ yielded best results.
+%
 The target sizes of the codebooks are again enforced by either splitting or
@@
 semi-continuous models, to reduce the number of parameters by selecting the
-Gaussians from a subspace spanned by a universal background model (codebook)
+Gaussians from a subspace spanned by a universal background model (UBM)
 and state specific transformations. In principle, the SGMM emission pdfs can
@@
 $\M_i$ and $\w_i$. The term ``subspace'' indicates that the parameters of the
-mixtures are limited to ta sub-space of the entire space of parameters of
+mixtures are limited to a subspace of the entire space of parameters of
 the underlying codebook.
@@
 their parent's using an interpolation weight $\rho$ in
-\begin{equation}
+\begin{equation} \label{eq:intra}
 \hat\gamma_{ji} \leftarrow 
-  \underbrace{\left(\gamma_{ji} + \frac{\rho \, \gamma_{p(j),i}}{\left( \sum_k \gamma_{p(j),k} \right) + \epsilon}\right)}_{\mathrel{\mathop{:}}= \bar\gamma_{ji}}
+  \underbrace{\left(\gamma_{ji} + \frac{\rho}{\left( \sum_k \gamma_{p(j),k} \right) + \epsilon} \, \gamma_{p(j),i}\right)}_{\mathrel{\mathop{:}}= \bar\gamma_{ji}}
   \cdot 
@@
 \end{equation}
-which leads to an expontially decreasing impact of the initial parameter set.
+which leads to an exponentially decreasing impact of the initial parameter set.
 
@@
 % TODO fix the WSJ train/dev/eval description
-For the WSJ data set, we trained on the si284 training set, tuned on dev
+For the WSJ data set, we trained on the SI-284 training set, tuned on dev
 and tested on eval92,93.
+As the classic semi-continuous system did not yield good performance on the
+RM data both in terms of run-time and performance, we omit it for the WSJ
+experiments.
+%
 % TODO Korbinian: There are more expts ongoing, they should be done by 
@@
   Gaussians in 1576 tree leaves.
-\item
-  {\em semi}: semi-continuous triphone system using 768 full covariance
-  Gaussians in 2500 tree leaves, but no smoothing (explanations below).
+%\item
+%  {\em semi}: semi-continuous triphone system using 768 full covariance
+%  Gaussians in 2500 tree leaves, but no smoothing (explanations below).
 \item
@@
 
+%------%<----------------------------------------------------------------------
 \subsection{RM}
 
-\begin{table}[tb]
+\begin{table}%[tb]
 \begin{center}
@@
 {\em semi} &  1.80 &  3.19 &  4.72 &  4.62 &  4.15 &  6.88 & 4.66 \\ \hline 
-{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & 2.90 \\ \hline
+{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & {\bf 2.90} \\ \hline
 {\em sgmm} &  0.48 &  2.20 &  2.62 &  2.50 &  1.93 &  5.12 & 2.78 \\ \hline
@@
 and the rather little training data.
-% TODO Korbinian: I have some numbere on diag vs. full; maybe not as a table
-Interestingly, the full covariances make the difference -- the same system with
-diagonal covariances performs significantly worse, but approximates the 
-continuous performance when using the same amount of Gaussians.
+
+%------%<----------------------------------------------------------------------
+% Now let's talk about the diag/full cov and interpolations
+
+\begin{table}%[tb]
+\begin{center}
+\begin{tabular}{|c|c||c|c|c|c|}
+\hline
+covariance & Gaussians & none & inter & intra & both \\ \hline\hline
+full       &      1024 & 3.70 & 3.64 & 3.79 & 3.74 \\ \hline
+full       &      3072 & 3.01 & 3.01 & 3.02 & 2.90 \\ \hline\hline
+diagonal   &      3072 & 4.13 & 4.15 & 4.25 & 4.35 \\ \hline
+diagonal   &      9216 & 3.22 & 3.09 & 3.28 & 3.20 \\ \hline
+\end{tabular}
+\end{center}
+\caption{\label{tab:rm_diagfull}
+Average \% WER of the multiple-codebook semi-continuous model using different 
+numbers diagonal and full covariance Gaussians, and different smoothings 
+on the RM data. 
+Settings are 208 codebooks, 2500 context dependent states; $\rho = 35$ 
+and $\varrho = 0.2$ if active.
+}
+\end{table}
+
+Keeping the number of codebooks and leaves as well as the smoothing parameters
+$rho$ and $\varrho$ of {\em 2lvl} constant, we experiment with the number of
+Gaussians and type of covariance; the results are displayed in 
+Tab.~\ref{tab:rm_diagfull}.
+Interestingly, the full covariances make a strong difference: Using the same
+number of Gaussians, full covariances lead to a significant improvement.
+On the other hand, a substantial increase of the number of diagonal Gaussians 
+leads to a similar performance as the regular continuous system.
+%
+Another observation from Tab.~\ref{tab:rm_diagfull} is that the smoothing
+parameters need to be carefully calibrated. While the inter-iteration smoothing
+helps in most cases, the intra-iteration smoothing coefficient $\rho$ strongly
+depends on the type and number of Gaussians, which is due to the direct influence
+of $\rho$ in Eq.~\ref{eq:intra}.
 
@@
 %                 and cmvn, and hopefully a good param combination for 2lvl...
-\begin{table}[tb]
+\begin{table}%[tb]
 \begin{center}
@@
 \hline
-~          & eval92 & eval93 & {\em avg} \\ \hline\hline
-{\em cont} &  13.17 &  18.61 &     15.23 \\ \hline
-{\em semi} &   \\ \hline 
+~          &    dev & eval92 & eval93 & {\em avg} \\ \hline\hline
+{\em cont} &        &  13.17 &  18.61 &     15.23 \\ \hline
+% Korbinian: no semi-continuous for WSJ, takes too long to compute; will do it
+%            for the journal, though.
+%{\em semi} &        &        &        &           \\ \hline
+
 % tri2-2lvl-208-3072-4000-0-0-0-0 !! no cmvn !!
-{\em 2lvl} &  13.95 &  21.61 &     16.85 \\ \hline
-{\em sgmm} &  10.76 &  17.82 &     13.44 \\ \hline
+%{\em 2lvl} &  13.95 &  21.61 &     16.85 \\ \hline
+
+% tri2-2lvl-208-4096-6000-0-1-35-0.2 !! no cmvn !!
+{\em 2lvl} &        &  12.83 &  21.61 &     16.16 \\ \hline
+
+{\em sgmm} &        &  10.76 &  17.82 &     13.44 \\ \hline
 \end{tabular}
@@
 \section{Summary}
-In this article, we compared two types of semi-continuous hidden Markov models, 
-one using a single codebook and the other using multiple codebooks based 
-on a two-level phonetic decision tree, with both continuous HMMs and SGMMs.
+In this article, we compared continuous models and SGMM to 
+two types of semi-continuous hidden Markov models, one using a single codebook 
+and the other using multiple codebooks based on a two-level phonetic decision tree.
 %
-While the first could not produce convincing results, the two-level tree 
-based multiple-codebook architecture shows promising performance, especially 
-with limited training data.
+While the first could not produce convincing results, the multiple-codebook 
+architecture shows promising performance, especially for limited training data.
 
 Although the current performance is below the state-of-the-art, the
-rather simple theory and computational complexity, paired with the possibility
+rather simple theory and low computational complexity, paired with the possibility
 of solely acoustic adaptation make two-level tree based semi-continuous acoustic

### 2011-09-27 misc/papers/icassp12_tied/icassp12.tex
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@522 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
   {\em semi}: semi-continuous triphone system using 768 full covariance
-  Gaussians in 2500 tree leaves, but no smoothing (explanations below).
+  Gaussians in 2500 tree leaves, no smoothing (explanations later).
 \item
@@
 depends on the type and number of Gaussians, which is due to the direct influence
-of $\rho$ in Eq.~\ref{eq:intra}.
+of $\rho$ on the counts in Eq.~\ref{eq:intra}.
 
@@
 
+% Korbinian: We might need that ;)
+%\footnotesize
 \bibliographystyle{IEEEbib}

### 2011-09-27 misc/papers/icassp12_tied/icassp12.tex
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@523 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 % TODO fix the WSJ train/dev/eval description
-For the WSJ data set, we trained on the SI-284 training set, tuned on dev
-and tested on eval92,93.
+%% Korbinian: For now, I inserted results on si84/half, using cmvn (tri2c),
+%             the si284 take just too long for the submission.
+%For the WSJ data set, we trained on the SI-284 training set, tuned on dev
+%and tested on eval92,93.
+To get an intuition of the performance on the WSJ data set, we trained on 
+half of the SI-84 training set using settings similar to the RM experiments; 
+we tested on the eval '92 and eval '93 data sets.
 As the classic semi-continuous system did not yield good performance on the
@@
 %
-% TODO Korbinian: There are more expts ongoing, they should be done by 
-%                 tomorrow! I will also fill in the details below then.
+% TODO Korbinian: I hope the SI-284 experiments finish soon enough while the 
+%                 submission is still open. For now, I describe SI-84/half.
 \begin{itemize}
@@
   Gaussians in 1576 tree leaves.
+%% TODO Korbinian: For now, I use the same setting as in RM, to evade the
+%                  tuning on the dev set. Also, no semi system, too slow
+%                  and bad performance on RM.
+\item
+    {\em 2lvl}: same configuration as for RM (no further tuning).
 %\item
-%  {\em semi}: semi-continuous triphone system using 768 full covariance
-%  Gaussians in 2500 tree leaves, but no smoothing (explanations below).
-\item
-  {\em 2lvl}: two-level tree based semi-continuous triphone system using
-  4096 full covariance Gaussians in 208 codebooks ( min/average/max
-  components),  tree leaves, $\rho = $ and $\varrho = $.
+%  {\em 2lvl}: two-level tree based semi-continuous triphone system using
+%  4096 full covariance Gaussians in 208 codebooks ( min/average/max
+%  components),  tree leaves, $\rho = $ and $\varrho = $.
 \item
@@
 performance of the four. Possible explanations are an insufficient number
-of Gaussians paired overfitting of the weights to a little number of 
+of Gaussians paired with overfitting of the weights to a little number of 
 components. Unfortunately, this could not be eased with the interpolations
@@
 Keeping the number of codebooks and leaves as well as the smoothing parameters
-$rho$ and $\varrho$ of {\em 2lvl} constant, we experiment with the number of
+$\rho$ and $\varrho$ of {\em 2lvl} constant, we experiment with the number of
 Gaussians and type of covariance; the results are displayed in 
@@
 \hline
-~          &    dev & eval92 & eval93 & {\em avg} \\ \hline\hline
-{\em cont} &        &  13.17 &  18.61 &     15.23 \\ \hline
-% Korbinian: no semi-continuous for WSJ, takes too long to compute; will do it
-%            for the journal, though.
-%{\em semi} &        &        &        &           \\ \hline
+~          & eval '92 & eval '93 & {\em avg} \\ \hline\hline
+{\em cont} &    12.75 &    17.10 &     14.39 \\ \hline % si-84/half, cmvn
 
@@
 % tri2-2lvl-208-4096-6000-0-1-35-0.2 !! no cmvn !!
-{\em 2lvl} &        &  12.83 &  21.61 &     16.16 \\ \hline
+{\em 2lvl} &    12.83 &    21.61 &     16.16 \\ \hline
 
-{\em sgmm} &        &  10.76 &  17.82 &     13.44 \\ \hline
+{\em sgmm} &    10.76 &    17.82 &     13.44 \\ \hline % si-84/half, cmvn
 \end{tabular}
@@
 
-% TODO Korbinian: This needs some more text once the above table is filled
 The results on the different test sets on the WSJ data are displayed in
 Tab.~\ref{tab:res_rm}.
+% TODO Korbinian: The following is for the si-84/half results, as currently
+%                 in the table.
+Using the parameters calibrated on the RM test data, the {\em 2lvl} system
+performance is not as good as of the continuous or GMM systems, but still
+in a similar range. Once the number of Gaussians and leaves, as well as the
+smoothing parameters are tuned on the development set, we expect the error
+rates to be in between the continuous and SGMM systems, as seen on the RM 
+data.
 

### 2011-09-27 misc/papers/icassp12_tied/icassp12.tex
submitted version
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@525 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 {\em semi} &  1.80 &  3.19 &  4.72 &  4.62 &  4.15 &  6.88 & 4.66 \\ \hline 
-{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & {\bf 2.90} \\ \hline
+{\em 2lvl} &  0.48 &  1.70 &  2.46 &  3.35 &  1.89 &  5.31 & 2.90 \\ \hline
 {\em sgmm} &  0.48 &  2.20 &  2.62 &  2.50 &  1.93 &  5.12 & 2.78 \\ \hline
@@
 ~                & eval '92 & eval '93 & {\em avg} \\ \hline\hline
-{\em cont} &    12.75 &    17.10 &     14.39 \\ \hline % si-84/half, cmvn
-
-% tri2-2lvl-208-3072-4000-0-0-0-0 !! no cmvn !!
-%{\em 2lvl} &  13.95 &  21.61 &     16.85 \\ \hline
-
-% tri2-2lvl-208-4096-6000-0-1-35-0.2 !! no cmvn !!
-{\em 2lvl} &    12.83 &    21.61 &     16.16 \\ \hline
-
-{\em sgmm} &    10.76 &    17.82 &     13.44 \\ \hline % si-84/half, cmvn
+% all experiments si-84/half, cmvn
+{\em cont}       &    12.75 &    17.10 &     14.39 \\ \hline
+{\em 2lvl/none}  &    12.92 &    17.85 &     14.79 \\ \hline
+{\em 2lvl/intra} &    13.03 &    17.56 &     14.61 \\ \hline
+{\em 2lvl/inter} &    12.80 &    17.59 &     14.61 \\ \hline
+{\em 2lvl/both}  &    13.01 &    17.59 &     14.75 \\ \hline
+{\em sgmm}       &    11.72 &    14.25 &     12.68 \\ \hline
 \end{tabular}
@@
 Detailed recognition results in \% WER for the '92 and '93 test sets using 
-different acoustic models on the WSJ data.
+different acoustic models on the WSJ data; the {\em 2lvl} system was trained
+with and without the different interpolations.
 }
@@
 performance is not as good as of the continuous or GMM systems, but still
-in a similar range. Once the number of Gaussians and leaves, as well as the
-smoothing parameters are tuned on the development set, we expect the error
-rates to be in between the continuous and SGMM systems, as seen on the RM 
-data.
+in a similar range while using about the same number of leaves but about a
+third of the Gaussians.
+Once the number of Gaussians and leaves, as well as the smoothing parameters 
+are tuned on the development set, we expect the error rates to be in between 
+the continuous and SGMM systems, as seen on the RM data.
 
 \section{Summary}
-In this article, we compared continuous models and SGMM to 
+In this article, we compared continuous and SGMM models to 
 two types of semi-continuous hidden Markov models, one using a single codebook 
@@
 architecture shows promising performance, especially for limited training data. 
+%
+% TODO This sentence only for SI-84/half experiments
+The experiments on the WSJ data need to be extended to the full training set, and
+the parameter settings calibrated on the development set.
 
@@
 rather simple theory and low computational complexity, paired with the possibility
-of solely acoustic adaptation make two-level tree based semi-continuous acoustic
-models an attractive alternative to low-resource applications -- both in terms
-of computational power and training data.
+of solely acoustic adaptation make multiple-codebook semi-continuous acoustic
+models an attractive alternative for low-resource applications -- both in terms
+of computational power and training data; especially for embedded applications,
+where the codebook evaluation can be implemented in hardware, thus eliminating
+the time difference in evaluating full and diagonal covariance Gaussians.
 %
@@
 % Korbinian: We might need that ;)
-%\footnotesize
+\footnotesize
 \bibliographystyle{IEEEbib}

### 2011-09-27 misc/papers/icassp12_tied/icassp12.tex
updates to semicont paper
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@526 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 %
-In this work, we compare continuous, classic and multiple-codebook semi-continuous,
-with full covariance matrices, and subspace Gaussian mixture models.
+In this work, we compare classic and multiple-codebook semi-continuous models 
+that use full covariance matrices with continuous HMMs and subspace Gaussian mixture models.
 %
@@
 \subsection{Acoustic-Phonetic Decision Tree Building}
-The acoustic-phonetic decision tree provides the link between phones
+The acoustic-phonetic decision tree provides a mapping between phones
 in context and emission probability density functions (pdfs). 
@@
 
-\subsection{Two-level Tree based Multiple-Codebook Semi-continuous Models}
+\subsection{Multiple-Codebook Semi-continuous Models}
 A style of system that lies between the semi-continuous and continuous types of
@@
 to first build the decision tree used for phonetic clustering to a relatively
-coarse level (e.g.~100); and to then split more finely (e.g.~2500), but for
-each new leaf, remember which leaf of the original tree it corresponds to.
-Each leaf of the first tree corresponds to a codebook of Gaussians, i.e.~the 
-Gaussian parameters are tied at this level.  For this type of system we do not
-apply any post-clustering.
-The leaves of the second tree contain the weights, and these leaves correspond 
-to the actual context-dependent HMM states.
+coarse level (e.g.~100 leaves). Each leaf of this tree corresponds to a codebook of Gaussians, i.e.~the Gaussian parameters are tied at this level. The leaves 
+of this tree are then split further to give a more fine-grained clustering (e.g.~2500 leaves), where for each new leaf we remember which codebook it corresponds to.
+The leaves of this second tree correspond to the actual context-dependent HMM states and contain the weights.
+For this type of system we do not apply any post-clustering.
 
@@
 Although an increasing number of tree leaves does not imply a huge increase in
-parameters, the sufficient statistics for the individual leaves will be
-sparser, especially for little training data.
+parameters, the estimation of weights at the individual leaves suffer from data 
+fragmentation, especially when training models with small amounts of data.
 The {\em intra-iteration} smoothing is motivated by the fact that tree leaves
@@
 Second, the statistics of each node and leaf are interpolated top-down with 
-their parent's using an interpolation weight $\rho$ in
+their parent's statistics using an interpolation weight $\tau$:
 \begin{equation} \label{eq:intra}
 \hat\gamma_{ji} \leftarrow 
-  \underbrace{\left(\gamma_{ji} + \frac{\rho}{\left( \sum_k \gamma_{p(j),k} \right) + \epsilon} \, \gamma_{p(j),i}\right)}_{\mathrel{\mathop{:}}= \bar\gamma_{ji}}
+  \underbrace{\left(\gamma_{ji} + \frac{\tau}{\left( \sum_k \gamma_{p(j),k} \right) + \epsilon} \, \gamma_{p(j),i}\right)}_{\mathrel{\mathop{:}}= \bar\gamma_{ji}}
   \cdot 
-  \underbrace{\frac{\sum_k \gamma_{jk}}{\sum_k \bar\gamma_{jk}}}_\text{normalization}
+  \underbrace{\frac{\sum_k \gamma_{jk}}{\sum_k \bar\gamma_{jk}}}_\text{normalization},
 \end{equation}
-where $\gamma_{ji}$ is the accumulator of weight $i$ in tree leaf $j$
-and $p(j)$ is the mapping to the parent node of $j$.
+where $\gamma_{ji}$ is the occupancy of the $i$ component in tree leaf $j$
+and $p(j)$ is the parent node of $j$.
 %
@@
 Another typical problem that arises when training semi-continuous models is
-that the weights tend to converge on a local optimum over the iterations. 
+that the weights tend to converge to a local optimum over the iterations. 
 This is due to the fact that the codebook parameters change rather slowly 
@@
 %
-To compensate for this, we smooth the newly estimated {\em parameters} 
-$\Theta^{(i+1)}$ with the ones from the prior iteration using an interpolation
-weight $\varrho$
+To compensate for this, we smooth the newly estimated parameters $\Theta^{(i+1)}$ with the ones from the prior iteration using an interpolation weight $\rho$
 \begin{equation}
-\hat\Theta^{(i+1)} \leftarrow (1 - \varrho) \, \Theta^{(i+1)} + \varrho \, \Theta^{(i)} % \quad .
+\hat\Theta^{(i+1)} \leftarrow (1 - \rho) \, \Theta^{(i+1)} + \rho \, \Theta^{(i)} % \quad .
 \end{equation}
@@
 Feb and Oct '89, Feb '91 and Sep '92 (about 1500 utterances total).  The parameters 
-were tuned to the test set.
+are tuned to the test set since RM has no hold-out data set. We use the following system identifiers in Table~\ref{tab:res_rm}:
 %
@@
   3072 full covariance Gaussians in 208 codebooks ($4/14.7/141$ min/average/max
-  components), 2500 tree leaves, $\rho = 35$ and $\varrho = 0.2$.
+  components), 2500 tree leaves, $\tau = 35$ and $\rho = 0.2$.
 \item
@@
 RM data both in terms of run-time and performance, we omit it for the WSJ
-experiments.
+experiments. Results corresponding to the following systems are presented in 
+Table~\ref{tab:res_wsj}:
 %
@@
 %  4096 full covariance Gaussians in 208 codebooks ( min/average/max
-%  components),  tree leaves, $\rho = $ and $\varrho = $.
+%  components),  tree leaves, $\tau = $ and $\rho = $.
 \item
@@
 The results on the different test sets on the RM data are displayed in 
-Tab.~\ref{tab:res_rm}. The classic semi-continuous system shows the worst 
+Table~\ref{tab:res_rm}. The classic semi-continuous system shows the worst 
 performance of the four. Possible explanations are an insufficient number
@@
 on the RM data. 
-Settings are 208 codebooks, 2500 context dependent states; $\rho = 35$ 
-and $\varrho = 0.2$ if active.
+Settings are 208 codebooks, 2500 context dependent states; $\tau = 35$ 
+and $\rho = 0.2$ if active.
 }
@@
 Keeping the number of codebooks and leaves as well as the smoothing parameters
-$\rho$ and $\varrho$ of {\em 2lvl} constant, we experiment with the number of
+$\tau$ and $\rho$ of {\em 2lvl} constant, we experiment with the number of
 Gaussians and type of covariance; the results are displayed in 
-Tab.~\ref{tab:rm_diagfull}.
+Table~\ref{tab:rm_diagfull}.
 Interestingly, the full covariances make a strong difference: Using the same
@@
 %
-Another observation from Tab.~\ref{tab:rm_diagfull} is that the smoothing
+Another observation from Table~\ref{tab:rm_diagfull} is that the smoothing
 parameters need to be carefully calibrated. While the inter-iteration smoothing
-helps in most cases, the intra-iteration smoothing coefficient $\rho$ strongly
+helps in most cases, the intra-iteration smoothing coefficient $\tau$ strongly
 depends on the type and number of Gaussians, which is due to the direct influence
-of $\rho$ on the counts in Eq.~\ref{eq:intra}.
+of $\tau$ on the counts in equation~\ref{eq:intra}.
 
@@
 The results on the different test sets on the WSJ data are displayed in
-Tab.~\ref{tab:res_rm}.
+Table~\ref{tab:res_rm}.
 % TODO Korbinian: The following is for the si-84/half results, as currently

### 2011-09-27 misc/papers/icassp12_lat/paper.tex
Adding figures to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@527 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
   \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
-  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Reidhammer}$^9$, {Karel Vesel\'{y}}$^2$
+  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
     \thanks{Thanks here.. remember Sanjeev.} }
@@
          $^5$ University of Edinburgh, U.K. \ \  $^6$  IDIAP, Martigny, Switzerland  \\ 
-         $^6$ Shanghai Jiao Tong University, China \ \ $^7$ Karlsruhe Institute of Technology, Germany \\ 
-         $^9$ Friedrich Alexander University of Erlangen-Nurnberg, Germany}
+         $^7$ Tsinghua University, Beijing, China \ \ $^8$ Karlsruhe Institute of Technology, Germany \\ 
+         $^9$ Pattern Recognition Lab, University of Erlangen-Nuremberg, Germany}
 

### 2011-09-27 misc/papers/icassp12_lat/paper.tex
Adding new results in paper.tex at bottom
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@528 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 running on blades, using 64bit version with double float and lattice-simple-decoder
+realtime factor running on pcburget, 22989 frames (decode3.sh)
 
@@
 
-experiment: changing lattice beam:
+1) experiment: changing lattice beam:
+(lattices 443c040w and 444c040i have been reduced from 10.0 to 8.1 because of determinization memory problems)
 
@@
 avg. link density: 23.56        16.0       10.82       7.26       4.91         3.38        2.38        1.77        1.4          1.17        1.0
-realtime factor:    3.61         3.17       2.59       2.41       2.34         2.98 ?      3.21 ?      1.89        3.12 ?       1.78        --
+realtime factor:    2.317        2.114      2.006      1.941     1.900         1.855       1.835       1.797       1.769        1.737       ---
 
@@
 rescore utt wrong: 205                                           204          204         204         202         205          212         227
+
+2) experiment: changing decoding beam:
+
+decoding beam:     16.0        15.0          14.0          13.0         12.0         11.0         10.0          9.0 (1 utt partial) 8.0 (2 partial)
+lattice beam:       7.0         7.0           7.0           7.0          7.0          7.0          7.0          7.0                7.0
+avg. link density:  7.4         7.26          7.04          6.74         6.33         5.71         4.9          4.08               3.09
+realtime factor:    2.646       1.941         1.428         1.052        0.775        0.555        0.388        0.267              0.177
+decoding WER:      11.51%       11.51%       11.54%        11.58%       11.75%       11.93%       12.32%        13.47%             15.41%
+ins/del/sub:       118/57/474   118/57/474   118/57/476    119/56/478   124/57/482   131/58/484   140/58/497    164/56/540         197/61/611
+utterances wrong:  227          227          227           227          228          228          229           236                246
+
+oracle WER:        3.03%        3.08%        3.24%         3.37%        3.56%        4.02%        4.72%         6.45%              9.18%
+ins/del/sub:       30/1/140     31/1/142     35/2/146      38/2/150     42/3/156     54/6/167     55/7/204      73/13/278          118/22/378
+oracle utt wrong:  100          101          105           108          112          119          128           153                184
+
+rescoring WER(15): 9.59%        9.59%        9.73%         9.80%        10.05%       10.26%       10.55%        11.82% (1 failed)  13.79% (2 failed)
+ins/del/sub:       116/36/389   117/36/388   115/37/397    117/38/398   123/39/405   132/39/408   141/37/417    ---                ---
+rescore utt wrong: 205          205          206           206          207          209          208           220                235
+
+3) bigram vs. trigram decoding:
+
+decoding lm:       bigram       trigram,pruned  bigram        trigram,pruned   bigram (just to see whether wider beams still help)
+decoding beam:     15.0         15.0            13.0          13.0             16.0 
+lattice beam:       9.0          9.0             9.2           9.2             10.0 (already quite big for lattice-oracle - more than 6GB RAM used)
+avg. link density: 16.0         14.98           14.86         13.58            24.84
+rt factor(old):     3.17         3.26            1.75         1.77              4.54
+
+decoding WER:      11.51%       10.57%          11.58%        10.80%           11.51%
+ins/del/sub:       118/57/474   116/44/436      119/56/478    120/47/442       118/57/474
+utterances wrong:  227          214             227           215              227
+
+oracle WER:        2.78%        2.84%           3.05%         3.19%            2.61%
+ins/del/sub:       25/0/132     30/0/130        30/2/140      33/1/146         21/0/126
+oracle utt wrong:  96           96              103           107              92
+
+rescoring WER(15):  9.59%       9.64%           9.80%         9.87%            9.59%
+ins/del/sub:       117/36/388   115/34/395      117/38/398    119/38/400       116/36/389
+rescore utt wrong: 205          205             206           209              205
+
+4) lattice vs. n-best lists
+decoding with trigram, pruned
+
+decoding:          lattice      n-best         n-best/beam
+decoding beam:     13.0         13.0           13.0  
+lattice beam:      9.2          --             9.35 [6.0-12.5] (few outliers below/above)
+N best:            --           <=20           206.96 [2-1600] (according to beam range and utterance)
+max tokens:        15000        150k,700k      150000-770000
+avg. link density: 13.58        14?            231.05
+realtime factor:   1.77         12             crap
+
+decoding WER:      10.80%       12.78% (why?)  14.96% (probably too low max tokens)
+ins/del/sub:       120/47/442
+utterances wrong:  215
+
+oracle WER:        3.19%
+ins/del/sub:       33/1/146
+oracle utt wrong:  107
+
+rescoring WER(15):  9.87%
+ins/del/sub:       119/38/400
+rescore utt wrong: 209

### 2011-09-27 misc/papers/icassp12_lat/paper.tex
Improvements to ICASSP paper.. nearly finished.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@529 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
   \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
-    \thanks{Thanks here.. remember Sanjeev.} }
+    \thanks{Thanks to Sanjeev Khudanpur for his help in preparing the paper}}
 
@@
   \caption{Acceptor $U$ describing the acoustic scores of an utterance}
+\vspace*{-0.2in}
 \label{fig:acceptor}
 \end{center}
+
 \end{figure}
@@
 a map from the minimal representation to the state index.  We can
-show that this algorithm is still correct (it will tend to give output
-with fewer states, i.e. more minimal).  As an optimization for speed, we also define
+show that this algorithm is still correct (it will tend to give more
+minimal output).   As an optimization for speed, we also define
 the {\em initial representation} to be the same type of subset, but prior
@@
 For these experiments we generate lattices with the bigram language model
-supplied with the WSJ database, and rescore them with the trigram language model.
+supplied with the WSJ database, and for rescoring experiments we
+use the trigram language model.  The acoustic scale was $1/16$ for first-pass
+decoding and $1/15$ for LM rescoring.
+
+For simplicity, our results were all obtained using a decoder that
+does not support a ``maximum active states'' option, so the only variables
+to consider are the beam used in the Viterbi beam search, and the separate
+beam $\alpha$ used for lattice generation.
+
+\begin{figure}
+\centering
+   \includegraphics[width=0.9\columnwidth]{figures/latbeam.eps} 
+ \vspace*{-0.1in}
+
+   \caption{\vspace*{-0.2in} Varying lattice beam $\alpha$ (Viterbi pruning beam fixed at 15) }
+  \label{fig:latbeam}
+\end{figure}
+
+Figure~\ref{fig:latbeam} shows how various quantities change as we vary
+$\alpha$, with the Viterbi beam fixed at 15.  Note that we get all the improvement from
+LM rescoring by increasing $\alpha$ to 4.  The time taken by our algorithm started to increase rapidly after about
+$\alpha=8$, so a value of $\alpha$ anywhere between about 4 and 8 is sufficient for LM rescoring 
+and still does not slow down decoding too much.
+Note that out of vocabularly words (OOVs) provide a floor on the lattice oracle error rate.
+Of the 333 test utterances, 87 contained an OOV word, yet only 93 sentences had oracle errors
+with $\alpha=10$.  Figure~\ref{fig:viterbibeam} shows the effect of varying the
+Viterbi decoding beam, while leaving $\alpha$ fixed at 7.
+
+\begin{figure}
+\centering
+   \includegraphics[width=0.9\columnwidth]{figures/decodebeam.eps} 
+ \vspace*{-0.1in}
+    \caption{\vspace*{-0.2in} Varying Viterbi pruning beam (lattice beam $\alpha$ fixed at 7) }
+  \label{fig:viterbibeam}
+\end{figure}
 
@@
 
-Conclusions here [mention does not rely on word-pair assumption]
+We have described a lattice generation method that is to our knowledge the first
+efficient method not to rely on the word-pair assumption of~\cite{ney_word_graph}.
+It includes an ingenious way of obtaining state-level alignment information
+via determinization in a specially designed semiring.
 

### 2011-09-28 misc/papers/icassp12_lat/paper.tex
Fixes to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@530 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
   \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
-    \thanks{Thanks to Sanjeev Khudanpur for his help in preparing the paper}}
+    \thanks{Thanks to Sanjeev Khudanpur for his help in preparing the paper, and to Honza Cernocky, Renata Kohlova, Martin Karafiat and Tomas Mikolov for their help relating to the Kaldi'11 workshop at BUT.}}
 
@@
 \end{equation}
-where $\circ$ is WFST composisition (note: view $\HCLG$ as a single symbol).
+where $\circ$ is WFST composition (note: view $\HCLG$ as a single symbol).
 For concreteness we will speak of ``costs'' rather
@@
 each state has a final-cost (or $\infty$ for non-final states);
-and there is a set of arcs, where each arc has a weight,
-weight (just think of this as a cost for now), an input label and an output
+and there is a set of arcs, where each arc has a weight
+(just think of this as a cost for now), an input label and an output
 label.  In $\HCLG$, the input labels are the identifiers of context-dependent
@@
 a beam $\alpha$; we do this periodically (every 25 frames) but this is equivalent to
-doing it just once at the end (similar to~\cite{efficient_general}).  Let the pruned
-state-level lattice that we get after pruning it at the end of the utterance be $P$.
+doing it just once at the end, as in~\cite{efficient_general}.  Let the 
+final pruned state-level lattice be $P$.
 
 Let $Q = \inv(P)$, and let $E$ be an encoded version of $Q$ as described above (with the
-state labels as part of the weights).  The final lattice we output is
+state labels as part of the weights).  The final lattice is
 \begin{equation}
@@
 The determinization and epsilon removal are done together by a single algorithm
-that we will describe below.
-
-The resulting lattice $L$ is a deterministic, acyclic weighted acceptor with the 
+that we will describe below.  $L$ is a deterministic, acyclic weighted acceptor with the 
 words as the labels, and the graph and acoustic costs and the alignments
@@
 decoding and $1/15$ for LM rescoring.  
-
-For simplicity, our results were all obtained using a decoder that
+For simplicity we used a decoder that
 does not support a ``maximum active states'' option, so the only variables
@@
 and still does not slow down decoding too much.
-Note that out of vocabularly words (OOVs) provide a floor on the lattice oracle error rate.
-Of the 333 test utterances, 87 contained an OOV word, yet only 93 sentences had oracle errors
-with $\alpha=10$.  Figure~\ref{fig:viterbibeam} shows the effect of varying the
+Out of vocabularly words (OOVs) provide a floor on the lattice oracle error rate:
+of 333 test utterances, 87 contained at least one OOV word, yet only 93 sentences (6 more) had oracle errors
+with $\alpha=10$.  Lattice density is defined as the average number of arcs crossing each frame.
+Figure~\ref{fig:viterbibeam} shows the effect of varying the
 Viterbi decoding beam, while leaving $\alpha$ fixed at 7.

### 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
Minor fixes; updates to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@531 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 can still yield competitive results while using fewer Gaussian components.
+Our best semi-continous systems included full-covariance codebook Gaussians,
+multiple codebooks, and smoothing of the weight parameters.
 \end{abstract}
@@
 \label{sec:intro}
-In the past years, semi-continuous hidden Markov models (SC-HMMs) \cite{huang1989shm}
-have not attracted much attention in the automatic speech recognition (ASR) community. 
-Most major frameworks (e.g.~CMU {\sc Sphinx} or HTK) more or less retired
+In the past years, semi-continuous hidden Markov models (SC-HMMs) \cite{huang1989shm},
+in which the Gaussian means and variances are shared among all the HMM states
+and only the weights differ, have not attracted much attention in the automatic 
+speech recognition (ASR) community. 
+Most major frameworks (e.g.~CMU {\sc Sphinx} or HTK) have more or less retired
 semi-continuous models in favor of continuous models, which performed 
 considerably better with the growing amount of available training data.
-%
-However, topics like cross-lingual model initialization and model training
-for languages with little training data became more and more popular recently,
-along with efforts on how to reduce the number of parameters of a continuous
-system while maintaining or even extending the modeling capabilities and thus
-the robustness and performance.
-%
-Typical techniques to reduce the number of parameters for continuous
-systems include generalizations of the phonetic context modeling to reduce the 
-number of states or state-tying to reduce the number of Gaussians.
 
-It is well known that context dependent states significantly improve the
-recognition performance (e.g.,~from monophone to triphone). However, 
-increasing the number of states of a continuous system not only requires more 
-Gaussians to be estimated, but also implies that the training data for each state 
-is reduced.
-%
-This is where semi-continuous models hold an advantage: the only state specific 
-variables are the weights of the Gaussians, while the codebook Gaussians 
-themselves are always estimated on the whole data set. 
-In other words, increasing the number of states leads to only modest increase 
-in the total number of parameters. The Gaussian means and variances are estimated 
-using all of the available data, making it possible to robustly estimate even full 
-covariance matrices.
-%
-This also allows to reliably estimate arbitrarily large phonetic context 
-(``polyphones''), which has been shown to achieve good performance for the 
-recognition of spontaneous  speech \cite{schukattalamazzini1994srf,schukattalamazzini1995as} 
-while having a relatively small number of parameters.
-%
-Further advantages of using codebook Gaussians are that they need to be 
-evaluated only once per frame, and that the codebook can be initialized, trained
-and adapted using even untranscribed speech.
+We were motivated by a number of factors to look again at semi-continuous models.
+One is that we are interested in applications where there may be a small amount
+of training data (or a small amount of in-domain training data), and since
+in semi-continuous models the amount of state-specific parameters is relatively
+small, we felt that they might have an advantage there.  We were also interested
+in improving the open-source speech recognition toolkit {\sc Kaldi}~\cite{povey2011tks},
+and felt that since a multiple-codebook form of semi-continous models is, to
+our knowledge, currently being used in a state-of-the-art system~\cite{prasad2004t2b},
+we should implement this type of model.  We also wanted to experiment with
+the style of semi-continous system described 
+in \cite{schukattalamazzini1994srf,schukattalamazzini1995as}, particularly the smoothing
+techniques, and see how these compare against a more conventional system.
+Lastly, we felt that it would be interesting to compare semi-continous
+models with Subspace Gaussian Mixture Models (SGMMs)~\cite{povey2011sgm}, since
+there are obvious similarities.
 
-Semi-continuous models can be extended to multiple codebooks; by
-assigning certain groups of states to specific codebooks, one can find a
-hybrid between a continuous and semi-continuous models that combine the strength
-of state-specific Gaussians with reliable parameter estimation \cite{prasad2004t2b}.
+In the experiments we report here we have not been able to match the performance of
+a standard diagonal GMM based system using a traditional semi-continuous system;
+however, using a two-level tree and multiple codebooks similar to~\cite{prasad2004t2b},
+we were able to get better performance than a traditional system (although not
+as good as SGMMs).  For both traditional and multiple-codebook configurations,
+we saw better results with full-covariance codebook Gaussians than with diagonal.
+We investigated smoothing the statistics for weight estimation using the decision
+tree, and saw only modest improvements.
+We note that the software used to do the experiments is part of the {\sc Kaldi} speech
+recognition toolkit~\cite{povey2011tks} and is freely available for download, along with the
+example scripts to reproduce these results.
 
-Extending the idea of sharing parameters, subspace Gaussian mixture models
-\cite{povey2011sgm} assign state specific weights and means, while the latter
-are derived from the means of a shared codebook (universal background model, UBM)
-using a state, phone or speaker specific transformation, thus limiting the state
-Gaussians to a subspace of the UBM.
+In Section~\ref{sec:am} we describe the traditional and tied-mixture models,
+as well as the multiple-codebook variant of the tied system, and describe some
+details of our implementation, such as how we build the two-level tree.
+In Section~\ref{sec:smooth} we describe two different forms of smoothing of the statistics,
+which we experiment with here.  
+In Section~\ref{sec:sys} we describe the systems we used (we are using
+the Resource Management and Wall Street Journal databases);
+in Section~\ref{sec:results} we give experimental results, and we
+summarize in Section~\ref{sec:summary}.
 
 % TODO Fix the number of hours for RM and WSJ SI-284
-In this article, we experiment with classic SC-HMMs and two-level tree based 
-multiple-codebook SC-HMMs which are, in terms of parameter sharing, somewhere in 
-between continuous and subspace Gaussian mixture models; 
-we compare the above four types of acoustic models on a small (Resource Management, RM, 5 hours)
-and medium sized corpus (Wall Street Journal, WSJ, 17 hours) of read English while keeping acoustic 
-frontend, training and decoding unchanged. 
-The software is part of the {\sc Kaldi} speech recognition toolkit 
-\cite{povey2011tks} and is freely available for download, along with the
-example scripts to reproduce the presented experiments.
+%In this article, we experiment with classic SC-HMMs and two-level tree based 
+%multiple-codebook SC-HMMs which are, in terms of parameter sharing, somewhere in 
+%between continuous and subspace Gaussian mixture models; 
+%we compare the above four types of acoustic models on a small (Resource Management, RM, 5 hours)
+%pand medium sized corpus (Wall Street Journal, WSJ, 17 hours) of read English while keeping acoustic 
+%frontend, training and decoding unchanged. 
+%The software is part of the {\sc Kaldi} speech recognition toolkit 
+%\cite{povey2011tks} and is freely available for download, along with the
+%example scripts to reproduce the presented experiments.
 
@@
 \subsection{Acoustic-Phonetic Decision Tree Building}
-The acoustic-phonetic decision tree provides a mapping between phones
-in context and emission probability density functions (pdfs). 
+The acoustic-phonetic decision tree provides a mapping from
+a phonetic context window (e.g. a triphone), together with a HMM-state index, 
+to an emission probability density function (pdf). 
 %
@@
 sufficient statistics needed to estimate a Gaussian for each seen combination 
-of (phonetic context, HMM-state).  These statistics are based on a Viterbi 
+of (phonetic context window, HMM-state).  These statistics are based on a Viterbi 
 alignment of a previously built system.  The roots
-of the decision can be configured; for the RM experiments, these correspond to
+of the decision trees can be configured; for the RM experiments, these correspond to
 the phones in the system, and on Wall Street Journal, they
@@
 The splitting procedure can ask questions not only about the context phones,
-but also about the central phone and the HMM state; the phonetic questions are 
+but also about the central phone and the HMM state; the questions about phones are 
 derived from an automatic clustering procedure.  The splitting procedure is
@@
 likelihood loss due to combining them is less than the likelihood improvement
-on the final split of the tree-building process; we typically restrict this 
-clustering to avoid combining leaves across different subtrees (e.g., from 
+on the final split of the tree-building process; and we restrict this 
+clustering to disallow combining leaves across different subtrees (e.g., from 
 different monophones).  The post-clustering process typically reduces the number 
@@
 \subsection{Continuous Models}
-For continuous models, every leaf of the tree is assigned an separate
+For continuous models, every leaf of the tree is assigned a separate
 Gaussian mixture model. Formally, the emission probability of tree leaf $j$ is 
@@
 \section{Smoothing Techniques for Semi-continuous Models}
+\label{sec:smooth}
 %
@@
 
-\section{Systems Description}
+\section{System Descriptions}
+\label{sec:sys}
+
 The experiments presented in this article are computed on the DARPA Resource 
@@
 
-On the acoustic frontend, we extract 13 Mel-frequency cepstral coefficients,
-compute deltas and delta-deltas, and apply cepstral mean normalization. For
-the decoding, we use a regular 3-gram estimated using the IRSTLM toolkit 
-\cite{federico2008iao}.
+On the acoustic frontend, we extract 13 Mel-frequency cepstral coefficients (MFCCs),
+apply cepstral mean normalization, and compute deltas and delta-deltas.  For
+decoding, we use the language models supplied with those corpora; for RM this
+is a word-pair grammar, and for WSJ we use the bigram version of the language
+models supplied with the data set.
 
@@
 Feb and Oct '89, Feb '91 and Sep '92 (about 1500 utterances total).  The parameters 
-are tuned to the test set since RM has no hold-out data set. We use the following system identifiers in Table~\ref{tab:res_rm}:
+are tuned to this joint test set since there was not enough data to hold out a
+development set and still get meaningful results.
+We use the following system identifiers in Table~\ref{tab:res_rm}:
 %
@@
 %and tested on eval92,93.
-To get an intuition of the performance on the WSJ data set, we trained on 
+
+Due to time constraints, we trained on 
 half of the SI-84 training set using settings similar to the RM experiments; 
@@
 As the classic semi-continuous system did not yield good performance on the
-RM data both in terms of run-time and performance, we omit it for the WSJ
+RM data both in terms of run-time and performance, we omitted it for the WSJ
 experiments. Results corresponding to the following systems are presented in 
@@
 \section{Results}
+\label{sec:results}
 
@@
 Table~\ref{tab:res_rm}. The classic semi-continuous system shows the worst 
-performance of the four. Possible explanations are an insufficient number
-of Gaussians paired with overfitting of the weights to a little number of 
-components. Unfortunately, this could not be eased with the interpolations
-and larger numbers of Gaussians would unacceptably increase the computational 
-load due to the full covariance matrices.
-
-The two-level tree based multiple-codebook system performance lies within the 
-continuous and SGMM system, which goes along with the modeling capabilities 
-and the rather little training data.
+performance of the four, which serves to confirm the conventional wisdom
+about the superiority of continous systems.
+The two-level tree based multiple-codebook system performance lies in between the 
+continuous and SGMM system.
 
@@
 %
-Another observation from Table~\ref{tab:rm_diagfull} is that the smoothing
-parameters need to be carefully calibrated. While the inter-iteration smoothing
-helps in most cases, the intra-iteration smoothing coefficient $\tau$ strongly
-depends on the type and number of Gaussians, which is due to the direct influence
-of $\tau$ on the counts in equation~\ref{eq:intra}.
+The effect of the two smoothing methods using $\tau$ and $\rho$ is not
+clear from Table~\ref{tab:rm_diagfull}, i.e. the results are not always
+consistent.  While the inter-iteration smoothing with $\rho$
+helps in most cases, the intra-iteration smoothing coefficient $\tau$ 
+shows an inconsistent effect.  In the future we may abandon this type of
+smoothing or investigate other smoothing methods.
 
@@
 Using the parameters calibrated on the RM test data, the {\em 2lvl} system
-performance is not as good as of the continuous or GMM systems, but still
+performance is not as good as the continuous or GMM systems, but still
 in a similar range while using about the same number of leaves but about a
@@
 \section{Summary}
-In this article, we compared continuous and SGMM models to 
+\label{sec:summary}
+
+In this article, we compared continuous and SGMM models with
 two types of semi-continuous hidden Markov models, one using a single codebook 
-and the other using multiple codebooks based on a two-level phonetic decision tree.
-%
-While the first could not produce convincing results, the multiple-codebook 
-architecture shows promising performance, especially for limited training data. 
-%
-% TODO This sentence only for SI-84/half experiments
-The experiments on the WSJ data need to be extended to the full training set, and
-the parameter settings calibrated on the development set.
+and the other using multiple codebooks based on a two-level phonetic decision tree,
+similar to~\cite{prasad2004t2b}.
+Using the single-codebook system we were not able to obtain competitive
+results, but using a multiple-codebook system 
+we were able to get better results than a traditional system.  Interestingly,
+for both the single-codebook or multiple-codebook systems, we saw better
+results with full rather than diagonal covariances.
 
-Although the current performance is below the state-of-the-art, the
-rather simple theory and low computational complexity, paired with the possibility
-of solely acoustic adaptation make multiple-codebook semi-continuous acoustic
-models an attractive alternative for low-resource applications -- both in terms
-of computational power and training data; especially for embedded applications,
-where the codebook evaluation can be implemented in hardware, thus eliminating
-the time difference in evaluating full and diagonal covariance Gaussians.
-%
-The take home message from the experiments is that the choice of acoustic model
-should be made based on a resource constraint (number of Gaussians, available
-compute power), and that both continuous and semi-continuous models can do the job.
+Our best multiple-codebook results reported here incorporate two different forms of 
+smoothing, one using the decision tree to interpolate count statistics among closely
+related leaves, and one smoothing the weights across iterations to encourage 
+convergence to a better local optimum.  In future work we intend to further investigate
+these smoothing procedures and clarify how important they are.
 
-For future work, effect of adaptation and especially discriminative training
-with respect to the codebook architecture need to be explored.
-Furthermore, the SGMM framework can be extended to two-level architecture to 
-reduce computational effort in both training and decoding.
+If we continue to see good results from these types of systems, 
+we may investigate ways to speed up the Gaussian computation for the 
+full-covariance multiple-codebook tied mixture systems (e.g. using
+diagonal Gaussians in a pre-pruning phase, as in SGMMs).  To get state-of-the-art
+error rates we would also have to 
+implement linear-transform estimation (e.g. Constrained MLLR for adaptation)
+for these types of models.  We also intend to investigate the interaction of these
+methods with discriminative training and system combination; and we will
+consider extending the SGMM framework to a two-level architecture similar to
+the one we used here.
 

### 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
Minor futher changes to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@532 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 In this work, we compare classic and multiple-codebook semi-continuous models
-that use full covariance matrices with continuous HMMs and subspace Gaussian mixture models.
-%
-Experiments using on the RM and WSJ corpora show that a semi-continuous system
-can still yield competitive results while using fewer Gaussian components.
-Our best semi-continous systems included full-covariance codebook Gaussians,
-multiple codebooks, and smoothing of the weight parameters.
+using diagonal and full covariance matrices with continuous HMMs and subspace Gaussian 
+mixture models.  
+Experiments using on the RM and WSJ corpora show that while a classical semi-continous
+system does not perform as well as a continuous one, multiple-codebook semi-continuous
+systems can perform better, particular when using full-covariance Gaussians.
 \end{abstract}

### 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
Further edits to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@533 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 \documentclass{article}
-\usepackage{spconf,amsmath,amssymb,graphicx,ngerman,bm}
+\usepackage{bm,spconf,amsmath,amssymb,graphicx,ngerman}
 \selectlanguage{USenglish}
@@
 We were motivated by a number of factors to look again at semi-continuous models.
-One is that we are interested in applications where there may be a small amount
+One is that we are interested in applications where there is a small amount
 of training data (or a small amount of in-domain training data), and since
@@
 allocate more components by splitting components at every iteration until a 
-pre-determined number of components is reached (in this case 9000). We allocate 
+pre-determined total number of components is reached. We allocate 
 the Gaussians proportional to a small power (e.g.~0.2) of the state's 
@@
 \begin{equation}
-p(\x | j) = \sum_{i=1}^{N} c_{ji} \nv(\x; \m_i, \k_i) 
+p(\x | j) = \sum_{i=1}^{N} c_{ji} \; \nv(\x; \m_i, \k_i) 
 \end{equation}
@@
 using the tree mapping to obtain preliminary codebooks.
-The target size $N_k$ of codebook $k$ is determined with respect to a power
-of the occupancies of the respective leaves as
+The target size $N_k$ of codebook $k$ is determined with respect to 
+the occupancies of the respective leaves as
 \begin{equation}
@@
 Gaussians from a subspace spanned by a universal background model (UBM)
-and state specific transformations. In principle, the SGMM emission pdfs can
+and state specific transformations.   The SGMM emission pdfs can
 be computed as
@@
 \end{eqnarray}
-where the covariance matrices $\k_i$ are shared between all leafs $j$. The
+where the covariance matrices $\k_i$ are shared between all leaves $j$. The
 weights $w_{ji}$ and means $\m_{ji}$ are derived from $\v_j$ together with
@@
 %
-A similar yet different interpolation scheme without normalization was used in 
+A similar interpolation scheme without normalization was used in 
 \cite{schukattalamazzini1994srf,schukattalamazzini1995as} along with a different
@@
 Another typical problem that arises when training semi-continuous models is
-that the weights tend to converge to a local optimum over the iterations. 
-This is due to the fact that the codebook parameters change rather slowly 
-as the statistics are collected from all respective tree leaves, but the 
-leaf specific weights converge much faster.
+that the weights tend to converge to a poor local optimum over the iterations. 
+Our intuition is that the weights tend to overfit to the Gaussian parameters
+in such a way that the model quickly converges with the
+GMM parameters very close to the initialization point.
 %
-To compensate for this, we smooth the newly estimated parameters $\Theta^{(i+1)}$ with the ones from the prior iteration using an interpolation weight $\rho$
+To try to counteract this effect, we smooth the newly estimated parameters 
+$\Theta^{(i+1)}$ with the ones from the prior iteration using an interpolation weight $\rho$
 \begin{equation}
-\hat\Theta^{(i+1)} \leftarrow (1 - \rho) \, \Theta^{(i+1)} + \rho \, \Theta^{(i)} % \quad .
+\hat\Theta^{(i)} \leftarrow (1 - \rho) \, \Theta^{(i)} + \rho \, \Theta^{(i-1)} % \quad .
 \end{equation}
@@
 
-The experiments presented in this article are computed on the DARPA Resource 
+The experiments presented in this article used the DARPA Resource 
 Management Continuous Speech Corpus (RM) \cite{price1993rm} and the Wall Street
@@
 
-On the acoustic frontend, we extract 13 Mel-frequency cepstral coefficients (MFCCs),
+For the acoustic frontend, we extract 13 Mel-frequency cepstral coefficients (MFCCs),
 apply cepstral mean normalization, and compute deltas and delta-deltas.  For

### 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
Further edits to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@534 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 which leads to an exponentially decreasing impact of the initial parameter set.
+We implemented this for all the parameters but found it only helped when
+applied to just the weights,
+which is consistent with our interpretation that stopping the weights from overfitting
+too fast allows the Gaussians to ``move farther'' than they otherwise would from
+the initial estimate.  We note that other practitioners have also used mechanisms that
+prevent the weights from changing too fast~\footnote{James Droppo, personal communication} and
+have found that this improved results.
 

### 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
Further edits to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@535 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

--- 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
@@
 which leads to an exponentially decreasing impact of the initial parameter set.
+We implemented this for all the parameters but found it only helped when
+applied to just the weights,
+which is consistent with our interpretation that stopping the weights from overfitting
+too fast allows the Gaussian parameters to move farther away from the initialization
+than they otherwise would.
+We note that other practitioners have also used mechanisms that
+prevent the weights from changing too fast and
+have found that this improved results\footnote{Jasha Droppo, personal communication.}.
 

### 2011-09-28 misc/papers/icassp12_tied/icassp12.tex
Final edits to paper.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/discrim@536 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 however, using a two-level tree and multiple codebooks similar to~\cite{prasad2004t2b},
-we were able to get better performance than a traditional system (although not
+in one setup we were able to get better performance than a traditional system (although not
 as good as SGMMs).  For both traditional and multiple-codebook configurations,
@@
 results, but using a multiple-codebook system 
-we were able to get better results than a traditional system.  Interestingly,
+we were able to get better results than a traditional system, for one
+database (RM).  Interestingly,
 for both the single-codebook or multiple-codebook systems, we saw better
@@
 % Korbinian: We might need that ;)
-\footnotesize
+%\footnotesize
+
 \bibliographystyle{IEEEbib}

### 2011-09-29 misc/papers/icassp12_lat/paper.tex
Final changes to paper.tex
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@539 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
+% Note: the original submission was rev. 530...
 % Template for SLT-2006 paper; to be used with:
@@
 % ------
-\title{Generating exact lattices in a WFST framework}
+\title{Generating exact lattices in the WFST framework}
 
@@
 \name{ \em Daniel Povey$^1$, Mirko Hannemann$^{1,2}$,  \\
-  \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^4$, {Arnab Ghoshal}$^5$, {Milos Janda}$^2$, {Stefan Kombrink}$^2$, \\
+ \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^{2,4}$, {Arnab Ghoshal}$^5$, {Milo\v{s} Janda}$^2$, {Martin Karafi\'{a}t}$^2$, {Stefan Kombrink}$^2$, \\
   \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
-    \thanks{Thanks to Sanjeev Khudanpur for his help in preparing the paper, and to Honza Cernocky, Renata Kohlova, Martin Karafiat and Tomas Mikolov for their help relating to the Kaldi'11 workshop at BUT.}}
+    \thanks{Thanks to Honza \v{C}ernock\'{y}, Renata Kohlov\'{a}, and Tom\'{a}\v{s} Ka\v{s}p\'{a}rek for their help relating to the Kaldi'11 workshop at BUT, and to Sanjeev Khudanpur for his help in preparing the paper.
+Researchers at BUT were partly supported by Technology Agency of the Czech Republic grant No.
+TA01011328, Czech Ministry of Education project No.
+MSM0021630528, and Grant Agency of the Czech Republic project No.
+102/08/0707.}}
 
@@
 
-\address{$^1$ Microsoft Research, Redmond, WA, {\normalsize \tt dpovey@microsoft.com} \\ 
-         $^2$ Brno University of Technology, Czech Republic, {\normalsize \tt ihannema@fit.vutbr.cz} \\ 
-         $^3$ CRIM, Montreal, Canada \ \  $^4$ SRI International, Menlo Park, CA, USA \\ 
-         $^5$ University of Edinburgh, U.K. \ \  $^6$  IDIAP, Martigny, Switzerland  \\ 
-         $^7$ Tsinghua University, Beijing, China \ \ $^8$ Karlsruhe Institute of Technology, Germany \\ 
-         $^9$ Pattern Recognition Lab, University of Erlangen-Nuremberg, Germany}
+\address{{ $^1$ Microsoft Research, Redmond, WA, {\normalsize \tt dpovey@microsoft.com} }\\ 
+         { $^2$ Brno University of Technology, Czech Republic, {\normalsize \tt ihannema@fit.vutbr.cz}} \\ 
+         { $^3$ CRIM, Montreal, Canada \ \  $^4$ SRI International, Menlo Park, CA, USA} \\ 
+         { $^5$ University of Edinburgh, U.K. \ \  $^6$  IDIAP, Martigny, Switzerland } \\ 
+         { $^7$ Tsinghua University, Beijing, China \ \ $^8$ Karlsruhe Institute of Technology, Germany} \\ 
+         { $^9$ Pattern Recognition Lab, University of Erlangen-Nuremberg, Germany}}
 
@@
 %
-\pagestyle{plain} % was set to empty in spconf.sty... this makes page number appear.
+%\pagestyle{plain} % was set to empty in spconf.sty... this makes page number appear.
 
@@
 \end{equation}
-where $\circ$ is WFST composition (note: view $\HCLG$ as a single symbol).
+where $H$, $C$, $L$ and $G$ represent the HMM structure, phonetic
+context-dependency, lexicon and grammar respectively, 
+and $\circ$ is WFST composition (note: view $\HCLG$ as a single symbol).
 For concreteness we will speak of ``costs'' rather
@@
 log-probability.  A WFST has a set of states with one distinguished
-start state\footnote{This is the formulation that corresponds best with the toolkit we use.},
+start state;\footnote{This is the formulation that corresponds best with the toolkit we use.},
 each state has a final-cost (or $\infty$ for non-final states);
-and there is a set of arcs, where each arc has a weight
+and there is a set of arcs between the states, where each arc has a weight
 (just think of this as a cost for now), an input label and an output
@@
 \includegraphics[height=2.1in,angle=270]{figures/acceptor_utterance.eps}
+\vspace*{-0.03in}
   \caption{Acceptor $U$ describing the acoustic scores of an utterance}
-\vspace*{-0.2in}
+\vspace*{-0.15in}
 \label{fig:acceptor}
@@
 
-Imagine we have an utterance, of length $T$, and want to ``decode'' it (i.e. find
-the most likely word sequence, and the corresponding alignment).  A WFST
+Imagine we want to ``decode'' an utterance of $T$ frames, i.e. we want to
+find the most likely word sequence, and the corresponding alignment.  A WFST
 interpretation of the decoding problem is as follows.  We construct an
@@
 WFST with identical input and output symbols).  It has $T{+}1$ states,
-and an arc for each combination of (time, context-dependent HMM state).  The
+with an arc for each combination of (time, context-dependent HMM state).  The
 costs on these arcs are the negated and scaled acoustic log-likelihoods.
@@
 When we do Viterbi decoding with beam-pruning, we are finding the best path through $B$. 
-
+%
 Since the beam pruning is a part of any practical search procedure and cannot
@@
 \end{itemize}
-Actually, this definition is not completely formalized-- we have left
-the second condition (accuracy of scores and alignments) a little vague.
-Let the lattice be $L$.  The way we would like to state this requirement is:
+We need to be a little more precise about what we mean by the scores
+and alignments being ``accurate''.  Let the lattice be $L$.  The 
+way we would like to state this requirement is:
 \begin{itemize}
   \item For every path in $L$, the score and alignment corresponds to 
-   the best-scoring path in in $B$ for the corresponding word
+   the best-scoring path in $B$ for the corresponding word
    sequence\footnote{Or one of the best-scoring paths, in case of a tie.}.
@@
 By ``word-sequence'' we mean a sequence of whatever symbols are on the
-output of $\HCLG$.  In our experiments these symbols represent words, but not including
-silence, which we represent via alternative paths in $L$.
+output of $\HCLG$.  In our experiments these output symbols represent words, but
+silences do not appear as output symbols (they are
+represented via alternative paths in $L$).
 
@@
 a different copy of the lexical tree for each preceding word, 
-assuming the word-pair assumption holds, in order to generate an accurate lattice it is
+assuming the word-pair assumption holds, in order to generate an accurate lattice, it is
 sufficient to store a single Viterbi back-pointer at the word level; the entire set of
@@
 it takes account of both the forward and backward parts of the cost.
-The paper also reported experiments with a different method that did not require any
-phone-pair assumption; these experiments showed that the more
-efficient method that relied on the phone-pair assumption
-had almost the same lattice oracle error rate as their more efficient method.  However, the 
+The paper also reported experiments with an accurate, ``reference'' method that did not require any
+phone-pair assumption; these experiments showed that the main method they were describing
+had almost the same lattice oracle error rate as the reference method.  However, the 
 experiments did not evaluate how much impact the assumption had on the accuracy of the scores, 
@@
 of not just a single back-pointer in each state, but the $N$ best back-pointers for the $N$ top-scoring distinct
-preceding.  Therefore, this algorithm has more in common with the sentence N-best
-algorithm than with the Viterbi algorithm.  By limiting $N$ to be quite small (e.g. $N{=}5$)
+word histories.  Therefore, this algorithm has more in common with the sentence N-best
+algorithm than with the Viterbi algorithm.  By limiting $N$ to be quite small (e.g. $N{=}5$),
 the algorithm was made efficient, but at the cost of losing word sequences
@@
 we have not implemented this simple version.
-In this case our algorithm would be quite similar
+In this case, our algorithm would be quite similar
 to~\cite{efficient_general}, except at the state level rather than the phone level.
@@
 the current frame, for $\epsilon$-input
-arcs).   As in~\cite{efficient_general}, at the end of the utterance
+arcs).   As in~\cite{efficient_general}, at the end of the utterance,
 we prune the resulting graph to discard any paths that are not within the beam $\alpha$ of the best cost. 
@@
 
-As in~\cite{efficient_general}, to save memory we actually do the pruning
+As in~\cite{efficient_general}, to save memory, we actually do the pruning
 periodically rather than waiting for the end of the file (we do it every 25 frames).
@@
 the determinization process can introduce a lot of unlikely arcs.  In fact, for particular
-utterances the determinization process can cause the lattice to expand enough to
+utterances, the determinization process can cause the lattice to expand enough to
 exhaust memory.  To deal with this, we currently just detect when determinization
 has produced more than a pre-set maximum number of states, then we prune with a tighter
-beam and try again.  In future we may try more sophisticated methods such as a
+beam and try again.  In future, we may try more sophisticated methods such as a
 determinization algorithm that does pruning itself.
@@
 We define the $\otimes$ operation as $(c, s) \otimes (c', s') = (c+c', (s,s'))$, where
-$(s,s')$ is $s$ and $s'$ appended together.  We define the $\oplus$ operation so that
+$(s,s')$ is a concatenation of $s$ and $s'$.  We define the $\oplus$ operation so that
 it returns whichever pair has the smallest cost: that is, $(c,s) \oplus (c',s')$ 
@@
 equal to the input symbol (word) on the corresponding arc of $Q$, and
-the weights on the arcs of $E$ containing both the the weight and the output symbol (p.d.f.),
+the weights on the arcs of $E$ containing both the weight and the output symbol (p.d.f.),
 if any, on the corresponding arcs of $Q$.  Let $D = \mathrm{det}(\mathrm{rmeps}(E))$.
@@
 We now summarize the lattice generation algorithm.  During decoding, we create a data-structure
-corresponding to a full state-level lattice.  That is, for every arc of $\HCLG$ we traverse
+corresponding to a full state-level lattice.  That is, for every arc of $\HCLG$, we traverse
 on every frame, we create a separate arc in the state-level lattice.  These arcs
@@
 We will briefly describe another unique aspect of our
-algorithm.  Determinization algorithms involve weighted subsets of states,
-of the form
+algorithm.  Determinization algorithms involve weighted subsets of states, e.g.:
 \begin{equation}
@@
 training data; we test on the November 1992 evaluation data.  
-For these experiments we generate lattices with the bigram language model
+For these experiments, we generate lattices with the bigram language model
 supplied with the WSJ database, and for rescoring experiments we
@@
 decoding and $1/15$ for LM rescoring.  
-For simplicity we used a decoder that
+For simplicity, we used a decoder that
 does not support a ``maximum active states'' option, so the only variables
@@
    \caption{\vspace*{-0.2in} Varying lattice beam $\alpha$ (Viterbi pruning beam fixed at 15) }
+  \vspace{-0.08in}
   \label{fig:latbeam}
@@
 $\alpha=8$, so a value of $\alpha$ anywhere between about 4 and 8 is sufficient for LM rescoring 
-and still does not slow down decoding too much.
+and still does not slow down decoding too much.  We do not display the real-time factor of the non-lattice-generating
+decoder on this data (2.26) as it was actually slower than the lattice generating decoder; this is
+presumably due to the overhead of reference counting.
 Out of vocabularly words (OOVs) provide a floor on the lattice oracle error rate:
@@
   \label{fig:viterbibeam}
+  \vspace{-0.08in}
 \end{figure}
@@
 rescore utt wrong: 209
+
+2.259 is baseline RT factor w/ decoding beam of 15, with simple-decoder.
+
+note: paper# is 3887
+passwd  CF8FECE1

### 2011-12-06 misc/papers/asru11_toolkit_poster/kaldi-poster.tex
Adding template for Kaldi poster
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@642 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}
\usepackage{grffile}
\mode<presentation>{\usetheme{kaldi1}}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{amsmath,amsthm, amssymb, latexsym}
%\usepackage{times}\usefonttheme{professionalfonts}  % obsolete
%\usefonttheme[onlymath]{serif}
\boldmath
\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}
% change list indention level
% \setdefaultleftmargin{3em}{}{}{}{}{}

%\usepackage{snapshot} % will write a .dep file with all dependencies, allows for easy bundling

\usepackage{array,booktabs,tabularx}
\newcolumntype{Z}{>{\centering\arraybackslash}X} % centered tabularx columns
\newcommand{\pphantom}{\textcolor{kaldiblack}} % phantom introduces a vertical space in p formatted table columns??!!

\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\graphicspath{{figures/}}

\title{\huge SURF-Face: Face Recognition Under Viewpoint Consistency Constraints}
\author{Philippe Dreuw, Pascal Steingrube, Harald Hanselmann and Hermann Ney}
\institute[RWTH Aachen University]{Human Language Technology and Pattern Recognition, RWTH Aachen University, Aachen, Germany}
\date[Sep. 8th, 2009]{Sep. 8th, 2009}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newlength{\columnheight}
\setlength{\columnheight}{105cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frame}
  \begin{columns}
    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth}  % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content            
            \begin{block}{Introduction}
              \begin{itemize}
              \item Most face recognition approaches are sensitive to registration errors
                \begin{itemize}
                \item rely on a very good initial alignment and illumination
                \end{itemize}
              \item We propose/analyze:
                \begin{itemize}
                \item grid-based and dense extraction of local features
                \item block-based matching accounting for different\\
	                  viewpoints and registration errors
                \end{itemize}
              \end{itemize}              
            \end{block}
            \vfill
            \begin{block}{Feature Extraction}
              \begin{columns}
                \begin{column}{.55\textwidth}
                  \begin{itemize}
                  \item Interest point based feature extraction
                    \begin{itemize}
                    \item SIFT or SURF interest point detector
                    \item leads to a \alert{very sparse} description
                    \end{itemize}
                  \item Grid-based feature extraction
                    \begin{itemize}
                    \item overlaid regular grid
                    \item leads to a \alert{dense} description
                    \end{itemize}
                  \end{itemize}
                \end{column}
                \begin{column}{.44\textwidth}
                  \centering
                  \begin{tabularx}{\linewidth}{ZZZ}
                     Orig.
                     &
                     IP
                     &
                     Grid
                     \\
                    \includegraphics[width=0.95\linewidth]{images/viola/m-012-1.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-ip/m-012-1.pngsurf.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-4/m-012-1.pngsurf.png}
                    \\
                    \includegraphics[width=0.95\linewidth]{images/viola/m-012-2.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-ip/m-012-2.pngsurf.png}
                    &
                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-4/m-012-2.pngsurf.png}
                    \\
                  \end{tabularx}
                \end{column}
              \end{columns}
              \vskip-1ex
            \end{block}
            \vfill
            \begin{block}{Feature Description}
              \begin{itemize}
              \item Scale Invariant Feature Transform (SIFT)
                \begin{itemize}
                \item 128-dimensional descriptor, histogram of gradients, scale invariant
                \end{itemize}
              \item Speeded Up Robust Features (SURF)
                \begin{itemize}
                \item 64-dimensional descriptor, histogram of gradients, scale invariant
                \end{itemize}
              \item face recognition: invariance w.r.t. rotation is often not necessary
                \begin{itemize}
                \item rotation dependent upright-versions U-SIFT, U-SURF-64, U-SURF-128
                \end{itemize}
              \end{itemize}
            \end{block}
            \vfill
            \begin{block}{Feature Matching}
              \begin{itemize}
              \item Recognition by Matching
                \begin{itemize}
                \item nearest neighbor matching strategy
                \item descriptor vectors extracted at keypoints in a test image $X$ are compared to all descriptor vectors extracted at keypoints from the reference images $Y_n, n=1,\cdots,N$ by the Euclidean distance
                \item decision rule:
                  \begin{align*}
                    X \rightarrow r(X) & = \arg \max_{c} \Big\{ \max_n \big\{ \sum_{x_i \in X} \delta(x_i,Y_{n,c})\big\} \Big\} 
                  \end{align*}
                \item additionally, a ratio constraint is applied in $\delta(x_i,Y_{n,c})$
                \end{itemize}
              \item Viewpoint Matching Constraints
                \begin{itemize}
                \item maximum matching: unconstrained
                \item grid-based matching: absolute box constraints
                \item grid-based best matching: absolute box constraints, overlapping
                \end{itemize}
              \item Postprocessing
                \begin{itemize}
                \item RANSAC-based outlier removal
                \item RANSAC-based system combination
                \end{itemize}
              \end{itemize}
            \end{block}
            \vfill
            \begin{block}{Matching Examples for the AR-Face and CMU-PIE Database}
              \begin{figure}
                \footnotesize
                \centering
                \begin{tabular}{p{.09\linewidth} | p{.12\linewidth} | p{.12\linewidth} | p{.12\linewidth} || p{.12\linewidth} | p{.12\linewidth} | p{.12\linewidth} | p{.09\linewidth} }
                  Feature
                  &
                  Maximum
                  & 
                  Grid
                  & 
                  Grid-Best
                  &
                  Maximum
                  & 
                  Grid
                  & 
                  Grid-Best
                  & 
                  Feature
                  \\
                  \hline
                  SIFT 
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/maximum_m-005-17.pgm--w-052-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/grid_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/grid-best_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/maximum_07-27-22.pgm--04-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/grid_07-27-22.pgm--07-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/grid-best_07-27-22.pgm--07-27-08}
                  &
                  SURF
                  \\
                  \hline
                  U-SIFT
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/maximum_m-005-17.pgm--w-052-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/grid_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/grid-best_m-005-17.pgm--m-005-4}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/maximum_07-27-22.pgm--04-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/grid_07-27-22.pgm--07-27-08}
                  &
                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/grid-best_07-27-22.pgm--07-27-08}
                  &
                  U-SURF
                  \\
                  \hline
                \end{tabular}
              \end{figure}
              \begin{itemize}
              \item Matching results for the AR-Face (left) and the CMU-PIE database (right)
                \begin{itemize}
                \item maximum matching show false classification examples
                \item grid matchings show correct classification examples
                \item upright descriptor versions reduce the number of false matches
                \end{itemize}
              \end{itemize}
            \end{block}
          }
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column

    % ---------------------------------------------------------%
    % Set up a column 
    \begin{column}{.49\textwidth}
      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
            % ---------------------------------------------------------%
            % fill each column with content

            \begin{block}{Databases}
              \begin{columns}
                \begin{column}{.6\textwidth}
                  \begin{itemize}
                  \item AR-Face 
                    \begin{itemize}
                    \item variations in illumination
                    \item many different facial expressions
                    \end{itemize}
                  \item CMU-PIE
                    \begin{itemize}
                    \item variations in illumination (frontal images from the illumination subset)
                    \end{itemize}
                  \end{itemize}
                \end{column}
                \begin{column}{.39\textwidth}
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/train/png/occlusions/arneutral/m-016-1}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/train/png/cropped/m-016-4}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/test/png/occlusions/ar1sun/m-016-8}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/test/png/occlusions/ar1scarf/m-016-11}
                  \vskip3ex

                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-02}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-17}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-18}
                  \-
                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-20}
                \end{column}
              \end{columns}
            \end{block}
            \vfill
            \begin{block}{Results: Manually Aligned Faces}
              \begin{itemize}
              \item AR-Face: 110 classes, 770 train, 770 test
              \end{itemize}
              \vskip-0.5ex
              \begin{table}
                \centering
                \small
                \begin{tabular}{@{} p{.2\linewidth} p{.18\linewidth} p{.25\linewidth} r r r @{}}
                  \toprule 
                  Descriptor &  Extraction        & \multicolumn{1}{p{.2\linewidth}}{\# Features}           & \multicolumn{3}{c @{}}{Error Rates [\%]}         \\
                  \cmidrule(l){4-6}
                             &                    &                                                      & \multicolumn{1}{c}{Maximum} & \multicolumn{1}{c}{Grid} & \multicolumn{1}{c @{}}{Grid-Best} \\
                  \cmidrule(r){1-1}  \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                               \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
                  SURF-64    & IPs                & \pphantom{1}64 $\times$ 5.6 (avg.)                    & 80.64     &  84.15  & 84.15     \\
                  SIFT       & IPs                & 128 $\times$ 633.78 (avg.)                           & 1.03      &  95.84  & 95.84      \\
                  \addlinespace
                  SURF-64    & 64x64-2 grid       & \pphantom{1}64 $\times$ 1024                          & 0.90      &  0.51   & 0.90      \\  
                  SURF-128   & 64x64-2 grid       & 128 $\times$ 1024                                    & 0.90      &  0.51   & 0.38       \\
                  SIFT       & 64x64-2 grid       & 128 $\times$ 1024                                    & 11.03     &  0.90   & 0.64       \\
                  \addlinespace
                  U-SURF-64  & 64x64-2 grid       & \pphantom{1}64 $\times$ 1024                          & 0.90      &  1.03   & 0.64      \\ 
                  U-SURF-128 & 64x64-2 grid       & 128 $\times$ 1024                                    & 1.55      &  1.29   & 1.03       \\ 
                  U-SIFT     & 64x64-2 grid       & 128 $\times$ 1024                                    & \textbf{0.25}  &  \textbf{0.25}   & \textbf{0.25}  \\
                  % \midrule
                  % \multicolumn{3}{l}{Modular PCA \cite{TanChen:subpatternPCA:Neurocomputing2005}}                         & 14.14 \\
                  % \multicolumn{3}{l}{Adaptively weighted Sub-Pattern PCA \cite{TanChen:subpatternPCA:Neurocomputing2005}} & 6.43  \\
                  % \multicolumn{3}{l}{DCT \cite{ekenel:LocalapFaceRecog:cvprbw2006}}                                       & 4.70  \\
                  \bottomrule
                \end{tabular}
              \end{table}

              \vskip1ex
              \begin{itemize}
              \item CMU-PIE: 68 classes, 68 train (``one-shot'' training), 1360 test
              \end{itemize}
              \vskip-0.5ex
              \begin{table}
                \centering
                \small
                \begin{tabular}{@{} p{.2\linewidth} p{.18\linewidth} p{.25\linewidth} r r r @{}}
                  \toprule 
                  Descriptor              &  Extraction    & \# Features                            & \multicolumn{3}{c @{}}{Error Rates [\%]}      \\
                  \cmidrule(l){4-6}
                                          &                    &                                    & \multicolumn{1}{c}{Maximum} & \multicolumn{1}{c}{Grid} & \multicolumn{1}{c @{}}{Grid-Best} \\
                  \cmidrule(r){1-1}  \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                          \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
                  SURF-64    & IPs           & \pphantom{1}64 $\times$ 6.80 (avg.)                  & 93.95         & 95.21             & 95.21     \\  
                  SIFT       & IPs           & 128 $\times$ 723.17 (avg.)                           & 43.47         & 99.33             & 99.33     \\  
                  \addlinespace
                  SURF-64    & 64x64-2 grid  & \pphantom{1}64  $\times$ 1024                        &   13.41       &  4.12             &  7.82     \\
                  SURF-128   & 64x64-2 grid  & 128 $\times$ 1024                                    &  12.45        &  3.68             & 3.24      \\  
                  SIFT       &  64x64-2 grid & 128 $\times$ 1024                                    & 27.92         &  7.00             & 9.80      \\  
                  \addlinespace
                  U-SURF-64  & 64x64-2 grid  & \pphantom{1}64  $\times$ 1024                        & \textbf{3.83} &   \textbf{0.51}   & \textbf{0.66}    \\  
                  U-SURF-128 & 64x64-2 grid  & 128 $\times$ 1024                                    & 5.67          &  0.95             & 0.88      \\  
                  U-SIFT     & 64x64-2 grid  & 128 $\times$ 1024                                    & 16.28         &  1.40             & 6.41      \\
                  % \midrule
                  % \midrule
                  % \multicolumn{3}{l}{Spherical Harmonics \cite{ZhangSamaras:SphericalHarmonics:pami2006}}                              & 1.80 \\
                  % \multicolumn{3}{l}{Modeling Phase Spectra using GMM \cite{MitraSavvidesBrockwell:ModelingPhaseSpectra:lncs2005}}    & 12.83 \\
                  \bottomrule
                \end{tabular}
              \end{table}              
            \end{block}
            \vfill
            \begin{block}{Results: Unaligned Faces}
              \begin{columns}
                \begin{column}{.59\textwidth}
                  \begin{itemize}
                  \item Automatically aligned by Viola \& Jones
                  \end{itemize}
                  \vskip-0.5ex
                  \begin{table}
                    \centering
                    \small
                    \begin{tabular}{@{} p{.4\linewidth} r r @{}}
                      \toprule 
                      Descriptor  &      \multicolumn{2}{c @{}}{Error Rates [\%]}      \\
                      \cmidrule(l){2-3}   
                      &   AR-Face       & CMU-PIE  \\
                      \cmidrule(r){1-1}  \cmidrule(lr){2-2}  \cmidrule(l){3-3}  
                      SURF-64     &   5.97          & 15.32    \\ 
                      SURF-128    &   5.71          & 11.42    \\ 
                      SIFT        &   5.45          & 8.32     \\  
                      \addlinespace
                      U-SURF-64   &   5.32          & 5.52     \\  
                      U-SURF-128  &   5.71          & \textbf{4.86}  \\ 
                      U-SIFT      &   \textbf{4.15} & 8.99     \\  
                      \bottomrule
                    \end{tabular}
                  \end{table}
                \end{column}
                \begin{column}{.39\textwidth}                
                  \vskip-3ex
                  \begin{itemize}
                  \item Manually aligned faces

                      \includegraphics[width=0.9\linewidth]{hanselmann-thesis/slides/figures/alignment_cmu}

                  \item Unaligned faces

                      \includegraphics[width=0.9\linewidth]{hanselmann-thesis/slides/figures/unalignment_cmu}

                  \end{itemize}
                \end{column}
              \end{columns}
            \end{block}
            \vfill
            \begin{block}{Results: Partially Occluded Faces}
              \begin{itemize}
              \item AR-Face: 110 classes, 110 train (``one-shot'' training), 550 test
              \end{itemize}
              \vskip-0.5ex
              \begin{table}
                \small
                \centering
                \begin{tabular}{@{} l @{} r r r r r r@{}}
                  \toprule 
                  Descriptor          & \multicolumn{6}{c @{}}{Error Rates [\%]}                                                                 \\
                  \cmidrule(l){2-7}
                  & \textit{AR1scarf}  & \textit{AR1sun}  & \textit{ARneutral} & \textit{AR2scarf} & \textit{AR2sun}  & Avg. \\
                  \cmidrule(r){1-1}     \cmidrule(lr){2-2}   \cmidrule(lr){3-3} \cmidrule(lr){4-4}   \cmidrule(lr){5-5}  \cmidrule(lr){6-6}  \cmidrule(l){7-7} 
                  SURF-64             & 2.72             & 30.00          & 0.00                & 4.54            & 47.27      & 16.90 \\        
                  SURF-128            & 1.81             & 23.63          & 0.00                & 3.63            & 40.90      & 13.99 \\
                  SIFT                & 1.81             & 24.54          & 0.00                & 2.72            & 44.54      & 14.72 \\
                  \addlinespace
                  U-SURF-64           & 4.54             & 23.63          & 0.00                & 4.54            & 47.27      & 15.99 \\
                  U-SURF-128          & 1.81             & \textbf{20.00} & 0.00                & 3.63            & 41.81      & 13.45 \\
                  U-SIFT              & \textbf{1.81}    & 20.90          & \textbf{0.00}       & \textbf{1.81}   & \textbf{38.18} & \textbf{12.54} \\
                  \cmidrule(r){1-1}     \cmidrule(lr){2-2}   \cmidrule(lr){3-3} \cmidrule(lr){4-4}   \cmidrule(lr){5-5}  \cmidrule(lr){6-6}  \cmidrule(l){7-7}
                  U-SURF-128+R        & 1.81             & 19.09          & 0.00                & 3.63             & 43.63     & 13.63 \\
                  U-SIFT+R            & 2.72             & \textbf{14.54} & 0.00                & \textbf{0.90}    & 35.45     & 10.72 \\
                  U-SURF-128+U-SIFT+R &  \textbf{0.90}   & 16.36          & \textbf{0.00}       & 2.72             & \textbf{32.72} & \textbf{10.54} \\       
                  % \midrule 
                  % \midrule 
                  % DCT \cite{ekenel:facialocclusion:icb2009}, baseline%
                  % & 8.2              & 61.8           & 7.3                & 16.4            & 62.7       & 31.28 \\
                  % DCT \cite{ekenel:facialocclusion:icb2009}, realigned%
                  % & 2.7              & 1.8            & 0.0                & 6.4             & 4.5        & 3.08  \\
                  \bottomrule
                \end{tabular}
              \end{table}
            \end{block}
            \vfill
            \begin{block}{Conclusions}
              \begin{itemize}
              \item Grid-based local feature extraction instead of interest points
              \item Local descriptors:
                \begin{itemize}
                \item upright descriptor versions achieved better results
                \item SURF-128 better than SURF-64
                \end{itemize}
              \item System robustness: manually aligned/unaligned/partially occluded faces
                \begin{itemize}
                \item SURF more robust to illumination
                \item SIFT more robust to changes in viewing conditions
                \end{itemize}
              \item RANSAC-based system combination and outlier removal
              \end{itemize}
            \end{block}
          }
          % ---------------------------------------------------------%
          % end the column
        \end{minipage}
      \end{beamercolorbox}
    \end{column}
    % ---------------------------------------------------------%
    % end the column
  \end{columns}
  \vskip1ex
  %\tiny\hfill\textcolor{ta2gray}{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php}}
  \tiny\hfill{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php} \hskip1em}
\end{frame}
\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Local Variables: 
%%% mode: latex
%%% TeX-PDF-mode: t
%%% End:

### 2011-12-07 misc/papers/asru11_toolkit_poster/kaldi-poster.tex
ASRU poster update
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@653 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 \boldmath
-\usepackage[orientation=portrait,size=a0,scale=1.4,debug]{beamerposter}
+\usepackage[orientation=landscape,size=a0,scale=1.4,debug]{beamerposter}
 % change list indention level
@@
 
-\title{\huge SURF-Face: Face Recognition Under Viewpoint Consistency Constraints}
-\author{Philippe Dreuw, Pascal Steingrube, Harald Hanselmann and Hermann Ney}
-\institute[RWTH Aachen University]{Human Language Technology and Pattern Recognition, RWTH Aachen University, Aachen, Germany}
-\date[Sep. 8th, 2009]{Sep. 8th, 2009}
+\title{\huge The Kaldi Speech Recognition Toolkit}
+\author{{Daniel Povey}$^1$, {Arnab Ghoshal}$^{2,3}$,
+  {Gilles Boulianne}$^4$, {Luk\'{a}\v{s} Burget}$^{5,6}$, {Ond\v{r}ej 
+    Glembek}$^5$, {Nagendra Goel}$^6$, \\
+  {Mirko Hannemann}$^5$, 
+  {Petr Motl\'{i}\v{c}ek}$^8$, {Yanmin Qian}$^9$, {Petr Schwarz}$^5$, 
+  {Jan Silovsk\'{y}}$^{10}$, {Georg Stemmer}$^{11}$, {Karel Vesel\'{y}}$^5$}
+\institute[]{   $^1$\,Microsoft Research, USA;
+   $^2$\,University of Edinburgh, UK;
+   $^3$\,Saarland University, Germany;
+   $^4$\,Centre de Recherche Informatique de Montr\'{e}al, Canada; 
+   $^5$\,Brno University of Technology, Czech Republic;
+   $^6$\,SRI International, USA;
+   $^7$\,Go-Vivace Inc., USA; $^8$\,IDIAP Research Institute, Switzerland; 
+   $^9$\,Tsinghua University, China;
+   $^{10}$\,Technical University of Liberec, Czech Republic; 
+   $^{11}$\,University of Erlangen-Nuremberg, Germany
+}
+\date[]{}
 
@@
 \newlength{\columnheight}
-\setlength{\columnheight}{105cm}
+\setlength{\columnheight}{65cm}
 
@@
             % fill each column with content            
-            \begin{block}{Introduction}
+            \begin{block}{Features of Kaldi}
+              \begin{columns}
+                \begin{column}{.44\textwidth}
               \begin{itemize}
-              \item Most face recognition approaches are sensitive to registration errors
-                \begin{itemize}
-                \item rely on a very good initial alignment and illumination
-                \end{itemize}
-              \item We propose/analyze:
-                \begin{itemize}
-                \item grid-based and dense extraction of local features
-                \item block-based matching accounting for different\\
-	                  viewpoints and registration errors
-                \end{itemize}
-              \end{itemize}              
-            \end{block}
-            \vfill
-            \begin{block}{Feature Extraction}
-              \begin{columns}
-                \begin{column}{.55\textwidth}
-                  \begin{itemize}
-                  \item Interest point based feature extraction
-                    \begin{itemize}
-                    \item SIFT or SURF interest point detector
-                    \item leads to a \alert{very sparse} description
-                    \end{itemize}
-                  \item Grid-based feature extraction
-                    \begin{itemize}
-                    \item overlaid regular grid
-                    \item leads to a \alert{dense} description
-                    \end{itemize}
+              \item Integration with Finite State Transducers
+              \item Extensive linear algebra support
+              \item Extensible design
+              \item Open license
+              \item Complete recipes
+              \item Thorough testing
               \end{itemize}
                 \end{column}
-                \begin{column}{.44\textwidth}
+                \begin{column}{.55\textwidth}
                   \centering
-                  \begin{tabularx}{\linewidth}{ZZZ}
-                     Orig.
-                     &
-                     IP
-                     &
-                     Grid
-                     \\
-                    \includegraphics[width=0.95\linewidth]{images/viola/m-012-1.png}
-                    &
-                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-ip/m-012-1.pngsurf.png}
-                    &
-                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-4/m-012-1.pngsurf.png}
-                    \\
-                    \includegraphics[width=0.95\linewidth]{images/viola/m-012-2.png}
-                    &
-                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-ip/m-012-2.pngsurf.png}
-                    &
-                    \includegraphics[width=0.95\linewidth]{images/viola/surf-64x64-4/m-012-2.pngsurf.png}
-                    \\
-                  \end{tabularx}
+                  \includegraphics[width=0.85\linewidth]{figures/kaldi-lib.pdf}
                 \end{column}
@@
             \vfill
-            \begin{block}{Feature Description}
+            \begin{block}{Feature Extraction}
               \begin{itemize}
-              \item Scale Invariant Feature Transform (SIFT)
-                \begin{itemize}
-                \item 128-dimensional descriptor, histogram of gradients, scale invariant
-                \end{itemize}
-              \item Speeded Up Robust Features (SURF)
-                \begin{itemize}
-                \item 64-dimensional descriptor, histogram of gradients, scale invariant
-                \end{itemize}
-              \item face recognition: invariance w.r.t. rotation is often not necessary
-                \begin{itemize}
-                \item rotation dependent upright-versions U-SIFT, U-SURF-64, U-SURF-128
-                \end{itemize}
+              \item 
               \end{itemize}
@@
             \vfill
-            \begin{block}{Feature Matching}
+            \begin{block}{Acoustic Modeling}
               \begin{itemize}
-              \item Recognition by Matching
-                \begin{itemize}
-                \item nearest neighbor matching strategy
-                \item descriptor vectors extracted at keypoints in a test image $X$ are compared to all descriptor vectors extracted at keypoints from the reference images $Y_n, n=1,\cdots,N$ by the Euclidean distance
-                \item decision rule:
-                  \begin{align*}
-                    X \rightarrow r(X) & = \arg \max_{c} \Big\{ \max_n \big\{ \sum_{x_i \in X} \delta(x_i,Y_{n,c})\big\} \Big\} 
-                  \end{align*}
-                \item additionally, a ratio constraint is applied in $\delta(x_i,Y_{n,c})$
-                \end{itemize}
-              \item Viewpoint Matching Constraints
-                \begin{itemize}
-                \item maximum matching: unconstrained
-                \item grid-based matching: absolute box constraints
-                \item grid-based best matching: absolute box constraints, overlapping
-                \end{itemize}
-              \item Postprocessing
-                \begin{itemize}
-                \item RANSAC-based outlier removal
-                \item RANSAC-based system combination
-                \end{itemize}
+              \item 
               \end{itemize}
@@
             \vfill
-            \begin{block}{Matching Examples for the AR-Face and CMU-PIE Database}
-              \begin{figure}
-                \footnotesize
-                \centering
-                \begin{tabular}{p{.09\linewidth} | p{.12\linewidth} | p{.12\linewidth} | p{.12\linewidth} || p{.12\linewidth} | p{.12\linewidth} | p{.12\linewidth} | p{.09\linewidth} }
-                  Feature
-                  &
-                  Maximum
-                  & 
-                  Grid
-                  & 
-                  Grid-Best
-                  &
-                  Maximum
-                  & 
-                  Grid
-                  & 
-                  Grid-Best
-                  & 
-                  Feature
-                  \\
-                  \hline
-                  SIFT 
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/maximum_m-005-17.pgm--w-052-4}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/grid_m-005-17.pgm--m-005-4}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-sift/grid-best_m-005-17.pgm--m-005-4}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/maximum_07-27-22.pgm--04-27-08}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/grid_07-27-22.pgm--07-27-08}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-surf/grid-best_07-27-22.pgm--07-27-08}
-                  &
-                  SURF
-                  \\
-                  \hline
-                  U-SIFT
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/maximum_m-005-17.pgm--w-052-4}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/grid_m-005-17.pgm--m-005-4}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/arface-usift/grid-best_m-005-17.pgm--m-005-4}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/maximum_07-27-22.pgm--04-27-08}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/grid_07-27-22.pgm--07-27-08}
-                  &
-                  \includegraphics[width=1.0\linewidth]{paper/bmvc09-surf/figures/matchings/cmupie-usurf/grid-best_07-27-22.pgm--07-27-08}
-                  &
-                  U-SURF
-                  \\
-                  \hline
-                \end{tabular}
-              \end{figure}
+            \begin{block}{Phonetic Decision Trees}
               \begin{itemize}
-              \item Matching results for the AR-Face (left) and the CMU-PIE database (right)
-                \begin{itemize}
-                \item maximum matching show false classification examples
-                \item grid matchings show correct classification examples
-                \item upright descriptor versions reduce the number of false matches
-                \end{itemize}
+              \item 
+              \item 
+              \item 
+              \item 
               \end{itemize}
@@
             \begin{block}{Databases}
-              \begin{columns}
-                \begin{column}{.6\textwidth}
               \begin{itemize}
-                  \item AR-Face 
+              \item Resource Management (RM)
                 \begin{itemize}
-                    \item variations in illumination
-                    \item many different facial expressions
+                \item 
+                \item 
                 \end{itemize}
-                  \item CMU-PIE
+              \item Wall Street Journal (WSJ)
                 \begin{itemize}
-                    \item variations in illumination (frontal images from the illumination subset)
+                \item
                 \end{itemize}
               \end{itemize}
+            \end{block}
+            \vfill
+            \begin{block}{Comparison with previously published results}
+              \begin{columns}
+                \begin{column}{.5\textwidth}
+                  \begin{table}
+                    \centering
+                    \small
+                    \begin{tabular}{l c c c c c} \toprule
+                      & \multicolumn{5}{c}{    Test set  }   \\ \cmidrule(l){2-6}
+                      &  Feb'89 &  Oct'89 & Feb'91 & Sep'92 & Avg   \\
+                      \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                               \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
+                      HTK    &  2.77    &  4.02   &  3.30  &  6.29  &  4.10 \\
+                      Kaldi  &  3.20    &  4.21   &  3.50  &  5.86  &  4.06 \\ 
+                      \bottomrule
+                    \end{tabular}            
+                  \end{table}
+                  \vskip1ex
                 \end{column}
-                \begin{column}{.39\textwidth}
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/train/png/occlusions/arneutral/m-016-1}
-                  \-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/train/png/cropped/m-016-4}
-                  \-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/test/png/occlusions/ar1sun/m-016-8}
-                  \-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/arface/test/png/occlusions/ar1scarf/m-016-11}
-                  \vskip3ex
-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-02}
-                  \-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-17}
-                  \-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-18}
-                  \-
-                  \includegraphics[width=0.22\linewidth]{hanselmann-databases/cmupie/test/png/cropped/00-27-20}
+                \begin{column}{.5\textwidth}
+                  \begin{table}
+                    \small
+                    \centering
+                    \begin{tabular}{@{} l @{} c c@{}}
+                      \toprule 
+                      & \multicolumn{2}{c}{    Test set  }   \\ \cmidrule(l){2-3}
+                      &  Nov'92      &    Nov'93  \\ 
+                      \cmidrule(lr){2-2} \cmidrule(r){3-3}
+                      Bell      &  11.9        &  15.4    \\
+                      HTK (+GD)  &  11.1        &  14.5   \\ 
+                      KALDI      &  11.8        &  15.0   \\ \bottomrule
+                    \end{tabular}
+                  \end{table}
                 \end{column}
@@
             \vfill
-            \begin{block}{Results: Manually Aligned Faces}
-              \begin{itemize}
-              \item AR-Face: 110 classes, 770 train, 770 test
-              \end{itemize}
-              \vskip-0.5ex
-              \begin{table}
-                \centering
-                \small
-                \begin{tabular}{@{} p{.2\linewidth} p{.18\linewidth} p{.25\linewidth} r r r @{}}
-                  \toprule 
-                  Descriptor &  Extraction        & \multicolumn{1}{p{.2\linewidth}}{\# Features}           & \multicolumn{3}{c @{}}{Error Rates [\%]}         \\
-                  \cmidrule(l){4-6}
-                             &                    &                                                      & \multicolumn{1}{c}{Maximum} & \multicolumn{1}{c}{Grid} & \multicolumn{1}{c @{}}{Grid-Best} \\
-                  \cmidrule(r){1-1}  \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                               \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
-                  SURF-64    & IPs                & \pphantom{1}64 $\times$ 5.6 (avg.)                    & 80.64     &  84.15  & 84.15     \\
-                  SIFT       & IPs                & 128 $\times$ 633.78 (avg.)                           & 1.03      &  95.84  & 95.84      \\
-                  \addlinespace
-                  SURF-64    & 64x64-2 grid       & \pphantom{1}64 $\times$ 1024                          & 0.90      &  0.51   & 0.90      \\  
-                  SURF-128   & 64x64-2 grid       & 128 $\times$ 1024                                    & 0.90      &  0.51   & 0.38       \\
-                  SIFT       & 64x64-2 grid       & 128 $\times$ 1024                                    & 11.03     &  0.90   & 0.64       \\
-                  \addlinespace
-                  U-SURF-64  & 64x64-2 grid       & \pphantom{1}64 $\times$ 1024                          & 0.90      &  1.03   & 0.64      \\ 
-                  U-SURF-128 & 64x64-2 grid       & 128 $\times$ 1024                                    & 1.55      &  1.29   & 1.03       \\ 
-                  U-SIFT     & 64x64-2 grid       & 128 $\times$ 1024                                    & \textbf{0.25}  &  \textbf{0.25}   & \textbf{0.25}  \\
-                  % \midrule
-                  % \multicolumn{3}{l}{Modular PCA \cite{TanChen:subpatternPCA:Neurocomputing2005}}                         & 14.14 \\
-                  % \multicolumn{3}{l}{Adaptively weighted Sub-Pattern PCA \cite{TanChen:subpatternPCA:Neurocomputing2005}} & 6.43  \\
-                  % \multicolumn{3}{l}{DCT \cite{ekenel:LocalapFaceRecog:cvprbw2006}}                                       & 4.70  \\
-                  \bottomrule
-                \end{tabular}
-              \end{table}
-
-              \vskip1ex
-              \begin{itemize}
-              \item CMU-PIE: 68 classes, 68 train (``one-shot'' training), 1360 test
-              \end{itemize}
-              \vskip-0.5ex
-              \begin{table}
-                \centering
-                \small
-                \begin{tabular}{@{} p{.2\linewidth} p{.18\linewidth} p{.25\linewidth} r r r @{}}
-                  \toprule 
-                  Descriptor              &  Extraction    & \# Features                            & \multicolumn{3}{c @{}}{Error Rates [\%]}      \\
-                  \cmidrule(l){4-6}
-                                          &                    &                                    & \multicolumn{1}{c}{Maximum} & \multicolumn{1}{c}{Grid} & \multicolumn{1}{c @{}}{Grid-Best} \\
-                  \cmidrule(r){1-1}  \cmidrule(lr){2-2}   \cmidrule(lr){3-3}                          \cmidrule(lr){4-4}            \cmidrule(lr){5-5}         \cmidrule(l){6-6}  
-                  SURF-64    & IPs           & \pphantom{1}64 $\times$ 6.80 (avg.)                  & 93.95         & 95.21             & 95.21     \\  
-                  SIFT       & IPs           & 128 $\times$ 723.17 (avg.)                           & 43.47         & 99.33             & 99.33     \\  
-                  \addlinespace
-                  SURF-64    & 64x64-2 grid  & \pphantom{1}64  $\times$ 1024                        &   13.41       &  4.12             &  7.82     \\
-                  SURF-128   & 64x64-2 grid  & 128 $\times$ 1024                                    &  12.45        &  3.68             & 3.24      \\  
-                  SIFT       &  64x64-2 grid & 128 $\times$ 1024                                    & 27.92         &  7.00             & 9.80      \\  
-                  \addlinespace
-                  U-SURF-64  & 64x64-2 grid  & \pphantom{1}64  $\times$ 1024                        & \textbf{3.83} &   \textbf{0.51}   & \textbf{0.66}    \\  
-                  U-SURF-128 & 64x64-2 grid  & 128 $\times$ 1024                                    & 5.67          &  0.95             & 0.88      \\  
-                  U-SIFT     & 64x64-2 grid  & 128 $\times$ 1024                                    & 16.28         &  1.40             & 6.41      \\
-                  % \midrule
-                  % \midrule
-                  % \multicolumn{3}{l}{Spherical Harmonics \cite{ZhangSamaras:SphericalHarmonics:pami2006}}                              & 1.80 \\
-                  % \multicolumn{3}{l}{Modeling Phase Spectra using GMM \cite{MitraSavvidesBrockwell:ModelingPhaseSpectra:lncs2005}}    & 12.83 \\
-                  \bottomrule
-                \end{tabular}
-              \end{table}              
-            \end{block}
-            \vfill
-            \begin{block}{Results: Unaligned Faces}
-              \begin{columns}
-                \begin{column}{.59\textwidth}
-                  \begin{itemize}
-                  \item Automatically aligned by Viola \& Jones
-                  \end{itemize}
-                  \vskip-0.5ex
-                  \begin{table}
-                    \centering
-                    \small
-                    \begin{tabular}{@{} p{.4\linewidth} r r @{}}
-                      \toprule 
-                      Descriptor  &      \multicolumn{2}{c @{}}{Error Rates [\%]}      \\
-                      \cmidrule(l){2-3}   
-                      &   AR-Face       & CMU-PIE  \\
-                      \cmidrule(r){1-1}  \cmidrule(lr){2-2}  \cmidrule(l){3-3}  
-                      SURF-64     &   5.97          & 15.32    \\ 
-                      SURF-128    &   5.71          & 11.42    \\ 
-                      SIFT        &   5.45          & 8.32     \\  
-                      \addlinespace
-                      U-SURF-64   &   5.32          & 5.52     \\  
-                      U-SURF-128  &   5.71          & \textbf{4.86}  \\ 
-                      U-SIFT      &   \textbf{4.15} & 8.99     \\  
-                      \bottomrule
-                    \end{tabular}
-                  \end{table}
-                \end{column}
-                \begin{column}{.39\textwidth}                
-                  \vskip-3ex
-                  \begin{itemize}
-                  \item Manually aligned faces
-
-                      \includegraphics[width=0.9\linewidth]{hanselmann-thesis/slides/figures/alignment_cmu}
-
-                  \item Unaligned faces
-
-                      \includegraphics[width=0.9\linewidth]{hanselmann-thesis/slides/figures/unalignment_cmu}
-
-                  \end{itemize}
-                \end{column}
-              \end{columns}
-            \end{block}
-            \vfill
-            \begin{block}{Results: Partially Occluded Faces}
+            \begin{block}{Other Results}
               \begin{itemize}
@@
                 \centering
-                \begin{tabular}{@{} l @{} r r r r r r@{}}
-                  \toprule 
-                  Descriptor          & \multicolumn{6}{c @{}}{Error Rates [\%]}                                                                 \\
-                  \cmidrule(l){2-7}
-                  & \textit{AR1scarf}  & \textit{AR1sun}  & \textit{ARneutral} & \textit{AR2scarf} & \textit{AR2sun}  & Avg. \\
-                  \cmidrule(r){1-1}     \cmidrule(lr){2-2}   \cmidrule(lr){3-3} \cmidrule(lr){4-4}   \cmidrule(lr){5-5}  \cmidrule(lr){6-6}  \cmidrule(l){7-7} 
-                  SURF-64             & 2.72             & 30.00          & 0.00                & 4.54            & 47.27      & 16.90 \\        
-                  SURF-128            & 1.81             & 23.63          & 0.00                & 3.63            & 40.90      & 13.99 \\
-                  SIFT                & 1.81             & 24.54          & 0.00                & 2.72            & 44.54      & 14.72 \\
-                  \addlinespace
-                  U-SURF-64           & 4.54             & 23.63          & 0.00                & 4.54            & 47.27      & 15.99 \\
-                  U-SURF-128          & 1.81             & \textbf{20.00} & 0.00                & 3.63            & 41.81      & 13.45 \\
-                  U-SIFT              & \textbf{1.81}    & 20.90          & \textbf{0.00}       & \textbf{1.81}   & \textbf{38.18} & \textbf{12.54} \\
-                  \cmidrule(r){1-1}     \cmidrule(lr){2-2}   \cmidrule(lr){3-3} \cmidrule(lr){4-4}   \cmidrule(lr){5-5}  \cmidrule(lr){6-6}  \cmidrule(l){7-7}
-                  U-SURF-128+R        & 1.81             & 19.09          & 0.00                & 3.63             & 43.63     & 13.63 \\
-                  U-SIFT+R            & 2.72             & \textbf{14.54} & 0.00                & \textbf{0.90}    & 35.45     & 10.72 \\
-                  U-SURF-128+U-SIFT+R &  \textbf{0.90}   & 16.36          & \textbf{0.00}       & 2.72             & \textbf{32.72} & \textbf{10.54} \\       
-                  % \midrule 
-                  % \midrule 
-                  % DCT \cite{ekenel:facialocclusion:icb2009}, baseline%
-                  % & 8.2              & 61.8           & 7.3                & 16.4            & 62.7       & 31.28 \\
-                  % DCT \cite{ekenel:facialocclusion:icb2009}, realigned%
-                  % & 2.7              & 1.8            & 0.0                & 6.4             & 4.5        & 3.08  \\
+                \begin{tabular}{l @{} c c c@{}} \toprule
+                  & RM (Avg) & WSJ Nov'92 & WSJ Nov'93 \\
+                  \cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(r){4-4}
+                  Triphone                    & 3.97     & 12.5       &  18.3   \\
+                  \,\, + fMLLR                & 3.59     & 11.4       &  15.5   \\ 
+                  \,\, + LVTLN                & 3.30     & 11.1       &  16.4   \\ 
+                  Splice-9 + LDA + MLLT       & 3.88     & 12.2       &  17.7   \\ 
+                  \,\, + SAT (fMLLR)          & 2.70     & 9.6        &  13.7   \\ 
+                  \,\, + SGMM + spk-vecs      & 2.45     & 10.0       &  13.4   \\ 
+                  \qquad + fMLLR              & 2.31     & 9.8        &  12.9   \\ 
+                  \qquad + ET                 & 2.15     & 9.0        &  12.3   \\
                   \bottomrule
@@
               \begin{itemize}
-              \item Grid-based local feature extraction instead of interest points
-              \item Local descriptors:
-                \begin{itemize}
-                \item upright descriptor versions achieved better results
-                \item SURF-128 better than SURF-64
-                \end{itemize}
-              \item System robustness: manually aligned/unaligned/partially occluded faces
-                \begin{itemize}
-                \item SURF more robust to illumination
-                \item SIFT more robust to changes in viewing conditions
-                \end{itemize}
-              \item RANSAC-based system combination and outlier removal
+              \item 
               \end{itemize}
@@
   %\tiny\hfill\textcolor{ta2gray}{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php}}
-  \tiny\hfill{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php} \hskip1em}
+%  \tiny\hfill{Created with \LaTeX \texttt{beamerposter}  \url{http://www-i6.informatik.rwth-aachen.de/~dreuw/latexbeamerposter.php} \hskip1em}
 \end{frame}

### 2011-12-09 misc/papers/asru11_toolkit_poster/kaldi-poster.tex
ASRU poster update
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@654 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8
Final Kaldi poster
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@655 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
 \usepackage[latin1]{inputenc}
-\usepackage{amsmath,amsthm, amssymb, latexsym}
+\usepackage{amsmath,amsthm, amssymb, latexsym, listings}
 %\usepackage{times}\usefonttheme{professionalfonts}  % obsolete
@@
 \begin{document}
-\begin{frame}
+\begin{frame}[fragile]
   \begin{columns}
@@
     % Set up a column 
-    \begin{column}{.49\textwidth}
+    \begin{column}{.33\textwidth}
       \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
@@
             % fill each column with content
-            \begin{block}{Features of Kaldi}
+
+            \begin{block}{Problem Statement}
+              \begin{itemize}
+              \item You are a researcher who wants to try out a new method.
+              \item You want other people to use your idea, if it works.
+              \item You find that older toolkits (HTK, CMUSphinx) are too hard to modify.
+              \item Also their license may not allow you to release your changes (HTK).
+              \end{itemize}
+            \end{block}
+            \vfill
+            \begin{block}{The Kaldi project}
+              \begin{itemize}
+              \item Open-source speech recognition toolkit, Apache-licensed.
+              \item Began at 2009 Johns Hopkins University CLSP summer workshop.
+              \item Further development included 2 summer workshops in Brno, Czech Republic, and ongoing work by the participants.
+              \item Uses OpenFst code for decoding graph construction.
+              \item Uses the BLAS and LAPACK libraries for linear algebra.
+              \item Our closest competitor is probably the RWTH Aachen toolkit.
+              \end{itemize}
+            \end{block}
+            \vfill
+            \begin{block}{Our vision}
+              \begin{itemize}
+              \item Distributed community of users and contributors.
+                \begin{itemize}
+                \item Not a free-for-all; original authors moderate contributions.
+                \item The Apache license allows you to fork the project.
+                \end{itemize}
+              \item Complete state-of-the-art recipes that run from public data
+                \begin{itemize}
+                \item These already exist for Resource Management and Wall Street Journal
+                \item Switchboard recipe exists but is not yet state-of-the-art.
+                \end{itemize}
+              \item Code that's well structured and simple to understand.
+              \item Thorough testing and documentation.
+              \end{itemize}
+            \end{block}
+            \vfill
+            \begin{block}{The structure of Kaldi}
               \begin{columns}
@@
               \begin{itemize}
-              \item Integration with Finite State Transducers
-              \item Extensive linear algebra support
-              \item Extensible design
-              \item Open license
-              \item Complete recipes
-              \item Thorough testing
+              \item OpenFst for FST functions
+              \item BLAS/LAPACK for linear algebra
+              \item Core functions in C++
+              \item Many simple command-line utilities
+              \item Example shell scripts
               \end{itemize}
@@
                   \centering
-                  \includegraphics[width=0.85\linewidth]{figures/kaldi-lib.pdf}
+                  \includegraphics[width=0.99\linewidth]{figures/kaldi-lib.pdf} \vspace*{0.3in}
                 \end{column}
@@
             \vfill
-            \begin{block}{Feature Extraction}
+          }
+        \end{minipage}
+      \end{beamercolorbox}
+    \end{column}
+    % ---------------------------------------------------------%
+    % end the column
+
+    % ---------------------------------------------------------%
+    % Set up a column 
+    \begin{column}{.33\textwidth}
+      \begin{beamercolorbox}[center,wd=\textwidth]{postercolumn}
+        \begin{minipage}[T]{.95\textwidth} % tweaks the width, makes a new \textwidth
+          \parbox[t][\columnheight]{\textwidth}{ % must be some better way to set the the height, width and textwidth simultaneously
+            % Since all columns are the same length, it is all nice and tidy.  You have to get the height empirically
+            % ---------------------------------------------------------%
+            % fill each column with content
+
+            \begin{block}{Example of Kaldi code}
+              %% We couldn't get verbatim to work in here, so used a temporary tex file.
+                  \includegraphics[width=0.9\linewidth,bb=128 517 486 725]{code-samples.pdf} 
+            \end{block}
+            \vfill
+            \begin{block}{Acoustic modeling techniques supported in Kaldi}
               \begin{itemize}
-              \item 
+              \item Acoustic front-end supports MFCC and PLP features, with 
+                cepstral mean and variance normalization, LDA, STC/MLLT, HLDA,
+                VTLN, etc.
+              \item HMM/GMM acoustic models; phonetic decision trees.
+              \item Also SGMMs, exponential transform.
+              \item No language modeling code, but support converting ARPA 
+                format LMs to FSTSs.
+              \item WFST-based decoders, lattice generation.
+              \item Discriminative training with MMI, boosted MMI (fMPE unfinished).
               \end{itemize}
@@
             \vfill
-            \begin{block}{Acoustic Modeling}
-              \begin{itemize}
-              \item 
-              \end{itemize}
-            \end{block}
-            \vfill
-            \begin{block}{Phonetic Decision Trees}
-              \begin{itemize}
-              \item 
-              \item 
-              \item 
-              \item 
-              \end{itemize}
+            \begin{block}{Example of top-level system building script}
+              %% We couldn't get verbatim to work in here, so used a temporary tex file.
+                  \includegraphics[width=0.9\linewidth,bb=128 472 495 725]{script-toplevel.pdf} 
             \end{block}
@@
 
-            \begin{block}{Databases}
-              \begin{itemize}
-              \item Resource Management (RM)
-                \begin{itemize}
-                \item 
-                \item 
-                \end{itemize}
-              \item Wall Street Journal (WSJ)
-                \begin{itemize}
-                \item
-                \end{itemize}
-              \end{itemize}
+            \begin{block}{Segment of triphone training sript}
+              %% We couldn't get verbatim to work in here, so used a temporary tex file.
+                  \includegraphics[width=0.95\linewidth,bb=134 284 552 713]{script-lowlevel.pdf} 
             \end{block}
@@
               \end{columns}
-            \end{block}
-            \vfill
-            \begin{block}{Other Results}
               \begin{itemize}
-              \item AR-Face: 110 classes, 110 train (``one-shot'' training), 550 test
+              \item These are not our best results: they just show that with similar system setups, we get similar results.
+                \begin{itemize}
+                \item WSJ results in table use bigram LM.
                 \end{itemize}
-              \vskip-0.5ex
-              \begin{table}
-                \small
-                \centering
-                \begin{tabular}{l @{} c c c@{}} \toprule
-                  & RM (Avg) & WSJ Nov'92 & WSJ Nov'93 \\
-                  \cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(r){4-4}
-                  Triphone                    & 3.97     & 12.5       &  18.3   \\
-                  \,\, + fMLLR                & 3.59     & 11.4       &  15.5   \\ 
-                  \,\, + LVTLN                & 3.30     & 11.1       &  16.4   \\ 
-                  Splice-9 + LDA + MLLT       & 3.88     & 12.2       &  17.7   \\ 
-                  \,\, + SAT (fMLLR)          & 2.70     & 9.6        &  13.7   \\ 
-                  \,\, + SGMM + spk-vecs      & 2.45     & 10.0       &  13.4   \\ 
-                  \qquad + fMLLR              & 2.31     & 9.8        &  12.9   \\ 
-                  \qquad + ET                 & 2.15     & 9.0        &  12.3   \\
-                  \bottomrule
-                \end{tabular}
-              \end{table}
-            \end{block}
-            \vfill
-            \begin{block}{Conclusions}
+              \item Current best RM result: 1.78\% 
                 \begin{itemize}
-              \item 
+                \item System combination: LDA+MLLT+SAT+MMI with LDA+MLLT+SAT+SGMM+fMLLR).
+                \end{itemize}
+              \item Current best WSJ result: 4.39\% on eval'92 open-vocabulary test set.
+                \begin{itemize}
+                \item Train on SI-284, LDA+MLLT+SAT+SGMM, extended vocabulary, 4-gram LM trained from supplied trainscripts.
+                \end{itemize}
               \end{itemize}
             \end{block}
+%%             \vfill
+%%             \begin{block}{Other Results}
+%%               \begin{itemize}
+%%               \item AR-Face: 110 classes, 110 train (``one-shot'' training), 550 test
+%%               \end{itemize}
+%%               \vskip-0.5ex
+%%               \begin{table}
+%%                 \small
+%%                 \centering
+%%                 \begin{tabular}{l @{} c c c@{}} \toprule
+%%                   & RM (Avg) & WSJ Nov'92 & WSJ Nov'93 \\
+%%                   \cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(r){4-4}
+%%                   Triphone                    & 3.97     & 12.5       &  18.3   \\
+%%                   \,\, + fMLLR                & 3.59     & 11.4       &  15.5   \\ 
+%%                   \,\, + LVTLN                & 3.30     & 11.1       &  16.4   \\ 
+%%                   Splice-9 + LDA + MLLT       & 3.88     & 12.2       &  17.7   \\ 
+%%                   \,\, + SAT (fMLLR)          & 2.70     & 9.6        &  13.7   \\ 
+%%                   \,\, + SGMM + spk-vecs      & 2.45     & 10.0       &  13.4   \\ 
+%%                   \qquad + fMLLR              & 2.31     & 9.8        &  12.9   \\ 
+%%                   \qquad + ET                 & 2.15     & 9.0        &  12.3   \\
+%%                   \bottomrule
+%%                 \end{tabular}
+%%               \end{table}
+%%             \end{block}
+%%             \vfill
+%%             \begin{block}{Conclusions}
+%%               \begin{itemize}
+%%               \item 
+%%               \end{itemize}
+%%             \end{block}
           }

### 2011-12-09 misc/papers/asru11_toolkit_poster/script-lowlevel.tex
Final Kaldi poster
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@655 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\documentclass{article}
\begin{document}

\begin{verbatim}
for n in `get_splits.pl $nj`; do
  featspart[$n]="ark:apply-cmvn --norm-vars=false \
   --utt2spk=ark:$data/split$nj/$n/utt2spk ark:$alidir/$n.cmvn \
   scp:$data/split$nj/$n/feats.scp ark:- | add-deltas ark:- ark:- |"
done
# tree building, graph compilation omitted.
while [ $x -lt $numiters ]; do
   echo "Iteration $x"
   if echo $realign_iters | grep -w $x >/dev/null; then
     echo "Aligning data"
     for n in `get_splits.pl $nj`; do
       $cmd $dir/log/align.$x.$n.log \
         gmm-align-compiled $scale_opts --beam=10 --retry-beam=40 \
           $dir/$x.mdl "ark:gunzip -c $dir/$n.fsts.gz|" "${featspart[$n]}" \
           "ark:|gzip -c >$dir/$n.ali.gz" || touch $dir/.error &
     done
     wait;
     [ -f $dir/.error ] && echo "Alignment error on iteration $x" && exit 1;
   fi
   for n in `get_splits.pl $nj`; do
     $cmd $dir/log/acc.$x.$n.log \
     gmm-acc-stats-ali  $dir/$x.mdl "${featspart[$n]}" \
      "ark,s,cs:gunzip -c $dir/$n.ali.gz|" $dir/$x.$n.acc || touch $dir/.error &
   done
   wait;
   [ -f $dir/.error ] && echo "Accumulation error on iteration $x" && exit 1;
   $cmd $dir/log/update.$x.log \
     gmm-est --write-occs=$dir/$[$x+1].occs --mix-up=$numgauss $dir/$x.mdl \
       "gmm-sum-accs - $dir/$x.*.acc |" $dir/$[$x+1].mdl || exit 1;
   rm $dir/$x.mdl $dir/$x.*.acc
   rm $dir/$x.occs 
   if [[ $x -le $maxiterinc ]]; then 
      numgauss=$[$numgauss+$incgauss];
   fi
   x=$[$x+1];
done
\end{verbatim}
\vspace{0.4in}

\end{document}

### 2011-12-09 misc/papers/asru11_toolkit_poster/script-toplevel.tex
Final Kaldi poster
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@655 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\documentclass{article}
\begin{document}

\begin{verbatim}
## Example of WSJ system building.
local/wsj_data_prep.sh /mnt/speech_data/WSJ?/??-{?,??}.?
local/wsj_prepare_dict.sh
local/wsj_format_data.sh
mfccdir=/mnt/my_storage/kaldi_wsj_mfcc
for x in test_eval92 test_eval93 test_dev93 train_si284; do 
 steps/make_mfcc.sh data/$x exp/make_mfcc/$x $mfccdir 4
done
# skipped some data-subsetting commands here.
# This setup would use GridEngine.
decode_cmd="queue.pl -q queue_name -l ram_free=1200M,mem_free=1200M"
train_cmd="queue.pl -q queue_name -l ram_free=700M,mem_free=700M"

steps/train_mono.sh --num-jobs 10 --cmd "$train_cmd" \
  data/train_si84_2kshort data/lang exp/mono0a
steps/align_deltas.sh --num-jobs 10 --cmd "$train_cmd" \
   data/train_si84_half data/lang exp/mono0a exp/mono0a_ali
steps/train_deltas.sh --num-jobs 10 --cmd "$train_cmd" \
    2000 10000 data/train_si84_half data/lang exp/mono0a_ali exp/tri1
# ... 
\end{verbatim}
\vspace{0.4in}

\end{document}

### 2012-02-02 misc/papers/icassp12_lat/paper.tex
Adding ASRU'11 toolkit paper to repository; making dir for poster
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@641 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8
Final changes to lattice paper; other very minor issues.
git-svn-id: https://svn.code.sf.net/p/kaldi/code/trunk@711 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

@@
+% paper # 3887, pw CF8FECE1
 % Note: the original submission was rev. 530... 
+% submitted again at 539 before review..
 % Template for SLT-2006 paper; to be used with:
@@
  \em {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^{2,4}$, {Arnab Ghoshal}$^5$, {Milo\v{s} Janda}$^2$, {Martin Karafi\'{a}t}$^2$, {Stefan Kombrink}$^2$, \\
-  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Ngoc Thang Vu}$^8$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$
+  \em {Petr Motl\'{i}\v{c}ek}$^6$, {Yanmin Qian}$^7$, {Korbinian Riedhammer}$^9$, {Karel Vesel\'{y}}$^2$, {Ngoc Thang Vu}$^8$
     \thanks{Thanks to Honza \v{C}ernock\'{y}, Renata Kohlov\'{a}, and Tom\'{a}\v{s} Ka\v{s}p\'{a}rek for their help relating to the Kaldi'11 workshop at BUT, and to Sanjeev Khudanpur for his help in preparing the paper.
@@
 MSM0021630528, and Grant Agency of the Czech Republic project No.
-102/08/0707.}}
+102/08/0707.
+Arnab Ghoshal was supported by EC FP7 grant 213850 (SCALE), and by EPSRC grant EP/I031022/1 (NST).}}
 
@@
 most directly applicable when using WFST decoders where the WFST is
-expanded down to the HMM-state level.  It outputs
-lattices that include state-level alignments as well as word labels.
+``fully expanded'', i.e. where the arcs correspond to HMM transitions.  It outputs
+lattices that include HMM-state-level alignments as well as word labels.
 The general idea is to create a state-level lattice during decoding, and to do
@@
 for each word sequence.  
+This special determinization algorithm is a solution to the following problem:
+Given a WFST A, compute a WFST B that, for each input-symbol-sequence of A,
+contains just the lowest-cost path through A.
 \end{abstract}
@@
 
+\vspace*{-0.075in}
 \section{Introduction}
+\vspace*{-0.05in}
 
@@
 
+\vspace*{-0.075in}
 \section{WFSTs and the decoding problem}
+\vspace*{-0.05in}
 \label{sec:wfst}
@@
 log-probability.  A WFST has a set of states with one distinguished
-start state;\footnote{This is the formulation that corresponds best with the toolkit we use.},
+start state\footnote{This is the formulation that corresponds best with the toolkit we use.};
 each state has a final-cost (or $\infty$ for non-final states);
-and there is a set of arcs between the states, where each arc has a weight
-(just think of this as a cost for now), an input label and an output
-label.  In $\HCLG$, the input labels are the identifiers of context-dependent
+and there is a set of arcs between the states, where each arc has an
+input label, an output label, and a weight
+(just think of this as a cost for now).
+In $\HCLG$, the input labels are the identifiers of context-dependent
 HMM states, and the output labels represent words.  For both the input and output
-symbols, the special label $\epsilon$ may appear meaning ``no label is present.''
+labels, the special symbol $\epsilon$ may appear, meaning ``no label is present.''
 
@@
 Imagine we want to ``decode'' an utterance of $T$ frames, i.e. we want to
-find the most likely word sequence, and the corresponding alignment.  A WFST
+find the most likely word sequence and its corresponding state-level alignment.  A WFST
 interpretation of the decoding problem is as follows.  We construct an
@@
 with an arc for each combination of (time, context-dependent HMM state).  The
-costs on these arcs are the negated and scaled acoustic log-likelihoods.
+costs on these arcs correspond to negated and scaled acoustic log-likelihoods.
 Call this acceptor $U$.  Define
@@
 \end{equation}
-which is the {\em search graph} for this utterance.  It has approximately $T+1$ times
+which we call the {\em search graph} of the utterance.  It has approximately $T{+}1$ times
 more states than $\HCLG$ itself.  The decoding problem is equivalent to finding
@@
 easily be avoided, we will define the desired outcome of lattice generation in terms
-of the visited subset $B$ of the search graph $S$.
+of the visited subset $B$ of $S$.
 
+\vspace*{-0.075in}
 \section{The lattice generation problem}
+\vspace*{-0.05in}
 \label{sec:lattices}
@@
 are also supported (along with separate language model, acoustic and pronunciation-probability 
-scores), and in~\cite{saon2005anatomy}, state-level alignments are also produced.
+scores), and in~\cite{saon2005anatomy}, HMM-state-level alignments are also produced.
 In our work here we will be producing state-level alignments; in fact, the input-symbols
@@
 
+\vspace*{-0.075in}
 \section{Previous lattice generation methods}
+\vspace*{-0.05in}
 \label{sec:previous}
@@
 not been able to evaluate how correct the word-pair assumption is in practice, but it seems
-unlikely to cause problems.  Such methods are not applicable for us anyway, as we
-use a WFST based decoder in which each copy of the lexical tree does not have a unique
-one-word history.
+unlikely to cause problems.  Such methods are not applicable for WFST based decoders anyway.
 
@@
 
+\vspace*{-0.075in}
 \section{Overview of our algorithm}
+\vspace*{-0.05in}
 \label{sec:overview}
@@
 \subsection{Version without alignments}
+\vspace*{-0.05in}
 
@@
 has produced more than a pre-set maximum number of states, then we prune with a tighter
-beam and try again.  In future, we may try more sophisticated methods such as a
-determinization algorithm that does pruning itself.
+beam and try again.
 
@@
 
+\vspace*{-0.075in}
 \subsection{Keeping separate graph and acoustic costs}
+\vspace*{-0.05in}
 
@@
 and $(a,b) \oplus (c,d)$ is equal to $(a,b)$ if $a{+}b < c{+}d$ or if $a{+}b = c{+}d$ and $a{-}b < c{-}d$,
-and otherwise is equal to $(c,d)$.  This is equivalent to the lexicographic semiring 
-of~\cite{roark2011lexicographic}, on the pair $((a{+}b),(a{-}b))$.
+and otherwise is equal to $(c,d)$.  This is equivalent to the normal lexicographic semiring (see~\cite{roark2011lexicographic}) on the pair $((a{+}b),(a{-}b))$.
 
+\vspace*{-0.075in}
 \subsection{Keeping state-level alignments}
+\vspace*{-0.05in}
 
@@
 
+\vspace*{-0.075in}
 \subsection{Summary of our algorithm}
+\vspace*{-0.05in}
 
-We now summarize the lattice generation algorithm.  During decoding, we create a data-structure
+During decoding, we create a data-structure
 corresponding to a full state-level lattice.  That is, for every arc of $\HCLG$, we traverse
@@
 final pruned state-level lattice be $P$.
-
 Let $Q = \inv(P)$, and let $E$ be an encoded version of $Q$ as described above (with the
 state labels as part of the weights).  The final lattice is
+\vspace*{-0.075in}
 \begin{equation}
-   L = \mathrm{prune}(\mathrm{det}(\mathrm{rmeps}(E)), \alpha) .
+   L = \mathrm{prune}(\mathrm{det}(\mathrm{rmeps}(E)), \alpha) . \vspace*{-0.075in}
 \end{equation}
@@
 
+\vspace*{-0.075in}
 \section{Details of our determinization algorithm}
+\vspace*{-0.05in}
 \label{sec:details}
@@
 avoid the expense of following $\epsilon$ arcs.
+
+%Although we have not written a proof of correctness of this algorithm, we
+%have tested it thoroughly.  This is done by generating random WFSTS,
+%applying this algorithm, and if it terminates, checking that the result
+%is deterministic and is equivalent (in this semiring) to the original WFST
+%represented as an acceptor in this semiring.
+
+Since submitting this paper, we have become aware of~\cite{shafran2011efficient},
+which solves the exact same problem for a different purpose.  They use a semiring
+which is more complicated than ours (the string part of the semiring becomes a
+structured object with parentheses).  They use this semiring instead of the one
+ we describe here, because in our semiring the $\oplus$-sum of two weights 
+does not necessarily left-divide the weights, and this is a problem for 
+a typical determinization algorithm.  We bypass
+this problem by defining a ``common divisor'' operation $\boxplus$ with the
+right properties (it $\oplus$-adds the weight part and returns the longest
+common prefix of the string part).  We use this instead of
+$\oplus$ when finding divisors in the determinization algorithm.
 
@@
 
+\vspace*{-0.1in}
 \section{Experimental results}
+\vspace*{-0.075in}
 \label{sec:exp}
+
+We do not compare with any other algorithms,
+as~\cite{ney_word_graph,odell_thesis,efficient_general} are designed for
+different types of decoders than ours, and the lattices contain less information,
+making comparisons hard to interpret; the algorithm of~\cite{saon2005anatomy} has
+similar requirements and outputs as ours, but besides being inexact, it is bound
+to be slower due to the need to store $N$ back-pointers, so we did not view it as
+worthwhile to do the experiment.
 
@@
 training data; we test on the November 1992 evaluation data.  
-For these experiments, we generate lattices with the bigram language model
+We generated lattices with the bigram language model
 supplied with the WSJ database, and for rescoring experiments we
@@
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
 
-   \caption{\vspace*{-0.2in} Varying lattice beam $\alpha$ (Viterbi pruning beam fixed at 15) }
-  \vspace{-0.08in}
+   \caption{ { Lattice properties, varying lattice beam $\alpha$ (Viterbi pruning beam fixed at 15)} }
+  \vspace{-0.15in}
   \label{fig:latbeam}
@@
 
-Figure~\ref{fig:latbeam} shows how various quantities change as we vary
-$\alpha$, with the Viterbi beam fixed at 15.  Note that we get all the improvement from
-LM rescoring by increasing $\alpha$ to 4.  The time taken by our algorithm started to increase rapidly after about
-$\alpha=8$, so a value of $\alpha$ anywhere between about 4 and 8 is sufficient for LM rescoring 
-and still does not slow down decoding too much.  We do not display the real-time factor of the non-lattice-generating
-decoder on this data (2.26) as it was actually slower than the lattice generating decoder; this is
-presumably due to the overhead of reference counting.
-Out of vocabularly words (OOVs) provide a floor on the lattice oracle error rate:
-of 333 test utterances, 87 contained at least one OOV word, yet only 93 sentences (6 more) had oracle errors
-with $\alpha=10$.  Lattice density is defined as the average number of arcs crossing each frame.
-Figure~\ref{fig:viterbibeam} shows the effect of varying the
-Viterbi decoding beam, while leaving $\alpha$ fixed at 7.
+Figure~\ref{fig:latbeam} shows how the lattice properties change as we vary
+$\alpha$, with the Viterbi beam fixed at 15; Figure~\ref{fig:viterbibeam} varies
+the Viterbi decoding beam, leaving $\alpha$ fixed at 7.  Lattice density is
+defined as the average number of arcs crossing each frame.  We get all the
+improvement from LM rescoring by increasing $\alpha$ to 4, and time taken
+increases rapidly when $\alpha > 8$, so we recommend roughly $4 < \alpha < 8$ for
+LM rescoring purposes.  We do not display the real-time factor of the
+non-lattice-generating decoder on this data (2.26xRT) as it was actually slower
+than the lattice generating decoder; this is possibly due to the overhead of
+reference counting.  Out of vocabularly words (OOVs) provide a floor on the
+lattice oracle error rate: of 333 test utterances, 87 contained at least one OOV
+word, yet only 93 sentences (6 more) had oracle errors with $\alpha=10$.
 
-\begin{figure}
+\begin{figure}[t]
 \centering
@@
  \vspace*{-0.1in}
-    \caption{\vspace*{-0.2in} Varying Viterbi pruning beam (lattice beam $\alpha$ fixed at 7) }
+    \caption{ {Lattice properties, varying Viterbi pruning beam (lattice beam $\alpha$ fixed at 7)}}
   \label{fig:viterbibeam}
-  \vspace{-0.08in}
+  \vspace{-0.15in}
 \end{figure}
 
-%\vspace*{-0.075in}
+\vspace*{-0.1in}
 \section{Conclusions}
 \label{sec:conc}
-%\vspace*{-0.05in}
+\vspace*{-0.075in}
 
 We have described a lattice generation method that is to our knowledge the first
-efficient method not to rely on the word-pair assumption of~\cite{ney_word_graph}.
-It includes an ingenious way of obtaining state-level alignment information
+efficient method that does not rely on the word-pair assumption of~\cite{ney_word_graph}.
+It includes an ingenious way of obtaining HMM-state-level alignment information
 via determinization in a specially designed semiring.
 
-%\vspace*{-0.1in}
+\vspace*{-0.04in}
+\footnotesize {
+\vspace*{-0.1in}
 \bibliographystyle{IEEEbib}
 \bibliography{refs}
+}
 

### 2012-02-10 misc/papers/asru11_toolkit/kaldi_asru.tex
sync with trunk
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/karel@753 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\documentclass[10pt,conference,letterpaper]{IEEEtran}
\usepackage{times,amsmath,amssymb,mathrsfs,bm}
\usepackage{epsfig,graphics,subfigure}
\usepackage{multirow,url}
\usepackage[usenames,dvipsnames]{color}

\def \st {\small \tt}

\title{The Kaldi Speech Recognition Toolkit}
%
\makeatletter
\def\author#1{\gdef\@author{#1\\}}
\makeatother
\author{{Daniel Povey}$^1$, {Arnab Ghoshal}$^2$, \\
  {Gilles Boulianne}$^3$, {Luk\'{a}\v{s} Burget}$^{4,5}$, {Ond\v{r}ej 
    Glembek}$^4$, {Nagendra Goel}$^6$, {Mirko Hannemann}$^4$, \\
  {Petr Motl\'{i}\v{c}ek}$^7$, {Yanmin Qian}$^8$, {Petr Schwarz}$^4$, 
  {Jan Silovsk\'{y}}$^9$, {Georg Stemmer}$^{10}$, {Karel Vesel\'{y}}$^4$
\vspace{1.6mm}\\
\fontsize{10}{10}\selectfont\itshape
   $^1$\,Microsoft Research, USA, {\st dpovey@microsoft.com}; \\
   $^2$\,Saarland University, Germany, {\st aghoshal@lsv.uni-saarland.de}; \\
   $^3$\,Centre de Recherche Informatique de Montr\'{e}al, Canada; 
   $^4$\,Brno University of Technology, Czech Republic; \\
   $^5$\,SRI International, USA;
   $^6$\,Go-Vivace Inc., USA; $^7$\,IDIAP Research Institute, Switzerland; 
   $^8$\,Tsinghua University, China; \\
   $^9$\,Technical University of Liberec, Czech Republic; 
   $^{10}$\,University of Erlangen-Nuremberg, Germany
}

% \author{%
% {First~Author{\small $~^{\#1}$}, Second~Author{\small $~^{*2}$} }%
% \vspace{1.6mm}\\
% \fontsize{10}{10}\selectfont\itshape
% $^{\#}$\,First Author Affiliation\\
% City, Country\\
% \fontsize{9}{9}\selectfont\ttfamily\upshape
% $^{1}$\,address@domain.name%
% \vspace{1.2mm}\\
% \fontsize{10}{10}\selectfont\rmfamily\itshape
% $^{*}$\,Second Author Affiliation\\
% City, Country\\
% \fontsize{9}{9}\selectfont\ttfamily\upshape
% $^{2}$\,address@domain.name%
% }

\begin{document}
\maketitle
%
\begin{abstract} 
We describe the design of Kaldi, a free, open-source toolkit for speech 
recognition research. Kaldi provides a speech recognition system based on 
finite-state transducers (using the freely available OpenFst), together with 
detailed documentation and scripts for building complete 
recognition systems. Kaldi is written is C++, and the core library supports 
modeling of arbitrary phonetic-context sizes, acoustic modeling with subspace 
Gaussian mixture models (SGMM) as well as standard Gaussian mixture models, 
together with all commonly used linear and affine transforms. Kaldi is released 
under the Apache License v2.0, which is highly nonrestrictive, making it 
suitable for a wide community of users.
\end{abstract}

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Kaldi\footnote{According to legend, Kaldi was the Ethiopian goatherd who 
discovered the coffee plant.} is an open-source toolkit for speech recognition 
written in C++ and licensed under the Apache License v2.0. 
% It is intended for use by speech recognition researchers. 
The goal of Kaldi is to have modern and flexible code that is easy to 
understand, modify and extend. Kaldi is available on SourceForge (see 
http://kaldi.sf.net/).  The tools compile on the commonly used Unix-like 
systems and on Microsoft Windows.

Researchers on automatic speech recognition (ASR) have several potential 
choices of open-source toolkits for building a recognition system. Notable 
among these are: HTK \cite{htkbook}, Julius \cite{julius} (both written in C), 
Sphinx-4 \cite{sphinx} (written in Java), and the RWTH ASR toolkit \cite{rwth}
(written in C++). Yet, our specific requirements|a finite-state transducer 
(FST) based framework, extensive linear algebra support, and a non-restrictive 
license|led to the development of Kaldi. Important features of Kaldi include:

% The work on Kaldi started during the 2009 Johns Hopkins University summer 
% workshop,
% % project titled ``Low Development Cost, High Quality Speech Recognition 
% % for New Languages and Domains,'' 
% where we were working on acoustic modeling 
% using subspace Gaussian mixture model (SGMM) \cite{sgmm_csl} and automated 
% lexicon learning.

\paragraph*{Integration with Finite State Transducers} We compile against the 
OpenFst toolkit \cite{openfst} (using it as a library).
\paragraph*{Extensive linear algebra support} We include a matrix library that wraps 
standard BLAS and LAPACK routines.
\paragraph*{Extensible design} We attempt to provide our algorithms in the most 
generic form possible. For instance, our decoders work with an interface that 
provides a score for a particular frame and FST input symbol. Thus the decoder 
could work from any suitable source of scores.
\paragraph*{Open license} The code is licensed under Apache v2.0, which is one 
of the least restrictive licenses available.
\paragraph*{Complete recipes} We make available complete recipes for 
building speech recognition systems, that work from widely available databases 
such as those provided by the Linguistic Data Consortium (LDC).
\paragraph*{Thorough testing}
The goal is for all or nearly all the code to have corresponding test routines.

The main intended use for Kaldi is acoustic modeling research; thus, we view
the closest competitors as being HTK and the RWTH ASR toolkit (RASR).  The chief
advantage versus HTK is modern, flexible, cleanly structured code and better
WFST and math support; also, our license terms are more open than either HTK
or RASR.

The paper is organized as follows: we start by describing the structure of the 
code and design choices (section \ref{sec:flavor}). This is followed by 
describing the individual components of a speech recognition system that the 
toolkit supports: feature extraction (section \ref{sec:feats}), acoustic 
modeling (section \ref{sec:am}), phonetic decision trees (section 
\ref{sec:tree}), language modeling (section \ref{sec:lm}), and decoders 
(section \ref{sec:decoder}). Finally, we provide some benchmarking results in 
section \ref{sec:expt}.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Overview of the toolkit}
\label{sec:flavor}
We give a schematic overview of the Kaldi toolkit in figure \ref{fig:kaldi-lib}.
The toolkit depends on two external libraries that are also freely available: 
one is OpenFst \cite{openfst} for the finite-state framework, and the other is 
numerical algebra libraries. We use the standard ``Basic Linear Algebra 
Subroutines'' (BLAS)%\footnote{http://www.netlib.org/blas/}%\cite{blas} 
and ``Linear Algebra PACKage'' (LAPACK)\footnote{Available from: http://www.netlib.org/blas/ and http://www.netlib.org/lapack/ respectively.} 
%\cite{lapack} 
routines for the latter.
% purpose, the details of which are described in section \ref{sec:matrix}.

\begin{figure}
  \begin{center}
    \includegraphics[width=0.75\columnwidth]{figs/kaldi-lib}\\
    \caption{A simplified view of the different components of Kaldi. The 
      library modules can be grouped into those that depend on linear algebra 
      libraries and those that depend on OpenFst. The {\em decodable} class 
      bridges these two halves. Modules that are lower down in the schematic 
      depend on one or more modules that are higher up.}
    \label{fig:kaldi-lib}
  \end{center}
\end{figure}

% We aim for the toolkit to be as loosely coupled as possible to make it easy 
% to reuse and refactor. This is reflected in the structure of the toolkit, where 
The library modules can be grouped into two distinct halves, each depending 
on only one of the external libraries (c.f. Figure \ref{fig:kaldi-lib}). A 
single module, the {\st DecodableInterface} (section \ref{sec:decoder}), 
bridges these two halves.

Access to the library functionalities is provided through command-line tools 
written in C++, which are then called from a scripting language for building 
and running a speech recognizer. 
% While this is similar to the traditional approach followed in several toolkits (e.g. HTK), the Kaldi approach differs fundamentally in how we view the tools. 
Each tool has very specific functionality with a small set of command line 
arguments: for example, there are separate executables for accumulating 
statistics, summing accumulators, and updating a GMM-based acoustic model using 
maximum likelihood estimation. 
% As such the code for the executables tend to be very simple with only a small set of command line arguments. 
Moreover, all the tools can read from and write to pipes which makes it 
easy to chain together different tools.

To avoid ``code rot'', We have tried to structure the toolkit in such a way that
implementing a new feature will generally involve adding new code and
command-line tools rather than modifying existing ones.

% An approach that has recently become popular is to have a scripting
% language such as Python call the C++ code directly, and to have the outer-level
% control flow and system design done in this scripting language.  This is
% the approach used in IBM's Attila toolkit~\cite{attila}.  The design of
% Kaldi does not preclude doing this in future, but for now we have avoided
% this approach because it requires users to be proficient in two different
% programming languages.

%A final point to mention is that we tend to prefer provably correct 
%algorithms. There has been an effort to avoid recipes and algorithms that 
%could possibly fail, even if they don't fail in the ``normal case'' (for 
%example, FST weight-pushing, which normally helps but which can fail or make things 
%much worse in certain cases).

% % -------------------------------------------------------------------------
% % -------------------------------------------------------------------------
% \section{The Kaldi Matrix Library}
% \label{sec:matrix}
% The Kaldi matrix library provides C++ classes for vector and different types of 
% matrices (description follows), as well as methods for linear algebra routines, 
% particularly inner and outer products, matrix-vector and matrix-matrix 
% products, matrix inversions and various matrix factorizations like Cholesky 
% and singular value decomposition (SVD) that are required by 
% various parts of the toolkit. The library avoids operators in favor of function 
% calls, and requires the matrices and vectors to have correct sizes instead of 
% automatically resizing the outputs.

% The matrix library does not call the Fortran interface of BLAS and LAPACK
% directly, but instead calls their C-language interface in form of CBLAS and 
% CLAPACK. In particular, we have tested Kaldi using the ``Automatically Tuned 
% Linear Algebra Software'' (ATLAS) \cite{atlas} library, the 
% Intel Math Kernel Library (MKL) library, and the Accelerate 
% Framework on OS X. It is possible to compile Kaldi with only ATLAS, even though
% ATLAS does not provide some necessary LAPACK routines, like SVD and eigenvalue 
% decompositions. For those routines we use a C++ implementation of code from the 
% ``Java Matrix Package'' (JAMA) \cite{jama} project.

% % -------------------------------------------------------------------------
% \subsection{Matrix and vector types}
% The matrix library defines the basic {\st Vector} and {\st Matrix} classes, 
% which are templated on the floating point precision type ({\st float} or {\st 
% double}). 
% {\scriptsize
% \begin{verbatim}
% template<typename Real> class Vector;
% template<typename Real> class Matrix;
% \end{verbatim}}
% % // Symmetric packed matrix class:
% % template<typename Real> class SpMatrix;
% % // Triangular packed matrix class:
% % template<typename Real> class TpMatrix;
% The {\st Matrix} class corresponds to the general matrix (GE) in BLAS and 
% LAPACK. There are also special classes for symmetric matrices ({\st SpMatrix}) 
% and triangular matrices ({\st TpMatrix}) (for Cholesky factors). Both of these 
% are represented in memory as a ``packed'' lower-triangular matrix. Currently, 
% we only support linear algebra operations with real numbers, although it is 
% possible to extend the functionality to include complex numbers as well, since 
% the underlying BLAS and LAPACK routines support complex numbers also.

% For matrix operations that involve only part of a vector or matrix, the {\st 
% SubVector} and {\st SubMatrix} classes are provided, which are treated as a 
% pointer into the underlying vector or matrix. These inherit from a common base 
% class of {\st Vector} and {\st Matrix}, respectively, and can be used in any 
% operation that does not involve memory allocation or deallocation.

% % -------------------------------------------------------------------------
% % -------------------------------------------------------------------------
% \section{Finite State Transducer (FST) library}
% \label{sec:fstext}

% We compile and link against an OpenFst~\cite{openfst}, which is an open-source
% weighted finite-state transducer library.  Both our training and decoding code
% accesses WFSTs, which are simply OpenFst's C++ objects (we will sometimes
% refer to these just as FSTs).

% We also provide code for various extensions to the OpenFst library, such as a
% constructed-on-demand context FST ($C$) which allows our toolkit to work
% efficiently for wide phonetic context.  There are also different versions of or
% extensions to some of the fundamental FST algorithms such as determinization,
% which we implement with epsilon removal and mechanisms to preserve stochasticity
% (discussed in Section~\ref{sec:graphs}); minimization, which we modify to work
% with non-deterministic input; and composition, where we provide a more efficient
% version.  We provide command-line tools with interfaces similar to OpenFst's
% command line tools, to allow these algorithms to be used from the shell.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
% \section{Finite State Transducer Algorithms}
% \label{sec:fstext}

% % -------------------------------------------------------------------------
% \subsection{Determinization}

% We use a different determinization algorithm from the one in OpenFst. Our 
% determinization algorithm is actually closer to the standard FST 
% determinization algorithm than the one in OpenFst, in that it does epsilon 
% removal along with determinization (thus, like many other FST algorithms, we do 
% not consider epsilon to be a ``real symbol'').

% Our determinization algorithm has a different way of handling what happens when 
% a transition in the initial determinized output has more than one output symbol 
% on it. The OpenFst determinization algorithm uses a function that moves the 
% output symbols (encoded as weights) around while maintaining equivalence, 
% in order to ensure that no arc has more than one (encoded) output symbol; it 
% does not introduce new states with epsilon symbols on their input, which is the 
% most ``obvious'' thing to do. However, this algorithm can fail for the output 
% of our determinization algorithm, because there can be cycles with more outputs 
% than states on the cycle (because it does epsilon removal). Instead, whenever 
% we encounter a link with more than one output symbols, we create a chain with a 
% sufficient number of intermediate states to accommodate all the output symbols. 
% The weight and input symbol go on the first link of this chain. The output of 
% our DeterminizeStar algorithm is deterministic according to the definitions 
% OpenFst uses, i.e. treating epsilons as a normal symbols. Its output does have 
% epsilons on the input side, which is against the normal definition of 
% determinism, but this is to be viewed as an encoding mechanism for allowing 
% more than one output symbol on a link, and in any case it only happens in quite 
% specific circumstances (an epsilon arc is always the only arc out of a state).

% % One other difference is that our program fstdeterminizestar does not require 
% % the input FST to have its output symbols encoded as weights.

% We supply a function which casts an FST in the tropical semiring to the log 
% semiring before determinizing, and then converts back. This is the form of 
% determinization used in our algorithms, as it preserves stochasticity (see 
% Section \ref{sec:fst:stoch}).

% % -------------------------------------------------------------------------
% \subsection{Removing epsilons}

% We supply an epsilon-removal algorithm called {\st RemoveEpsLocal()} that is 
% guaranteed to never blow up the FST, but on the other hand is not guaranteed to 
% remove all epsilons. This function has slightly different behaviour from 
% OpenFst's {\st RemoveEps()} function, because it will combine two arcs if one 
% has an input epsilon and one an output epsilon. The function preserves FST 
% equivalence.

% There is also a function {\st RemoveEpsLocalSpecial()} that preserves 
% equivalence in the tropical semiring while preserving stochasticity in the log 
% semiring (for more on stochasticity see next section). This is a case where the 
% usefulness of the semiring formalism breaks down a little bit, as we have to 
% consider two semirings simultaneously.

% % -------------------------------------------------------------------------
% \subsection{Preserving stochasticity and testing it}
% \label{sec:fst:stoch}
% We define a stochastic FST as an FST in which, in the FST's semiring the sum of 
% the weights of the arcs out of any given state (plus the final-weight) equals 
% one (in the semiring). This concept is most useful and natural in the log 
% semiring; essentially, a stochastic FST is one where the sum of the weights out 
% of a given arc is one (e.g. a properly normalized HMM would correspond to a 
% stochastic FST).

% We aim for most of the FST algorithms we use to preserve stochasticity, in the 
% sense that they will produce stochastic outputs given stochastic inputs. For 
% non-stochastic inputs, we aim that the minimum and maximum range of the weights 
% will not get larger. In order to preserve stochasticity, the FSTs that we 
% compose with have to have certain properties. Consider L, for example. 
% We require that for any linear FST corresponding to a path through G, call this 
% FST F, the product L o F must be stochastic. This basically means that L have 
% properly normalized pronunciation probabilities. The actual property formally 
% required may be a little stronger than this (this relates to ensuring that the 
% probabilities appear at the ``right time'').

% % -------------------------------------------------------------------------
% \subsection{Minimization}

% We use the minimization algorithm supplied by OpenFst, but we apply a patch 
% before compiling OpenFst so that minimization can be applied to 
% non-deterministic FSTs. The reason for this is so that we can remove 
% disambiguation symbols before minimizing, which is more optimal (it allows 
% minimization to combine more states). Fundamentally, OpenFst's minimization 
% algorithm is applicable to non-deterministic FSTs. This is 
% the same thing the fstminimize program does except that it does not do weight 
% pushing. This is desirable for us because the way we ensure stochasticity 
% entails avoiding any weight pushing.

% % -------------------------------------------------------------------------
% \subsection{Composition}

% For the most part we use OpenFst's own composition algorithms, but we do make 
% use of a more efficient composition algorithm for certain common cases. It uses 
% the ``Matcher'' concept of OpenFst; a Matcher is a kind of helper class used 
% during composition that performs lookup on the arcs out of a state to find any 
% arcs with a particular input or output symbol. The normal matcher that OpenFst 
% uses is SortedMatcher, which relies on arcs being sorted on the relevant label, 
% and does a binary search. TableMatcher detects cases where it would be 
% efficient to build a table indexed by label, and for those states it avoids the 
% overhead of binary search. This leads to a speedup when composing with things 
% like lexicons that have a very high out-degree.

% % -------------------------------------------------------------------------
% \subsection{Adding and removing disambiguation symbols}

% Our FST recipes (like other transducer-based recipes) rely on disambiguation 
% symbols. In the normal recipes, these are added to the input side of the 
% lexicon FST (L) to make it determizable. We also add disambiguation symbols to 
% G and C (see Disambiguation symbols). Whenever we do a composition and the FST 
% on the right has disambiguation symbols on its input, we (in theory) add to 
% each state in the left-hand FST a self-loop for each of the disambiguation 
% symbols, which has that symbol on both its input and output. Note that the 
% actual integer symbols id's for the disambiguation symbols on the left and 
% right may not be the same. For instance, we have a special symbol \#0 in G 
% (where epsilon would normally be). The symbol-id for this would generally be 
% the highest-numbered word plus one. But when we want to pass this symbol 
% through L, we need a symbol in L's input symbol table (which mainly contains 
% phones), to represent \#0. We have a function AddSelfLoops() that takes a 
% mutable FST and two vectors of labels (a label is an integer id for a symbol). 
% The vectors are the same size, and represent corresponding input and output 
% labels for the disambiguation symbols. This function adds self-loops to each 
% final state and each state with non-epsilon output symbols on at least one arc 
% out of it.

% We remove disambiguation symbols with the function DeleteISymbols(), accessible 
% on the command line with the program fstrmsymbols.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Feature Extraction}
\label{sec:feats}
Our feature extraction and waveform-reading code aims to create standard MFCC 
and PLP features, setting reasonable defaults but leaving available the options 
that people are most likely to want to tweak (for example, the number of mel 
bins, minimum and maximum frequency cutoffs, etc.).  We support most commonly
used feature extraction approaches: e.g. VTLN, cepstral mean and variance normalization,
LDA, STC/MLLT, HLDA, and so on.

%The feature extraction 
%pipeline is implemented as a series of functions, each of which output a 
%matrix of floating point numbers and take a matrix as input, except the 
%windowing function, which reads the waveform samples as a vector.

% The windowing function can optionally dither (add random Gaussian noise to) 
% the waveform, remove DC offset and pre-emphasize the higher frequencies. 
% Our FFT implementation~\cite{rico_book} works for window lengths that are not 
% powers of 2, and we also provide an implementation of split-radix FFT for 
% 0-padded windows whose lengths are powers of 2. 
% We support cepstral liftering 
% and optional warping of the Mel filter banks using vocal tract length 
% normalization (VTLN).

%Cepstral mean and variance normalization, dynamic (i.e. 
%delta) features of arbitrary order, splicing of arbitrary number of frames 
% to the left or right of the current frame 
%and dimensionality reduction using linear discriminant analysis (LDA) or 
%heteroscedastic linear discriminant analysis (HLDA) \cite{hlda} are supported 
%at the executable layer. 
% through simple command line tools. 
%Additionally, we support reading and writing of features in the format used by 
%HTK \cite{htkbook}.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Acoustic Modeling}
\label{sec:am}
Our aim is for Kaldi to support conventional models (i.e. diagonal GMMs) and 
Subspace Gaussian Mixture Models (SGMMs), but to also be easily extensible to 
new kinds of model. 
% Following the general design philosophy of Kaldi, the 
% acoustic modeling code is made up of classes with very specific functionality 
% that do not ``know anything'' about how they get used. For example, the {\st 
% DiagGmm} class just stores the parameters of a diagonal-covariance Gaussian 
% mixture model (together with accessors and mutators for the parameters) and 
% provides methods for likelihood computation. Estimation of GMMs is handled by a 
% separate class\footnote{In fact, different estimation classes are responsible 
% for different estimation algorithms, e.g. ML or MMI.} that accumulates the 
% sufficient statistics.

% -------------------------------------------------------------------------
\subsection{Gaussian mixture models}
We support GMMs with diagonal and full covariance structures. Rather than 
representing individual Gaussian densities separately, we directly implement a 
GMM class that is parametrized by the {\em natural parameters}, i.e. means 
times inverse covariances and inverse covariances. The GMM classes also store 
the {\em constant} term in likelihood computation, which consist of all the 
terms that do not depend on the data vector. 
% In other words, the constant term is the likelihood of the zero vector.
Such an implementation is suitable for efficient log-likelihood computation 
with simple dot-products.

% -------------------------------------------------------------------------
\subsection{GMM-based acoustic model}
The ``acoustic model'' class {\st AmDiagGmm} represents a collection of {\st DiagGmm} objects, indexed by ``pdf-ids'' that correspond to context-dependent
HMM states.   This class does not represent any HMM structure, but just a collection of densities (i.e. GMMs).
% with a slightly richer interface that supports, among other things, setting the number of Gaussian components in each pdf proportional to the occupancy of the corresponding HMM state. 
There are separate classes that represent the HMM structure, principally the 
topology and transition-modeling code and the code responsible for compiling 
decoding graphs, which provide a mapping between the HMM states and the pdf 
index of the acoustic model class.  
% The classes for estimating the acoustic model parameters are likewise implemented as a collection of GMM estimators. 
Speaker adaptation and other linear transforms like maximum likelihood linear 
transform (MLLT) \cite{mllt} or semi-tied covariance (STC) \cite{stc} are 
implemented by separate classes.

% -------------------------------------------------------------------------
\subsection{HMM Topology}
It is possible in Kaldi to separately specify the HMM topology for each 
context-independent phone.  The topology format allows nonemitting states, and 
allows the user to pre-specify tying of the p.d.f.'s in different HMM states.

% -------------------------------------------------------------------------
\subsection{Speaker adaptation}
We support both model-space adaptation using maximum likelihood linear 
regression (MLLR) \cite{mllr} and feature-space adaptation using feature-space
MLLR (fMLLR), also known as constrained MLLR \cite{gales_linxform}. For both 
MLLR and fMLLR, multiple transforms can be estimated using a regression tree 
\cite{regtree}. When a single fMLLR transform is needed, it can be used as an 
additional processing step in the feature pipeline. 
% It is also possible to only estimate the bias vector of an affine transform, 
% or the bias vector and a diagonal transform matrix, which are suitable when 
% the amount of adaptation data is small\footnote{This is currently implemented only for fMLLR.}. 
The toolkit also supports speaker normalization using a linear approximation 
to VTLN, similar to \cite{lvtln}, or conventional feature-level VTLN, or
a more generic approach for gender normalization which we call the ``exponential 
transform''~\cite{asru_et}.  Both fMLLR and VTLN can be used for speaker adaptive training (SAT) of the acoustic models.

% -------------------------------------------------------------------------
\subsection{Subspace Gaussian Mixture Models}
For subspace Gaussian mixture models (SGMMs), the toolkit provides an 
implementation of the approach described in \cite{sgmm_csl}. There is a single 
class {\st AmSgmm} that represents a whole collection of pdf's; unlike the 
GMM case there is no class that represents a single pdf of the SGMM. Similar to 
the GMM case, however, separate classes handle model estimation and speaker 
adaptation using fMLLR.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Phonetic Decision Trees}
\label{sec:tree}
Our goals in building the phonetic decision tree code were to make it
efficient for arbitrary context sizes (i.e. we avoided enumerating
contexts), and also to make it general enough to support a wide range of
approaches.  The conventional approach is, in each HMM-state of each
monophone, to have a decision tree that asks questions about, say, 
the left and right phones.  In our framework, the decision-tree
roots can be shared among the phones and among the states of the
phones, and questions can be asked about any phone in the context window,
and about the HMM state.  Phonetic questions can be supplied based on
linguistic knowledge, but in our recipes the questions are generated
automatically based on a tree-clustering of the phones. 
Questions about things like phonetic stress (if marked in the dictionary)
and word start/end information are supported via an extended phone set;
in this case we share the decision-tree roots among the different versions of the
same phone.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Language Modeling}
\label{sec:lm}
Since Kaldi uses an FST-based framework, it is possible, in principle, to use 
any language model that can be represented as an FST. 
% We are working on mechanisms that are able to handle LMs that would get too 
% large when represented this way. 
We provide tools for converting LMs in the standard ARPA 
format to FSTs.  In our recipes, we have used the IRSTLM toolkit
\footnote{Available from: http://hlt.fbk.eu/en/irstlm} for 
purposes like LM pruning.  For building LMs from raw text, users may use the 
IRSTLM toolkit, for which we provide installation help, or a more fully-featured
toolkit such as SRILM~\footnote{Available from: http://www.speech.sri.com/projects/srilm/}.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Creating Decoding Graphs}
\label{sec:graphs}

All our training and decoding algorithms use Weighted Finite State Transducers 
(WFSTs).  In the conventional 
recipe~\cite{wfst}, the input symbols on the decoding graph correspond to 
context-dependent states (in our toolkit, these symbols are 
numeric and we call them pdf-ids).  However, because we allow different phones 
to share the same pdf-ids, we would have a number of problems with this 
approach, including not being able to determinize the FSTs, and not having 
sufficient information from the Viterbi path through an FST to work out the 
phone sequence or to train the transition probabilities.  In order to fix these 
problems, we put on the input of the FSTs a slightly more fine-grained integer 
identifier that we call a ``transition-id'', that encodes the pdf-id, the phone 
it is a member of, and the arc (transition) within the topology specification 
for that phone.  There is a one-to-one mapping between the ``transition-ids'' 
and the transition-probability parameters in the model: we decided 
make transitions as fine-grained as we could without increasing the 
size of the decoding graph.  
%An advantage of having the transition-ids as the 
%graph input symbols is that all we need in Viterbi-based model training is the 
%sequence of input symbols (transition-ids) in the Viterbi path through the 
%FST.  We call this sequence an alignment.  A set of alignments gives us all the 
%information we need in order to train the p.d.f.'s and the transition 
%probabilities.  Since an alignment encodes the complete phone sequence, it is 
%possible to convert alignments between different decision trees.

Our decoding-graph construction process is based on the recipe described
in~\cite{wfst}; however, there are a number of differences.  One important one
relates to the way we handle ``weight-pushing'', which is the operation that is
supposed to ensure that the FST is stochastic.  ``Stochastic'' means that the
weights in the FST sum to one in the appropriate sense, for each state (like a
properly normalized HMM).  Weight pushing may fail or may lead to bad pruning 
behavior if the FST representing the grammar or language model ($G$) is not 
stochastic, e.g. for backoff language models.  Our approach is to
avoid weight-pushing altogether, but to ensure that each stage of graph creation
``preserves stochasticity'' in an appropriate sense.  Informally, what this 
means is that the ``non-sum-to-one-ness'' (the failure to sum to one) will 
never get worse than what was originally present in $G$.

% This requires changes 
%to some algorithms, e.g. determinization.

%There are other 
%differences too: we minimize after removing disambiguation symbols, which is 
%more optimal but requires changes to the minimization code of OpenFst; and we 
%use a version of determinization that removes input epsilon symbols, which 
%requires certain changes in other parts of the recipe (chiefly: introducing 
%disambiguation symbols on the input of $G$).

% The graph creation process required in test time is put together at the 
% shell-script level.  In training 
% time, the graph creation is done as a C++ program which can be made more 
% efficient.  The aim is for all the C++ tools to be quite generic, and not to 
% have to know about ``special'' things like silence.  For instance, alternative 
% pronunciations and optional silence are supplied as part of the lexicon FST 
% ($L$) which is generally produced by a script.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Decoders}
\label{sec:decoder}
We have several decoders, from simple to highly optimized; more will be added 
to handle things like on-the-fly language model rescoring and lattice 
generation.  By ``decoder'' we mean a C++ class that implements the core 
decoding algorithm.  The decoders do not require a particular type of acoustic 
model: they need an object satisfying a very simple interface with a function 
that provides some kind of acoustic model score for a particular (input-symbol 
and frame).

{\scriptsize
\begin{verbatim}
class DecodableInterface {
 public:
  virtual float LogLikelihood(int frame, int index) = 0;
  virtual bool IsLastFrame(int frame) = 0;
  virtual int NumIndices() = 0;
  virtual ~DecodableInterface() {}
};
\end{verbatim}}

Command-line decoding programs are all quite simple, do just one 
pass of decoding, and are all specialized for one decoder and one 
acoustic-model type.  Multi-pass decoding is implemented at the script level.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Experiments}
\label{sec:expt}
We report experimental results on the Resource Management (RM) corpus and on 
Wall Street Journal.  
% We note that the experiments reported here should be fully reproducible, 
% except for minor differences in WER due to differences in compiler behavior 
% and random number generation algorithms.  
The results reported here correspond to version 1.0 of Kaldi; the scripts that 
correspond to these experiments may be found in {\st egs/rm/s1} and 
{\st egs/wsj/s1}.

% and we will provide ``system identifiers'' (corresponding to 
%training runs) to help locate particular experiments in our scripts.

% The scripts include all data preparation stages, and require only the 
% original datasets as distributed by the Linguistic Data Consortium (LDC).

% %% Arnab, if you don't like this format, we can discuss other options.
% {\scriptsize
% \begin{verbatim}
% svn co \
%  https://kaldi.svn.sourceforge.net/svnroot/kaldi/kaldi-v1.0
% \end{verbatim}
% }

% ; see
% {\scriptsize \verb|egs/rm/README.txt|} and {\scriptsize \verb|egs/wsj/README.txt|}.
% In the experimental results in this section we will provide system identifiers
% to make it easier to locate the corresponding experiments.

\subsection{Comparison with previously published results}

% We first report some results intended to demonstrate that the basic
% algorithms included in the toolkit give results comparable to those
% previously reported in the literature.

\begin{table}
\centering{
  \caption{Basic triphone system on Resource Management: \%WERs}
  \label{rm:baseline}
  \begin{tabular}{|c|c|c|c|c|c|} \hline
           & \multicolumn{5}{|c|}{    Test set  }   \\ \cline{2-6}
           &  Feb'89 &  Oct'89 & Feb'91 & Sep'92 & Avg   \\ \hline
    HTK    &  2.77    &  4.02   &  3.30  &  6.29  &  4.10 \\
    Kaldi  &  3.20    &  4.21   &  3.50  &  5.86  &  4.06 \\ \hline 
  \end{tabular}}
\end{table}

Table~\ref{rm:baseline} shows the results of a context-dependent triphone system
with mixture-of-Gaussian densities; the HTK baseline numbers are taken 
from~\cite{Povey:ICASSP99} and the systems use essentially the same algorithms.
The features are MFCCs with per-speaker cepstral mean subtraction.  The language
model is the word-pair bigram language model supplied with the RM corpus.
The WERs are essentially the same.  Decoding time was about 0.13$\times$RT, measured
on an Intel Xeon CPU at 2.27GHz.  The system identifier for the Kaldi results is
tri3c.

Table~\ref{wsj:baseline} shows similar results for the Wall Street Journal 
system, this time without cepstral mean subtraction.  The WSJ corpus comes with
bigram and trigram language models.
% and most of our experiments use a pruned version of the trigram language model (with the number of entries reduced from 6.7 million to 1.5 million) since our 
% fully-expanded FST gets too large with the full language model (we are working on decoding strategies that can work with large language models).  
% For comparison with published results, we report bigram
% decoding in Table~\ref{wsj:baseline}, 
and we compare with published numbers using the bigram language model. The 
baseline results are reported in~\cite{Reichl:ITSAP00}, which we refer to as 
``Bell Labs'' (for the authors' affiliation), and a HTK system described 
in~\cite{Woodland:ICASSP94}.  The HTK system was gender-dependent (a gender-independent
baseline was not reported), so the HTK results are slightly better.  Our decoding
time was about 0.5$\times$RT.

\begin{table}
\centering{
 \caption{Basic triphone system, WSJ, 20k open vocabulary, bigram LM, SI-284 train: \%WERs}
 \label{wsj:baseline}
\begin{tabular}{|c|c|c|} \hline
            & \multicolumn{2}{|c|}{    Test set  }   \\ \cline{2-3}
            &  Nov'92      &    Nov'93  \\ \hline
  Bell      &  11.9        &  15.4    \\
 HTK (+GD)  &  11.1        &  14.5   \\ 
 KALDI      &  11.8        &  15.0   \\ \hline
\end{tabular}}
\end{table}

\subsection{Other experiments}

Here we report some more results on both the WSJ test sets (Nov'92 and Nov'93) 
using systems trained on just the SI-84 part of the training data, that
demonstrate different features that are supported by Kaldi. 
% Note that the triphone results for the WSJ sets are worse than those in Table 
% \ref{wsj:baseline} due to the smaller training set. 
We also report results on the RM task, averaged over 6 test sets: the 4 
mentioned in table \ref{rm:baseline} together with Mar'87 and Oct'87. The best 
result for a conventional GMM system is achieved by a SAT 
% a speaker-adaptively trained 
system that splices 9 frames (4 on each side of the 
current frame) and uses LDA to project down to 40 dimensions, together with 
MLLT.  We achieve better performance on average, with an SGMM system trained
on the same features, with speaker vectors and fMLLR adaptation.  The last
line, with the best results, includes the ``exponential transform''~\cite{asru_et} in
the features.

\begin{table}
\centering{
 \caption{Results on RM and on WSJ, 20k open vocabulary, bigram LM, trained on half of SI-84: \%WERs}
 \label{results}
\begin{tabular}{|l|c|c|c|} \hline
                             & RM (Avg) & WSJ Nov'92 & WSJ Nov'93 \\ \hline
 Triphone                    & 3.97     & 12.5       &  18.3   \\
 \,\, + fMLLR                & 3.59     & 11.4       &  15.5   \\ 
 \,\, + LVTLN                & 3.30     & 11.1       &  16.4   \\ 
 Splice-9 + LDA + MLLT       & 3.88     & 12.2       &  17.7   \\ 
 \,\, + SAT (fMLLR)          & 2.70     & 9.6        &  13.7   \\ 
 \,\, + SGMM + spk-vecs      & 2.45     & 10.0       &  13.4   \\ 
 \qquad + fMLLR              & 2.31     & 9.8        &  12.9   \\ 
 \qquad + ET                 & 2.15     & 9.0        &  12.3   \\
\hline
\end{tabular}}
\end{table}

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusion}
We described the design of Kaldi, a free and open-source speech recognition 
toolkit. The toolkit currently supports modeling of context-dependent phones of 
arbitrary context lengths, and all commonly used techniques that can be 
estimated using maximum likelihood. It also supports the recently proposed 
SGMMs. Development of Kaldi is continuing and we are working on using large 
language models in the FST framework, lattice generation and discriminative 
training.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section*{Acknowledgments}

{\footnotesize
We would like to acknowledge participants and collaborators in the 2009 Johns
Hopkins University Workshop, including Mohit Agarwal, Pinar Akyazi, Martin
Karafiat, Feng Kai, Ariya Rastrow, Richard C. Rose and Samuel Thomas; Patrick
Nguyen, for introducing the participants in that workshop and for help with WSJ
recipes, and faculty and staff at JHU for their help during that workshop, 
including Sanjeev Khudanpur, Desir\'{e}e Cleves, and the late Fred Jelinek.

We would like to acknowledge the support of Geoffrey Zweig and Alex Acero at 
Microsoft Research.  We are grateful to Jan (Honza) \v{C}ernock\'{y} 
for helping us organize the workshop at the Brno University of Technology 
during August 2010 and 2011. Thanks to Tomas Ka\v{s}p\'{a}rek for system 
support and Renata Kohlov\'{a} for administrative support.

We would like to thank Michael Riley, who visited us in Brno to deliver 
lectures on finite state transducers and helped us understand OpenFst; Henrique 
(Rico) Malvar of Microsoft Research for allowing the use of his FFT code; and 
Patrick Nguyen for help with WSJ recipes. 
We would like to acknowledge the help with coding and documentation from 
Sandeep Boda and Sandeep Reddy (sponsored by Go-Vivace Inc.) and Haihua Xu.  
We thank Pavel Matejka (and Phonexia s.r.o.) for allowing the use of feature 
processing code.

% It is possible that this list of contributors contains oversights; any 
% important omissions are unlikely to be intentional.

During the development of Kaldi, Arnab Ghoshal was supported by the European 
Community's Seventh Framework Programme under grant agreement no. 213850 
(SCALE); the BUT researchers were supported by the Technology Agency of the 
Czech Republic under project No. TA01011328, and partially by Czech MPO project 
No. FR-TI1/034.

The JHU 2009 workshop was supported by National Science Foundation Grant Number
IIS-0833652, with supplemental funding from Google Research, DARPA's GALE program
and the Johns Hopkins University Human Language Technology Center of Excellence.

% Finally, it is quite possible that someone else who helped us significantly was
% inadvertently omitted here; if so, we apologize.
}

\bibliographystyle{IEEEtran}
\bibliography{refs}

\end{document}

% LocalWords:  Kaldi automata OpenFst SGMM SourceForge toolkits HTK RWTH FST LM
% LocalWords:  BLAS LAPACK PACKage decodable refactor DecodableInterface GMM Xu
% LocalWords:  functionalities executables Cholesky SVD Fortran CBLAS CLAPACK
% LocalWords:  MKL JAMA templated SpMatrix TpMatrix SubVector SubMatrix WFSTs
% LocalWords:  deallocation OpenFst's FSTs determinization stochasticity MFCC
% LocalWords:  PLP mel pre FFT cepstral liftering VTLN LDA heteroscedastic HLDA
% LocalWords:  GMMs SGMMs DiagGmm accessors mutators MMI AmDiagGmm pdf MLLT STC
% LocalWords:  nonemitting fMLLR AmSgmm pdf's triphones monophone biphone LMs
% LocalWords:  ARPA IRSTLM SRILM determinize backoff rescoring WER egs triphone
% LocalWords:  WERs MFCCs bigram xRT spk LVTLN vecs WSJ Mohit Agarwal Pinar Ka
% LocalWords:  Akyazi Karafiat Feng Ariya Rastrow Sanjeev Khudanpur Desir Boda
% LocalWords:  Cleves Henrique Malvar Sandeep Reddy Haihua Matejka Phonexia rek
% LocalWords:  Zweig Acero Honza ernock Kohlov MPO IIS DARPA's

### 2012-02-10 misc/papers/asru11_toolkit_poster/code-samples.tex
sync with trunk
git-svn-id: https://svn.code.sf.net/p/kaldi/code/sandbox/karel@753 5e6a8d80-dfce-4ca6-a32a-6e07a63d50c8

\documentclass{article}
\begin{document}

\begin{verbatim}
void DiagGmm::LogLikelihoods(const VectorBase<BaseFloat> &data,
                             Vector<BaseFloat> *loglikes) const {
  loglikes->Resize(gconsts_.Dim(), kUndefined);
  loglikes->CopyFromVec(gconsts_);
  if (static_cast<int32>(data.Dim()) != Dim()) {
    KALDI_ERR << "DiagGmm::ComponentLogLikelihood, dimension "
        << "mismatch" << (data.Dim()) << " vs. "<< (Dim());
  }
  Vector<BaseFloat> data_sq(data);
  data_sq.ApplyPow(2.0);

  // loglikes +=  means * inv(vars) * data.
  loglikes->AddMatVec(1.0, means_invvars_, kNoTrans, data, 1.0);
  // loglikes += -0.5 * inv(vars) * data_sq.
  loglikes->AddMatVec(-0.5, inv_vars_, kNoTrans, data_sq, 1.0);
}
\end{verbatim}
\vspace{0.4in}

\end{document}

### 2015-09-03 misc/papers/icassp12_tied/icassp12.tex
changes to cudamatrix code to upgrade to the CuBLAS v2 API

@@
 mixture models.  
-Experiments using on the RM and WSJ corpora show that while a classical semi-continous
+Experiments on the RM and WSJ corpora show that while a classical semi-continuous
 system does not perform as well as a continuous one, multiple-codebook semi-continuous
@@
 in improving the open-source speech recognition toolkit {\sc Kaldi}~\cite{povey2011tks},
-and felt that since a multiple-codebook form of semi-continous models is, to
+and felt that since a multiple-codebook form of semi-continuous models is, to
 our knowledge, currently being used in a state-of-the-art system~\cite{prasad2004t2b},
 we should implement this type of model.  We also wanted to experiment with
-the style of semi-continous system described 
+the style of semi-continuous system described 
 in \cite{schukattalamazzini1994srf,schukattalamazzini1995as}, particularly the smoothing
 techniques, and see how these compare against a more conventional system.
-Lastly, we felt that it would be interesting to compare semi-continous
+Lastly, we felt that it would be interesting to compare semi-continuous
 models with Subspace Gaussian Mixture Models (SGMMs)~\cite{povey2011sgm}, since
@@
 the continuous, semi-continuous, multiple-codebook semi-continuous, and 
-subspace forms of Gaussian Mixture Models. 
-We also describe the phonetic decision-tree building process, which is 
+subspace forms of Gaussian mixture models. 
+We also describe the phonetic decision tree building process, which is 
 necessary background for how we build the multiple-codebook systems.
@@
 the phones in the system, and on Wall Street Journal, they
-correspond to the ``real phones'', i.e. grouping different stress and
+correspond to the ``real phones'', i.e., grouping different stress and
 position-marked versions of the same phone together.
@@
 likelihood loss due to combining them is less than the likelihood improvement
-on the final split of the tree-building process; and we restrict this 
+on the final split of the tree building process; and we restrict this 
 clustering to disallow combining leaves across different subtrees (e.g., from 
@@
 pre-determined total number of components is reached. We allocate 
-the Gaussians proportional to a small power (e.g.~0.2) of the state's 
+the Gaussians proportional to a small power (e.g.,~0.2) of the state's 
 occupation count.
@@
 For better performance, we initialize the codebook using the tree statistics
-collected on a prior phone alignment. For each tree leaf, we include
+collected on a prior Viterbi alignment. For each tree leaf, we include
 the respective Gaussian as a component in the mixture.
@@
 to first build the decision tree used for phonetic clustering to a relatively
-coarse level (e.g.~100 leaves). Each leaf of this tree corresponds to a codebook of Gaussians, i.e.~the Gaussian parameters are tied at this level. The leaves 
-of this tree are then split further to give a more fine-grained clustering (e.g.~2500 leaves), where for each new leaf we remember which codebook it corresponds to.
-The leaves of this second tree correspond to the actual context-dependent HMM states and contain the weights.
+coarse level (e.g.~100 leaves). Each leaf of this tree corresponds to a codebook 
+of Gaussians, i.e.,~the Gaussian parameters are tied at this level. The leaves 
+of this tree are then split further to give a more fine-grained clustering 
+(e.g.,~2500 leaves), where for each new leaf we remember which codebook it corresponds to. The leaves of this second tree correspond to the actual context-dependent HMM states and contain the weights.
 For this type of system we do not apply any post-clustering.
@@
   { \sum_{l \in \{m(l) = k\}} \text{occ}(l) }
-  { \sum_r \sum_{t \in \{m(t) = r\}} \text{occ}(l) } 
+  { \sum_t \text{occ}(t) } 
   \left( N - K \cdot N_0 \right)
@@
 \end{eqnarray}
-where the covariance matrices $\k_i$ are shared between all leaves $j$. The
+where the covariance matrices $\k_i$ are shared between all states $j$. The
 weights $w_{ji}$ and means $\m_{ji}$ are derived from $\v_j$ together with
@@
 performance of the four, which serves to confirm the conventional wisdom
-about the superiority of continous systems.
+about the superiority of continuous systems.
 The two-level tree based multiple-codebook system performance lies in between the 
@@
 The effect of the two smoothing methods using $\tau$ and $\rho$ is not
-clear from Table~\ref{tab:rm_diagfull}, i.e. the results are not always
+clear from Table~\ref{tab:rm_diagfull}, i.e., the results are not always
 consistent.  While the inter-iteration smoothing with $\rho$
@@
 Using the parameters calibrated on the RM test data, the {\em 2lvl} system
-performance is not as good as the continuous or GMM systems, but still
+performance is not as good as the continuous or SGMM systems, but still
 in a similar range while using about the same number of leaves but about a

### 2015-09-03 egs/aurora4/s5/local/aurora2flist.pl

@@ -1,0 +1,55 @@
+#!/usr/bin/env perl
+
+# Copyright 2010-2011 Microsoft Corporation
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+# This is modified from the script in standard Kaldi recipe to account
+# for the way the WSJ data is structured on the Edinburgh systems. 
+# - Arnab Ghoshal, 12/1/12
+
+# This program takes as its standard input an .ndx file from the WSJ corpus that looks
+# like this:
+#;; File: tr_s_wv1.ndx, updated 04/26/94
+#;;
+#;; Index for WSJ0 SI-short Sennheiser training data
+#;; Data is read WSJ sentences, Sennheiser mic.
+#;; Contains 84 speakers X (~100 utts per speaker MIT/SRI and ~50 utts 
+#;; per speaker TI) = 7236 utts
+#;;
+#11_1_1:wsj0/si_tr_s/01i/01ic0201.wv1
+#11_1_1:wsj0/si_tr_s/01i/01ic0202.wv1
+#11_1_1:wsj0/si_tr_s/01i/01ic0203.wv1
+
+# and as command-line argument it takes the names of the WSJ disk locations, e.g.:
+# /group/corpora/public/wsjcam0/data on DICE machines.
+# It outputs a list of absolute pathnames.
+
+$wsj_dir = $ARGV[0];
+
+while(<STDIN>){
+  if(m/^;/){ next; } # Comment.  Ignore it.
+  else {
+    #m/^([0-9_]+):\s*(\S+)$/  || die "Could not parse line $_";
+    @A=split(" ", $_);
+    $filename = $A[0]; # as a subdirectory of the distributed disk.
+    #if ($filename !~ m/\.wv1$/) { $filename .= ".wv1"; }
[…]

### 2015-09-03 egs/aurora4/s5/local/run_rnnlms_sgmm5b.sh

@@ -1,0 +1,42 @@
+#!/bin/bash
+
+for test in dev93 eval92; do
+
+  steps/lmrescore.sh --cmd "$decode_cmd" data/lang_test_bd_tgpr data/lang_test_bd_fg \
+    data/test_${test} exp/sgmm5b_mmi_b0.1/decode_bd_tgpr_${test}_it4 exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4 || exit 1;
+
+
+# Note: for N-best-list generation, choosing the acoustic scale (12) that gave
+# the best WER on this test set.  Ideally we should do this on a dev set.
+
+ # This step interpolates a small RNNLM (with weight 0.25) with the 4-gram LM.
+  steps/rnnlmrescore.sh \
+    --N 100 --cmd "$decode_cmd" --inv-acwt 12 \
+    0.25 data/lang_test_bd_fg data/local/rnnlm.h30.voc10k data/test_${test} \
+    exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4 exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4_rnnlm30_0.25  \
+    || exit 1;
+
+  steps/rnnlmrescore.sh \
+    --N 100 --cmd "$decode_cmd" --inv-acwt 12 \
+    0.5 data/lang_test_bd_fg data/local/rnnlm.h100.voc20k data/test_${test} \
+    exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4 exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4_rnnlm100_0.5 \
+    || exit 1;
+
+  steps/rnnlmrescore.sh \
+    --N 100 --cmd "$decode_cmd" --inv-acwt 12 \
+    0.5 data/lang_test_bd_fg data/local/rnnlm.h200.voc30k data/test_${test} \
+    exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4 exp/sgmm5b_mmi_b0.1/decode_bd_fg_${test}_it4_rnnlm200_0.5 \
+    || exit 1;
+
+  steps/rnnlmrescore.sh \
+    --N 100 --cmd "$decode_cmd" --inv-acwt 12 \
[…]

### 2015-09-03 egs/babel/s5/local/prepare_kaldi_lm_from_training_text.sh

@@ -1,0 +1,223 @@
+#!/bin/bash
+
+# This script was copied from ../10hSystem/local (Author: Guoguo Chen?)
+# It will be modified to make it somewhat more reusable
+# But the current goal is just to have some LM while training AMs, so that
+# intermediate models may be evaluated ... (Sanejev Khudanpur)
+
+# This script trains LMs on the WSJ LM-training data.
+# It requires that you have already run wsj_extend_dict.sh,
+# to get the larger-size dictionary including all of CMUdict
+# plus any OOVs and possible acronyms that we could easily 
+# derive pronunciations for.
+
+# This script takes as command-line arguments the relevant data/lang
+# and data/train directory.  We just train an LM on that.
+# the output is G.fst, in the lang directory.
+
+# Begin configuration section.
+LMtoCompileIntoFST=4gram-mincount/lm_pr5.0.gz  # This is the LM to compile into G.fst
+# End configuration section.
+
+if [ -f path.sh ]; then . ./path.sh; fi
+. parse_options.sh || exit 1;
+
+if [ $# -ne 3 ]; then
+  echo "Usage: $0 lang-dir data-dir temp-dir";
+  echo "  --LMtoCompileIntoFST ([34]gram[-mincount]*/lm_(unpruned|pr[12345].0).gz"
+  exit 1;
+fi
+
+lang=$1
+data=$2
+dir=$3
+
+mkdir -p $dir
+
+export PATH=$PATH:/export/babel/sanjeev/kaldi-trunk/tools/kaldi_lm
+
+##################################################################
+# This portion just makes sure the kaldi-lm tools are available
+##################################################################
+
+( # First make sure the kaldi_lm toolkit is installed.
+ cd /export/babel/sanjeev/kaldi-trunk/tools || exit 1;
+ if [ -d kaldi_lm ]; then
+   echo Not installing the kaldi_lm toolkit since it is already there.
+ else
+   echo Downloading and installing the kaldi_lm tools
+   if [ ! -f kaldi_lm.tar.gz ]; then
+     wget http://www.danielpovey.com/files/kaldi/kaldi_lm.tar.gz || exit 1;
+   fi
+   tar -xvzf kaldi_lm.tar.gz || exit 1;
+   cd kaldi_lm
+   make || exit 1;
[…]

### 2015-09-03 egs/babel/s5/local/uem_ctm2segments.pl

@@ -1,0 +1,232 @@
+#!/usr/bin/env perl
+use Getopt::Long;
+
+################################################################################
+# Convert a CTM file produced by decoding a long segment, typically several min
+# long, into a sequence of shorter segments of duration 10-15 seconds.  Produce
+# a segments file of the form used for Kaldi training/decoding
+#
+#       utteranceID recordingID startTime endTime
+#
+# The desired outcome is that the long (input) segment will be recursively cut
+# into shorter segments at the location of long silences, leaving (say) 0.5 sec
+# of silence at each end of the two resulting shorter segments, until all the
+# segments are of the desired duration.
+#
+# NOTE: It is assumed that the CTM file provides time information at 0.01 sec
+#       resolution, and that any missing segments in the CTM correspond to the
+#       optional silence model, whose output token was removed by the sequence
+#
+#       lattice-align-words --> lattice-to-ctm-conf --> raw CTM file
+#
+        $ctmTimeStep = 0.01; # Could be changed if needed by --timeStep
+#
+#       It is further assumed that the explicit silence token (word) is
+#
+        $silence = "<silence>";
+#
+#       This could be changed using the --silence option if needed.
+#
+#       Another option is the minimum silence duration to permit segmentation
+#
+        $minSilence = 1.02; # seconds
+#
+#       Maximum allowed segment length, could be changed through --maxSegLen
+#
+        $maxSegLen = 30; #seconds
+#
+#       Default segment length, used when the ctm segment is too long
+#
+        $defaultSegLen = 10; # seconds
+################################################################################
+
+GetOptions("ctmTimeStep=f" => \$ctmTimeStep, 
+  "minSilence=f" => \$minSilence, 
+  "silence=s" => \$silence, 
+  "maxSegLen=f" => \$maxSegLen, 
+  "defaultSegLen=f" => \$defaultSegLen);
+
+if ($#ARGV == 1) {
+  $ctmFile  = $ARGV[0];
+  $segmentsFile = $ARGV[1];
+  print STDERR ("$0: $ctmFile $segmentsFile\n");
[…]

### 2015-09-03 egs/babel/s5/run-5-dev-bnf.sh

@@ -1,0 +1,181 @@
+#!/bin/bash 
+set -e
+set -o pipefail
+
+. conf/common_vars.sh || exit 1;
+. ./lang.conf || exit 1;
+
+
+type=dev10h
+data_only=false
+fast_path=true
+skip_kws=false
+skip_stt=false
+tmpdir=`pwd`
+
+. utils/parse_options.sh
+
+if [ $# -ne 0 ]; then
+  echo "Usage: $(basename $0) --type (dev10h|dev2h|eval|shadow)"
+  exit 1
+fi
+
+if [[ "$type" != "dev10h" && "$type" != "dev2h" && "$type" != "eval" && "$type" != "shadow" ]] ; then
+  echo "Warning: invalid variable type=${type}, valid values are dev10h|dev2h|eval"
+  echo "Hope you know what your ar doing!"
+fi
+
+
+datadir=data/app_${type}
+dirid=${type}
+
+[ ! -d data/${dirid} ] && echo "No such directory data/${dirid}" && exit 1;
+[ ! -d exp/tri5/decode_${dirid} ] && echo "No such directory exp/tri5/decode_${dirid}" && exit 1;
+
+# Set my_nj; typically 64.
+my_nj=`cat exp/tri5/decode_${dirid}/num_jobs` || exit 1;
+
+
+
+if [ ! -f data/bnf_${dirid}/.done ]; then
+  [ ! -d data/bnf_${dirid} ] && \
+    mkdir -p $tmpdir/data/bnf_${dirid} && \
+    ln -s $tmpdir/data/bnf_${dirid} data/bnf_${dirid}
+
+  steps_BNF/make_bnf_feat.sh --nj $my_nj --cmd "$decode_cmd" \
+    --transform_dir exp/tri5/decode_${dirid}/ \
+    data/${dirid} data/bnf_${dirid} exp_BNF/bnf_dnn \
+    exp/tri5_ali exp_BNF/make_bnf_${dirid}
+
+  touch data/bnf_${dirid}/.done
+fi
+
+if [ ! -f data/sat_${dirid}/.done ]; then
+  [ ! -d data/sat_${dirid} ] && \
+    mkdir -p $tmpdir/data/sat_${dirid} && \
+    ln -s $tmpdir/data/sat_${dirid} data/sat_${dirid}
+
+  steps/make_fmllr_feats.sh --cmd "$decode_cmd -tc 10" --nj $my_nj \
[…]

### 2015-09-03 egs/babel/s5b/local/kws_oracle.sh

@@ -1,0 +1,136 @@
+#!/bin/bash
+
+# Copyright 2012  Johns Hopkins University (Author: Guoguo Chen, Jan Trmal)
+#           2013  Johns Hopkins University 
+# Apache 2.0.
+
+. ./path.sh
+. ./cmd.sh
+
+# Begin configuration section.  
+cmd=run.pl
+acwt=0.09091 #Acoustic weight -- should not be necessary for oracle lattices 
+duptime=0.6  #Max time difference in which the occurences of the same KW will be seen as duplicates
+text=     # an alternative reference text to use. when not specified, the <data-dir>/text will be used
+model=    # acoustic model to use 
+extraid=  # kws setup extra ID (kws task was setup using kws_setup.sh --extraid <id>
+stage=0   # to resume the computation from different stage
+# End configuration section.
+
+set -e 
+set -o pipefail
+
+echo "$0 $@"  # Print the command line for logging
+
+
+[ -f ./path.sh ] && . ./path.sh; # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# != 3 ]; then
+  echo "Usage $0 [options] <lang-dir> <data-dir> <decode-dir>"
+  echo ""
+  echo "Main options (for others, see top of script file)"
+  echo "  --cmd (utils/run.pl|utils/queue.pl <queue opts>) # how to run jobs."
+  echo "  --text <text-file> #The alternative text file in the format SEGMENT W1 W2 W3..., "
+  echo "                     #The default text file is taken from <data-dir>/text"
+  echo ""
+  exit 1;
+fi
+
+lang=$1;
+data=$2;
+decodedir=$3;
+
+if [ -z $text ] ; then
+  text=$data/text
+fi
+
+if [ -z "$model" ]; then # if --model <mdl> was not specified on the command line...
+  srcdir=`dirname $decodedir`; # The model directory is one level up from decoding directory.
+  model=$srcdir/final.mdl; 
+fi
+
+if [ -z $extraid ] ; then # the same logic as with kws_setup.sh
+  kwsdatadir=$data/kws
+else
+  kwsdatadir=$data/${extraid}_kws
+fi
+
+nj=`cat $decodedir/num_jobs`;
+
+oracledir=$decodedir/kws_oracle
+mkdir -p $oracledir 
+mkdir -p $oracledir/log
[…]

### 2015-09-03 egs/babel/s5c/local/kaldi_dir2uem.py

@@ -1,0 +1,101 @@
+#! /usr/bin/env python
+
+import argparse, sys
+from argparse import ArgumentParser
+import re
+
+def main():
+  parser = ArgumentParser(description='Convert kaldi data directory to uem dat files',
+      formatter_class=argparse.ArgumentDefaultsHelpFormatter)
+  parser.add_argument('--verbose', type=int, \
+      dest='verbose', default=0, \
+      help='Give higher verbose for more logging')
+  parser.add_argument('--get-text', action='store_true', \
+      help='Get text in dat file')
+  parser.add_argument('--prefix', type=str, \
+      help='Add db file name as db-<prefix>-{utt/spk}.dat')
+  parser.add_argument('kaldi_dir', \
+      help='Kaldi data directory')
+  parser.add_argument('output_dir', \
+      help='Directory to store uem dat files')
+  parser.usage=':'.join(parser.format_usage().split(':')[1:]) \
+      + 'e.g. :  %(prog)s --prefix 203-lao-v0 data/dev10h.seg CMU_db'
+  options = parser.parse_args()
+
+  if options.get_text:
+    try:
+      text_file = open(options.kaldi_dir+'/text', 'r')
+    except IOError as e:
+      repr(e)
+      sys.stderr.write("%s: No such file %s\n" % (sys.argv[0], options.kaldi_dir+'/text'))
+      sys.exit(1)
+
+  try:
+    segments_file = open(options.kaldi_dir+'/segments', 'r')
+  except IOError as e:
+    repr(e)
+    sys.stderr.write("%s: No such file %s\n" % (sys.argv[0], options.kaldi_dir+'/segments'))
+    sys.exit(1)
+
+  try:
+    scp_file = open(options.kaldi_dir+'/wav.scp', 'r')
+  except IOError as e:
+    repr(e)
+    sys.stderr.write("%s: No such file %s\n" % (sys.argv[0], options.kaldi_dir+'/wav.scp'))
+    sys.exit(1)
+
+  reco2file_map = {}
+  for line in scp_file.readlines():
+    splits = line.strip().split()
+    m = re.search(r".*/(?P<file_name>[0-9A-Za-z_]*\.(sph|wav)).*", line)
+    if not m:
+      sys.stderr.write("%s does not contain a valid speech file (.wav or .sph)\n" % line.strip())
+      sys.exit(1)
[…]

### 2015-09-03 egs/gp/s1/local/gp_norm_trans_SW.pl

@@ -1,0 +1,54 @@
+#!/usr/bin/env perl
+use warnings; #sed replacement for -w perl parameter
+
+# Copyright 2012  Arnab Ghoshal
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+
+# This script normalizes the GlobalPhone Swedish transcripts that have been 
+# extracted in a format where each line contains an utterance ID followed by
+# the transcript, e.g:
+# SW002_6 sextio <#60> procent a^r enbilsaOkerier 
+# The normalization is similar to that in 'gp_format_dict_SW.pl' script.
+
+my $usage = "Usage: gp_format_trans_SW.pl -i transcript > formatted\
+Normalizes transcriptions for GlobalPhone Swedish. The input format is \
+assumed to be utterance ID followed by transcript on the same line.\n";
+
+use strict;
+use Getopt::Long;
+die "$usage" unless(@ARGV >= 1);
+my ($in_trans);
+GetOptions ("i=s" => \$in_trans);  # Input transcription
+
+open(T, "<$in_trans") or die "Cannot open transcription file '$in_trans': $!";
+while (<T>) {
+  s/\r//g;  # Since files could be in DOS format!
+  chomp;
+  $_ =~ m:^(\S+)\s+(.+): or die "Bad line: $_";
+  my $utt_id = $1;
+  my $trans = $2;
+
+  $trans =~ s/^\s*//; $trans =~ s/\s*$//;  # Normalize spaces
+  $trans =~ s/\`(.*?)\'/$1/g;  # Remove quotation marks.
+
+  print $utt_id;
+  for my $word (split(/\s+/, $trans)) {
[…]

### 2015-09-03 egs/gp/s1/steps/train_mono.sh

@@ -1,0 +1,202 @@
+#!/bin/bash
+
+# Copyright 2012  Arnab Ghoshal
+# Copyright 2010-2011  Microsoft Corporation
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+# To be run from ..
+# Flat start and monophone training, with delta-delta features.
+# This script applies cepstral mean normalization (per speaker).
+
+function error_exit () {
+  echo -e "$@" >&2; exit 1;
+}
+
+function readint () {
+  local retval=${1/#*=/};  # In case --switch=ARG format was used
+  retval=${retval#0*}      # Strip any leading 0's
+  [[ "$retval" =~ ^-?[1-9][0-9]*$ ]] \
+    || error_exit "Argument \"$retval\" not an integer."
+  echo $retval
+}
+
+nj=4       # Default number of jobs
+stage=-4   # Default starting stage (start with calculating CMN/CVN stats)
+qcmd=""    # Options for the submit_jobs.sh script
+sjopts=""  # Options for the submit_jobs.sh script
+
+PROG=`basename $0`;
+usage="Usage: $PROG [options] <data-dir> <lang-dir> <exp-dir>\n
+e.g.: $PROG data/train.1k data/lang exp/mono\n\n
+Options:\n
+  --help\t\tPrint this message and exit\n
+  --num-jobs INT\tNumber of parallel jobs to run (default=$nj).\n
+  --qcmd STRING\tCommand for submitting a job to a grid engine (e.g. qsub) including switches.\n
[…]

### 2015-09-03 egs/gp/s1/utils/int2sym.pl

@@ -1,0 +1,90 @@
+#!/usr/bin/env perl
+# Copyright 2010-2011 Microsoft Corporation
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+
+$ignore_noninteger = 0;
+$ignore_first_field = 0;
+$field = -1;
+for($x = 0; $x < 2; $x++) {
+    if($ARGV[0] eq "--ignore-noninteger") { $ignore_noninteger = 1; shift @ARGV; }
+    if($ARGV[0] eq "--ignore-first-field") { $ignore_first_field = 1; shift @ARGV; }
+    if($ARGV[0] eq "--field") { 
+       shift @ARGV; $field = $ARGV[0]+0; shift @ARGV;
+       if ($field < 1) { die "Bad argument to --field option: $field"; }
+    }
+}
+
+if ($ignore_first_field && $field > 0) { die "Incompatible options ignore-first-field and field"; }
+$zfield = $field-1; # Change to zero-based indexing.
+
+$symtab = shift @ARGV;
+if(!defined $symtab) {
+    die "Usage: sym2int.pl symtab [input] > output\n";
+}
+open(F, "<$symtab") || die "Error opening symbol table file $symtab";
+while(<F>) {
+    @A = split(" ", $_);
+    @A == 2 || die "bad line in symbol table file: $_";
+    $int2sym{$A[1]} = $A[0];
+}
+
+sub int2sym {
+    my $a = shift @_;
+    my $pos = shift @_;
+    if($a !~  m:^\d+$:) { # not all digits..
+        if($ignore_noninteger) {
+            print $a . " ";
+            next;
+        } else {
+            if($pos == 0) {
+                die "int2sym.pl: found noninteger token $a (try --ignore-first-field)\n";
+            } else {
[…]

### 2015-09-03 egs/gp/s1/utils/make_lexicon_fst.pl

@@ -1,0 +1,122 @@
+#!/usr/bin/env perl
+# Copyright 2010-2011 Microsoft Corporation
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+
+# makes lexicon FST (no pron-probs involved).
+
+if(@ARGV != 1 && @ARGV != 3 && @ARGV != 4) {
+    die "Usage: make_lexicon_fst.pl lexicon.txt [silprob silphone [sil_disambig_sym]] lexiconfst.txt"
+}
+
+$lexfn = shift @ARGV;
+if(@ARGV == 0) {
+    $silprob = 0.0;
+} elsif (@ARGV == 2){ 
+    ($silprob,$silphone) = @ARGV;
+} else {
+    ($silprob,$silphone,$sildisambig) = @ARGV;
+}
+if($silprob != 0.0) {
+    $silprob < 1.0 || die "Sil prob cannot be >= 1.0";
+    $silcost = -log($silprob);
+    $nosilcost = -log(1.0 - $silprob);
+}
+
+
+open(L, "<$lexfn") || die "Error opening lexicon $lexfn";
+
+
+
+sub is_sil {
+    # Return true (1) if provided with a phone-sequence
+    # that means silence.
+    # @_ is the parameters of the function
+    # This function returns true if @_ equals ( $silphone )
+    # or something of the form ( "#0", $silphone, "#1" )
+    # where the "#0" and "#1" are disambiguation symbols.
+    return ( @_ == 1 && $_[0] eq $silphone ||
+             (@_ == 3 && $_[1] eq $silphone &&
+              $_[0] =~ m/^\#\d+$/ &&
+              $_[0] =~ m/^\#\d+$/));
+}
+
[…]

### 2015-09-03 egs/gp/s1/utils/score_text.sh

@@ -1,0 +1,28 @@
+#!/bin/bash
+# Copyright 2010-2011 Microsoft Corporation
+
+if [ -f ./path.sh ]; then . ./path.sh; fi
+
+if [ $# -ne 2 ]; then
+   echo "Usage: scripts/score_text.sh <decode-dir> <data-dir>"
+   exit 1;
+fi
+
+dir=$1
+data=$2
+
+if [ ! -f $data/text ]; then
+  echo Could not find transcriptions in $data/text
+  exit 1
+fi
+
+
+cat $data/text | sed 's:<NOISE>::g' |  sed 's:<SPOKEN_NOISE>::g' > $dir/test_trans.filt
+
+# We assume the transcripts are already in integer form.
+cat $dir/*.txt |  sed 's:<UNK>::g' > $dir/text
+
+compute-wer --text --mode=present ark:$dir/test_trans.filt ark,p:$dir/text >& $dir/wer
+
+grep WER $dir/wer
+

### 2015-09-03 egs/hkust/s5/cmd.sh

@@ -1,0 +1,13 @@
+# "queue.pl" uses qsub.  The options to it are
+# options to qsub.  If you have GridEngine installed,
+# change this to a queue you have access to.
+# Otherwise, use "run.pl", which will run jobs locally
+# (make sure your --num-jobs options are no more than
+# the number of cpus on your machine.
+
+export train_cmd="queue.pl -q all.q@a*.clsp.jhu.edu"
+export decode_cmd="queue.pl -q all.q@a*.clsp.jhu.edu"
+#export train_cmd=run.pl
+#export decode_cmd=run.pl
+
+

### 2015-09-03 egs/hkust/s5/local/ext/score_basic_ext.sh

@@ -1,0 +1,105 @@
+#!/bin/bash
+# Copyright Johns Hopkins University (Author: Daniel Povey) 2012-2013. 
+#           Hong Kong University of Science and Technology (Author: Ricky Chan Ho Yin). Apache 2.0.
+#
+# A Chinese character error rate (CER) scoring script for Kaldi; 
+#
+# By default the script assumes the input Chinese words/characters are 3 bytes UTF8 code \
+# and continuous printable English/ASCII characters without space are treated as single token.
+#
+# When --useword2charmap option is applied, an input Chinese words to Chinese characters mapping table \
+# (e.g. a word2char_map likes "195k_chinese_word2char_map" - 1st column of 195k_chinese_word2char_map contains chinese words, \
+# 2nd to end column fields are corresponding seperated chinese characters) \
+# is used for converting the corresponding Chinese words to seperate Chinese characters for scoring. \
+# Please make sure the Chinese words in the mapping table is a superset of the Chinese words for decoding if you apply this option.
+
+# begin configuration section.
+cmd=run.pl
+min_lmwt=9
+max_lmwt=20
+useword2charmap=
+#end configuration section.
+
+[ -f ./path.sh ] && . ./path.sh
+. parse_options.sh || exit 1;
+
+if [ $# -ne 3 ]; then
+  echo "Usage: local/score_basic_ext.sh [--cmd (run.pl|queue.pl...)] <data-dir> <lang-dir|graph-dir> <decode-dir>"
+  echo " Options:"
+  echo "    --cmd (run.pl|queue.pl...)      # specify how to run the sub-processes."
+  echo "    --min_lmwt <int>                # minumum LM-weight for lattice rescoring "
+  echo "    --max_lmwt <int>                # maximum LM-weight for lattice rescoring "
+  echo "    --useword2charmap word2charmap  # use a chinese word to chinese characters mapping "
+  exit 1;
+fi
+
+data=$1
[…]

### 2015-09-03 egs/librispeech/s5/local/online/run_nnet2_disc.sh

@@ -1,0 +1,162 @@
+#!/bin/bash
+
+
+# This script does discriminative training on top of the online,
+# system trained in run_nnet2.sh.
+# note: this relies on having a cluster that has plenty of CPUs as well as GPUs,
+# since the lattice generation runs in about real-time, so takes of the order of
+# 1000 hours of CPU time.
+# 
+# Note: rather than using any features we have dumped on disk, this script
+# regenerates them from the wav data three times-- when we do lattice
+# generation, numerator alignment and discriminative training.  This made the
+# script easier to write and more generic, because we don't have to know where
+# the features and the iVectors are, but of course it's a little inefficient.
+# The time taken is dominated by the lattice generation anyway, so this isn't
+# a huge deal.
+
+. cmd.sh
+
+
+stage=0
+train_stage=-10
+use_gpu=true
+srcdir=exp/nnet2_online/nnet_a
+criterion=smbr
+drop_frames=false  # only relevant for MMI actually.
+learning_rate=0.0001
+train_stage=-10 # can be used to start training in the middle.
+decode_start_epoch=0 # can be used to avoid decoding all epochs, e.g. if we decided to run more.
+num_epochs=4
+cleanup=false  # run with --cleanup true --stage 6 to clean up (remove large things like denlats,
+               # alignments and degs).
+
+set -e
+. cmd.sh
+. ./path.sh
+. ./utils/parse_options.sh
+
+
+if $use_gpu; then
+  if ! cuda-compiled; then
+    cat <<EOF && exit 1 
+This script is intended to be used with GPUs but you have not compiled Kaldi with CUDA 
+If you want to use GPUs (and have them), go to src/, and configure and make on a machine
+where "nvcc" is installed.  Otherwise, call this script with --use-gpu false
+EOF
+  fi
+  parallel_opts="-l gpu=1" 
+  num_threads=1
+else
+  # Use 4 nnet jobs just like run_4d_gpu.sh so the results should be
+  # almost the same, but this may be a little bit slow.
+  num_threads=16
+  parallel_opts="-pe smp $num_threads" 
+fi
+
[…]

### 2015-09-03 egs/librispeech/s5/run.sh

@@ -1,0 +1,355 @@
+#!/bin/bash
+
+
+# Set this to somewhere where you want to put your data, or where
+# someone else has already put it.  You'll want to change this 
+# if you're not on the CLSP grid.
+data=/export/a15/vpanayotov/data
+
+# base url for downloads.
+data_url=www.openslr.org/resources/12
+lm_url=www.openslr.org/resources/11
+
+. cmd.sh
+. path.sh
+
+# you might not want to do this for interactive shells.
+set -e
+
+# download the data.  Note: we're using the 100 hour setup for
+# now; later in the script we'll download more and use it to train neural
+# nets.
+for part in dev-clean test-clean dev-other test-other train-clean-100; do
+  local/download_and_untar.sh $data $data_url $part
+done
+
+# download the LM resources
+local/download_lm.sh $lm_url data/local/lm || exit 1
+
+# format the data as Kaldi data directories
+for part in dev-clean test-clean dev-other test-other train-clean-100; do
+  # use underscore-separated names in data directories.
+  local/data_prep.sh $data/LibriSpeech/$part data/$(echo $part | sed s/-/_/g) || exit 1
+done
+
+## Optional text corpus normalization and LM training
+## These scripts are here primarily as a documentation of the process that has been
+## used to build the LM. Most users of this recipe will NOT need/want to run
+## this step. The pre-built language models and the pronunciation lexicon, as
+## well as some intermediate data(e.g. the normalized text used for LM training),
+## are available for download at http://www.openslr.org/11/
+#local/lm/train_lm.sh $LM_CORPUS_ROOT \
+#  data/local/lm/norm/tmp data/local/lm/norm/norm_texts data/local/lm || exit 1
+
+## Optional G2P training scripts.
+## As the LM training scripts above, this script is intended primarily to
+## document our G2P model creation process
+#local/g2p/train_g2p.sh data/local/dict/cmudict data/local/lm
+
[…]

### 2015-09-03 egs/lre07/v1/local/make_lre96.pl

@@ -1,0 +1,80 @@
+#!/usr/bin/env perl
+#
+# Copyright 2014  David Snyder  Daniel Povey
+# Apache 2.0.
+
+use File::Basename;
+if (@ARGV != 2) {
+  print STDERR "Usage: $0 <path-to-LDC96S*> <path-to-output>\n";
+  print STDERR "e.g. $0 /export/corpora/NIST/lid96e1 data\n";
+  exit(1);
+}
+
+($db_base, $out_base_dir) = @ARGV;
+
+$ldc_code = lc basename($db_base);
+
+foreach $set ('3', '10', '30') {
+  $tmp_dir = "$out_base_dir/tmp";
+  if (system("mkdir -p $tmp_dir") != 0) {
+    die "Error making directory $tmp_dir"; 
+  }
+  
+  if (system("find $db_base/test/$set -name '*.sph' | grep '$set' > $tmp_dir/sph.list")
+    != 0) {
+    die "Error getting list of sph files";
+  }
+  
+  $tmp_dir = "$out_base_dir/tmp";
+  open(WAVLIST, "<", "$tmp_dir/sph.list") or die "cannot open wav list";
+
+  %wav = ();
+  while($sph = <WAVLIST>) {
+    chomp($sph);
+    @A = split("/", $sph);
+    $basename = $A[$#A];
+    $raw_basename = $basename;
+    $raw_basename =~ s/\.sph$// || die "bad basename $basename";
+    $wav{$raw_basename} = $sph;
+  }
+
+  close(WAVLIST) || die;
+
+  $out_dir = $out_base_dir . "/" . $ldc_code . '_' . $set;
+  if (system("mkdir -p $out_dir") != 0) {
+    die "Error making directory $out_dir"; 
+  }
+
+  %utt2lang = ();
+  open(SEG, "<$db_base/test/$set/seg_lang.ndx") || die "Failed to open $db_base/test/$set/seg_lang.ndx";
+  while ($line = <SEG>) {
+    chomp($line);
+    ($utt, $lang) = split(" ", $line);
+    $utt2lang{$utt} = $lang;
+  }
+
+  open(WAV, ">$out_dir" . '/wav.scp') 
+    || die "Failed opening output file $out_dir/wav.scp";
+  open(UTT2LANG, ">$out_dir" . '/utt2lang') 
+    || die "Failed opening output file $out_dir/utt2lang";
+  open(UTT2SPK, ">$out_dir" . '/utt2spk') 
[…]

### 2015-09-03 egs/reverb/s5/local/merge_dict.pl

@@ -1,0 +1,57 @@
+#!/usr/bin/env perl
+# Copyright 2013 MERL (author: Felix Weninger)
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+use strict;
+
+if ($#ARGV < 2) {
+  print "Usage: $0 <d1> <d2> <target>\n";
+  exit 1;
+}
+
+my ($dict1, $dict2, $target) = @ARGV;
+
+open(D1, $dict1) or die "$dict1: $!";
+open(D2, $dict2) or die "$dict2: $!";
+open(T, '>', $target) or die "$target: $!";
+
+# read all pronunciations from d1
+my %d;
+
+while (<D1>) {
+  next if (/^#/);
+  chomp;
+  my @els = split(/\s+/, $_, 2);
+  push(@{$d{$els[0]}}, uc $els[1]);
+}
+
+while (<D2>) {
+  next if (/^;;/);
+  chomp;
+  my @els = split(/\s+/, $_, 2);
+  $els[0] =~ s/\(\d+\)//;
+  if (!defined $d{$els[0]}) {
+    #print "Adding pronunciation from d2 for $els[0]\n";
+    my $ptmp = $els[1];
+    $ptmp =~ s/[0-2]//g;
+    push(@{$d{$els[0]}}, $ptmp);
+  }
+}
+
+for my $w (sort keys %d) {
+  for my $p (0..$#{$d{$w}}) {
+    print T "$w  $d{$w}[$p]\n"; #, join(" ", @{$d{$w}}), "\n";
+  }
+}

### 2015-09-03 egs/rm/s5/local/nnet/run_multilingual.sh

@@ -1,0 +1,160 @@
+#!/bin/bash
+
+# Copyright 2015  University of Illinois (Author: Amit Das)
+# Copyright 2012-2015  Brno University of Technology (Author: Karel Vesely)
+
+# Apache 2.0
+
+# This example script trains Multi-lingual DNN with <BlockSoftmax> output, using FBANK features.
+# The network is trained on multiple languages simultaneously, creating a separate softmax layer
+# per language while sharing hidden layers across all languages.
+# The script supports arbitrary number of languages.
+
+. ./cmd.sh
+. ./path.sh
+
+# Example setup, the options are in 'csl' format, they must have same number of elements,
+lang_code_csl="rm,wsj" # One label for each language,
+lang_weight_csl="1.0,0.1" # Per-language weights, they scale loss-function and gradient, 1.0 for each language is good,
+ali_dir_csl="exp/tri3b_ali,../../wsj/s5/exp/tri4b_ali_si284" # One ali-dir per language,
+data_dir_csl="data/train,../../wsj/s5/data/train_si284" # One train-data-dir per language (features will be re-computed),
+
+nnet_type=dnn_small # dnn_small | dnn | bn
+
+stage=0
+. utils/parse_options.sh || exit 1;
+
+set -euxo pipefail
+
+# Convert 'csl' to bash array (accept separators ',' ':'),
+lang_code=($(echo $lang_code_csl | tr ',:' ' ')) 
+ali_dir=($(echo $ali_dir_csl | tr ',:' ' '))
+data_dir=($(echo $data_dir_csl | tr ',:' ' '))
+
+# Make sure we have same number of items in lists,
+! [ ${#lang_code[@]} -eq ${#ali_dir[@]} -a ${#lang_code[@]} -eq ${#data_dir[@]} ] && \
+  echo "Non-matching number of 'csl' items: lang_code ${#lang_code[@]}, ali_dir ${ali_dir[@]}, data_dir ${#data_dir[@]}" && \
+  exit 1
+num_langs=${#lang_code[@]}
+
+# Check if all the input directories exist,
+for i in $(seq 0 $[num_langs-1]); do
+  echo "lang = ${lang_code[$i]}, alidir = ${ali_dir[$i]}, datadir = ${data_dir[$i]}"
[…]

### 2015-09-03 egs/rm/s5/local/nnet2/run_4a.sh

@@ -1,0 +1,27 @@
+#!/bin/bash
+
+. cmd.sh
+
+
+# This example runs on top of "raw-fMLLR" features.
+# reducing splice-width from 7 to 5
+
+(  steps/nnet2/train_tanh.sh --splice-width 5 --num-epochs 20 \
+     --cleanup false \
+     --initial-learning-rate 0.02 --final-learning-rate 0.004 \
+     --num-hidden-layers 2 \
+     --num-epochs-extra 10 --add-layers-period 1 \
+     --mix-up 4000 \
+     --cmd "$decode_cmd" \
+     --hidden-layer-dim 375 \
+      data/train data/lang exp/tri3c_ali exp/nnet4a  || exit 1
+
+   steps/nnet2/decode.sh --config conf/decode.config --cmd "$decode_cmd" --nj 20 --feat-type raw \
+     --transform-dir exp/tri3c/decode \
+     exp/tri3c/graph data/test exp/nnet4a/decode
+
+   steps/nnet2/decode.sh --config conf/decode.config --cmd "$decode_cmd" --nj 20 --feat-type raw \
+     --transform-dir exp/tri3c/decode_ug \
+     exp/tri3c/graph_ug data/test exp/nnet4a/decode_ug
+ ) 
+

### 2015-09-03 egs/sre08/v1/local/make_swbd_cellular2.pl

@@ -1,0 +1,83 @@
+#!/usr/bin/env perl
+#
+# Copyright   2013   Daniel Povey
+# Apache 2.0
+
+if (@ARGV != 2) {
+  print STDERR "Usage: $0 <path-to-LDC2004S07> <path-to-output>\n";
+  print STDERR "e.g. $0 /export/corpora5/LDC/LDC2004S07 data/swbd_cellular2_train\n";
+  exit(1);
+}
+($db_base, $out_dir) = @ARGV;
+
+if (system("mkdir -p $out_dir")) {
+  die "Error making directory $out_dir";
+}
+
+open(CS, "<$db_base/docs/swb_callstats.tbl") || die  "Could not open $db_base/docs/swb_callstats.tbl";
+open(GNDR, ">$out_dir/spk2gender") || die "Could not open the output file $out_dir/spk2gender";
+open(SPKR, ">$out_dir/utt2spk") || die "Could not open the output file $out_dir/utt2spk";
+open(WAV, ">$out_dir/wav.scp") || die "Could not open the output file $out_dir/wav.scp";
+
+@badAudio=("45024", "40022");
+
+while (<CS>) {
+  $line = $_ ;
+  @A = split(",", $line);
+  if (/$A[0]/i ~~ @badAudio) {
+    # do nothing 
+  } else {
+    $wav = "sw_" . $A[0];
+    $spkr1= "swc" . $A[1];
+    $spkr2= "swc" . $A[2];
+    $gender1 = $A[3];
+    $gender2 = $A[4];
+    if ($A[3] eq "M") {
+      $gender1 = "m";
+    } elsif ($A[3] eq "F") {
+      $gender1 = "f";
+    } else {
+      die "Unknown Gender in $line";
+    }
+    if ($A[4] eq "M") {
+      $gender2 = "m";
+    } elsif ($A[4] eq "F") {
+      $gender2 = "f";
+    } else {
+      die "Unknown Gender in $line";
+    }
+    if (-e "$db_base/data/$wav.sph") {
+      $uttId = $spkr1 . "-swbdc_" . $wav ."_1";
+      if (!$spk2gender{$spkr1}) {
+        $spk2gender{$spkr1} = $gender1;
+        print GNDR "$spkr1"," $gender1\n";
+      }
+      print WAV "$uttId"," sph2pipe -f wav -p -c 1 $db_base/data/$wav.sph |\n";
+      print SPKR "$uttId"," $spkr1","\n";
+
[…]

### 2015-09-03 egs/swbd/s5/local/swbd_p1_data_prep.sh

@@ -1,0 +1,113 @@
+#!/bin/bash
+#
+
+# To be run from one directory above this script.
+
+## The input is some directory containing the switchboard-1 release 2
+## corpus (LDC97S62).  Note: we don't make many assumptions about how
+## you unpacked this.  We are just doing a "find" command to locate
+## the .sph files.
+
+# for example /mnt/matylda2/data/SWITCHBOARD_1R2
+
+. path.sh
+
+#check existing directories
+if [ $# != 1 ]; then
+   echo "Usage: swbd_p1_data_prep.sh /path/to/SWBD"
+   exit 1; 
+fi 
+
+SWBD_DIR=$1
+
+dir=data/local/train
+mkdir -p $dir
+
+
+# Audio data directory check
+if [ ! -d $SWBD_DIR ]; then
+  echo "Error: run.sh requires a directory argument"
+  exit 1; 
+fi  
+
+# Trans directory check
+if [ ! -d $dir/swb_ms98_transcriptions ]; then
+   # To get the SWBD transcriptions and dict, do:
+   echo " *** Downloading trascriptions and dictionary ***"   
+   ( 
+    cd $dir;
+    wget http://www.isip.piconepress.com/projects/switchboard/releases/switchboard_word_alignments.tar.gz
+    tar -xf switchboard_word_alignments.tar.gz
+   )
+else
+  echo "Directory with transcriptions exists, skipping downloading"
+fi
+
+
+# Option A: SWBD dictionary file check
+[ ! -f $dir/swb_ms98_transcriptions/sw-ms98-dict.text ] && \
+     echo  "SWBD dictionary file does not exist" &&  exit 1;
+
+# find sph audio files
+find $SWBD_DIR -iname '*.sph' > $dir/sph.flist
+
+n=`cat $dir/sph.flist | wc -l`
+[ $n -ne 2435 ] && \
+  echo Warning: expected 2435 data data files, found $n
+
+
+# (1a) Transcriptions preparation
+# make basic transcription file (add segments info)
+awk '{name=substr($1,1,6);gsub("^sw","sw0",name); side=substr($1,7,1);stime=$2;etime=$3;
+ printf("%s-%s_%06.0f-%06.0f", name, side, int(100*stime+0.5), int(100*etime+0.5));
[…]

### 2015-09-03 egs/swbd/s5b/local/nnet2/run_5c_gpu.sh

@@ -1,0 +1,35 @@
+#!/bin/bash
+
+
+# This is on the full training set (with duplicates removed).  This version of the recipe runs on GPUs.
+# We assume you have 8 GPU cards.  You have to use --num-threads 1 so it will
+# use the version of the code that can use GPUs (the -parallel training code
+# cannot use GPUs unless we make further modifications as the CUDA model assumes
+# a single thread per GPU context, and we're not currently set up to create multiple
+# GPU contexts.  We assume the queue is set up as in JHU (or
+# as in the "Kluster" project on Sourceforge) where "gpu" is a consumable
+# resource that you can set to number of GPU cards a machine has.
+
+. ./cmd.sh
+. ./path.sh
+! cuda-compiled && cat <<EOF && exit 1 
+This script is intended to be used with GPUs but you have not compiled Kaldi with CUDA 
+If you want to use GPUs (and have them), go to src/, and configure and make on a machine
+where "nvcc" is installed.
+EOF
+
+( 
+  if [ ! -f exp/nnet5c_gpu/final.mdl ]; then
+    steps/nnet2/train_tanh.sh --cmd "$decode_cmd" --parallel-opts "-l gpu=1" --io-opts "-tc 5" \
+      --num-threads 1 --minibatch-size 512 --max-change 40.0 --mix-up 20000 --samples-per-iter 300000 \
+      --num-epochs 10 --num-epochs-extra 3 --initial-learning-rate 0.0067 --final-learning-rate 0.00067 \
+      --num-jobs-nnet 10 --num-hidden-layers 5 --hidden-layer-dim 1536 data/train_nodup data/lang \
+        exp/tri4b exp/nnet5c_gpu || exit 1;
+  fi
+
+  for lm_suffix in tg fsh_tgpr; do
+    steps/nnet2/decode.sh --cmd "$decode_cmd" --nj 30 \
+      --config conf/decode.config --transform-dir exp/tri4b/decode_eval2000_sw1_${lm_suffix} \
+      exp/tri4b/graph_sw1_${lm_suffix} data/eval2000 exp/nnet5c_gpu/decode_eval2000_sw1_${lm_suffix} &
+  done
+)

### 2015-09-03 egs/swbd/s5b/run.sh

@@ -1,0 +1,408 @@
+#!/bin/bash
+
+# Warning: this recipe is now out of date.  See ../s5c/
+
+exit 1;
+
+
+
+# This is a shell script, but it's recommended that you run the commands one by
+# one by copying and pasting into the shell.
+# Caution: some of the graph creation steps use quite a bit of memory, so you
+# should run this on a machine that has sufficient memory.
+
+. cmd.sh
+. path.sh
+set -e # exit on error
+# mfccdir should be some place with a largish disk where you
+# want to store MFCC features. 
+mfccdir=mfcc
+
+if [ -z $IRSTLM ] ; then
+  export IRSTLM=$KALDI_ROOT/tools/irstlm/
+fi
+export PATH=${PATH}:$IRSTLM/bin
+if ! command -v prune-lm >/dev/null 2>&1 ; then
+  echo "$0: Error: the IRSTLM is not available or compiled" >&2
+  echo "$0: Error: We used to install it by default, but." >&2
+  echo "$0: Error: this is no longer the case." >&2
+  echo "$0: Error: To install it, go to $KALDI_ROOT/tools" >&2
+  echo "$0: Error: and run extras/install_irstlm.sh" >&2
+  exit 1
+fi
+
+
+# Prepare Switchboard data. This command can also take a second optional argument
+# which specifies the directory to Switchboard documentations. Specifically, if
+# this argument is given, the script will look for the conv.tab file and correct
+# speaker IDs to the actual speaker personal identification numbers released in
+# the documentations. The documentations can be found here: 
+# https://catalog.ldc.upenn.edu/docs/LDC97S62/
+# Note: if you are using this link, make sure you rename conv_tab.csv to conv.tab
+# after downloading.
+# Usage: local/swbd1_data_prep.sh /path/to/SWBD [/path/to/SWBD_docs]
+local/swbd1_data_prep.sh /export/corpora3/LDC/LDC97S62
+# local/swbd1_data_prep.sh /home/dpovey/data/LDC97S62
+# local/swbd1_data_prep.sh /data/corpora0/LDC97S62
[…]

### 2015-09-03 egs/swbd/s5c/local/nnet/run_dnn.sh

@@ -1,0 +1,153 @@
+#!/bin/bash
+
+# Copyright 2012-2014  Brno University of Technology (Author: Karel Vesely)
+#                2014  Guoguo Chen
+# Apache 2.0
+
+# This example script trains a DNN on top of fMLLR features. 
+# The training is done in 3 stages,
+#
+# 1) RBM pre-training:
+#    in this unsupervised stage we train stack of RBMs, 
+#    a good starting point for frame cross-entropy trainig.
+# 2) frame cross-entropy training:
+#    the objective is to classify frames to correct pdfs.
+# 3) sequence-training optimizing sMBR: 
+#    the objective is to emphasize state-sequences with better 
+#    frame accuracy w.r.t. reference alignment.
+
+. ./cmd.sh ## You'll want to change cmd.sh to something that will work on your system.
+           ## This relates to the queue.
+
+. ./path.sh ## Source the tools/utils (import the queue.pl)
+
+# Config:
+gmmdir=exp/tri4
+data_fmllr=data-fmllr-tri4
+stage=0 # resume training with --stage=N
+has_fisher=true
+# End of config.
+. utils/parse_options.sh || exit 1;
+#
+
+if [ $stage -le 0 ]; then
+  # Store fMLLR features, so we can train on them easily,
+  # eval2000
+  dir=$data_fmllr/eval2000
+  steps/nnet/make_fmllr_feats.sh --nj 10 --cmd "$train_cmd" \
+     --transform-dir $gmmdir/decode_eval2000_sw1_tg \
+     $dir data/eval2000 $gmmdir $dir/log $dir/data || exit 1
+  # train
+  dir=$data_fmllr/train_nodup
+  steps/nnet/make_fmllr_feats.sh --nj 10 --cmd "$train_cmd" \
+     --transform-dir ${gmmdir}_ali_nodup \
+     $dir data/train_nodup $gmmdir $dir/log $dir/data || exit 1
+  # split the data : 90% train 10% cross-validation (held-out)
+  utils/subset_data_dir_tr_cv.sh $dir ${dir}_tr90 ${dir}_cv10 || exit 1
+fi
+
+if [ $stage -le 1 ]; then
[…]

### 2015-09-03 egs/swbd/s5c/local/score_sclite.sh

@@ -1,0 +1,127 @@
+#!/bin/bash
+# Copyright Johns Hopkins University (Author: Daniel Povey) 2012.  Apache 2.0.
+
+# begin configuration section.
+cmd=run.pl
+stage=0
+min_lmwt=5
+max_lmwt=20
+reverse=false
+word_ins_penalty=0.0,0.5,1.0
+#end configuration section.
+
+[ -f ./path.sh ] && . ./path.sh
+. parse_options.sh || exit 1;
+
+if [ $# -ne 3 ]; then
+  echo "Usage: local/score_sclite.sh [--cmd (run.pl|queue.pl...)] <data-dir> <lang-dir|graph-dir> <decode-dir>"
+  echo " Options:"
+  echo "    --cmd (run.pl|queue.pl...)      # specify how to run the sub-processes."
+  echo "    --stage (0|1|2)                 # start scoring script from part-way through."
+  echo "    --min_lmwt <int>                # minumum LM-weight for lattice rescoring "
+  echo "    --max_lmwt <int>                # maximum LM-weight for lattice rescoring "
+  echo "    --reverse (true/false)          # score with time reversed features "
+  exit 1;
+fi
+
+data=$1
+lang=$2 # Note: may be graph directory not lang directory, but has the necessary stuff copied.
+dir=$3
+
+model=$dir/../final.mdl # assume model one level up from decoding dir.
+
+hubscr=$KALDI_ROOT/tools/sctk/bin/hubscr.pl 
+[ ! -f $hubscr ] && echo "Cannot find scoring program at $hubscr" && exit 1;
+hubdir=`dirname $hubscr`
+
+for f in $data/stm $data/glm $lang/words.txt $lang/phones/word_boundary.int \
+     $model $data/segments $data/reco2file_and_channel $dir/lat.1.gz; do
+  [ ! -f $f ] && echo "$0: expecting file $f to exist" && exit 1;
+done
+
+name=`basename $data`; # e.g. eval2000
+
+mkdir -p $dir/scoring/log
+
+if [ $stage -le 0 ]; then
+  for wip in $(echo $word_ins_penalty | sed 's/,/ /g'); do
+    if $reverse; then
+      $cmd LMWT=$min_lmwt:$max_lmwt $dir/scoring/log/get_ctm.LMWT.${wip}.log \
[…]

### 2015-09-03 egs/wsj/s5/local/online/run_nnet2.sh

@@ -1,0 +1,220 @@
+#!/bin/bash
+
+# this is our online-nnet2 build.  it's a "multi-splice" system (i.e. we have
+# splicing at various layers), with p-norm nonlinearities.  We use the "accel2"
+# script which uses between 2 and 14 GPUs depending how far through training it
+# is.  You can safely reduce the --num-jobs-final to however many GPUs you have
+# on your system.
+
+# For joint training with RM, this script is run using the following command line,
+# and note that the --stage 8 option is only needed in case you already ran the
+# earlier stages.
+# local/online/run_nnet2.sh --stage 8 --dir exp/nnet2_online/nnet_ms_a_partial --exit-train-stage 15
+
+. cmd.sh
+
+
+stage=0
+train_stage=-10
+use_gpu=true
+dir=exp/nnet2_online/nnet_ms_a
+exit_train_stage=-100
+. cmd.sh
+. ./path.sh
+. ./utils/parse_options.sh
+
+if $use_gpu; then
+  if ! cuda-compiled; then
+    cat <<EOF && exit 1 
+This script is intended to be used with GPUs but you have not compiled Kaldi with CUDA 
+If you want to use GPUs (and have them), go to src/, and configure and make on a machine
+where "nvcc" is installed.  Otherwise, call this script with --use-gpu false
+EOF
+  fi
+  parallel_opts="--gpu 1" 
+  num_threads=1
+  minibatch_size=512
+  # the _a is in case I want to change the parameters.
+else
+  num_threads=16
+  minibatch_size=128
+  parallel_opts="--num-threads $num_threads" 
+fi
+
+local/online/run_nnet2_common.sh --stage $stage || exit 1;
+
+if [ $stage -le 8 ]; then
+  # last splicing was instead: layer3/-4:2" 
+  steps/nnet2/train_multisplice_accel2.sh --stage $train_stage \
+    --exit-stage $exit_train_stage \
+    --num-epochs 8 --num-jobs-initial 2 --num-jobs-final 14 \
+    --num-hidden-layers 4 \
+    --splice-indexes "layer0/-1:0:1 layer1/-2:1 layer2/-4:2" \
+    --feat-type raw \
[…]

### 2015-09-03 egs/wsj/s5/local/run_nnet2.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+
+. ./cmd.sh
+
+# This shows what you can potentially run; you'd probably want to pick and choose.
+
+use_gpu=true
+
+if $use_gpu; then
+  local/nnet2/run_5b_gpu.sh # various VTLN combinations,  Mel-filterbank features, si284 train (multiplied by 5).
+  local/nnet2/run_5c.sh --use-gpu true # this is on top of fMLLR features.
+  local/nnet2/run_6c_gpu.sh # this is discriminative training of tanh neural nets on top of run_5c_gpu.sh
+  local/nnet2/run_5d.sh --use-gpu true # this is p-norm training on top of fMLLR features.  <THIS IS THE MAIN RECIPE>
+  local/nnet2/run_5e_gpu.sh # this is ensemble training of p-norm nnets on top of fMLLR features.
+  local/nnet2/run_6d_gpu.sh # this is discriminative training of p-norm neural nets on top of run_5d_gpu.sh
+else
+  local/nnet2/run_5b.sh # various VTLN combinations, Mel-filterbank features,  si284 train (multiplied by 5).
+  local/nnet2/run_5c.sh --use-gpu false # this is on top of fMLLR features.
+  local/nnet2/run_5d.sh --use-gpu false # this is p-norm on top of fMLLR features.  <THIS IS THE MAIN RECIPE>
+fi
+
+

### 2015-09-03 egs/wsj/s5/steps/align_lvtln.sh

@@ -1,0 +1,196 @@
+#!/bin/bash
+
+# Copyright 2014  Vimal Manohar
+
+# Computes training alignments; assumes features are (LDA+MLLT or delta+delta-delta)
+# Will ignore fMLLR.
+# Will estimate VTLN warping factors
+# as a by product, which can be used to extract VTLN-warped features.
+
+# Begin configuration section
+stage=0
+nj=4
+cmd=run.pl
+use_graphs=false
+# Begin configuration.
+scale_opts="--transition-scale=1.0 --acoustic-scale=0.1 --self-loop-scale=0.1"
+beam=10.0
+retry_beam=40
+boost_silence=1.0 # factor by which to boost silence during alignment.
+logdet_scale=1.0
+cleanup=false
+
+# End configuration section
+echo "$0 $@"  # Print the command line for logging
+
+[ -f ./path.sh ] && . ./path.sh; # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# != 4 ]; then
+   echo "Wrong #arguments ($#, expected 4)"
+   echo "Usage: steps/align_lvtln.sh [options] <data-dir> <lang-dir> <src-dir>  <align-dir>"
+   echo " e.g.: steps/align_lvtln.sh data/train data/lang exp/tri2c exp/tri2c_ali"
+   echo "main options (for others, see top of script file)"
+   echo "  --config <config-file>                   # config containing options"
+   echo "  --nj <nj>                                # number of parallel jobs"
+   echo "  --use-graphs true                                # use graphs in src-dir"
+   echo "  --cmd <cmd>                              # Command to run in parallel with"
+   exit 1;
+fi
+
+data=$1
+lang=$2
+srcdir=$3
+dir=$4
+
+oov=`cat $lang/oov.int` || exit 1;
+silphonelist=`cat $lang/phones/silence.csl` || exit 1;
+sdata=$data/split$nj
+
+if [ -f $data/spk2warp ]; then
+  echo "$0: file $data/spk2warp exists.  This script expects non-VTLN features"
+  exit 1;
+fi
+
[…]

### 2015-09-03 egs/wsj/s5/steps/align_raw_fmllr.sh

@@ -1,0 +1,146 @@
+#!/bin/bash
+# Copyright 2012-2013  Johns Hopkins University (Author: Daniel Povey)
+# Apache 2.0
+
+# Computes training alignments; assumes features are (LDA+MLLT or delta+delta-delta)
+# + fMLLR (probably with SAT models).
+# It first computes an alignment with the final.alimdl (or the final.mdl if final.alimdl
+# is not present), then does 2 iterations of fMLLR estimation.
+
+# If you supply the --use-graphs option, it will use the training
+# graphs from the source directory (where the model is).  In this
+# case the number of jobs must match the source directory.
+
+
+# Begin configuration section.  
+stage=0
+nj=4
+cmd=run.pl
+use_graphs=false
+# Begin configuration.
+scale_opts="--transition-scale=1.0 --acoustic-scale=0.1 --self-loop-scale=0.1"
+beam=10
+retry_beam=40
+boost_silence=1.0 # factor by which to boost silence during alignment.
+# End configuration options.
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f path.sh ] && . ./path.sh # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# != 4 ]; then
+   echo "usage: steps/align_fmllr.sh <data-dir> <lang-dir> <src-dir> <align-dir>"
+   echo "e.g.:  steps/align_fmllr.sh data/train data/lang exp/tri1 exp/tri1_ali"
+   echo "main options (for others, see top of script file)"
+   echo "  --config <config-file>                           # config containing options"
+   echo "  --nj <nj>                                        # number of parallel jobs"
+   echo "  --use-graphs true                                # use graphs in src-dir"
+   echo "  --cmd (utils/run.pl|utils/queue.pl <queue opts>) # how to run jobs."
+   exit 1;
+fi
+
+data=$1
+lang=$2
+srcdir=$3
+dir=$4
+
+oov=`cat $lang/oov.int` || exit 1;
[…]

### 2015-09-03 egs/wsj/s5/steps/nnet/make_priors.sh

@@ -1,0 +1,76 @@
+#!/bin/bash 
+
+# Copyright 2012-2014 Brno University of Technology (author: Karel Vesely)
+# Apache 2.0
+# To be run from .. (one directory up from here)
+# see ../run.sh for example
+
+# Begin configuration section.
+nj=4
+cmd=run.pl
+use_gpu=no
+# End configuration section.
+
+echo "$0 $@"  # Print the command line for logging
+
+if [ -f path.sh ]; then . ./path.sh; fi
+. parse_options.sh || exit 1;
+
+if [ $# != 2 ]; then
+   echo "usage: $0 [options] <data-dir> <nnet-dir>";
+   echo "options: "
+   echo "  --cmd 'queue.pl <queue opts>'   # how to run jobs."
+   echo "  --nj <nj>                       # number of parallel jobs"
+   echo "  --remove-last-components <N>    # number of NNet Components to remove from the end"
+   echo "  --use-gpu (no|yes|optional)     # forwarding on GPU"
+   exit 1;
+fi
+
+if [ -f path.sh ]; then . path.sh; fi
+
+data=$1
+nndir=$2
+
+######## CONFIGURATION
+
+required="$data/feats.scp $nndir/final.nnet $nndir/final.feature_transform"
+for f in $required; do
+  [ ! -f $f ] && echo "$0: Missing $f" && exit 1;
+done
+
+sdata=$data/split$nj
+[[ -d $sdata && $data/feats.scp -ot $sdata ]] || split_data.sh $data $nj || exit 1;
+
+echo "Accumulating prior stats by forwarding '$data' with '$nndir'"
+
+# PREPARE FEATURE EXTRACTION PIPELINE
+# import config,
+cmvn_opts=
+delta_opts=
+D=$nndir
+[ -e $D/norm_vars ] && cmvn_opts="--norm-means=true --norm-vars=$(cat $D/norm_vars)" # Bwd-compatibility,
+[ -e $D/cmvn_opts ] && cmvn_opts=$(cat $D/cmvn_opts)
+[ -e $D/delta_order ] && delta_opts="--delta-order=$(cat $D/delta_order)" # Bwd-compatibility,
+[ -e $D/delta_opts ] && delta_opts=$(cat $D/delta_opts)
+#
+# Create the feature stream,
[…]

### 2015-09-03 egs/wsj/s5/steps/nnet2/remove_egs.sh

@@ -1,0 +1,48 @@
+#!/bin/bash
+
+# Copyright 2014  Johns Hopkins University (Author: Daniel Povey).  
+# Apache 2.0.
+
+# This script removes the examples in an egs/ directory, e.g.
+# steps/nnet2/remove_egs.sh exp/nnet4b/egs/
+# We give it its own script because we need to be careful about
+# things that are soft links to something in storage/ (i.e. remove the
+# data that's linked to as well as the soft link), and we want to not
+# delete the examples if someone has done "touch $dir/egs/.nodelete".
+
+
+if [ $# != 1 ]; then
+  echo "Usage: $0 <egs-dir>"
+  echo "e.g.: $0 data/nnet4b/egs/"
+  echo "e.g.: $0 data/nnet4b_mpe/degs/"
+  echo "This script is usually equivalent to 'rm <egs-dir>/egs.* <egs-dir>/degs.*' but it follows"
+  echo "soft links to <egs-dir>/storage/; and it avoids deleting anything in the directory if"
+  echo "someone did 'touch <egs-dir>/.nodelete"
+  exit 1;
+fi
+
+egs=$1
+
+if [ ! -d $egs ]; then
+  echo "$0: expected directory $egs to exist"
+  exit 1;
+fi
+
+if [ -f $egs/.nodelete ]; then
+  echo "$0: not deleting egs in $egs since $egs/.nodelete exists"
+  exit 0;
+fi
+
+
+flist=$egs/egs.*.ark
+
+
+for f in $egs/egs.*.ark $egs/degs.*.ark; do
+  if [ -L $f ]; then
+    rm $(dirname $f)/$(readlink $f)  # this will print a warning if it fails.
+  fi
+  rm $f 2>/dev/null
+done
+
+
+echo "$0: Finished deleting examples in $egs"

### 2015-09-03 egs/wsj/s5/steps/nnet3/decode.sh

@@ -1,0 +1,154 @@
+#!/bin/bash
+
+# Copyright 2012-2015  Johns Hopkins University (Author: Daniel Povey).
+# Apache 2.0.
+
+# This script does decoding with a neural-net.  If the neural net was built on
+# top of fMLLR transforms from a conventional system, you should provide the
+# --transform-dir option.
+
+# Begin configuration section.
+stage=1
+transform_dir=    # dir to find fMLLR transforms.
+nj=4 # number of decoding jobs.  If --transform-dir set, must match that number!
+acwt=0.1  # Just a default value, used for adaptation and beam-pruning..
+cmd=run.pl
+beam=15.0
+max_active=7000
+min_active=200
+ivector_scale=1.0
+lattice_beam=8.0 # Beam we use in lattice generation.
+iter=final
+num_threads=1 # if >1, will use gmm-latgen-faster-parallel
+parallel_opts=  # ignored now.
+scoring_opts=
+skip_scoring=false
+feat_type=
+online_ivector_dir=
+minimize=false
+# End configuration section.
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f ./path.sh ] && . ./path.sh; # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# -ne 3 ]; then
+  echo "Usage: $0 [options] <graph-dir> <data-dir> <decode-dir>"
+  echo "e.g.:   steps/nnet3/decode.sh --nj 8 \\"
+  echo "--online-ivector-dir exp/nnet2_online/ivectors_test_eval92 \\"
+  echo "    exp/tri4b/graph_bg data/test_eval92_hires $dir/decode_bg_eval92"
+  echo "main options (for others, see top of script file)"
+  echo "  --transform-dir <decoding-dir>           # directory of previous decoding"
+  echo "                                           # where we can find transforms for SAT systems."
+  echo "  --config <config-file>                   # config containing options"
+  echo "  --nj <nj>                                # number of parallel jobs"
+  echo "  --cmd <cmd>                              # Command to run in parallel with"
+  echo "  --beam <beam>                            # Decoding beam; default 15.0"
[…]

### 2015-09-03 egs/wsj/s5/steps/online/nnet2/extract_ivectors_online.sh

@@ -1,0 +1,127 @@
+#!/bin/bash
+
+# Copyright     2013  Daniel Povey
+# Apache 2.0.
+
+
+# This script extracts iVectors for a set of utterances, given
+# features and a trained iVector extractor.
+
+# The script is based on ^/egs/sre08/v1/sid/extract_ivectors.sh.  Instead of
+# extracting a single iVector per utterance, it extracts one every few frames
+# (controlled by the --ivector-period option, e.g. 10, which is to save compute).
+# This is used in training (and not-really-online testing) of neural networks
+# for online decoding.
+
+# Rather than treating each utterance separately, it carries forward
+# information from one utterance to the next, within the speaker. 
+
+
+# Begin configuration section.
+nj=30
+cmd="run.pl"
+stage=0
+num_gselect=5 # Gaussian-selection using diagonal model: number of Gaussians to select
+min_post=0.025 # Minimum posterior to use (posteriors below this are pruned out)
+ivector_period=10
+posterior_scale=0.1 # Scale on the acoustic posteriors, intended to account for
+                    # inter-frame correlations.  Making this small during iVector
+                    # extraction is equivalent to scaling up the prior, and will
+                    # will tend to produce smaller iVectors where data-counts are
+                    # small.  It's not so important that this match the value
+                    # used when training the iVector extractor, but more important
+                    # that this match the value used when you do real online decoding
+                    # with the neural nets trained with these iVectors.
+compress=true       # If true, compress the iVectors stored on disk (it's lossy
+                    # compression, as used for feature matrices).
+max_count=0         # The use of this option (e.g. --max-count 100) can make
+                    # iVectors more consistent for different lengths of
+                    # utterance, by scaling up the prior term when the
+                    # data-count exceeds this value.  The data-count is after
+                    # posterior-scaling, so assuming the posterior-scale is 0.1,
+                    # --max-count 100 starts having effect after 1000 frames, or
+                    # 10 seconds of data.
+
[…]

### 2015-09-03 egs/wsj/s5/steps/tandem/train_mmi_sgmm2.sh

@@ -1,0 +1,190 @@
+#!/bin/bash
+# Copyright 2012  Johns Hopkins University (Author: Daniel Povey).  Apache 2.0.
+#                 Korbinian Riedhammer
+
+# MMI training (or optionally boosted MMI, if you give the --boost option),
+# for SGMMs.  4 iterations (by default) of Extended Baum-Welch update.
+#
+# Begin configuration section.
+cmd=run.pl
+num_iters=4
+boost=0.0
+cancel=true # if true, cancel num and den counts on each frame.
+acwt=0.1
+stage=0
+update_opts=
+transform_dir=
+# End configuration section
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f ./path.sh ] && . ./path.sh; # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# -ne 6 ]; then
+  echo "Usage: steps/tandem/train_mmi_sgmm2.sh <data1> <data2> <lang> <ali> <denlats> <exp>"
+  echo " e.g.: steps/tandem/train_mmi_sgmm2.sh {mfcc,bottleneck}/data1/train_si84 data1/lang exp/tri2b_ali_si84 exp/tri2b_denlats_si84 exp/tri2b_mmi"
+  echo "Main options (for others, see top of script file)"
+  echo "  --boost <boost-weight>                           # (e.g. 0.1), for boosted MMI.  (default 0)"
+  echo "  --cancel (true|false)                            # cancel stats (true by default)"
+  echo "  --cmd (utils/run.pl|utils/queue.pl <queue opts>) # how to run jobs."
+  echo "  --config <config-file>                           # config containing options"
+  echo "  --stage <stage>                                  # stage to do partial re-run from."  
+  echo "  --transform-dir <transform-dir>                  # directory to find fMLLR transforms."
+  exit 1;
+fi
+
+data1=$1
+data2=$2
+lang=$3
+alidir=$4
+denlatdir=$5
+dir=$6
+mkdir -p $dir/log
+
+for f in $data1/feats.scp $alidir/{tree,final.mdl,ali.1.gz} $denlatdir/lat.1.gz; do
[…]

### 2015-09-03 egs/wsj/s5/steps/train_deltas.sh

@@ -1,0 +1,165 @@
+#!/bin/bash
+
+# Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+# Apache 2.0
+
+# Begin configuration.
+stage=-4 #  This allows restarting after partway, when something when wrong.
+config=
+cmd=run.pl
+scale_opts="--transition-scale=1.0 --acoustic-scale=0.1 --self-loop-scale=0.1"
+realign_iters="10 20 30";
+num_iters=35    # Number of iterations of training
+max_iter_inc=25 # Last iter to increase #Gauss on.
+beam=10
+careful=false
+retry_beam=40
+boost_silence=1.0 # Factor by which to boost silence likelihoods in alignment
+power=0.25 # Exponent for number of gaussians according to occurrence counts
+cluster_thresh=-1  # for build-tree control final bottom-up clustering of leaves
+norm_vars=false # deprecated.  Prefer --cmvn-opts "--norm-vars=true"
+                # use the option --cmvn-opts "--norm-means=false"
+cmvn_opts=
+delta_opts=
+context_opts=   # use"--context-width=5 --central-position=2" for quinphone
+# End configuration.
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f path.sh ] && . ./path.sh;
+. parse_options.sh || exit 1;
+
+if [ $# != 6 ]; then
+   echo "Usage: steps/train_deltas.sh <num-leaves> <tot-gauss> <data-dir> <lang-dir> <alignment-dir> <exp-dir>"
+   echo "e.g.: steps/train_deltas.sh 2000 10000 data/train_si84_half data/lang exp/mono_ali exp/tri1"
+   echo "main options (for others, see top of script file)"
+   echo "  --cmd (utils/run.pl|utils/queue.pl <queue opts>) # how to run jobs."
+   echo "  --config <config-file>                           # config containing options"
+   echo "  --stage <stage>                                  # stage to do partial re-run from."
+   exit 1;
+fi
+
+numleaves=$1
+totgauss=$2
+data=$3
+lang=$4
+alidir=$5
+dir=$6
+
[…]

### 2015-09-03 egs/wsj/s5/steps/train_segmenter.sh

@@ -1,0 +1,144 @@
+#!/bin/bash
+
+# Copyright 2013  Johns Hopkins University (Author: Daniel Povey)
+# Apache 2.0
+
+# Begin configuration.
+stage=-4 # For restarting a process that went part way.
+config=
+cmd=run.pl
+
+scale_opts="--transition-scale=1.0 --acoustic-scale=0.1 --self-loop-scale=0.1"
+realign_iters="10 20 30";
+num_iters=35    # Number of iterations of training
+max_iter_inc=25 # Last iter to increase #Gauss on.
+beam=10
+retry_beam=40
+boost_silence=1.0 # Factor by which to boost silence likelihoods in alignment
+power=0.25 # Exponent for number of gaussians according to occurrence counts
+cluster_thresh=-1  # for build-tree control final bottom-up clustering of leaves
+# End configuration.
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f path.sh ] && . ./path.sh;
+. parse_options.sh || exit 1;
+
+if [ $# != 6 ]; then
+   echo "Usage: steps/train_segmenter.sh deltas.sh <num-leaves> <tot-gauss> <data-dir> <lang-dir> <alignment-dir> <exp-dir>"
+   echo "e.g.: steps/train_deltas.sh 2000 10000 data/train_si84_half data/lang exp/mono_ali exp/tri1"
+   echo "main options (for others, see top of script file)"
+   echo "  --cmd (utils/run.pl|utils/queue.pl <queue opts>) # how to run jobs."
+   echo "  --config <config-file>                           # config containing options"
+   echo "  --stage <stage>                                  # stage to do partial re-run from."
+   exit 1;
+fi
+
+numleaves=$1
+totgauss=$2
+data=$3
+lang=$4
+alidir=$5
+dir=$6
+
+for f in $alidir/final.mdl $alidir/ali.1.gz $data/feats.scp $lang/phones.txt; do
+  [ ! -f $f ] && echo "train_deltas.sh: no such file $f" && exit 1;
+done
+
+numgauss=$numleaves
[…]

### 2015-09-03 egs/wsj/s5/utils/format_lm.sh

@@ -1,0 +1,84 @@
+#!/bin/bash -u
+
+# Copyright 2012  Arnab Ghoshal
+# Copyright 2010-2011  Microsoft Corporation
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+set -o errexit
+
+if [ $# -ne 4 ]; then
+  printf "Usage: %s lang_dir LM lexicon out_dir\n" `basename $0`
+  echo "  Convert ARPA-format language models to FSTs.";
+  exit 1;
+fi
+
+lang_dir=$1
+lm=$2
+lexicon=$3
+out_dir=$4
+mkdir -p $out_dir
+
+[ -f ./path.sh ] && . ./path.sh
+
+echo "Converting '$lm' to FST"
+
+for f in phones.txt words.txt L.fst L_disambig.fst phones/; do
+  cp -r $lang_dir/$f $out_dir
+done
+
+lm_base=$(basename $lm '.gz')
+gunzip -c $lm | utils/find_arpa_oovs.pl $out_dir/words.txt \
+  > $out_dir/oovs_${lm_base}.txt
+
+# Removing all "illegal" combinations of <s> and </s>, which are supposed to 
+# occur only at being/end of utt.  These can cause determinization failures 
+# of CLG [ends up being epsilon cycles].
+gunzip -c $lm \
+  | egrep -v '<s> <s>|</s> <s>|</s> </s>' \
+  | arpa2fst - | fstprint \
+  | utils/remove_oovs.pl $out_dir/oovs_${lm_base}.txt \
+  | utils/eps2disambig.pl | utils/s2eps.pl \
+  | fstcompile --isymbols=$out_dir/words.txt --osymbols=$out_dir/words.txt \
+    --keep_isymbols=false --keep_osymbols=false \
[…]

### 2015-09-03 egs/wsj/s5/utils/nnet/make_lstm_proto.py

@@ -1,0 +1,76 @@
+#!/usr/bin/env python
+
+# Copyright 2015  Brno University of Technology (author: Karel Vesely)
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+# Generated Nnet prototype, to be initialized by 'nnet-initialize'.
+
+import sys
+
+###
+### Parse options
+###
+from optparse import OptionParser
+usage="%prog [options] <feat-dim> <num-leaves> >nnet-proto-file"
+parser = OptionParser(usage)
+#
+parser.add_option('--num-cells', dest='num_cells', type='int', default=800, 
+                   help='Number of LSTM cells [default: %default]');
+parser.add_option('--num-recurrent', dest='num_recurrent', type='int', default=512, 
+                   help='Number of LSTM recurrent units [default: %default]');
+parser.add_option('--num-layers', dest='num_layers', type='int', default=2, 
+                   help='Number of LSTM layers [default: %default]');
+parser.add_option('--lstm-stddev-factor', dest='lstm_stddev_factor', type='float', default=0.01, 
+                   help='Standard deviation of initialization [default: %default]');
+parser.add_option('--param-stddev-factor', dest='param_stddev_factor', type='float', default=0.04, 
+                   help='Standard deviation in output layer [default: %default]');
+parser.add_option('--clip-gradient', dest='clip_gradient', type='float', default=5.0, 
[…]

### 2015-09-03 egs/wsj/s5/utils/perturb_data_dir_speed.sh

@@ -1,0 +1,94 @@
+#!/bin/bash 
+
+# Copyright 2013  Johns Hopkins University (author: Daniel Povey)
+#           2014  Tom Ko
+# Apache 2.0
+
+# This script operates on a directory, such as in data/train/,
+# that contains some subset of the following files:
+#  wav.scp
+#  spk2utt
+#  utt2spk
+#  text
+#
+# It generates the files which are used for perturbing the speed of the original data.
+
+. utils/parse_options.sh
+
+if [ $# != 3 ]; then
+  echo "Usage: perturb_data_dir_speed.sh <warping-factor> <srcdir> <destdir>"
+  echo "e.g.:"
+  echo " $0 0.9 data/train_si284 data/train_si284p"
+  exit 1
+fi
+
+export LC_ALL=C
+
+factor=$1
+srcdir=$2
+destdir=$3
+label="sp"
+spk_prefix=$label$factor"-"
+utt_prefix=$label$factor"-"
+
+#check is sox on the path
+which sox &>/dev/null
+! [ $? -eq 0 ] && echo "sox: command not found" && exit 1;
+
+if [ ! -f $srcdir/utt2spk ]; then
+  echo "$0: no such file $srcdir/utt2spk" 
+  exit 1;
+fi
+
+set -e;
+set -o pipefail
+
+mkdir -p $destdir
+
+cat $srcdir/utt2spk | awk -v p=$utt_prefix '{printf("%s %s%s\n", $1, p, $1);}' > $destdir/utt_map
+cat $srcdir/spk2utt | awk -v p=$spk_prefix '{printf("%s %s%s\n", $1, p, $1);}' > $destdir/spk_map
+cat $srcdir/utt2spk | awk -v p=$utt_prefix '{printf("%s%s %s\n", p, $1, $1);}' > $destdir/utt2uniq
+
+cat $srcdir/utt2spk | utils/apply_map.pl -f 1 $destdir/utt_map  | \
+  utils/apply_map.pl -f 2 $destdir/spk_map >$destdir/utt2spk
+
+utils/utt2spk_to_spk2utt.pl <$destdir/utt2spk >$destdir/spk2utt
+
+if [ -f $srcdir/segments ]; then
+  # also apply the spk_prefix to the recording-ids.
+  cat $srcdir/wav.scp | awk -v p=$spk_prefix '{printf("%s %s%s\n", $1, p, $1);}' > $destdir/reco_map
+
[…]

### 2015-09-03 egs/wsj/s5/utils/split_data.sh

@@ -1,0 +1,135 @@
+#!/bin/bash
+# Copyright 2010-2013 Microsoft Corporation 
+#                     Johns Hopkins University (Author: Daniel Povey)
+
+# Licensed under the Apache License, Version 2.0 (the "License");
+# you may not use this file except in compliance with the License.
+# You may obtain a copy of the License at
+#
+#  http://www.apache.org/licenses/LICENSE-2.0
+#
+# THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+# KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+# WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+# MERCHANTABLITY OR NON-INFRINGEMENT.
+# See the Apache 2 License for the specific language governing permissions and
+# limitations under the License.
+
+split_per_spk=true
+if [ "$1" == "--per-utt" ]; then
+  split_per_spk=false
+  shift
+fi
+
+if [ $# != 2 ]; then
+  echo "Usage: split_data.sh [--per-utt] <data-dir> <num-to-split>"
+  echo "This script will not split the data-dir if it detects that the output is newer than the input."
+  echo "By default it splits per speaker (so each speaker is in only one split dir),"
+  echo "but with the --per-utt option it will ignore the speaker information while splitting."
+  exit 1
+fi
+
+data=$1
+numsplit=$2
+
+if [ $numsplit -le 0 ]; then
+  echo "Invalid num-split argument $numsplit";
+  exit 1;
+fi
+
+n=0;
+feats=""
+wavs=""
+utt2spks=""
+texts=""
+
+nu=`cat $data/utt2spk | wc -l`
+nf=`cat $data/feats.scp 2>/dev/null | wc -l`
+nt=`cat $data/text 2>/dev/null | wc -l` # take it as zero if no such file
+if [ -f $data/feats.scp ] && [ $nu -ne $nf ]; then
+  echo "** split_data.sh: warning, #lines is (utt2spk,feats.scp) is ($nu,$nf); you can "
+  echo "**  use utils/fix_data_dir.sh $data to fix this."
+fi
+if [ -f $data/text ] && [ $nu -ne $nt ]; then
[…]

### 2015-09-03 misc/maintenance/fix_apache_headers.sh

@@ -1,0 +1,27 @@
+#!/bin/bash
+
+# makes sure the line See ../../COPYING for clarification regarding multiple
+# authors appears in the apache headers in the source, and that source files
+# have their Apache headers.  Including this mainly for documentation, as I
+# doubt the issue will occur much in future.
+
+# run this from the top level of the repo, as
+# misc/maintenance/fix_apache_headers.sh
+
+set -e
+cd src
+rm -rf tmp
+for x in */*.{h,cc,dox}; do 
+  if [ $x != "util/basic-filebuf.h" ]; then
+    if ! grep 'COPYING for clarification' $x >/dev/null; then
+      echo Fixing $x;
+      if ! grep "Apache License" $x >/dev/null; then
+        echo "$0: warning: file $x may not have an Apache license header"
+      else
+        cp $x tmp; cat tmp | perl -ape ' if (m/Licensed under the Apache License/) { 
+        print "// See ../../COPYING for clarification regarding multiple authors\n"; 
+        print "//\n";} ' > $x;
+      fi
+    fi
+  fi
+done

### 2015-09-03 src/bin/align-equal.cc

@@ -1,0 +1,147 @@
+// bin/align-equal.cc
+
+// Copyright 2009-2013  Microsoft Corporation
+//                      Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "tree/context-dep.h"
+#include "hmm/transition-model.h"
+#include "fstext/fstext-lib.h"
+#include "decoder/training-graph-compiler.h"
+
+
+/** @brief Write equally spaced alignments of utterances (to get training started).
+*/
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    typedef kaldi::int32 int32;
+    using fst::SymbolTable;
+    using fst::VectorFst;
+    using fst::StdArc;
+
+    const char *usage =  "Write equally spaced alignments of utterances (to get training started)\n"
+        "Usage:  align-equal <tree-in> <model-in> <lexicon-fst-in> <features-rspecifier> <transcriptions-rspecifier> <alignments-wspecifier>\n"
+        "e.g.: \n"
+        " align-equal 1.tree 1.mdl lex.fst scp:train.scp ark:train.tra ark:equal.ali\n";
+
+    ParseOptions po(usage);
+    std::string disambig_rxfilename;
+    po.Register("read-disambig-syms", &disambig_rxfilename, "File containing "
+                "list of disambiguation symbols in phone symbol table");
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 6) {
+      po.PrintUsage();
+      exit(1);
+    }
+
[…]

### 2015-09-03 src/bin/prons-to-wordali.cc

@@ -1,0 +1,145 @@
+// bin/prons-to-wordali.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "hmm/transition-model.h"
+#include "hmm/hmm-utils.h"
+#include "util/common-utils.h"
+#include "fst/fstlib.h"
+#include "fstext/fstext-utils.h"
+
+
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  using fst::VectorFst;
+  using fst::StdArc;
+  typedef kaldi::int32 int32;
+  try {
+    const char *usage =
+        "Caution: this program relates to older scripts and is deprecated,\n"
+        "for modern scripts see egs/wsj/s5/steps/{get_ctm,get_train_ctm}.sh\n"
+        "Given per-utterance pronunciation information as output by \n"
+        "words-to-prons, and per-utterance phone alignment information\n"
+        "as output by ali-to-phones --write-lengths, output word alignment\n"
+        "information that can be turned into the ctm format.\n"
+        "Outputs is pairs of (word, #frames), or if --per-frame is given,\n"
+        "just the word for each frame.\n"
+        "Note: zero word-id usually means optional silence.\n"
+        "Format is standard format for archives of vector<pair<int32, int32> >\n"
+        "i.e. :\n"
+        "utt-id  600 22 ; 1028 32 ; 0 41\n"
+        "where 600, 1028 and 0 are the word-ids, and 22, 32 and 41 are the\n"
+        "lengths.\n"
+        "\n"
+        "Usage:  prons-to-wordali [options] <prons-rspecifier>"
[…]

### 2015-09-03 src/cudamatrix/cu-sp-matrix.h

@@ -1,0 +1,163 @@
+// cudamatrix/cu-sp-matrix.h
+
+// Copyright 2009-2013  Karel Vesely
+//                2014  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#ifndef KALDI_CUDAMATRIX_CU_SP_MATRIX_H_
+#define KALDI_CUDAMATRIX_CU_SP_MATRIX_H_
+
+#include <sstream>
+
+#include "cudamatrix/cu-common.h"
+#include "matrix/matrix-common.h"
+#include "matrix/sp-matrix.h"
+#include "cudamatrix/cu-array.h"
+#include "cudamatrix/cu-math.h"
+#include "cudamatrix/cu-packed-matrix.h"
+#include "cudamatrix/cu-matrix.h"
+
+namespace kaldi {
+
+/// TraceSpSp returns tr(A B)
+template<typename Real, typename OtherReal>
+Real TraceSpSp(const CuSpMatrix<Real> &A, const CuSpMatrix<OtherReal> &B);
+
+template<typename Real>
+class CuSpMatrix : public CuPackedMatrix<Real> {
+  friend class CuMatrixBase<Real>;
+  friend class CuVectorBase<Real>;
+  friend class CuTpMatrix<Real>;
+  friend class CuSubMatrix<Real>;
+  friend class CuRand<Real>;
+
+  template<class R, class S>
+  friend R TraceSpSp(const CuSpMatrix<R> &A, const CuSpMatrix<S> &B);
+ public:
+  
+  CuSpMatrix(): CuPackedMatrix<Real>() {}
+  
[…]

### 2015-09-03 src/decoder/lattice-faster-decoder.h

@@ -1,0 +1,419 @@
+// decoder/lattice-faster-decoder.h
+
+// Copyright 2009-2013  Microsoft Corporation;  Mirko Hannemann;
+//           2013-2014  Johns Hopkins University (Author: Daniel Povey)
+//                2014  Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+// Note: this file is "upstream" from lattice-faster-online-decoder.h,
+// and changes in this file should be made to lattice-faster-online-decoder.h,
+// if applicable.
+
+#ifndef KALDI_DECODER_LATTICE_FASTER_DECODER_H_
+#define KALDI_DECODER_LATTICE_FASTER_DECODER_H_
+
+
+#include "util/stl-utils.h"
+#include "util/hash-list.h"
+#include "fst/fstlib.h"
+#include "itf/decodable-itf.h"
+#include "fstext/fstext-lib.h"
+#include "lat/determinize-lattice-pruned.h"
+#include "lat/kaldi-lattice.h"
+
+namespace kaldi {
+
+struct LatticeFasterDecoderConfig {
+  BaseFloat beam;
+  int32 max_active;
+  int32 min_active;
+  BaseFloat lattice_beam;
+  int32 prune_interval;
+  bool determinize_lattice; // not inspected by this class... used in
+                            // command-line program.
+  BaseFloat beam_delta; // has nothing to do with beam_ratio
+  BaseFloat hash_ratio;
+  BaseFloat prune_scale;   // Note: we don't make this configurable on the command line,
+                           // it's not a very important parameter.  It affects the
+                           // algorithm that prunes the tokens as we go.
[…]

### 2015-09-03 src/decoder/training-graph-compiler.h

@@ -1,0 +1,110 @@
+// decoder/training-graph-compiler.h
+
+// Copyright 2009-2011 Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+
+//  http://www.apache.org/licenses/LICENSE-2.0
+
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+#ifndef KALDI_DECODER_TRAINING_GRAPH_COMPILER_H_
+#define KALDI_DECODER_TRAINING_GRAPH_COMPILER_H_
+
+#include "base/kaldi-common.h"
+#include "hmm/transition-model.h"
+#include "fst/fstlib.h"
+#include "fstext/fstext-lib.h"
+
+
+namespace kaldi {
+
+struct TrainingGraphCompilerOptions {
+
+  BaseFloat transition_scale;
+  BaseFloat self_loop_scale;
+  bool rm_eps;
+  bool reorder;  // (Dan-style graphs)
+
+  explicit TrainingGraphCompilerOptions(BaseFloat transition_scale = 1.0,
+                                        BaseFloat self_loop_scale = 1.0,
+                                        bool b = true) :
+      transition_scale(transition_scale),
+      self_loop_scale(self_loop_scale),
+      rm_eps(false),
+      reorder(b) { }
+
+  void Register(OptionsItf *opts) {
+    opts->Register("transition-scale", &transition_scale, "Scale of transition "
+                   "probabilities (excluding self-loops)");
+    opts->Register("self-loop-scale", &self_loop_scale, "Scale of self-loop vs. "
+                   "non-self-loop probability mass ");
+    opts->Register("reorder", &reorder, "Reorder transition ids for greater decoding efficiency.");
+    opts->Register("rm-eps", &rm_eps,  "Remove [most] epsilons before minimization (only applicable "
[…]

### 2015-09-03 src/featbin/compute-cmvn-stats.cc

@@ -1,0 +1,180 @@
+// featbin/compute-cmvn-stats.cc
+
+// Copyright 2009-2012  Microsoft Corporation
+//                      Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "matrix/kaldi-matrix.h"
+#include "transform/cmvn.h"
+
+namespace kaldi {
+
+bool AccCmvnStatsWrapper(std::string utt,
+                         const MatrixBase<BaseFloat> &feats,
+                         RandomAccessBaseFloatVectorReader *weights_reader,
+                         Matrix<double> *cmvn_stats) {
+  if (!weights_reader->IsOpen()) {
+    AccCmvnStats(feats, NULL, cmvn_stats);
+    return true;
+  } else {
+    if (!weights_reader->HasKey(utt)) {
+      KALDI_WARN << "No weights available for utterance " << utt;
+      return false;
+    }
+    const Vector<BaseFloat> &weights = weights_reader->Value(utt);
+    if (weights.Dim() != feats.NumRows()) {
+      KALDI_WARN << "Weights for utterance " << utt << " have wrong dimension "
+                 << weights.Dim() << " vs. " << feats.NumRows();
+      return false;
+    }
+    AccCmvnStats(feats, &weights, cmvn_stats);
+    return true;
+  }
+}
+                  
+
+}
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    using kaldi::int32;
+
+    const char *usage =
+        "Compute cepstral mean and variance normalization statistics\n"
[…]

### 2015-09-03 src/featbin/paste-feats.cc

@@ -1,0 +1,203 @@
+// featbin/paste-feats.cc
+
+// Copyright 2012 Korbinian Riedhammer
+//           2013 Brno University of Technology (Author: Karel Vesely)
+//           2013 Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "matrix/kaldi-matrix.h"
+
+namespace kaldi {
+
+// returns true if successfully appended.
+bool AppendFeats(const std::vector<Matrix<BaseFloat> > &in,
+                 std::string utt,
+                 int32 tolerance,
+                 Matrix<BaseFloat> *out) {
+  // Check the lengths
+  int32 min_len = in[0].NumRows(),
+      max_len = in[0].NumRows(),
+      tot_dim = in[0].NumCols();
+  for (int32 i = 1; i < in.size(); i++) {
+    int32 len = in[i].NumRows(), dim = in[i].NumCols();
+    tot_dim += dim;
+    if(len < min_len) min_len = len;
+    if(len > max_len) max_len = len;
+  }
+  if (max_len - min_len > tolerance || min_len == 0) {
+    KALDI_WARN << "Length mismatch " << max_len << " vs. " << min_len
+               << (utt.empty() ? "" : " for utt ") << utt
+               << " exceeds tolerance " << tolerance;
+    out->Resize(0, 0);
+    return false;
+  }
+  if (max_len - min_len > 0) {
+    KALDI_VLOG(2) << "Length mismatch " << max_len << " vs. " << min_len
[…]

### 2015-09-03 src/featbin/process-pitch-feats.cc

@@ -1,0 +1,185 @@
+// featbin/process-pitch-feats.cc
+
+// Copyright 2013   Bagher BabaAli
+//                  Johns Hopkins University (author: Daniel Povey)
+//
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+
+namespace kaldi {
+
+// For the probability-of-voicing features (the first element of
+// each row), change p -> log ((p + 0.0001) / (1.00001 - p))
+// This makes it more Gaussian; otherwise it clumps up near
+// the edges of the range [0, 1].
+void ProcessPovFeatures(Matrix<BaseFloat> *mat) {
+  int32 num_frames = mat->NumRows();
+  for (int32 i = 0; i < num_frames; i++) {
+    BaseFloat p = (*mat)(i, 0);
+    KALDI_ASSERT(p >= 0.0 && p <= 1.0);
+    (*mat)(i, 0) = Log((p + 0.0001) / (1.0001 - p));
+  }
+}
+
+void TakeLogOfPitch(Matrix<BaseFloat> *mat) {
+  int32 num_frames = mat->NumRows();
+  for (int32 i = 0; i < num_frames; i++) {
+    KALDI_ASSERT((*mat)(i, 1) > 0.0);
+    (*mat)(i, 1) = Log((*mat)(i, 1));
+  }
+}
+
+// Subtract the moving average over a largish window
+// (e.g. 151 frames)
+void SubtractMovingAverage(int32 normalization_window_size,
+                           Matrix<BaseFloat> *mat) {
+  int32 num_frames = mat->NumRows();
[…]

### 2015-09-03 src/fstext/determinize-lattice-test.cc

@@ -1,0 +1,196 @@
+// fstext/determinize-lattice-test.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "fstext/determinize-lattice.h"
+#include "fstext/lattice-utils.h"
+#include "fstext/fst-test-utils.h"
+#include "base/kaldi-math.h"
+
+namespace fst {
+
+void TestLatticeStringRepository() {
+  typedef int32 IntType;
+
+  LatticeStringRepository<IntType> sr;
+  typedef LatticeStringRepository<IntType>::Entry Entry;
+
+  for(int i = 0; i < 100; i++) {
+    int len = kaldi::Rand() % 5;
+    vector<IntType> str(len), str2(kaldi::Rand() % 4);
+    const Entry *e = NULL;
+    for(int i = 0; i < len; i++) {
+      str[i] = kaldi::Rand() % 5;
+      e = sr.Successor(e, str[i]);
+    }
+    sr.ConvertToVector(e, &str2);
+    assert(str == str2);
+
+    int len2 = kaldi::Rand() % 5;
+    str2.resize(len2);
+    const Entry *f = sr.EmptyString(); // NULL
+    for(int i = 0; i < len2; i++) {
+      str2[i] = kaldi::Rand() % 5;
+      f = sr.Successor(f, str2[i]);
+    }
+    vector<IntType> prefix, prefix2(kaldi::Rand() % 10),
+        prefix3;
+    for(int i = 0; i < len && i < len2; i++) {
+      if (str[i] == str2[i]) prefix.push_back(str[i]);
+      else break;
+    }
+    const Entry *g = sr.CommonPrefix(e, f);
[…]

### 2015-09-03 src/fstext/kaldi-fst-io.cc

@@ -1,0 +1,55 @@
+// fstext/kaldi-fst-io.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+//           2012-2015  Johns Hopkins University (Author: Daniel Povey)
+//                2013  Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "fstext/kaldi-fst-io.h"
+#include "base/kaldi-error.h"
+#include "base/kaldi-math.h"
+#include "util/kaldi-io.h"
+
+namespace fst {
+
+VectorFst<StdArc> *ReadFstKaldi(std::string rxfilename) {
+  if (rxfilename == "") rxfilename = "-"; // interpret "" as stdin,
+  // for compatibility with OpenFst conventions.
+  kaldi::Input ki(rxfilename);
+  fst::FstHeader hdr;
+  if (!hdr.Read(ki.Stream(), rxfilename))
+    KALDI_ERR << "Reading FST: error reading FST header from "
+              << kaldi::PrintableRxfilename(rxfilename);
+  FstReadOptions ropts("<unspecified>", &hdr);
+  VectorFst<StdArc> *fst = VectorFst<StdArc>::Read(ki.Stream(), ropts);
+  if (!fst)
+    KALDI_ERR << "Could not read fst from "
+              << kaldi::PrintableRxfilename(rxfilename);
+  return fst;
+}
+
+void WriteFstKaldi(const VectorFst<StdArc> &fst,
+                   std::string wxfilename) {
+  if (wxfilename == "") wxfilename = "-"; // interpret "" as stdout,
+  // for compatibility with OpenFst conventions.
[…]

### 2015-09-03 src/fstext/trivial-factor-weight-test.cc

@@ -1,0 +1,255 @@
+// fstext/trivial-factor-weight-test.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-math.h"
+#include "fstext/pre-determinize.h"
+#include "fstext/determinize-star.h"
+#include "fstext/trivial-factor-weight.h"
+#include "fstext/fst-test-utils.h"
+// Just check that it compiles, for now.
+
+namespace fst
+{
+
+// Don't instantiate with log semiring, as RandEquivalent may fail.
+template<class Arc>  void TestFactor() {
+  typedef typename Arc::Label Label;
+  typedef typename Arc::StateId StateId;
+  typedef typename Arc::Weight Weight;
+
+  VectorFst<Arc> *fst = new VectorFst<Arc>();
+  int n_syms = 2 + kaldi::Rand() % 5, n_states = 3 + kaldi::Rand() % 10, n_arcs = 5 + kaldi::Rand() % 30, n_final = 1 + kaldi::Rand()%3;  // Up to 2 unique symbols.
+  cout << "Testing pre-determinize with "<<n_syms<<" symbols, "<<n_states<<" states and "<<n_arcs<<" arcs and "<<n_final<<" final states.\n";
+  SymbolTable *sptr = NULL;
+
+  vector<Label> all_syms;  // including epsilon.
+  // Put symbols in the symbol table from 1..n_syms-1.
[…]

### 2015-09-03 src/gmm/full-gmm.h

@@ -1,0 +1,232 @@
+// gmm/full-gmm.h
+
+// Copyright 2009-2011  Jan Silovsky;
+//                      Saarland University (Author: Arnab Ghoshal);
+//                      Microsoft Corporation
+//           2012       Arnab Ghoshal
+//           2013       Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#ifndef KALDI_GMM_FULL_GMM_H_
+#define KALDI_GMM_FULL_GMM_H_
+
+#include <utility>
+#include <vector>
+
+#include "base/kaldi-common.h"
+#include "gmm/model-common.h"
+#include "matrix/matrix-lib.h"
+
+namespace kaldi {
+
+class DiagGmm;
+class FullGmmNormal;  // a simplified representation, see full-gmm-normal.h
+
+/** \class Definition for Gaussian Mixture Model with full covariances
+  */
+class FullGmm {
+  /// this makes it a little easier to modify the internals
+  friend class FullGmmNormal;
+
+ public:
+  /// Empty constructor.
+  FullGmm() : valid_gconsts_(false) {}
+
+  explicit FullGmm(const FullGmm &gmm): valid_gconsts_(false) {
+    CopyFromFullGmm(gmm);
+  }
+
+  FullGmm(int32 nMix, int32 dim): valid_gconsts_(false) { Resize(nMix, dim); }
+
+  /// Resizes arrays to this dim. Does not initialize data.
+  void Resize(int32 nMix, int32 dim);
+
+  /// Returns the number of mixture components in the GMM
+  int32 NumGauss() const { return weights_.Dim(); }
[…]

### 2015-09-03 src/gmmbin/gmm-diff-accs.cc

@@ -1,0 +1,87 @@
+// gmmbin/gmm-diff-accs.cc
+
+// Copyright 2012  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "util/common-utils.h"
+#include "gmm/mle-am-diag-gmm.h"
+#include "hmm/transition-model.h"
+
+
+int main(int argc, char *argv[]) {
+  typedef kaldi::int32 int32;
+  try {
+    const char *usage =
+        "Compute difference between accumulators\n"
+        "Usage: gmm-diff-accs [options] plus-stats minus-stats diff-stats-out\n";
+        
+    bool binary = true;
+    kaldi::ParseOptions po(usage);
+    po.Register("binary", &binary, "Write output in binary mode");
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 3) {
+      po.PrintUsage();
+      exit(1);
+    }
+
+    std::string plus_stats_rxfilename = po.GetArg(1),
+        minus_stats_rxfilename = po.GetArg(2),
+        diff_stats_wxfilename = po.GetArg(3);
+
+    kaldi::Vector<double> transition_accs_plus;
+    kaldi::AccumAmDiagGmm gmm_accs_plus;
+    {
+      bool binary_read;
+      kaldi::Input ki(plus_stats_rxfilename, &binary_read);
+      transition_accs_plus.Read(ki.Stream(), binary_read, false);
+      gmm_accs_plus.Read(ki.Stream(), binary_read, false);
+    }
+    kaldi::Vector<double> transition_accs_minus;
+    kaldi::AccumAmDiagGmm gmm_accs_minus;
+    {
[…]

### 2015-09-03 src/gmmbin/gmm-est-basis-fmllr.cc

@@ -1,0 +1,224 @@
+// gmmbin/gmm-est-basis-fmllr.cc
+
+// Copyright 2012  Carnegie Mellon University (author: Yajie Miao)
+//           2014  Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include <string>
+using std::string;
+#include <vector>
+using std::vector;
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "gmm/am-diag-gmm.h"
+#include "hmm/transition-model.h"
+#include "transform/fmllr-diag-gmm.h"
+#include "transform/basis-fmllr-diag-gmm.h"
+#include "hmm/posterior.h"
+
+namespace kaldi {
+void AccumulateForUtterance(const Matrix<BaseFloat> &feats,
+                            const Posterior &post,
+                            const TransitionModel &trans_model,
+                            const AmDiagGmm &am_gmm,
+                            FmllrDiagGmmAccs *spk_stats) {
+  Posterior pdf_post;
+  ConvertPosteriorToPdfs(trans_model, post, &pdf_post);
+  for (size_t i = 0; i < post.size(); i++) {
+    for (size_t j = 0; j < pdf_post[i].size(); j++) {
+      int32 pdf_id = pdf_post[i][j].first;
+      spk_stats->AccumulateForGmm(am_gmm.GetPdf(pdf_id),
+                                  feats.Row(i),
+                                  pdf_post[i][j].second);
+    }
+  }
+}
+
+
+}
+
+int main(int argc, char *argv[]) {
+  try {
+    typedef kaldi::int32 int32;
+    using namespace kaldi;
+    const char *usage =
[…]

### 2015-09-03 src/gmmbin/gmm-init-model-flat.cc

@@ -1,0 +1,153 @@
+// gmmbin/gmm-init-model-flat.cc
+
+// Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "gmm/am-diag-gmm.h"
+#include "hmm/transition-model.h"
+#include "gmm/mle-am-diag-gmm.h"
+#include "tree/build-tree-utils.h"
+#include "tree/context-dep.h"
+#include "tree/clusterable-classes.h"
+#include "util/text-utils.h"
+
+namespace kaldi {
+
+void GetFeatureMeanAndVariance(const std::string &feat_rspecifier,
+                               Vector<BaseFloat> *inv_var_out,                               
+                               Vector<BaseFloat> *mean_out) {
+  double count = 0.0;
+  Vector<double> x_stats, x2_stats;
+
+  SequentialDoubleMatrixReader feat_reader(feat_rspecifier);
+  for (; !feat_reader.Done(); feat_reader.Next()) {
+    const Matrix<double> &mat = feat_reader.Value();
+    if (x_stats.Dim() == 0) {
+      int32 dim = mat.NumCols();
+      x_stats.Resize(dim);
+      x2_stats.Resize(dim);
+    }
+    for (int32 i = 0; i < mat.NumRows(); i++) {
+      count += 1.0;
+      x_stats.AddVec(1.0, mat.Row(i));
+      x2_stats.AddVec2(1.0, mat.Row(i));
+    }
+  }
[…]

### 2015-09-03 src/gmmbin/gmm-scale-accs.cc

@@ -1,0 +1,80 @@
+// gmmbin/gmm-scale-accs.cc
+
+// Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "util/common-utils.h"
+#include "gmm/mle-am-diag-gmm.h"
+#include "hmm/transition-model.h"
+
+
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  try {
+    typedef kaldi::int32 int32;
+
+    const char *usage =
+        "Scale GMM accumulators\n"
+        "Usage: gmm-scale-accs [options] scale stats-in stats-out\n"
+        "e.g.: gmm-scale-accs 0.5 1.stats half.stats\n";
+
+    bool binary = true;
+    kaldi::ParseOptions po(usage);
+    po.Register("binary", &binary, "Write output in binary mode");
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 3) {
+      po.PrintUsage();
+      exit(1);
+    }
+
+    std::string
+        scale_string = po.GetArg(1),
+        stats_rxfilename = po.GetArg(2),
+        stats_wxfilename = po.GetArg(3);
+
+    BaseFloat scale;
+    if (!ConvertStringToReal(scale_string, &scale))
+      KALDI_ERR << "Invalid first argument to gmm-scale-accs: expect a number: "
+                << scale_string;
+          
+    kaldi::Vector<double> transition_accs;
+    kaldi::AccumAmDiagGmm gmm_accs;
+
+    {
+      bool binary_read;
[…]

### 2015-09-03 src/hmm/tree-accu.cc

@@ -1,0 +1,142 @@
+// hmm/tree-accu.cc
+
+// Copyright 2009-2011 Microsoft Corporation
+//                2013 Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+
+//  http://www.apache.org/licenses/LICENSE-2.0
+
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+#include "util/kaldi-io.h"
+#include "hmm/tree-accu.h"
+#include "hmm/hmm-utils.h"
+
+namespace kaldi {
+
+static int32 MapPhone(const std::vector<int32> *phone_map,
+                      int32 phone) {
+  if (phone == 0 || phone_map == NULL) return phone;
+  else if (phone < 0 || phone >= phone_map->size()) {
+    KALDI_ERR << "Out-of-range phone " << phone << " bad --phone-map option?";
+  }
+  return (*phone_map)[phone];
+}
+
+
+void AccumulateTreeStats(const TransitionModel &trans_model,
+                         BaseFloat var_floor,
+                         int N,  // context window size.
+                         int P,  // central position.
+                         const std::vector<int32> &ci_phones,
+                         const std::vector<int32> &alignment,
+                         const Matrix<BaseFloat> &features,
+                         const std::vector<int32> *phone_map,
+                         std::map<EventType, GaussClusterable*> *stats) {
+
+  KALDI_ASSERT(IsSortedAndUniq(ci_phones));
+  std::vector<std::vector<int32> > split_alignment;
+  bool ans = SplitToPhones(trans_model, alignment, &split_alignment);
+  if (!ans) {
+    KALDI_WARN << "AccumulateTreeStats: alignment appears to be bad, not using it";
+    return;
+  }
+  int cur_pos = 0;
+  int dim = features.NumCols();
[…]

### 2015-09-03 src/ivector/voice-activity-detection.h

@@ -1,0 +1,89 @@
+// ivector/voice-activity-detection.h
+
+// Copyright  2013   Daniel Povey
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#ifndef KALDI_IVECTOR_VOICE_ACTIVITY_DETECTION_H_
+#define KALDI_IVECTOR_VOICE_ACTIVITY_DETECTION_H_
+
+#include <cassert>
+#include <cstdlib>
+#include <string>
+#include <vector>
+
+#include "matrix/matrix-lib.h"
+#include "util/common-utils.h"
+#include "base/kaldi-error.h"
+
+namespace kaldi {
+
+/*
+  Note: we may move the location of this file in the future, e.g. to feat/
+  This code is geared toward speaker-id applications and is not suitable
+  for automatic speech recognition (ASR) because it makes independent
+  decisions for each frame without imposing any notion of continuity.
+*/
+ 
+struct VadEnergyOptions {
+  BaseFloat vad_energy_threshold;
+  BaseFloat vad_energy_mean_scale;
+  int32 vad_frames_context;
+  BaseFloat vad_proportion_threshold;
+  
+  VadEnergyOptions(): vad_energy_threshold(5.0),
+                      vad_energy_mean_scale(0.5),
+                      vad_frames_context(5),
+                      vad_proportion_threshold(0.6) { }
+  void Register(OptionsItf *opts) {
+    opts->Register("vad-energy-threshold", &vad_energy_threshold,
+                   "Constant term in energy threshold for MFCC0 for VAD (also see "
+                   "--vad-energy-mean-scale)");
+    opts->Register("vad-energy-mean-scale", &vad_energy_mean_scale,
[…]

### 2015-09-03 src/ivectorbin/ivector-mean.cc

@@ -1,0 +1,156 @@
+// ivectorbin/ivector-mean.cc
+
+// Copyright 2013-2014  Daniel Povey
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+
+
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  typedef kaldi::int32 int32;
+  try {
+    const char *usage =
+        "With 3 or 4 arguments, averages iVectors over all the\n"
+        "utterances of each speaker using the spk2utt file.\n"
+        "Input the spk2utt file and a set of iVectors indexed by\n"
+        "utterance; output is iVectors indexed by speaker.  If 4\n"
+        "arguments are given, extra argument is a table for the number\n"
+        "of utterances per speaker (can be useful for PLDA).  If 2\n"
+        "arguments are given, computes the mean of all input files and\n"
+        "writes out the mean vector.\n"
+        "\n"
+        "Usage: ivector-mean <spk2utt-rspecifier> <ivector-rspecifier> "
+        "<ivector-wspecifier> [<num-utt-wspecifier>]\n"
+        "or: ivector-mean <ivector-rspecifier> <mean-wxfilename>\n"
+        "e.g.: ivector-mean data/spk2utt exp/ivectors.ark exp/spk_ivectors.ark exp/spk_num_utts.ark\n"
+        "or: ivector-mean exp/ivectors.ark exp/mean.vec\n"
+        "See also: ivector-subtract-global-mean\n";
+    
+    ParseOptions po(usage);
+    bool binary_write = false;
+    po.Register("binary", &binary_write, "If true, write output in binary "
[…]

### 2015-09-03 src/lat/minimize-lattice-test.cc

@@ -1,0 +1,73 @@
+// lat/minimize-lattice-test.cc
+
+// Copyright 2013  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "lat/kaldi-lattice.h"
+#include "lat/minimize-lattice.h"
+#include "lat/push-lattice.h"
+#include "fstext/rand-fst.h"
+
+
+namespace kaldi {
+using namespace fst;
+
+CompactLattice *RandDeterministicCompactLattice() {
+  RandFstOptions opts;
+  opts.acyclic = true;
+  while (1) {
+    Lattice *fst = fst::RandPairFst<LatticeArc>(opts);
+    CompactLattice *cfst = new CompactLattice;
+    if (!DeterminizeLattice(*fst, cfst)) {
+      delete fst;
+      delete cfst;
+      KALDI_WARN << "Determinization failed, trying again.";
+    } else {
+      delete fst;
+      return cfst;
+    }
+  }
+}
+
+void TestMinimizeCompactLattice() {
+  CompactLattice *clat = RandDeterministicCompactLattice();
+  CompactLattice clat2(*clat);
+  BaseFloat delta = (Rand() % 2 == 0 ? 1.0 : 1.0e-05);
+
+  // Minimization will only work well on determinized and pushed lattices.
+  PushCompactLatticeStrings(&clat2);
+  PushCompactLatticeWeights(&clat2);
+  
+  MinimizeCompactLattice(&clat2, delta);
[…]

### 2015-09-03 src/lat/push-lattice.h

@@ -1,0 +1,62 @@
+// lat/push-lattice.h
+
+// Copyright 2013        Johns Hopkins University (Author: Daniel Povey)
+//           2014        Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#ifndef KALDI_LAT_PUSH_LATTICE_H_
+#define KALDI_LAT_PUSH_LATTICE_H_
+
+#include <vector>
+#include <map>
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "fstext/fstext-lib.h"
+#include "hmm/transition-model.h"
+#include "lat/kaldi-lattice.h"
+
+namespace fst {
+
+
+/// This function pushes the transition-ids as far towards the start as they
+/// will go.  It can be useful prior to lattice-align-words (for non-linear
+/// lattices).  We can't use the generic OpenFst "push" function because
+/// it uses the sum as the divisor, which is not appropriate in this case
+/// (a+b generally won't divide a or b in this semiring).
+/// It returns true on success, false if it failed due to TopSort failing,
+/// which should never happen, but we handle it gracefully by just leaving the
+/// lattice the same.
+/// This function used to be called just PushCompactLattice.
+template<class Weight, class IntType>
+bool PushCompactLatticeStrings(
+    MutableFst<ArcTpl<CompactLatticeWeightTpl<Weight, IntType> > > *clat);
+
+/// This function pushes the weights in the CompactLattice so that all states
+/// except possibly the start state, have Weight components (of type
+/// LatticeWeight) that "sum to one" in the LatticeWeight (i.e. interpreting the
[…]

### 2015-09-03 src/lat/sausages.cc

@@ -1,0 +1,364 @@
+// lat/sausages.cc
+
+// Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+//           2015  Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "lat/sausages.h"
+#include "lat/lattice-functions.h"
+
+namespace kaldi {
+
+// this is Figure 6 in the paper.
+void MinimumBayesRisk::MbrDecode() {
+  
+  for (size_t counter = 0; ; counter++) {
+    NormalizeEps(&R_);
+    AccStats(); // writes to gamma_
+    double delta_Q = 0.0; // change in objective function.
+
+    one_best_times_.clear();
+    one_best_confidences_.clear();
+    
+    // Caution: q in the line below is (q-1) in the algorithm
+    // in the paper; both R_ and gamma_ are indexed by q-1.
+    for (size_t q = 0; q < R_.size(); q++) {
+      if (do_mbr_) { // This loop updates R_ [indexed same as gamma_]. 
+        // gamma_[i] is sorted in reverse order so most likely one is first.
+        const vector<pair<int32, BaseFloat> > &this_gamma = gamma_[q];
+        double old_gamma = 0, new_gamma = this_gamma[0].second;
+        int32 rq = R_[q], rhat = this_gamma[0].first; // rq: old word, rhat: new.
+        for (size_t j = 0; j < this_gamma.size(); j++)
+          if (this_gamma[j].first == rq) old_gamma = this_gamma[j].second;
[…]

### 2015-09-03 src/latbin/lattice-align-words.cc

@@ -1,0 +1,134 @@
+// latbin/lattice-align-words.cc
+
+// Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "lat/kaldi-lattice.h"
+#include "lat/word-align-lattice.h"
+#include "lat/lattice-functions.h"
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    using fst::StdArc;
+    using kaldi::int32;
+
+    const char *usage =
+        "Convert lattices so that the arcs in the CompactLattice format correspond with\n"
+        "words (i.e. aligned with word boundaries).  Note: it will generally be more\n"
+        "efficient if you apply 'lattice-push' before this program.\n"
+        "Usage: lattice-align-words [options] <word-boundary-file> <model> <lattice-rspecifier> <lattice-wspecifier>\n"
+        " e.g.: lattice-align-words  --silence-label=4320 --partial-word-label=4324 \\\n"
+        "   data/lang/phones/word_boundary.int final.mdl ark:1.lats ark:aligned.lats\n"
+        "Note: word-boundary file has format (on each line):\n"
+        "<integer-phone-id> [begin|end|singleton|internal|nonword]\n"
+        "See also: lattice-align-words-lexicon, for use in cases where phones\n"
+        "don't have word-position information.\n";
+    
+    ParseOptions po(usage);
+    BaseFloat max_expand = 0.0;
+    bool output_if_error = true;
+    bool do_test = false;
+    
+    po.Register("output-error-lats", &output_if_error, "Output lattices that aligned "
[…]

### 2015-09-03 src/latbin/lattice-best-path.cc

@@ -1,0 +1,137 @@
+// latbin/lattice-best-path.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "fstext/fstext-lib.h"
+#include "lat/kaldi-lattice.h"
+#include "lat/lattice-functions.h"
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+    using fst::SymbolTable;
+    using fst::VectorFst;
+    using fst::StdArc;
+
+    const char *usage =
+        "Generate 1-best path through lattices; output as transcriptions and alignments\n"
+        "Note: if you want output as FSTs, use lattice-1best; if you want output\n"
+        "with acoustic and LM scores, use lattice-1best | nbest-to-linear\n"
+        "Usage: lattice-best-path [options]  lattice-rspecifier [ transcriptions-wspecifier [ alignments-wspecifier] ]\n"
+        " e.g.: lattice-best-path --acoustic-scale=0.1 ark:1.lats ark:1.tra ark:1.ali\n";
+      
+    ParseOptions po(usage);
+    BaseFloat acoustic_scale = 1.0;
+    BaseFloat lm_scale = 1.0;
+
+    std::string word_syms_filename;
+    po.Register("acoustic-scale", &acoustic_scale, "Scaling factor for acoustic likelihoods");
+    po.Register("lm-scale", &lm_scale, "Scaling factor for LM probabilities. "
+                "Note: the ratio acoustic-scale/lm-scale is all that matters.");
[…]

### 2015-09-03 src/latbin/lattice-compose.cc

@@ -1,0 +1,166 @@
+// latbin/lattice-compose.cc
+
+// Copyright 2009-2011  Microsoft Corporation;  Saarland University
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "fstext/fstext-lib.h"
+#include "lat/kaldi-lattice.h"
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+    using fst::SymbolTable;
+    using fst::VectorFst;
+    using fst::StdArc;
+
+    const char *usage =
+        "Composes lattices (in transducer form, as type Lattice).  Depending\n"
+        "on the command-line arguments, either composes lattices with lattices,\n"
+        "or lattices with FSTs (rspecifiers are assumed to be lattices, and\n"
+        "rxfilenames are assumed to be FSTs, which have their weights interpreted\n"
+        "as \"graph weights\" when converted into the Lattice format.\n"
+        "\n"
+        "Usage: lattice-compose [options] lattice-rspecifier1 "
+        "(lattice-rspecifier2|fst-rxfilename2) lattice-wspecifier\n"
+        " e.g.: lattice-compose ark:1.lats ark:2.lats ark:composed.lats\n"
+        " or: lattice-compose ark:1.lats G.fst ark:composed.lats\n";
+    
+    ParseOptions po(usage);
+
+    int32 num_states_cache = 50000;
+    int32 phi_label = fst::kNoLabel; // == -1
+    po.Register("phi-label", &phi_label, "If >0, the label on backoff arcs of the LM");
[…]

### 2015-09-03 src/latbin/lattice-minimize.cc

@@ -1,0 +1,110 @@
+// latbin/lattice-minimize.cc
+
+// Copyright 2013  Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "fstext/fstext-lib.h"
+#include "lat/kaldi-lattice.h"
+#include "lat/minimize-lattice.h"
+#include "lat/push-lattice.h"
+
+
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+    using fst::SymbolTable;
+    using fst::VectorFst;
+    using fst::StdArc;
+
+    const char *usage =
+      "Minimize lattices, in CompactLattice format.  Should be applied to\n"
+      "determinized lattices (e.g. produced with --determinize-lattice=true)\n"
+      "Note: by default this program\n"
+      "pushes the strings and weights prior to minimization."
+      "Usage: lattice-minimize [options] lattice-rspecifier lattice-wspecifier\n"
+        " e.g.: lattice-minimize ark:1.lats ark:2.lats\n";
+
+    ParseOptions po(usage);
+
+    bool push_strings = true;
+    bool push_weights = true;
+
+    po.Register("push-strings", &push_strings, "If true, push the strings in the "
+                "lattice to the start.");
+    po.Register("push-weights", &push_weights, "If true, push the weights in the "
+                "lattice to the start.");
+    
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 2) {
+      po.PrintUsage();
[…]

### 2015-09-03 src/latbin/lattice-to-phone-lattice.cc

@@ -1,0 +1,100 @@
+// latbin/lattice-to-phone-lattice.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "fstext/fstext-lib.h"
+#include "lat/kaldi-lattice.h"
+#include "lat/lattice-functions.h"
+#include "hmm/transition-model.h"
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+    using fst::SymbolTable;
+    using fst::VectorFst;
+    using fst::StdArc;
+
+    const char *usage =
+        "Convert the words or transition-ids into phones, which are worked out\n"
+        "from the transition-ids.  If --replace-words=true (true by default),\n"
+        "replaces the words with phones, otherwise replaces the transition-ids.\n"
+        "If --replace-words=false, it will preserve the alignment of transition-ids/phones\n"
+        "to words, so that if you do \n"
+        "lattice-align-words | lattice-to-phone-lattice --replace-words=false,\n"
+        "you can get the phones corresponding to each word in the lattice.\n"
+        "\n"
+        "Usage: lattice-to-phone-lattice [options] model lattice-rspecifier lattice-wspecifier\n"
+        " e.g.: lattice-to-phone-lattice 1.mdl ark:1.lats ark:phones.lats\n"
+        "See also: lattice-align-words, lattice-align-phones\n";
+      
+    ParseOptions po(usage);
+    bool replace_words = true;
[…]

### 2015-09-03 src/matrix/kaldi-blas.h

@@ -1,0 +1,132 @@
+// matrix/kaldi-blas.h
+
+// Copyright 2009-2011  Ondrej Glembek;  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+
+//  http://www.apache.org/licenses/LICENSE-2.0
+
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+#ifndef KALDI_MATRIX_KALDI_BLAS_H_
+#define KALDI_MATRIX_KALDI_BLAS_H_
+
+// This file handles the #includes for BLAS, LAPACK and so on.
+// It manipulates the declarations into a common format that kaldi can handle.
+// However, the kaldi code will check whether HAVE_ATLAS is defined as that
+// code is called a bit differently from CLAPACK that comes from other sources.
+
+// There are three alternatives:
+//   (i) you have ATLAS, which includes the ATLAS implementation of CBLAS
+//   plus a subset of CLAPACK (but with clapack_ in the function declarations).
+//   In this case, define HAVE_ATLAS and make sure the relevant directories are
+//   in the include path.
+
+//   (ii) you have CBLAS (some implementation thereof) plus CLAPACK.
+//   In this case, define HAVE_CLAPACK.
+//   [Since CLAPACK depends on BLAS, the presence of BLAS is implicit].
+
+//  (iii) you have the MKL library, which includes CLAPACK and CBLAS.
+
+// Note that if we are using ATLAS, no Svd implementation is supplied,
+// so we define HAVE_Svd to be zero and this directs our implementation to
+// supply its own "by hand" implementation which is based on TNT code.
+
+
+
+
[…]

### 2015-09-03 src/nnet2/mixup-nnet.cc

@@ -1,0 +1,222 @@
+// nnet2/mixup-nnet.cc
+
+// Copyright 2012   Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "nnet2/mixup-nnet.h"
+#include "gmm/model-common.h" // for GetSplitTargets()
+#include <numeric> // for std::accumulate
+
+namespace kaldi {
+namespace nnet2 {
+
+
+/** This function makes sure the neural net ends with a
+    SumGroupComponent.  If it doesn't, it adds one
+    (with a single mixture/matrix corresponding to each
+    output element.)  [Before doing so, it makes sure
+    that the last layer is a SoftmaxLayer, which is what
+    we expect.  You can remove this check if there is some
+    use-case that makes sense where the type of the previous
+    layer is different.
+ */
+static void GiveNnetCorrectTopology(Nnet *nnet,
+                                    AffineComponent **affine_component,
+                                    SoftmaxComponent **softmax_component,
+                                    SumGroupComponent **sum_group_component) {
+  int32 nc = nnet->NumComponents();
+  KALDI_ASSERT(nc > 0);
+  Component* component = &(nnet->GetComponent(nc - 1));
+  if ((*sum_group_component =
+       dynamic_cast<SumGroupComponent*>(component)) == NULL) {
+    KALDI_LOG << "Adding SumGroupComponent to neural net.";
+    int32 dim = component->OutputDim();
+    // Give it the same learning rate as the first updatable layer we have.
+    std::vector<int32> sizes(dim, 1); // a vector of all ones, of dimension "dim".
+  
+    *sum_group_component = new SumGroupComponent();
[…]

### 2015-09-03 src/nnet2/nnet-component-test.cc

@@ -1,0 +1,907 @@
+// nnet2/nnet-component-test.cc
+
+// Copyright 2012-2014  Johns Hopkins University (author:  Daniel Povey)
+//                2015  Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "nnet2/nnet-component.h"
+#include "util/common-utils.h"
+
+namespace kaldi {
+namespace nnet2 {
+
+
+void UnitTestGenericComponentInternal(const Component &component,
+                                      const ChunkInfo in_info,
+                                      const ChunkInfo out_info)  { 
+  
+  CuMatrix<BaseFloat> input(in_info.NumRows(), in_info.NumCols()),
+      output(1, out_info.NumRows() * out_info.NumCols());
+  input.SetRandn();
+  CuVector<BaseFloat> objf_vec(out_info.NumCols()); // objective function is linear function of output.
+  objf_vec.SetRandn(); // set to Gaussian noise.
+
+  int32 rand_seed = Rand();
+
+  RandomComponent *rand_component =
+      const_cast<RandomComponent*>(dynamic_cast<const RandomComponent*>(&component));
+  if (rand_component != NULL) {
+    srand(rand_seed);
+    rand_component->ResetGenerator();
+  }
+  component.Propagate(in_info, out_info, input, &output);
+  {
+    bool binary = (Rand() % 2 == 0);
+    Output ko("tmpf", binary);
+    component.Write(ko.Stream(), binary);
+  }
+  Component *component_copy;
+  {
+    bool binary_in;
+    Input ki("tmpf", &binary_in);
[…]

### 2015-09-03 src/nnet2/nnet-functions.cc

@@ -1,0 +1,78 @@
+// nnet2/nnet-functions.cc
+
+// Copyright 2011-2012  Karel Vesely
+//                      Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "nnet2/nnet-nnet.h"
+#include "util/stl-utils.h"
+
+namespace kaldi {
+namespace nnet2 {
+
+int32 IndexOfSoftmaxLayer(const Nnet &nnet) {
+  int32 index = -1, nc = nnet.NumComponents();
+  for (int32 c = 0; c < nc; c++) {
+    const Component *component = &(nnet.GetComponent(c));
+    if (dynamic_cast<const SoftmaxComponent*>(component) != NULL) {
+      if (index != -1) return -1; // >1 softmax components.
+      else index = c;
+    }
+  }
+  return index;
+}
+
+void InsertComponents(const Nnet &src_nnet,
+                      int32 c_to_insert, // component-index before which to insert.
+                      Nnet *dest_nnet) {
+  KALDI_ASSERT(c_to_insert >= 0 && c_to_insert <= dest_nnet->NumComponents());
+  int32 c_tot = dest_nnet->NumComponents() + src_nnet.NumComponents();
+  std::vector<Component*> components(c_tot);
+  for (int32 c = 0; c < c_to_insert; c++)
+    components[c] = dest_nnet->GetComponent(c).Copy();
+  for (int32 c = 0; c < src_nnet.NumComponents(); c++)
+    components[c + c_to_insert] = src_nnet.GetComponent(c).Copy();
+  for (int32 c = c_to_insert; c < dest_nnet->NumComponents(); c++)
[…]

### 2015-09-03 src/nnet2/shrink-nnet.h

@@ -1,0 +1,59 @@
+// nnet2/shrink-nnet.h
+
+// Copyright 2012  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#ifndef KALDI_NNET2_SHRINK_NNET_H_
+#define KALDI_NNET2_SHRINK_NNET_H_
+
+#include "nnet2/nnet-update.h"
+#include "nnet2/nnet-compute.h"
+#include "itf/options-itf.h"
+
+namespace kaldi {
+namespace nnet2 {
+
+/** Configuration class that controls neural net "shrinkage" which is actually a
+    scaling on the parameters of each of the updatable layers.
+ */
+struct NnetShrinkConfig {
+  int32 num_bfgs_iters; // The dimension is small (e.g. 3 to 5) so we do
+  // BFGS.  We actually implement this as L-BFGS but setting the number of
+  // vectors to be the same as the dimension of the space.  Note: this
+  // num-iters is in reality the number of function evaluations.
+
+  BaseFloat initial_step;
+  
+  NnetShrinkConfig(): num_bfgs_iters(10), initial_step(0.1) { }
+  void Register(OptionsItf *opts) {
+    opts->Register("num-bfgs-iters", &num_bfgs_iters, "Number of iterations of "
+                   "BFGS to use when optimizing shrinkage parameters");
+    opts->Register("initial-step", &initial_step, "Parameter in the optimization, "
+                   "used to set the initial step length");
+  }  
+};
+
+void ShrinkNnet(const NnetShrinkConfig &shrink_config,
+                const std::vector<NnetExample> &validation_set,
[…]

### 2015-09-03 src/nnet2bin/nnet-logprob2-parallel.cc

@@ -1,0 +1,164 @@
+// nnet2bin/nnet-logprob2-parallel.cc
+
+// Copyright 2012  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "hmm/transition-model.h"
+#include "nnet2/nnet-update-parallel.h"
+#include "nnet2/am-nnet.h"
+#include "nnet2/nnet-compute.h"
+#include "thread/kaldi-task-sequence.h"
+
+namespace kaldi {
+namespace nnet2 {
+
+struct NnetLogprobTask {
+  NnetLogprobTask(const AmNnet &am_nnet,
+                  const CuVector<BaseFloat> &inv_priors,
+                  const std::string &key,
+                  const Matrix<BaseFloat> &feats,
+                  BaseFloatMatrixWriter *prob_writer_nodiv,
+                  BaseFloatMatrixWriter *logprob_writer_divided):
+      am_nnet_(am_nnet), inv_priors_(inv_priors), key_(key), feats_(feats),
+      prob_writer_nodiv_(prob_writer_nodiv),
+      logprob_writer_divided_(logprob_writer_divided) { }
+  void operator () () {
+    log_probs_.Resize(feats_.NumRows(), am_nnet_.NumPdfs());
+    bool pad_input = true;
+    NnetComputation(am_nnet_.GetNnet(), feats_, pad_input,
+                    &log_probs_);
+  }
+
+  ~NnetLogprobTask() { // Produces output.  Run sequentially.
+    // at this point they are probabilities, not log-probs, without prior division.
[…]

### 2015-09-03 src/nnet2bin/nnet-show-progress.cc

@@ -1,0 +1,164 @@
+// nnet2bin/nnet-show-progress.cc
+
+// Copyright 2012-2013  Johns Hopkins University (author:  Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "hmm/transition-model.h"
+#include "nnet2/train-nnet.h"
+#include "nnet2/am-nnet.h"
+
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    using namespace kaldi::nnet2;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+
+    const char *usage =
+        "Given an old and a new model and some training examples (possibly held-out),\n"
+        "show the average objective function given the mean of the two models,\n"
+        "and the breakdown by component of why this happened (computed from\n"
+        "derivative information).  Also shows parameter differences per layer.\n"
+        "If training examples not provided, only shows parameter differences per\n"
+        "layer.\n"
+        "\n"
+        "Usage:  nnet-show-progress [options] <old-model-in> <new-model-in> [<training-examples-in>]\n"
+        "e.g.: nnet-show-progress 1.nnet 2.nnet ark:valid.egs\n";
+    
+    ParseOptions po(usage);
+
+    int32 num_segments = 1;
+    int32 batch_size = 1024;
+    std::string use_gpu = "optional";
+    
+    po.Register("num-segments", &num_segments,
[…]

### 2015-09-03 src/nnet2bin/nnet-train-ensemble.cc

@@ -1,0 +1,145 @@
+// nnet2bin/nnet-train-simple.cc
+
+// Copyright 2012  Johns Hopkins University (author: Daniel Povey)
+//           2014  Xiaohui Zhang
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "hmm/transition-model.h"
+#include "nnet2/train-nnet-ensemble.h"
+#include "nnet2/am-nnet.h"
+
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    using namespace kaldi::nnet2;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+
+    const char *usage =
+        "Train an ensemble of neural networks with backprop and stochastic\n"
+        "gradient descent using minibatches.  Modified version of nnet-train-simple.\n"
+        "Implements parallel gradient descent with a term that encourages the nnets to\n"
+        "produce similar outputs.\n"
+        "\n"
+        "Usage:  nnet-train-ensemble [options] <model-in-1> <model-in-2> ... <model-in-n>"
+        "  <training-examples-in> <model-out-1> <model-out-2> ... <model-out-n> \n"
+        "\n"
+        "e.g.:\n"
+        " nnet-train-ensemble 1.1.nnet 2.1.nnet ark:egs.ark 2.1.nnet 2.2.nnet \n";
+    
+    bool binary_write = true;
+    bool zero_stats = true;
+    int32 srand_seed = 0;
+    std::string use_gpu = "yes";
+    NnetEnsembleTrainerConfig train_config;
+    
+    ParseOptions po(usage);
[…]

### 2015-09-03 src/nnet3/nnet-parse.cc

@@ -1,0 +1,484 @@
+// Copyright      2015  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include <iterator>
+#include <sstream>
+#include <iomanip>
+#include "nnet3/nnet-parse.h"
+
+namespace kaldi {
+namespace nnet3 {
+
+bool ConfigLine::ParseLine(const std::string &line) {
+  if (line.size() == 0) return false;   // Empty line
+
+  // Line ends or begins with space -> remove it and recurse.
+  if (isspace(line[line.size()-1]) || isspace(line[0])) {
+    size_t initial_pos = line.find_first_not_of(" \t\r\n"),
+        final_pos = line.find_last_not_of(" \t\r\n");
+    if (initial_pos == std::string::npos || final_pos <= initial_pos)
+      return false;
+    std::string processed_line(line, initial_pos, final_pos - initial_pos + 1);
+    return ParseLine(processed_line);
+  }
+
+  size_t pos = 0;
+  size_t found_eq = line.find_first_of("=", pos + 1);
+  if (found_eq == std::string::npos) return false; // Could not find '='
+
+
+  while (found_eq < line.size()) {
+    std::string key(line, pos, found_eq - pos);
+    if (!IsValidName(key)) return false;
+    if (found_eq == std::string::npos) return false; // Could not find '='
+    if (found_eq == line.size() - 1 || line[found_eq+1] == ' ' || line[found_eq+1] == '\t') {
+      // Empty value for key
+      data_.insert(std::make_pair(key, std::make_pair("", false)));
+      pos = line.find_first_not_of(" \t", found_eq + 1);
[…]

### 2015-09-03 src/nnet3/nnet-training.cc

@@ -1,0 +1,214 @@
+// nnet3/nnet-training.cc
+
+// Copyright      2015    Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "nnet3/nnet-training.h"
+#include "nnet3/nnet-utils.h"
+
+namespace kaldi {
+namespace nnet3 {
+
+NnetTrainer::NnetTrainer(const NnetTrainerOptions &config,
+                         Nnet *nnet):
+    config_(config),
+    nnet_(nnet),
+    compiler_(*nnet, config_.optimize_config),
+    num_minibatches_processed_(0) {
+  if (config.zero_component_stats)
+    ZeroComponentStats(nnet);
+}
+
+
+void NnetTrainer::Train(const NnetExample &eg) {
+  bool need_model_derivative = true;
+  ComputationRequest request;
+  GetComputationRequest(*nnet_, eg, need_model_derivative,
+                        config_.store_component_stats,
+                        &request);
+  const NnetComputation *computation = compiler_.Compile(request);
+
+  const Nnet *const_nnet = (config_.update_per_minibatch ?
+                            static_cast<const Nnet*>(nnet_->Copy()) :
+                            nnet_);
+  NnetComputer computer(config_.compute_config, *computation,
+                        *const_nnet, nnet_);
+  // give the inputs to the computer object.
+  computer.AcceptInputs(*nnet_, eg);
+  computer.Forward();
+
+  this->ProcessOutputs(eg, &computer);
+  computer.Backward();
+
+  if (config_.update_per_minibatch)
+    delete const_nnet;
+}
+
[…]

### 2015-09-03 src/nnet3bin/nnet3-merge-egs.cc

@@ -1,0 +1,129 @@
+// nnet3bin/nnet3-merge-egs.cc
+
+// Copyright 2012-2015  Johns Hopkins University (author:  Daniel Povey)
+//                2014  Vimal Manohar
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "hmm/transition-model.h"
+#include "nnet3/nnet-example.h"
+#include "nnet3/nnet-example-utils.h"
+
+namespace kaldi {
+namespace nnet3 {
+// returns the number of indexes/frames in the NnetIo named "output" in the eg,
+// or crashes if it is not there.
+int32 NumOutputIndexes(const NnetExample &eg) {
+  for (size_t i = 0; i < eg.io.size(); i++)
+    if (eg.io[i].name == "output")
+      return eg.io[i].indexes.size();
+  KALDI_ERR << "No output named 'output' in the eg.";
+  return 0;  // Suppress compiler warning.
+}
+
+}
+}
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    using namespace kaldi::nnet3;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+
+    const char *usage =
+        "This copies nnet training examples from input to output, but while doing so it\n"
+        "merges many NnetExample objects into one, forming a minibatch consisting of a\n"
+        "single NnetExample.  Note: if --measure-output-frames=true, which it is by default,\n"
+        "the --minibatch-size option will be interpreted as a target number of output frames;\n"
+        "otherwise as a number of input examples to combine.  This makes a difference\n"
[…]

### 2015-09-03 src/nnet3bin/nnet3-shuffle-egs.cc

@@ -1,0 +1,114 @@
+// nnet3bin/nnet3-shuffle-egs.cc
+
+// Copyright 2012-2015  Johns Hopkins University (author:  Daniel Povey)
+//                2014  Vimal Manohar
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "hmm/transition-model.h"
+#include "nnet3/nnet-example.h"
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    using namespace kaldi::nnet3;
+    typedef kaldi::int32 int32;
+    typedef kaldi::int64 int64;
+
+    const char *usage =
+        "Copy examples (typically single frames or small groups of frames) for\n"
+        "neural network training, from the input to output, but randomly shuffle the order.\n"
+        "This program will keep all of the examples in memory at once, unless you\n"
+        "use the --buffer-size option\n"
+        "\n"
+        "Usage:  nnet3-shuffle-egs [options] <egs-rspecifier> <egs-wspecifier>\n"
+        "\n"
+        "nnet3-shuffle-egs --srand=1 ark:train.egs ark:shuffled.egs\n";
+    
+    int32 srand_seed = 0;
+    int32 buffer_size = 0;
+    ParseOptions po(usage);
+    po.Register("srand", &srand_seed, "Seed for random number generator ");
+    po.Register("buffer-size", &buffer_size, "If >0, size of a buffer we use "
+                "to do limited-memory partial randomization.  Otherwise, do "
+                "full randomization.");
+    
+    po.Read(argc, argv);
+
[…]

### 2015-09-03 src/nnetbin/feat-to-post.cc

@@ -1,0 +1,79 @@
+// nnetbin/feat-to-post.cc
+
+// Copyright 2014       Brno University of Technology (Author: Karel Vesely)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "hmm/posterior.h"
+
+/** @brief Convert features into posterior format, is used to specify NN training targets. */
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  typedef kaldi::int32 int32;
+  try {
+    const char *usage =
+        "Convert features into posterior format, which is used for NN training targets\n"
+        "in Karel's nnet1 tools.\n"
+        "(speed is not an issue for reasonably low NN-output dimensions)\n"
+        "Usage:  feat-to-post [options] feat-rspecifier posteriors-wspecifier\n"
+        "e.g.:\n"
+        " feat-to-post scp:feats.scp ark:feats.post\n";
+
+    ParseOptions po(usage);
+
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 2) {
+      po.PrintUsage();
+      exit(1);
+    }
+
+    std::string feats_rspecifier = po.GetArg(1);
+    std::string posteriors_wspecifier = po.GetArg(2);
+
+    int32 num_done = 0;
+    SequentialBaseFloatMatrixReader feats_reader(feats_rspecifier);
+    PosteriorWriter posterior_writer(posteriors_wspecifier);
+
+    for (; !feats_reader.Done(); feats_reader.Next()) {
+      num_done++;
[…]

### 2015-09-03 src/nnetbin/nnet-kl-hmm-acc.cc

@@ -1,0 +1,110 @@
+// nnetbin/nnet-kl-hmm-acc.cc
+
+// Copyright 2013  Idiap Research Institute (Author: David Imseng)
+//                 Karlsruhe Institute of Technology (Author: Ngoc Thang Vu)
+//                 Brno University of Technology (Author: Karel Vesely)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "nnet/nnet-nnet.h"
+#include "nnet/nnet-kl-hmm.h"
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+
+
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  using namespace kaldi::nnet1;
+  typedef kaldi::int32 int32;
+  try {
+    const char *usage =
+        "Collect the statistics for the Kl-HMM trainign.\n"
+        "Usage:  nnet-kl-hmm-acc [options] <feature-rspecifier> <alignments-rspecifier> <kl-hmm-accumulator>\n"
+        "e.g.: \n"
+        " nnet-kl-hmm-acc scp:train.scp ark:train.ali kl-hmm.acc\n";
+
+    ParseOptions po(usage);
+
+    bool binary = false;
+    int32 n_kl_states = 0;
+    po.Register("binary", &binary, "Write output in binary mode");
+    po.Register("nkl-states", &n_kl_states, "Number of states in Kl-HMM");
+
+    
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 3) {
+      po.PrintUsage();
+      exit(1);
+    }
+
+    std::string feature_rspecifier = po.GetArg(1),
[…]

### 2015-09-03 src/nnetbin/nnet-kl-hmm-sum-accs.cc

@@ -1,0 +1,76 @@
+// nnetbin/nnet-kl-gmm-sum-accs.cc
+
+// Copyright 2013  Idiap Research Institute (Author: David Imseng)
+//                 Karlsruhe Institute of Technology (Author: Ngoc Thang Vu)
+//                 Brno University of Technology (Author: Karel Vesely)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "nnet/nnet-nnet.h"
+#include "nnet/nnet-kl-hmm.h"
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+
+
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  using namespace kaldi::nnet1;
+
+  try {
+    typedef int32 int32;
+    const char *usage =
+        "Sum multiple accumulated stats files for KL-HMM training.\n"
+        "Usage: nnet-kl-hmm-sum-accs [options] nnet-component stats-in1 stats-in2 ...\n";
+
+    bool binary = true;
+    int32 n_kl_states = 0;
+    int32 n_posterior_dim = 0;
+    ParseOptions po(usage);
+    po.Register("binary", &binary, "Write output in binary mode");
+    po.Register("nkl-states", &n_kl_states, "Number of states in Kl-HMM");
+    po.Register("posterior-dim", &n_posterior_dim, "Dimensionality of posterior features");
+    po.Read(argc, argv);
+
+    if (po.NumArgs() < 2) {
+      po.PrintUsage();
+      exit(1);
+    }
+
+    std::string nnet_component_filename = po.GetArg(1);
[…]

### 2015-09-03 src/online2bin/ivector-extract-online2.cc

@@ -1,0 +1,159 @@
+// online2bin/ivector-extract-online2.cc
+
+// Copyright 2014  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "gmm/am-diag-gmm.h"
+#include "online2/online-ivector-feature.h"
+#include "thread/kaldi-task-sequence.h"
+
+int main(int argc, char *argv[]) {
+  using namespace kaldi;
+  typedef kaldi::int32 int32;
+  typedef kaldi::int64 int64;
+  try {
+    const char *usage =
+        "Extract iVectors for utterances every --ivector-period frames, using a trained\n"
+        "iVector extractor and features and Gaussian-level posteriors.  Similar to\n"
+        "ivector-extract-online but uses the actual online decoder code to do it,\n"
+        "and does everything in-memory instead of using multiple processes.\n"
+        "Note: the value of the --use-most-recent-ivector config variable is ignored\n"
+        "it's set to false.  The <spk2utt-rspecifier> is mandatory, to simplify the code;\n"
+        "if you want to do it separately per utterance, just make it of the form\n"
+        "<utterance-id> <utterance-id>.\n"
+        "The iVectors are output as an archive of matrices, indexed by utterance-id;\n"
+        "each row corresponds to an iVector.  If --repeat=true, outputs the whole matrix\n"
+        "of iVectors, not just every (ivector-period)'th frame\n"
+        "The input features are the raw, non-cepstral-mean-normalized features, e.g. MFCC.\n"
+        "\n"
[…]

### 2015-09-03 src/sgmm2/estimate-am-sgmm2-ebw.h

@@ -1,0 +1,242 @@
+// sgmm2/estimate-am-sgmm2-ebw.h
+
+// Copyright 2012  Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#ifndef KALDI_SGMM2_ESTIMATE_AM_SGMM2_EBW_H_
+#define KALDI_SGMM2_ESTIMATE_AM_SGMM2_EBW_H_ 1
+
+#include <string>
+#include <vector>
+
+#include "gmm/model-common.h"
+#include "itf/options-itf.h"
+#include "sgmm2/estimate-am-sgmm2.h"
+
+namespace kaldi {
+
+/**
+   This header implements a form of Extended Baum-Welch training for SGMMs.
+   If you are confused by this comment, see Dan Povey's thesis for an explanation of
+   Extended Baum-Welch.
+   A note on the EBW (Extended Baum-Welch) updates for the SGMMs... In general there is
+   a parameter-specific value D that is similar to the D in EBW for GMMs.  The value of
+   D is generally set to:
+     E * (denominator-count for that parameter)   +   tau-value for that parameter
+   where the tau-values are user-specified parameters that are specific to the type of
+   the parameter (e.g. phonetic vector, subspace projection, etc.).  Things are a bit
+   more complex for this update than for GMMs, because it's not just a question of picking
+   a tau-value for smoothing: there is sometimes a scatter-matrix of some kind (e.g.
+   an outer product of vectors, or something) that defines a quadratic objective function
+   that we'll add as smoothing.  We have to pick where to get this scatter-matrix from.
+   We feel that it's appropriate for the "E" part of the D to get its scatter-matrix from
[…]

### 2015-09-03 src/sgmm2bin/sgmm2-gselect.cc

@@ -1,0 +1,110 @@
+// sgmm2bin/sgmm2-gselect.cc
+
+// Copyright 2009-2012   Saarland University  Microsoft Corporation
+//                       Johns Hopkins University (author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "sgmm2/am-sgmm2.h"
+#include "hmm/transition-model.h"
+
+int main(int argc, char *argv[]) {
+  try {
+    using namespace kaldi;
+    const char *usage =
+        "Precompute Gaussian indices for SGMM training "
+        "Usage: sgmm2-gselect [options] <model-in> <feature-rspecifier> <gselect-wspecifier>\n"
+        "e.g.: sgmm2-gselect 1.sgmm \"ark:feature-command |\" ark:1.gs\n"
+        "Note: you can do the same thing by combining the programs sgmm2-write-ubm, fgmm-global-to-gmm,\n"
+        "gmm-gselect and fgmm-gselect\n";
+
+    ParseOptions po(usage);
+    kaldi::Sgmm2GselectConfig sgmm_opts;
+    std::string preselect_rspecifier;
+    std::string likelihood_wspecifier;
+    po.Register("write-likes", &likelihood_wspecifier, "Wspecifier for likelihoods per "
+                "utterance");
+    sgmm_opts.Register(&po);
+    po.Read(argc, argv);
+
+    if (po.NumArgs() != 3) {
+      po.PrintUsage();
+      exit(1);
+    }
+
+    std::string model_filename = po.GetArg(1),
+        feature_rspecifier = po.GetArg(2),
[…]

### 2015-09-03 src/sgmmbin/sgmm-latgen-faster.cc

@@ -1,0 +1,271 @@
+// sgmmbin/sgmm-latgen-faster.cc
+
+// Copyright 2009-2011  Saarland University;  Microsoft Corporation;
+//                      Johns Hopkins University (author: Daniel Povey)
+//                2014  Guoguo Chen
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include <string>
+using std::string;
+
+#include "base/kaldi-common.h"
+#include "util/common-utils.h"
+#include "sgmm/am-sgmm.h"
+#include "hmm/transition-model.h"
+#include "fstext/fstext-lib.h"
+#include "decoder/decoder-wrappers.h"
+#include "sgmm/decodable-am-sgmm.h"
+#include "base/timer.h"
+
+namespace kaldi {
+
+// the reference arguments at the beginning are not const as the style guide
+// requires, but are best viewed as inputs.
+bool ProcessUtterance(LatticeFasterDecoder &decoder,
+                      const AmSgmm &am_sgmm,
+                      const TransitionModel &trans_model,
+                      const SgmmGselectConfig &sgmm_opts,
+                      double log_prune,
+                      double acoustic_scale,
+                      const Matrix<BaseFloat> &features,
+                      RandomAccessInt32VectorVectorReader &gselect_reader,
+                      RandomAccessBaseFloatVectorReaderMapped &spkvecs_reader,
+                      const fst::SymbolTable *word_syms,
+                      const std::string &utt,
+                      bool determinize,
+                      bool allow_partial,
+                      Int32VectorWriter *alignments_writer,
+                      Int32VectorWriter *words_writer,
+                      CompactLatticeWriter *compact_lattice_writer,
[…]

### 2015-09-03 src/thread/kaldi-task-sequence.h

@@ -1,0 +1,197 @@
+// thread/kaldi-task-sequence.h
+
+// Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#ifndef KALDI_THREAD_KALDI_TASK_SEQUENCE_H_
+#define KALDI_THREAD_KALDI_TASK_SEQUENCE_H_ 1
+
+#include <pthread.h>
+#include "thread/kaldi-thread.h"
+#include "itf/options-itf.h"
+#include "thread/kaldi-semaphore.h"
+
+
+namespace kaldi {
+
+/**
+   In kaldi-thread.h, the class MultiThreader, and the function
+   RunMultiThreaded, provided a mechanism to run a specified number of jobs
+   simultaneously, in parellel, and wait for them all to finish.  This file
+   addresses a different problem, typically encountered in Kaldi command-line
+   programs that process a sequence of items.  The problem is where
+   items to be processed are coming in, all of different sizes (e.g. utterances
+   with different numbers of frames) and we would like them to be run in
+   parallel somehow, using multiple threads; but they must still make good use
+   of the number of threads available; and they must be output in the same order
+   as they came in.
+
+   Here, we will still accept objects of some class C with an operator () that
+   takes no arguments.  C may also have a constructor and a destructor that do
+   something (typically the constructor just sets variables, and the destructor
+   does some kind of output).  We have a templated class TaskSequencer<C> which
+   is responsible for running the jobs in parallel.  It has a function Run()
+   that will accept a new object of class C; this will block until a thread is
+   free, at which time it will spawn a thread that starts running the operator
+   () of the class.  When classes are finished running, the objects will be
[…]

### 2015-09-03 src/thread/kaldi-thread-test.cc

@@ -1,0 +1,110 @@
+// thread/kaldi-thread-test.cc
+
+// Copyright 2012  Johns Hopkins University (Author: Daniel Povey)
+//                 Frantisek Skala
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include <algorithm>
+#include "base/kaldi-common.h"
+#include "thread/kaldi-thread.h"
+#include "thread/kaldi-mutex.h"
+
+namespace kaldi {
+
+class MyThreadClass {  // Sums up integers from 0 to max_to_count-1.
+ public:
+  MyThreadClass(int32 max_to_count, int32 *i): max_to_count_(max_to_count),
+                                               iptr_(i),
+                                               private_counter_(0) { }
+  // Use default copy constructor and assignment operators.
+  void operator() () {
+    int32 block_size = (max_to_count_+ (num_threads_-1) ) / num_threads_;
+    int32 start = block_size * thread_id_,
+        end = std::min(max_to_count_, start + block_size);
+    for (int32 j = start; j < end; j++)
+      private_counter_ += j;
+  }
+  ~MyThreadClass() {
+    *iptr_ += private_counter_;
+  }
+
+  static void *run(void *c_in) {
+    MyThreadClass *c = static_cast<MyThreadClass*>(c_in);
+    (*c)();  // call operator () on it.
+    return NULL;
+  }
+
+ public:
+  int32 thread_id_;  // 0 <= thread_number < num_threads
+  int32 num_threads_;
+
+ private:
[…]

### 2015-09-03 src/tree/build-tree-questions.cc

@@ -1,0 +1,127 @@
+// tree/build-tree-questions.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+//
+//  http://www.apache.org/licenses/LICENSE-2.0
+//
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "base/kaldi-math.h"
+#include "util/stl-utils.h"
+#include "tree/build-tree-questions.h"
+#include "tree/build-tree-utils.h"
+
+namespace kaldi {
+
+void QuestionsForKey::Write(std::ostream &os, bool binary) const {
+  WriteToken(os, binary, "<QuestionsForKey>");
+  int32 size = initial_questions.size();
+  WriteBasicType(os, binary, size);
+  for (int32 i = 0; i < size; i++)
+    WriteIntegerVector(os, binary, initial_questions[i]);
+  refine_opts.Write(os, binary);
+  WriteToken(os, binary, "</QuestionsForKey>");
+}
+
+void QuestionsForKey::Read(std::istream &is, bool binary) {
+  int32 size;
+  ExpectToken(is, binary, "<QuestionsForKey>");
+  ReadBasicType(is, binary, &size);
+  initial_questions.resize(size);
+  for (int32 i = 0; i < size; i++)
+    ReadIntegerVector(is, binary, &(initial_questions[i]));
+  refine_opts.Read(is, binary);
+  ExpectToken(is, binary, "</QuestionsForKey>");
+}
+
+void Questions::Write(std::ostream &os, bool binary) const {
+  WriteToken(os, binary, "<Questions>");
+
+  std::vector<EventKeyType> keys_with_options;
+  this->GetKeysWithQuestions(&keys_with_options);
[…]

### 2015-09-03 src/util/kaldi-table.cc

@@ -1,0 +1,300 @@
+// util/kaldi-table.cc
+
+// Copyright 2009-2011  Microsoft Corporation
+
+// See ../../COPYING for clarification regarding multiple authors
+//
+// Licensed under the Apache License, Version 2.0 (the "License");
+// you may not use this file except in compliance with the License.
+// You may obtain a copy of the License at
+
+//  http://www.apache.org/licenses/LICENSE-2.0
+
+// THIS CODE IS PROVIDED *AS IS* BASIS, WITHOUT WARRANTIES OR CONDITIONS OF ANY
+// KIND, EITHER EXPRESS OR IMPLIED, INCLUDING WITHOUT LIMITATION ANY IMPLIED
+// WARRANTIES OR CONDITIONS OF TITLE, FITNESS FOR A PARTICULAR PURPOSE,
+// MERCHANTABLITY OR NON-INFRINGEMENT.
+// See the Apache 2 License for the specific language governing permissions and
+// limitations under the License.
+
+#include "util/kaldi-table.h"
+#include "util/text-utils.h"
+
+namespace kaldi {
+
+
+bool ReadScriptFile(const std::string &rxfilename,
+                    bool warn,
+                    std::vector<std::pair<std::string, std::string> > *script_out) {
+  bool is_binary;
+  Input input;
+
+  if (!input.Open(rxfilename, &is_binary)) {
+    if (warn) KALDI_WARN << "Error opening script file: " <<
+                 PrintableRxfilename(rxfilename);
+    return false;
+  }
+  if (is_binary) {
+    if (warn) KALDI_WARN << "Error: script file appears to be binary: " <<
+                 PrintableRxfilename(rxfilename);
+    return false;
+  }
+
+  bool ans = ReadScriptFile(input.Stream(), warn, script_out);
+  if (warn && !ans)
+    KALDI_WARN << "[script file was: " << PrintableRxfilename(rxfilename) << "]";
+  return ans;
+}
+
+bool ReadScriptFile(std::istream &is,
+                    bool warn,
+                    std::vector<std::pair<std::string, std::string> > *script_out) {
+  KALDI_ASSERT(script_out != NULL);
+  std::string line;
+  int line_number = 0;
+  while (getline(is, line)) {
+    line_number++;
+    const char *c = line.c_str();
+    if (*c == '\0') {
+      if (warn) KALDI_WARN << "Empty "<<line_number<<"'th line in script file";
+      return false;  // Empty line so invalid scp file format..
+    }
+
[…]

### 2015-09-03 tools/extras/travis_script.sh

@@ -1,0 +1,67 @@
+#!/bin/bash
+
+# We take into account dependency pointers optionally set in the environment.
+# Typical usage shown below; any one can be safely left unset.
+#   INCDIRS="~/xroot/usr/include"
+#   LIBDIRS="~/xroot/usr/lib /usr/lib/openblas-base"
+#   CXX=gcc++-4.9
+#   CFLAGS="-march=native -O2"
+#   LDFLAGS="-llapack"
+
+# Maximum make parallelism. Simply -j runs out of memory on Travis VM.
+MAXPAR=3
+
+# Directories with code that can be tested with Travis (space-separated)
+TESTABLE_DIRS="src/"
+
+# Run verbose (run and echo) and exit if failed.
+runvx() {
+  echo "\$ $@"
+  "$@" || exit 1
+}
+
+# $(addsw -L foo bar) => "-Lfoo -Lbar".
+addsw() {
+  local v=() s=$1; shift;
+  for d; do v+=("$s$d"); done
+  echo ${v[@]};
+}
+
+# $(mtoken CXX gcc) => "CXX=gcc"; # $(mtoken CXX ) => "".
+mtoken() { echo ${2+$1=$2}; }
+
+# Print machine info and environment.
+runvx uname -a
+runvx env
+
+# Check for changes in src/ only; report success right away if none.
+# However, do run tests if TRAVIS_COMMIT_RANGE does not parse. This
+# most likely means the branch was reset by --force; re-run tests then.
+if git rev-parse "${TRAVIS_COMMIT_RANGE}" >/dev/null 2>&1 && \
+   ! git diff --name-only "${TRAVIS_COMMIT_RANGE}" -- ${TESTABLE_DIRS} | read REPLY
+then
+  echo; echo "No changes outside ${TESTABLE_DIRS} in the commit" \
+             "range ${TRAVIS_COMMIT_RANGE}; reporting success."
+  exit 0;
+fi
+
+# Prepare make command fragments.
+CF="$CFLAGS -g $(addsw -I $INCDIRS)"
+LDF="$LDFLAGS $(addsw -L $LIBDIRS)"
+CCC="$(mtoken CC $CXX) $(mtoken CXX $CXX)"
+
+runvx cd tools
+runvx make openfst $CCC CXXFLAGS="$CF" -j$MAXPAR
+cd ..
+runvx cd src
[…]

### 2017-05-16 tools/extras/irstlm/doc/LMAdaptation.tex
Digit Recognition and Yes/No Recognition

Language model adaptation can be  applied when little training data is given for the 
task at hand, but much more data from other less related sources is available.  {\tt tlm} supports two adaptation methods.

\subsection{Minimum Discriminative Information Adaptation}
MDI adaptation  is used  when domain related  data is very  little but
enough to  estimate a  unigram LM.  Basically,  the n-gram probs  of a
general  purpose (background)  LM are  scaled so  that they  match the
target unigram distribution.

\noindent	 
Relevant parameters:
\begin{itemize}
\item {\tt -ar=value}: the adaptation {\tt rate},  a real number ranging 
 from 0 (=no adaptation) to 1 (=strong adaptation).

\item {\tt -ad=file}: the  adaptation file,  either a text  or a
  unigram table.

\item {\tt -ao=y}: open vocabulary mode, which  must be set if the adaptation file
 might contain new words to be added to the basic dictionary.
\end{itemize}

\noindent
As an example, we apply MDI adaptation on the ``adapt'' file:
\begin{small}
\begin{verbatim}
$> tlm -tr=train.www -lm=wb -n=3 -te=test -dub=1000000 -ad=adapt -ar=0.8 -ao=yes
   n=49984 LP=326327.8053 PP=684.470312 OVVRate=0.04193341869
\end{verbatim}
\end{small}

\noindent
\paragraph{Warning:}  modified shift-beta  smoothing  cannot  be applied  in  open
vocabulary mode  ({\tt -ao=yes}).  If  this is the  case, you  should either
change  smoothing method  or simply  add  the adaptation  text to  the
background LM (use {\tt -aug} parameter  of {\tt ngt}). In
general, this solution should  provide better performance.
\begin{small}
\begin{verbatim}
$> ngt -i=train.www -aug=adapt -o=train-adapt.www -n=3 -b=yes
$> tlm -tr=train-adapt.www -lm=msb -n=3 -te=test -dub=1000000 -ad=adapt -ar=0.8
  n=49984 LP=312276.1746 PP=516.7311396 OVVRate=0.04193341869
\end{verbatim}
\end{small}

\subsection{Mixture Adaptation}

\noindent
Mixture adaptation  is useful  when you have  enough training  data to
estimate a  bigram or  trigram LM and  you also have  data collections
from other domains.

\noindent
Relevant parameters:
\begin{itemize}
\item {\tt-lm=mix} : specifies mixture smoothing method
\item {\tt -slmi=<filename>}: specifies filename with information about LMs to combine.
\end{itemize}

\noindent
In the example directory, the file {\tt sublmi} contains the following lines:
\begin{verbatim}
2
-slm=msb -str=adapt -sp=0
-slm=msb -str=train.www -sp=0
\end{verbatim}

\noindent
This means  that we use train a  mixture model on the  {\tt adapt} data set and
combine it  with the train data. For each data  set the desired
smoothing method is specified  (disregard the parameter {\tt -sp}). The file
used for adaptation is the one in FIRST position.

\begin{verbatim}
$> tlm -tr=train.www -lm=mix -slmi=sublm -n=3 -te=test -dub=1000000
  n=49984 LP=307199.3273 PP=466.8244383 OVVRate=0.04193341869
\end{verbatim}

\noindent
{\bf Warning}: for  computational reasons it  is expected that  the $n$-gram
table  specified by {\tt -tr}  contains AT  LEAST the  $n$-grams of  the last
table specified in the slmi file, i.e. {\tt train.www} in  the example.
Faster computations are achieved by putting the largest dataset as the
last sub-model in the list and the union of all data sets as training
file.

\noindent
It is  also IMPORTANT  that a  large {\tt -dub} value  is specified  so that
probabilities  of  sub-LMs  can  be  correctly  computed  in  case  of
out-of-vocabulary words.

### 2017-05-16 tools/extras/irstlm/doc/LMCompilation.tex
Digit Recognition and Yes/No Recognition

LMs in ARPA, iARPA, and qARPA format can be stored in a compact binary table through the command:

\begin{verbatim}
$> compile-lm train.lm train.blm
\end{verbatim}

\noindent
which generates the binary file {\tt train.blm} that can be quickly loaded in memory.  If the LM
is really very large, {\tt compile-lm} can avoid to create the binary LM directly in memory through the 
option {\tt -memmap=1}, which exploits the {\em Memory Mapping} mechanism in order to work as 
much as possible on disk rather than in RAM. \\

\begin{verbatim}
$> compile-lm train.lm train.blm --memmap=1
\end{verbatim}
\noindent
This option clearly pays a fee  in terms of speed, but  is often the only way to proceed. It is also recommended 
that the hard disk for the LM storage belongs to the computer on which the compilation is performed.

\noindent
Notice that most of the functionalities of {\tt compile-lm} (see below) apply to binary and quantized models.

\noindent
By default, the command uses the directory ``/tmp'' for storing
intermediate results.  For huge LMs, the temporary files can grow
dramatically causing a ``disk full'' system error.  It is possible to
explicitly set the directory used for temporary computation through the
parameter ``--tmpdir''.
\begin{verbatim}
$> compile-lm train.lm train.blm --tmpdir=<mytmpdir>
\end{verbatim}

\subsection{Inverted order of ngrams}
\label{sec:inverted-lm}
For a faster access, the ngrams can be stored in inverted order with the following two commands:
\begin{verbatim}
$> sort-lm.pl -inv -ilm train.lm -olm train.inv.lm
$> compile-lm train.inv.lm train.inv.blm --invert=yes
\end{verbatim}

\paragraph{Warning:} The following pipeline is no more allowed!!

\COMMENT{
or with the following pipeline:
}
\begin{verbatim}
$> cat train.lm | sort-lm.pl -inv | \
   compile-lm /dev/stdin train.inv.blm --invert=yes
\end{verbatim}

### 2017-05-16 tools/extras/irstlm/doc/LMInterpolation.tex
Digit Recognition and Yes/No Recognition

We provide a convenient tool to estimate mixtures of LMs that have been already 
created in one of the available formats.  The tool permits to estimate interpolation
weights through the EM algorithm, to compute the perplexity, and to query the interpolated
LM.

\noindent
Data used in those examples can be found in the directory {\tt example/interpolateLM/},
which represents the relative path for all the parameters of the referred commands.

\noindent
Interpolated LMs are defined by a configuration file in the following format:
\begin{verbatim}
3
0.3 lm-file1
0.3 lm-file2
0.4 lm-file3
\end{verbatim}

\noindent
The first number indicates the number of LMs to be interpolated, then each LM is specified
by its weight and its file (either in ARPA or binary format). Notice that you can interpolate
LMs with different orders\\

\noindent
Given an initial configuration file {\tt lmlist.init} (with arbitrary weights), new weights can be estimated
through Expectation-Maximization on some text sample {\tt test} by running the command:
\begin{verbatim}
$> interpolate-lm lmlist.init --learn=test
\end{verbatim}
\noindent
New weights will be written in the updated configuration file, called by default {\tt lmlist.init.out}.
You can also specify the name of the updated configuration file as follows:

\begin{verbatim}
$> interpolate-lm lmlist.init lmlist.final --learn=test
\end{verbatim}

\noindent
Similarly to {\tt compile-lm}, interpolated LMs can be queried through the option {\tt --score}

\begin{verbatim}
$> interpolate-lm lmlist.final --score=yes < test
\end{verbatim}

\noindent
and can return the perplexity of a given input text (``{\tt --eval text-file}''), optionally  at sentence level  by enabling the option ``{\tt --sentence yes}'',

\begin{verbatim}
$> interpolate-lm lmlist.final --eval=test 
$> interpolate-lm lmlist.final --eval=test --sentence=yes
\end{verbatim}

\bigskip
\noindent
If there are binary LMs in the list,  {\tt interpolate-lm} can avoid to load them in memory through the memory 
mapping option {\tt -memmap=1}.

\noindent
The full list of options is:

\begin{verbatim}
--learn text-file   learn optimal interpolation for text-file
--order n           order of n-grams used in --learn (optional)
--eval text-file    compute perplexity on text-file
--dub dict-size     dictionary upper bound (default 10^7)
--score [yes|no]    compute log-probs of n-grams from stdin
--debug [1-3]       verbose output for --eval option (see compile-lm)
--sentence [yes|no] (compute perplexity at sentence level (identified
                    through the end symbol)
--memmap 1          use memory map to read a binary LM
\end{verbatim}

### 2017-05-16 tools/extras/irstlm/doc/LMPruning.tex
Digit Recognition and Yes/No Recognition

Large LMs files can be pruned in a smart way by means of the command 
{\tt prune-lm} that removes $n$-grams for which resorting to the back-off 
results in a small loss. {\IRSTLM} implements a method similar to the 
Weighted Difference Method described in the paper {\em Scalable Backoff
Language Models} by Seymore and Rosenfeld.

\noindent
The syntax is as follows:
\begin{verbatim}
$> prune-lm --threshold=1e-6,1e-6  train.lm.gz  train.plm
\end{verbatim}
Thresholds for each n-gram level, up from 2-grams, are based on empirical 
evidence. Threshold zero results in no pruning. If less thresholds are specified,
the right most is applied to the higher levels. Hence, in the above example we
could have just specified one threshold, namely {\tt --threshold=1e-6}. 
The effect of pruning is shown in the following messages of {\tt prune-lm}:

\begin{verbatim}1-grams: reading 15059 entries
2-grams: reading 142684 entries
3-grams: reading 293685 entries
done
OOV code is 15058
OOV code is 15058
pruning LM with thresholds: 
 1e-06 1e-06
savetxt: train.plm
save: 15059 1-grams
save: 138252 2-grams
save: 194194 3-grams
\end{verbatim}

\noindent
The saved LM table {\tt train.plm}  contains about 3\% less bigrams, and 34\%  
less trigrams.
Notice that the output of prune-lm is an ARPA LM file, while the input can be 
either an ARPA or binary LM. 
In order to measure the loss in accuracy introduced
by pruning, perplexity of the resulting LM can be computed (see below).

\paragraph{Warning:} the possible quantization should be performed after pruning.

\paragraph{Warning:} 
{\IRSTLM} does not provide a reliable probability for the special
1-gram composed by the ``sentence start symbol'' ({\tt <s>}) , because none
should ever ask for it.  However, this pruning method requires the
computation of the probability of this 1-gram.  Hence, (only) in this case
the probability of this special 1-gram is arbitrarily set to 1.

### 2017-05-16 tools/extras/irstlm/doc/dict.tex
Digit Recognition and Yes/No Recognition

{\tt dict} is the command which copes with the dictionaries.

\begin{itemize}
\item It extracts the dictionary from a corpus or a dictionary;
\item It computes and shows the dictionary growth curve;
\item It computes and shows the out-of-vocabulary rate on a test corpus.
\end{itemize}

\subsubsection{Synopsis}

\begin{tabular}{llll}
\multicolumn{4}{l}{USAGE}\\
    & \multicolumn{3}{l}{\tt dict -i=$<$inputfile$>$ [options]} \\
    \\
\multicolumn{4}{l}{OPTIONS} \\
    & {\tt Curve}& {\tt c} &      show dictionary growth curve; default is false\\
    & {\tt CurveSize} & {\tt cs} &    default 10\\
    & {\tt Freq} & {\tt f} &    output word frequencies; default is false\\
    & {\tt Help} & {\tt h} &    print this help\\
    & {\tt InputFile} & {\tt i} &    input file (Mandatory)\\
    & {\tt IntSymb} & {\tt is} &    interruption symbol\\
    & {\tt ListOOV} & {\tt oov} &    print OOV words to stderr; default is false\\
    & {\tt LoadFactor} & {\tt lf} &    set the load factor for cache; it should be a positive real value; default is 0\\
    & {\tt OutputFile} & {\tt o} &    output file\\
    & {\tt PruneFreq} & {\tt pf} &    prune words with frequency below the specified value\\
    & {\tt PruneRank} & {\tt pr} &    prune words with frequency rank above the specified value\\
    & {\tt Size} & {\tt s} &    initial dictionary size; default is $10^6$\\
    & {\tt sort} & & sort dictionary by frequency; default is false\\
    & {\tt TestFile} & {\tt t} &    compute OOV rates on the specified test corpus\\
\end{tabular}

\subsubsection{Extraction of a dictionary}
To extract the dictionary from a given a text and store it in a file, run the following command:

\begin{verbatim}
$> dict -i=train.txt.se -o=train.dict -f=true
\end{verbatim}

The input text can be also generated on the fly by passing a command as value of the parameter{\tt InputFile }; in this case the single or double quotation marks are required.  
\begin{verbatim}
$> dict -i="cat train.txt | add-start-end.sh" -o=train.dict -f=true
\end{verbatim}

\noindent
For some applications like speech recognition, it  can be useful to limit the LM dictionary.
You can obtain such a pruned list either by means of the parameter {\tt PruneRank}, which only stores the top frequent, let us say, 10K words:
\begin{verbatim}
$> dict -i=train.txt.se -o=train.dict.pr10k -pr=10000
\end{verbatim}

\noindent
or by means of the parameter {\tt PruneFreq}, which only store the terms occurring more than a given amount of times, let us say, 5:
\begin{verbatim}
$> dict -i=train.txt.se -o=train.dict.pf5 -pf=5
\end{verbatim}

\noindent
The two pruning strategies can be combined.

\subsubsection{Dictionary growth curve}
{\tt dict} can display the distribution of the terms according to their frequency in a text or in a pre-computed dictionary. This facility is enabled by the parameter {\tt Curve}; the maximum frequency taken into account is specified by the parameter {\tt CurveSize}.

\begin{verbatim}
dict -i=train.dict -c=yes -cs=50
\end{verbatim}

\noindent
The output looks as follows
\begin{verbatim}
Dict size: 7893
**************** DICTIONARY GROWTH CURVE ****************
Freq  Entries  Percent
>0    7893     100.00%
>1    4880      61.83%
>2    3721      47.14%
>3    2990      37.88%
...
>47   271        3.43%
>48   264        3.34%
>49   258        3.27%
*********************************************************
\end{verbatim}
\noindent
Each row of the table reports, given the value in the first column, the amount of terms (second column) having at least the given frequency (first column), and its percentage  (third column) with respect to the total amount of entries.

\subsubsection{Out-of-vocabulary rate statistics}
{\tt dict} can display the distribution of the terms according to their frequency in a text or in a pre-computed dictionary; the maximum frequency taken into account is specified by the parameter {\tt CurveSize}.
\begin{verbatim}
$> dict -i=train.dict -t=test.txt.se -cs=50
\end{verbatim}

\noindent
The output looks as follows
\begin{verbatim}
Dict size: 7893
Words of test: 1009
**************** OOV RATE STATISTICS ****************
Freq  OOV_Entries  OOV_Rate
<1    119          11.79%
<2    151          14.97%
<3    191          18.93%
...
<48   457          45.29%
<49   457          45.29%
<50   457          45.29%
*********************************************************

\end{verbatim}
\noindent
Each row of the table reports, given the value in the first column, the out-of-vocabulary rate on the test set, assuming to prune the dictionary at the given frequency. In other words, 191 (18.93\%) of the running terms in the test set has a frequency smaller than 3 in the dictionary.

### 2017-05-16 tools/extras/irstlm/doc/ngt.tex
Digit Recognition and Yes/No Recognition

{\tt ngt} is the command which copes with the $n$-gram counts.

\begin{itemize}
\item It extracts the $n$-gram counts and stores into a $n$-gram table.
\item It prunes $n$-gram table.
\item It merges $n$-gram tables.
\item It transforms $n$-gram table formats.
\end{itemize}

\noindent
A new  $n$-gram table for the  limited dictionary can  be computed with {\tt ngt} by specifying 
the sub-dictionary:
\begin{verbatim}
$> ngt -i=train.www -sd=top10k -n=3 -o=train.10k.www -b=yes
\end{verbatim}
The command replaces  all words outside  top10K with  the special
out-of-vocabulary symbol {\tt \_unk\_}.{\tt dict} is the command which copes with the dictionaries.

\noindent
Another useful feature of ngt is the merging of two $n$-gram tables. Assume that we have 
split our training corpus into files  {\tt text-a} and file {\tt text-b} and have computed $n$-gram 
tables for both files, we can merge them with the option {\tt -aug}:
\begin{verbatim}
$> ngt -i="gunzip -c text-a.gz" -n=3 -o=text-a.www -b=yes
$> ngt -i="gunzip -c text-b.gz" -n=3 -o=text-b.www -b=yes
$> ngt -i=text-a.www -aug=text-b.www -n=3 -o=text.www -b=yes
\end{verbatim}

\paragraph{Warning:} Note that if the concatenation of {\tt text-a.gz} and {\tt text-b.gz} is equal to {\tt train.gz} the resulting $n$-gram tables
{\tt text.www} and {\tt train.www} can slightly differ. This happens because during the construction of each single $n$-gram table few $n$-grams are automatically added to make it consistent for further computation.

### 2017-05-16 tools/extras/irstlm/doc/parallelComputation.tex
Digit Recognition and Yes/No Recognition

This package provides facilities to build a gigantic LM in parallel in order to reduce computation time.
The script implementing this feature is based on the {\tt SUN Grid Engine} software\footnote{http://www.sun.com/software/gridware}.

\noindent
To apply the parallel computation run the following script (instead of {\tt build-lm.sh}):

\begin{verbatim}
$> build-lm-qsub.sh -i "gunzip -c train.gz" -n 3  -o train.ilm.gz -k 5
\end{verbatim}
Besides the options of {\tt build-lm.sh}, parameters for the SGE manager can be provided through the following one:

\begin{verbatim}
   -q      parameters for qsub, e.g. "-q <queue>", "-l <resources>"
\end{verbatim}

\noindent
The script performs the same {\em split-and-merge} policy described in Section~\ref{sec:giganticLM}, but some computation is performed in parallel (instead of sequentially) distributing the tasks on several machines.

### 2017-05-16 tools/extras/irstlm/doc/tlm.tex
Digit Recognition and Yes/No Recognition

Language models have to  cope with out-of-vocabulary words, that is internally represented
with the word class  {\tt \_unk\_}.  In order  to
compare perplexity of LMs having  different vocabulary size it is better
to define  a conventional dictionary  size, or dictionary  upper bound
size,  trough the  parameter  ({\tt -dub}).  In  the  following example,  we
compare the perplexity of the full vocabulary LM against the perplexity of the
LM estimated over the more frequent 10K-words. In our comparison, we assume a dictionary 
upper bound of one million words.

\begin{verbatim}
$>tlm -tr=train.10k.www -n=3 -lm=wb -te=test -dub=1000000
  n=49984 LP=342160.8721 PP=939.5565162 OVVRate=0.07666453265

$>tlm -tr=train.www -n=3 -lm=wb -te=test -dub=1000000
  n=49984 LP=336276.7842 PP=835.2144716 OVVRate=0.05007602433
\end{verbatim}

\noindent
The  large  difference  in  perplexity  between the two LMs is   explained  by  the 
significantly higher  OOV rate of the 10K-word LM.

\noindent
N-gram LMs generally apply frequency smoothing techniques, and combine
smoothed frequencies according to  two main schemes: interpolation and
back-off.  The  toolkit assumes interpolation
as default.  The back-off  scheme is computationally more costly but
often provides better performance. It  can be activated with the option
{\tt -bo=yes}, e.g.:

\begin{verbatim}
$>tlm -tr=train.10k.www -n=3 -lm=wb -te=test -dub=1000000 -bo=yes
  n=49984 LP=337278.3227 PP=852.1186066 OVVRate=0.07666453265
\end{verbatim}

\noindent
This toolkit implements several frequency smoothing methods, which are
specified  by  the  parameter  {\tt -lm}.  Three  methods  are  particularly
recommended:
\begin{itemize}
\item [a)] {\bf Modified shift-beta}, also known as  ``improved kneser-ney smoothing''.  
This smoothing scheme gives top performance when training data is not 
very sparse but it is more time and memory consuming during the estimation phase:

\begin{verbatim}
$>tlm -tr=train.www -n=3 -lm=msb -te=test -dub=1000000 -bo=yes
  n=49984 LP=321877.3411 PP=626.1609806 OVVRate=0.05007602433
\end{verbatim}

\item [b)] {\bf Witten Bell smoothing}. This is an excellent smoothing
   method which works well in every data condition and is much less time and memory consuming:

\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=wb -te=test -dub=1000000  -bo=yes
  n=49984 LP=331577.2279 PP=760.2652095 OVVRate=0.05007602433
\end{verbatim}

\item [c)] {\bf Shift-beta smoothing}. This smoothing method is a simpler and cheaper version
of the Modified shift-beta method and works sometimes better than Witten-Bell method:

\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=sb -te=test -dub=1000000  -bo=yes
  n=49984 LP=334724.5032 PP=809.6750442 OVVRate=0.05007602433
\end{verbatim}

\noindent
Moreover, the non linear smoothing parameter $\beta$ can be specified with the option {\tt -beta}:
\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=sb -beta=0.001 -te=test -dub=1000000  
       -bo=yes
  n=49984 LP=449339.8282 PP=8019.836058 OVVRate=0.05007602433
\end{verbatim}
This could be helpful in case we need to use language models with very limited frequency smoothing.

\end{itemize}
\subsection*{Limited Vocabulary}
\noindent
Using an  n-gram table  with a fixed  or limited  dictionary  will cause
some performance  degradation, as LM smoothing  statistics result
slightly distorted. A  valid alternative is to estimate  the LM on the
full dictionary of the training corpus and to use a limited dictionary
just when  saving the  LM on a  file.  This  can be achieved  with the
option {\tt -d} (or {\tt -dictionary}):
\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=msb -bo=y -te=test -o=train.lm -d=top10k
\end{verbatim}

### 2017-05-16 egs/0_MP_Digits/steps/cleanup/make_biased_lm_graphs.sh

@@ -1,0 +1,167 @@
+#!/bin/bash
+# Copyright 2012-2016     Johns Hopkins University (Author: Daniel Povey)
+#                2016     Vimal Manohar
+# Apache 2.0
+
+
+# This script creates biased decoding graphs based on the data transcripts as
+# HCLG.fsts.scp, in the specified directory; this can be consumed by
+# decode_segmentation.sh.
+# This is for use in data-cleanup and data-filtering.
+
+
+set -u
+set -o pipefail
+set -e
+
+# Begin configuration section.
+nj=10
+cmd=run.pl
+scale_opts="--transition-scale=1.0 --self-loop-scale=0.1"
+top_n_words=100 # Number of common words that we compile into each graph (most frequent
+                # in $data/text.orig.
+top_n_words_weight=1.0  # this weight is before renormalization; it can be more
+                        # or less than 1.
+min_words_per_graph=100  # Utterances will be grouped so that they have at least
+                         # this many words, before making the graph.
+stage=0
+
+### options for make_one_biased_lm.py.
+ngram_order=4  # maximum n-gram order to use (but see also --min-lm-state-cout).
+min_lm_state_count=10  # make this smaller (e.g. 2) for more strongly biased LM.
+discounting_constant=0.3  # strictly between 0 and 1.  Make this closer to 0 for
+                          # more strongly biased LM.
+
+# End configuration options.
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f path.sh ] && . ./path.sh # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# != 4 ]; then
+   echo "usage: $0 <data-dir|text> <lang-dir> <dir> <graph-dir>"
+   echo "e.g.:  $0 data/train data/lang exp/tri3_cleanup exp/tri3_cleanup/graphs"
+   echo "  This script creates biased decoding graphs per utterance (or possibly"
+   echo "  groups of utterances, depending on --min-words-per-graph).  Its output"
+   echo "  goes to <dir>/HCLG.fsts.scp, indexed by utterance.  Directory <dir> is"
+   echo "  required to be a model or alignment directory, containing 'tree' and 'final.mdl'."
+   echo "Main options (for others, see top of script file)"
+   echo "  --scale-opts <scale-opts>                 # Options relating to language"
[…]

### 2017-05-16 egs/0_MP_Digits/utils/pinyin_map.pl

@@ -1,0 +1,79 @@
+#!/usr/bin/env perl
+use warnings; #sed replacement for -w perl parameter
+
+$num_args = $#ARGV + 1;
+if ($num_args != 1) {
+  print "\nUsage: pinyin2phone.pl pinyin2phone\n";
+  exit;
+}
+
+open(MAPS, $ARGV[0]) or die("Could not open pinyin map file.");
+my %py2ph; foreach $line (<MAPS>) { @A = split(" ", $line);
+  $py = shift(@A);
+  $py2ph{$py} = [@A];
+}
+
+#foreach $word ( keys %py2ph ) {
+     #foreach $i ( 0 .. $#{ $py2ph{$word} } ) {
+     #    print " $word = $py2ph{$word}[$i]";
+     #}
+     #print " $#{ $py2ph{$word} }";
+     #print "\n";
+#}
+
+my @entry;
+
+while (<STDIN>) {
+  @A = split(" ", $_);
+  @entry = ();
+  $W = shift(@A);
+  push(@entry, $W);
+  for($i = 0; $i < @A; $i++) {
+    $initial= $A[$i]; $final = $A[$i];
+    #print $initial, " ", $final, "\n";
+    if ($A[$i] =~ /^CH[A-Z0-9]+$/) {$initial =~ s:(CH)[A-Z0-9]+:$1:; $final =~ s:CH([A-Z0-9]+):$1:;}
+    elsif ($A[$i] =~ /^SH[A-Z0-9]+$/) {$initial =~ s:(SH)[A-Z0-9]+:$1:; $final =~ s:SH([A-Z0-9]+):$1:;}
+    elsif ($A[$i] =~ /^ZH[A-Z0-9]+$/) {$initial =~ s:(ZH)[A-Z0-9]+:$1:; $final =~ s:ZH([A-Z0-9]+):$1:;}
+    elsif ($A[$i] =~ /^B[A-Z0-9]+$/) {$initial =~ s:(B)[A-Z0-9]+:$1:; $final =~ s:B([A-Z0-9]+):$1:;}
+    elsif ($A[$i] =~ /^C[A-Z0-9]+$/) {$initial =~ s:(C)[A-Z0-9]+:$1:; $final =~ s:C([A-Z0-9]+):$1:;}
[…]

### 2017-05-16 egs/1_MP_YesNo/steps/decode_sgmm2.sh

@@ -1,0 +1,226 @@
+#!/bin/bash
+
+# Copyright 2012  Johns Hopkins University (Author: Daniel Povey).  Apache 2.0.
+
+# This script does decoding with an SGMM system, with speaker vectors.
+# If the SGMM system was
+# built on top of fMLLR transforms from a conventional system, you should
+# provide the --transform-dir option.
+
+# Begin configuration section.
+stage=1
+transform_dir=    # dir to find fMLLR transforms.
+nj=4 # number of decoding jobs.
+acwt=0.1  # Just a default value, used for adaptation and beam-pruning..
+cmd=run.pl
+beam=13.0
+gselect=15  # Number of Gaussian-selection indices for SGMMs.  [Note:
+            # the first_pass_gselect variable is used for the 1st pass of
+            # decoding and can be tighter.
+first_pass_gselect=3 # Use a smaller number of Gaussian-selection indices in
+            # the 1st pass of decoding (lattice generation).
+max_active=7000
+max_mem=50000000
+#WARNING: This option is renamed lattice_beam (it was renamed to follow the naming
+#         in the other scripts
+lattice_beam=6.0 # Beam we use in lattice generation.
+vecs_beam=4.0 # Beam we use to prune lattices while getting posteriors for
+    # speaker-vector computation.  Can be quite tight (actually we could
+    # probably just do best-path.
+use_fmllr=false
+fmllr_iters=10
+fmllr_min_count=1000
+num_threads=1 # if >1, will use gmm-latgen-faster-parallel
+parallel_opts=  # ignored now.
+skip_scoring=false
+scoring_opts=
+# note: there are no more min-lmwt and max-lmwt options, instead use
+# e.g. --scoring-opts "--min-lmwt 1 --max-lmwt 20"
+# End configuration section.
+
+echo "$0 $@"  # Print the command line for logging
+
+[ -f ./path.sh ] && . ./path.sh; # source the path.
+. parse_options.sh || exit 1;
+
+if [ $# -ne 3 ]; then
[…]

### 2017-05-16 egs/1_MP_YesNo/steps/nnet3/get_egs_targets.sh

@@ -1,0 +1,415 @@
+#!/bin/bash
+
+# Copyright 2012-2015 Johns Hopkins University (Author: Daniel Povey).
+#           2015-2016 Vimal Manohar
+# Apache 2.0.
+
+# This script is similar to steps/nnet3/get_egs.sh but used
+# when getting general targets (not from alignment directory) for raw nnet
+#
+# This script, which will generally be called from other neural-net training
+# scripts, extracts the training examples used to train the neural net (and also
+# the validation examples used for diagnostics), and puts them in separate archives.
+#
+# This script dumps egs with several frames of labels, controlled by the
+# frames_per_eg config variable (default: 8).  This takes many times less disk
+# space because typically we have 4 to 7 frames of context on the left and
+# right, and this ends up getting shared.  This is at the expense of slightly
+# higher disk I/O while training.
+
+
+# Begin configuration section.
+cmd=run.pl
+feat_type=raw       # set it to 'lda' to use LDA features.
+target_type=sparse  # dense to have dense targets,
+                    # sparse to have posteriors targets
+num_targets=        # required for target-type=sparse with raw nnet
+frames_per_eg=8   # number of frames of labels per example.  more->less disk space and
+                  # less time preparing egs, but more I/O during training.
+                  # Note: may in general be a comma-separated string of alternative
+                  # durations (more useful when using large chunks, e.g. for BLSTMs);
+                  # the first one (the principal num-frames) is preferred.
+left_context=4    # amount of left-context per eg (i.e. extra frames of input features
+                  # not present in the output supervision).
+right_context=4   # amount of right-context per eg.
+left_context_initial=-1    # if >=0, left-context for first chunk of an utterance
+right_context_final=-1     # if >=0, right-context for last chunk of an utterance
+compress=true   # set this to false to disable compression (e.g. if you want to see whether
+                # results are affected).
+num_utts_subset=300     # number of utterances in validation and training
[…]

### 2017-05-16 egs/1_MP_YesNo/utils/build_const_arpa_lm.sh

@@ -1,0 +1,49 @@
+#!/bin/bash
+
+# Copyright 2014  Guoguo Chen
+# Apache 2.0
+
+# This script reads in an Arpa format language model, and converts it into the
+# ConstArpaLm format language model.
+
+# begin configuration section
+# end configuration section
+
+[ -f path.sh ] && . ./path.sh;
+
+. utils/parse_options.sh
+
+if [ $# != 3 ]; then
+  echo "Usage: "
+  echo "  $0 [options] <arpa-lm-path> <old-lang-dir> <new-lang-dir>"
+  echo "e.g.:"
+  echo "  $0 data/local/lm/3-gram.full.arpa.gz data/lang/ data/lang_test_tgmed"
+  echo "Options"
+  exit 1;
+fi
+
+export LC_ALL=C
+
+arpa_lm=$1
+old_lang=$2
+new_lang=$3
+
+mkdir -p $new_lang
+
+mkdir -p $new_lang
+cp -r $old_lang/* $new_lang
+
+unk=`cat $new_lang/oov.int`
+bos=`grep -w "<s>" $new_lang/words.txt | awk '{print $2}'`
+eos=`grep "</s>" $new_lang/words.txt | awk '{print $2}'`
+if [[ -z $bos || -z $eos ]]; then
+  echo "$0: <s> and </s> symbols are not in $new_lang/words.txt"
+  exit 1
+fi
+
+
+arpa-to-const-arpa --bos-symbol=$bos \
+  --eos-symbol=$eos --unk-symbol=$unk \
+  "gunzip -c $arpa_lm | utils/map_arpa_lm.pl $new_lang/words.txt|"  $new_lang/G.carpa  || exit 1;
+
+exit 0;

### 2017-05-16 tools/extras/irstlm/src/ngramtable.cpp

@@ -1,0 +1,1870 @@
+// $Id: ngramtable.cpp 35 2010-07-19 14:52:11Z nicolabertoldi $
+
+/******************************************************************************
+IrstLM: IRST Language Model Toolkit, compile LM
+Copyright (C) 2006 Marcello Federico, ITC-irst Trento, Italy
+
+This library is free software; you can redistribute it and/or
+modify it under the terms of the GNU Lesser General Public
+License as published by the Free Software Foundation; either
+version 2.1 of the License, or (at your option) any later version.
+
+This library is distributed in the hope that it will be useful,
+but WITHOUT ANY WARRANTY; without even the implied warranty of
+MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+Lesser General Public License for more details.
+
+You should have received a copy of the GNU Lesser General Public
+License along with this library; if not, write to the Free Software
+Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA  02110-1301 USA
+
+******************************************************************************/
+
+#include <sstream>
+#include "util.h"
+#include "mfstream.h"
+#include "math.h"
+#include "mempool.h"
+#include "htable.h"
+#include "dictionary.h"
+#include "n_gram.h"
+#include "ngramtable.h"
+#include "crc.h"
+
+using namespace std;
+
+tabletype::tabletype(TABLETYPE tt,int codesize) {
+
+    if (codesize<=4 && codesize>0)
+      CODESIZE=codesize;
+    else {
+			exit_error(IRSTLM_ERROR_DATA,"ngramtable wrong codesize");
+    }
+
+    code_range[1]=255;
+    code_range[2]=65535;
+    code_range[3]=16777214;
+    code_range[4]=2147483640;
+    code_range[6]=140737488360000LL; //stay below true limit
[…]

### 2017-05-16 tools/extras/irstlm/src/plsa.cpp

@@ -1,0 +1,250 @@
+/******************************************************************************
+ IrstLM: IRST Language Model Toolkit, compile LM
+ Copyright (C) 2006 Marcello Federico, ITC-irst Trento, Italy
+ 
+ This library is free software; you can redistribute it and/or
+ modify it under the terms of the GNU Lesser General Public
+ License as published by the Free Software Foundation; either
+ version 2.1 of the License, or (at your option) any later version.
+ 
+ This library is distributed in the hope that it will be useful,
+ but WITHOUT ANY WARRANTY; without even the implied warranty of
+ MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU
+ Lesser General Public License for more details.
+ 
+ You should have received a copy of the GNU Lesser General Public
+ License along with this library; if not, write to the Free Software
+ Foundation, Inc., 51 Franklin Street, Fifth Floor, Boston, MA  02110-1301 USA
+ 
+ ******************************************************************************/
+
+
+#include <iostream>
+#include "cmd.h"
+#include <pthread.h>
+#include "thpool.h"
+#include "util.h"
+#include "mfstream.h"
+#include "mempool.h"
+#include "htable.h"
+#include "dictionary.h"
+#include "n_gram.h"
+#include "ngramtable.h"
+#include "doc.h"
+#include "cplsa.h"
+
+using namespace std;
+using namespace irstlm;
+
+void print_help(int TypeFlag=0){
+    std::cerr << std::endl << "plsa -  probabilistic latent semantic analysis modeling" << std::endl;
+    std::cerr << std::endl << "USAGE:"  << std::endl;
+    std::cerr << "       plsa -tr|te=<text>  -m=<model> -t=<n> [options]" << std::endl;
+    std::cerr << std::endl << "DESCRIPTION:" << std::endl;
+    std::cerr << "       Train a PLSA model from a corpus and test it to infer topic or word " << std::endl;
+    std::cerr << "       distributions from other texts." << std::endl;
+    std::cerr << "       Notice: multithreading is available both for training and inference." << std::endl;
+    
+    std::cerr << std::endl << "OPTIONS:" << std::endl;
+    
+    
+    FullPrintParams(TypeFlag, 0, 1, stderr);
+    
+    std::cerr << std::endl << "EXAMPLES:" << std::endl;
[…]

### 2017-06-08 tools/extras/irstlm/doc/ClassAndChunkLMs.tex

{\IRSTLM} allows the use of class and chunk LMs, and a special
handling of input tokens which are concatenation of $N \ge 1$ fields separated
by the character \#, e.g.

\begin{verbatim}
   word#lemma#part-of-speech#word-class
\end{verbatim}

\noindent The processing is guided by the format of the file passed to
Moses or {\tt compile-lm}: if it contains just the LM, either in textual or
binary format, it is treated as usual; otherwise, it is supposed to have
the following format:

\begin{verbatim}
LMMACRO <lmmacroSize> <selectedField> <collapse>
<lmfilename>
<mapfilename>
\end{verbatim}

\noindent where:
\begin{verbatim}
 LMMACRO is a reserved keyword
 <lmmacroSize> is a positive integer
 <selectedField> is an integer >=-1
 <collapse> is a boolean value (true, false)
 <lmfilename> is a file containing a LM (format compatible with {\IRSTLM})
 <mapfilename> is an (optional) file with a (one|many)-to-one map
\end{verbatim}

\noindent The various cases are discussed with examples in the
following. Data used in those examples can be found in the directory {\tt
example/chunkLM/} which represents the relative path for all the parameters
of the referred commands.  Note that texts with different tokens (words,
POS, word\#POS pairs...) used either as input or for training LMs are all
derived from the same multifield texts in order to allow direct comparison
of results.

\subsection{Field selection}

The simplest case is that of the LM in {\tt <lmfilename>} referring just to
one specific field of the input tokens. In this case, it is possible to
specify the field to be selected before querying the LM through the integer
{\tt <selectedField>} ($0$ for the first filed, $1$ for the
second...). With the value $-1$, no selection is applied and the LM is
queried with n-grams of whole strings.  The other parameters are set as:

\begin{verbatim}
 <lmmacroSize> : set to the size of the LM in <lmfilename>
 <collapse>    : false
\end{verbatim}

\noindent The third line optionally reserved to {\tt <mapfilename>} does not exist.

\bigskip
\noindent
Examples:

\bigskip
\noindent
\thesubsection.a) selection of the second field:
\begin{verbatim}
$> compile-lm --eval=test/test.w-micro cfgfile/cfg.2ndfield
%% Nw=126 PP=2.68 PPwp=0.00 Nbo=0 Noov=0 OOV=0.00%
\end{verbatim}

\noindent
\thesubsection.b) selection of the first field:
\begin{verbatim}
$> compile-lm --eval=test/test.w-micro cfgfile/cfg.1stfield
%% Nw=126 PP=9.71 PPwp=0.00 Nbo=76 Noov=0 OOV=0.00%
\end{verbatim}

\noindent The result of the latter case is identical to that obtained with
the standard configuration involving just words:

\bigskip
\noindent
\thesubsection.c) usual case on words:
\begin{verbatim}
$> compile-lm --eval=test/test.w lm/train.en.blm 
%% Nw=126 PP=9.71 PPwp=0.00 Nbo=76 Noov=0 OOV=0.00%
\end{verbatim}

\subsection{Class LMs}

Possibly, a many-to-one or one-to-one map can be passed through the
{\tt <mapfilename>} parameter which has the simple format:

\begin{verbatim}
w1 class(w1)
w2 class(w2)
 ...
wM class(wM)
\end{verbatim}

\noindent The map is applied to each component of ngrams before the LM
query. Examples:
\bigskip

\noindent \thesubsection.a) map applied to the second field:
\begin{verbatim}
$> compile-lm --eval=test/test.w-micro cfgfile/cfg.2ndfld-map
%% Nw=126 PP=16.40 PPwp=0.00 Nbo=33 Noov=0 OOV=0.00%
\end{verbatim}

\noindent \thesubsection.b) just to assess the correctness of the (16.2.a) result:
\begin{verbatim}
$> compile-lm --eval=test/test.macro lm/train.macro.blm
%% Nw=126 PP=16.40 PPwp=0.00 Nbo=33 Noov=0 OOV=0.00%

\end{verbatim}

\subsection{Chunk LMs}

A particular processing is performed whenever fields are supposed to
correspond to microtags, i.e. the per-word projections of chunk labels. By
means of the {\tt <collapse>} parameter, it is possible to activate a
processing aiming at collapsing the sequence of microtags defining a
chunk. The chunk LM is then queried with ngrams of chunk labels, in an
asynchronous manner with respect to the sequence of words, as in general
chunks consist of more words.

\noindent
The collapsing operation is automatically activated if the sequence of
microtags is:

\begin{verbatim}
 TAG( TAG+ TAG+ ... TAG+ TAG)
\end{verbatim}

\noindent
Such a sequence is collapsed into a single chunk label (let us say {\tt
CHNK}) as long as {\tt TAG(}, {\tt TAG+} and {\tt TAG)} are all mapped into
the same label {\tt CHNK}. The map into different labels or a different
use/position of characters $($, $+$ and $)$ in the lexicon of tags prevent
the collapsing operation even if {\tt <collapse>} is set to {\tt true}. Of
course, if {\tt <collapse>} is {\tt false}, no collapse is attempted.

\paragraph{Warning:} In this context, it assumes an important role the parameter {\tt
<lmmacroSize>}: it defines the size of the n-gram before the collapsing
operation, that is the number of microtags of the actually processed
sequence. {\tt <lmmacroSize>} should be large enough to ensure that after
the collapsing operation, the resulting n-gram of chunks is at least of the
size of the LM to be queried (the {\tt <lmfilename>}). As an example,
assuming {\tt <lmmacroSize>=6}, {\tt <selectedField>=1}, {\tt
<collapse>=true} and 3 the size of the chunk LM, the following input

\begin{verbatim}
 on#PP average#NP( 30#NP+ -#NP+ 40#NP+ cm#NP)
\end{verbatim}

\noindent will yield to query the LM with just the bigram {\tt (PP,NP)},
instead of a more informative trigram; for this particular case, the value
6 for {\tt <lmmacroSize>} is not enough.  On the other side, for efficiency
reasons, it cannot be set to an unlimited valued. A reasonable value could
derive from the average number of microtags per chunk (2-3), which means
setting {\tt <lmmacroSize>} to two-three times the size of the LM in {\tt
<lmfilename>}.  Examples:
\bigskip

\noindent \thesubsection.a) second field, micro$\rightarrow$macro map, collapse:
\begin{verbatim}
$> compile-lm --eval=test/test.w-micro cfgfile/cfg.2ndfld-map-cllps
%% Nw=126 PP=1.84 PPwp=0.00 Nbo=0 Noov=0 OOV=0.00%

$> compile-lm --eval=test/test.w-micro cfgfile/cfg.2ndfld-map-cllps -d=1
%% Nw=126 PP=1.83774013 ... OOV=0.00% logPr=-33.29979642

\end{verbatim}

\noindent
\thesubsection.b) whole token,  micro$\rightarrow$macro map, collapse:
\begin{verbatim}
$> compile-lm --eval=test/test.micro cfgfile/cfg.token-map-cllps
%% Nw=126 PP=1.84 PPwp=0.00 Nbo=0 Noov=0 OOV=0.00%
\end{verbatim}

\noindent
\thesubsection.c)  whole token,  micro$\rightarrow$macro map, NO collapse:
\begin{verbatim}
$> compile-lm --eval=test/test.micro cfgfile/cfg.token-map
%% Nw=126 PP=16.40 PPwp=0.00 Nbo=0 Noov=0 OOV=0.00%
\end{verbatim}
\noindent Note that the configuration (16.3.c) gives the same result of that in
example (16.2.b), as they are equivalent.

\bigskip
\noindent
\thesubsection.d) As an actual example related to the ``warning'' note
reported above, the following configuration with usual LM:

\begin{verbatim}
$> compile-lm --eval=test/test.chunk lm/train.macro.blm -d=1
Nw=73 PP=2.85754443 ... OOV=0.00000000% logPr=-33.28748842
\end{verbatim}

\noindent not necessarily yields the same log-likelihood ({\tt logPr}) nor the same perplexity ({\tt PP}) of case (16.3.a).
In fact, concerning {\tt PP}, the length of the input sequence is definitely different (126 tokens before collapsing, 73 after that).
Even the {\tt logPr} is different (-33.29979642 vs. -33.28748842) because in (16.3.a) some 6-grams ({\tt
<lmmacroSize>} is set to 6) after collapsing reduce to $n$-grams of size less
than 3 (the size of lm/train.macro.blm). By setting {\tt <lmmacroSize>} to
a larger value (e.g. 8), the same {\tt logPr} will be computed.

### 2017-06-08 tools/extras/irstlm/doc/LMFileFormats.tex

{\IRSTLM} supports several types of input and output formats for handling LMs, $n$-gram counts, dictionaries.

%{\IRSTLM} supports three output formats of LMs. These formats have the
%purpose of permitting  the use of LMs by  external programs.

\subsection{File Formats for Dictionary}
The dictionary is the data structure exploited by  {\IRSTLM} to store a set of terms.

{\IRSTLM} saves a dictionary in textual file format consisting of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "dictionary", a fixed value 0, and the amount of terms the dictionary contains;
\item a set of terms listed according to either their occurrence or their frequency in the data.
\end{itemize}
Here is an excerpt.
\begin{verbatim}
dictionary 0 7893
<s>
</s>
solemn
ceremony
marks
....
\end{verbatim}

\noindent
Optionally, the occurrence frequencies of each term can be stored as well; in this case the keyword is "DICTIONARY".

\noindent
Here is an excerpt.
\begin{verbatim}
DICTIONARY 0 7893
<s> 5000
</s> 5001
solemn 7
ceremony 59
....
\end{verbatim}

\IMPORTANT{The list order is used by {\IRSTLM} to define the internal codes of the terms.
In the vast majority of cases, it is completely transparent and irrelevant to the user.
Only in very few cases highlighted in this manual, this order is crucial.}

\subsection{File Formats for $n$-gram Table}
The $n$-gram table is the data structure exploited by  {\IRSTLM} to store a set of $n$-grams. {\IRSTLM} stores an $n$-gram table either in textual or binary formats.

\subsubsection{Textual format}
The textual format consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "nGrAm", the order $n$ of the $n$-grams, the amount of $n$-grams the $n$-gram table contains, and a second keyword representing the table type;
\item a second line reporting the size of the dictionary associated to the $n$-grams;
\item the terms of the dictionary (one term per line) with their frequency;
\item the list of all $n$-grams with their counts.
\end{itemize}
Here is an excerpt.
\begin{verbatim}
NgRaM 3 76857 ngram
7893
<s> 5000
</s> 5001
solemn 7
ceremony 59
...
<s> <s> <s>     2
<s> <s> </s>    1
<s> </s> </s>   1
<s> solemn ceremony     1
<s> a solemn    1
\end{verbatim}

\subsubsection{Binary format}
The binary format is similar, but its main keyword is "{\tt NgRaM}" (different caseing), and the list of $n$-grams is binarized; hence, the last portion of the binary $n$-gram table is not user-readable.

\subsubsection{Google $n$-gram format}
{\IRSTLM} supports the Google $n$-gram format as well both for input and output. This format, always textual, simply consists of the list of all $n$-grams with their counts.
Here is an excerpt.
\begin{verbatim}
<s> <s> <s>     2
<s> <s> </s>    1
<s> </s> </s>   1
<s> solemn ceremony     1
<s> a solemn    1
...
\end{verbatim}

\subsubsection{Table types}
The table type keyword represents the way the $n$-grams are collected and the way they are exploited for further computation:
\begin{itemize}
\item{\tt ngram}: each entry is a standard $n$-gram, i.e. a contiguous sequence of $n$ terms; they are usually used to estimate a standard $n$-gram LM;
\item{\tt co-occ$K$}: each entry is a xxxxxx, where $K$ is XXXXXX;
\item{\tt hm$S$}: each entry is a xxxxxx, where $K$ is XXXXXX;
\end{itemize}

\subsection{File Formats for LM}
{\IRSTLM} handles LM both in textual and binary formats.
It provides facilities to save disk space storing probabilities as quantized values instead of floating  point values, and to reduce access time saving $n$-grams in inverted order.

\subsubsection{Textual Format}
The textual format is the well-known ARPA format introduced in DARPA ASR evaluations  to exchange LMs.
ARPA format  is  supported by most third party LM toolkit, like SRILM and KenLM.

The ARPA format consists of:
\begin{itemize}
\item one block reporting the amount $n$-grams stored for level $m$ of the LM ($m<n$); this block starts with the keyword "{\tt \textbackslash{}data}";
\item one block per level reporting the set of $m$-grams for that level together with their log-probability (first field), and the backoff log-probability (last field); each block starts with the keyword "{\tt \textbackslash{}$m$-grams}", with the correct value for $m$;
\item the keyword  "{\tt \textbackslash{}end\textbackslash{}}" closes the file.
\end{itemize}

\noindent
Here is an excerpt.
\begin{verbatim}
\data\
ngram  1=      7894
ngram  2=     46269
ngram  3=     12188
\1-grams:
-4.79871        <s>     -0.826378
....
-1.20244        <unk>
\2-grams:
-3.29024        <s> <s> -0.221849
...
-0.289359       restructuring of
\3-grams:
-0.397606       <s> <s> <s>
-1.67881        <s> a hong_kong
...
-0.420213       seymf council ,
\end\
\end{verbatim}

\noindent
Empty lines can occur before and after each block.

\noindent
There is no limit to the order $n$ of $n$-grams.

\IMPORTANT{Backoff log-probabilities are not reported if equal to 0; backoff log-probabilities do not exist for the largest order.}

\subsubsection{Quantized Textual Format}
This textual format extends the ARPA textual format including codebooks that quantize 
probabilities and back-off weights of each $n$-gram level.

The quantized ARPA format consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "qARPA", the order $n$ of the LM, the size of the $n$ codebooks
\item one block reporting the amount $n$-grams stored for level $m$ of the LM ($m<n$); this block starts with the keyword "{\tt \textbackslash{}data}";
\item one block per level reporting first the codebooks for the level and then the set of $m$-grams for that level together with their quantized log probability (first field), and the backoff quantized  log-probability (last field); each block starts with the keyword "{\tt \textbackslash{}$m$-grams}", with the correct value for $m$;
\item the keyword  "{\tt \textbackslash{}end\textbackslash{}}" closes the file.
\end{itemize}

\noindent
Here is an excerpt.
\begin{verbatim}
qARPA 3 256 256 256
\data\
ngram 1= 7894
ngram 2= 46269
ngram 3= 12188
\1-grams:
256
-4.79885 -99
-4.62261 -1.75587
....
0       <s>     53
186     </s>    0
....
\2-grams:
256
-3.79901 -99
-3.62278 -3.01953
...
7       <s>     <s>     255
65      <s>     </s>    255
....
\end\
\end{verbatim}

\subsubsection{Intermediate Textual Format}
This is an {\em intermediate} ARPA format used by {\IRSTLM} for optimizing computation of huge LM. It differs from the ARPA format in two aspects:
\begin{itemize}
\item the header line contains only the keyword {\tt iARPA};
\item the first field of each $n$-gram entry is its smoothed frequency of instead of its log-probability.
\end{itemize}

\COMMENT{
\noindent
Nevertheless, iARPA format is properly managed by the {\tt compile-lm} command
in order to generate a binary version or a standard ARPA version.
}

\subsubsection{Binary Format}
The binary format supported by {\IRSTLM} allows for save disk space and upload the LM quicker.

\noindent 
The binary format consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "blmt", the order $n$ of the LM, and the amount $n$-grams stored for level $m$ of the LM ($m<n$);
\item a second line reporting the size of the dictionary associated to the $n$-grams;
\item the terms of the dictionary (one term per line) with their frequency, if available;
\item a binary section containing $n$-grams and their probabilities; this portion is not user-readable.
\end{itemize}

\begin{verbatim}
blmt 3       7894      46269      12188
7894
<s> 1
</s> 5001
solemn 7
...
_binary_data_
\end{verbatim}

\subsubsection{Quantized Binary Format}
The quantized binary format stores the quantized version of a LM.

\noindent 
It consists of:
\begin{itemize}
\item a header line specifying the most important information about the file itself: the keyword "Qblmt", the order $n$ of the LM, and the amount $n$-grams stored for level $m$ of the LM ($m<n$);
\item a second line specifying the most important information of the codebooks of each level: the keyword "NumCenters", and  the size of the $n$ codebooks;
\item a third line reporting the size of the dictionary associated to the $n$-grams;
\item the terms of the dictionary (one term per line) with their frequency, if available;
\item a binary section containing the codebooks, the $n$-grams and their quantized probabilities; this portion is not user-readable.
\end{itemize}

\begin{verbatim}
Qblmt 3 7894 46269 12188
NumCenters 256 256 256
7894
<s> 1
</s> 5001
solemn 7
...
_binary_data_
\end{verbatim}

\subsubsection{Inverted Binary Format}
{\IRSTLM} can store the $n$-grams in inverted order to speed up access time. This applies to both standard and quantized binary formats, namely {\tt blmt} or {\tt Qblmt}. The keywords are {\tt blmtI} or {\tt QblmtI}, respectively.

### 2017-06-08 tools/extras/irstlm/doc/LMFiltering.tex

A large LM can be filtered according to a word list through the command:

\begin{verbatim}
$> compile-lm  train.lm filtered.lm --filter=list
\end{verbatim}
The resulting LM will only contain n-grams inside the provided list of words,
with the exception of the 1-gram level, which by default is preserved identical
to the original LM. This behavior can be changed by setting the option 
{\tt --keepunigrams no}.  LM filtering can be useful once very large LMs can
be specialized in advance to work on a particular portion of language.
\noindent
If the original LM is in binary format and is very large, {\tt compile-lm} can avoid to load it in memory,
through the memory mapping option {\tt -memmap=1}.

### 2017-06-08 tools/extras/irstlm/doc/LMInterface.tex

LMs are useful when they can be queried through another application in order to compute 
perplexity scores or n-gram probabilities. {\IRSTLM} provides two possible interfaces: 
\begin{itemize}
\item at the command level, through  {\tt compile-lm}
\item at the c++ library level, mainly through methods of the class {\tt lmtable}
\end{itemize}

\noindent
In the following, we will only focus on the command level interface. Details about
the c++ library interface will be provided in a future version of this manual.

\subsection{Perplexity Computation}
Assume we have estimated and saved the following LM:

\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=wb -te=test -o=train.lm -ps=no
 n=49984 LP=308057.0419 PP=474.9041687 OVVRate=0.05007602433
\end{verbatim}

\noindent
To compute the perplexity directly from the LM on disk, we can use the command:

\begin{verbatim}
$> compile-lm train.lm  --eval=test
 %% Nw=49984 PP=1064.40 PPwp=589.50 Nbo=38071 Noov=2503 OOV=5.01%
\end{verbatim}
Notice that {\tt PPwp} reports the contribution of OOV words to the perplexity. Each OOV word is indeed penalized by dividing the
LM probability of the {\tt unk} word by  the quantity

\centerline{{\tt DictionaryUpperBound} - {\tt SizeOfDictionary}}

\noindent
The OOV penalty can be modify by changing the {\tt DictionaryUpperBound} with the parameter {\tt --dub} (whose default value is set to $10^7$). \\

\noindent
The perplexity of the pruned LM can be computed with the command:
\begin{verbatim}
$> compile-lm train.plm --eval=test --dub=10000000
%% Nw=49984 PP=1019.69 PPwp=564.73 Nbo=39907 Noov=2503 OOV=5.01%
\end{verbatim}
Interestingly, a slightly better value is obtained which could be explained by the 
fact that pruning has removed many unfrequent trigrams and has redistributed 
their probabilities over more frequent bigrams.

\noindent
Notice that {\tt PPwp} reports the perplexity with a fixed dictionary upper-bound of 10 million words. Indeed:
\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=wb -te=test -o=train.lm -ps=no -dub=10000000 
n=49984 LP=348396.8632 PP=1064.401254 OVVRate=0.05007602433
\end{verbatim}

\bigskip
\noindent
Again, if the LM is in binary format and is very large, {\tt compile-lm} can avoid to load it in memory,
through the memory mapping option {\tt -memmap=1}.

\bigskip
\noindent
By enabling the option ``{\tt --sentence yes}'', {\tt compile-lm} computes perplexity and related figures (OOV rate, number of backoffs, etc.) for each input sentence. The end of a sentence is identified by a given symbol ({\tt </s>} by default).
\begin{verbatim}
$> compile-lm train.plm --eval=test --dub=10000000 --sentence=yes	
\end{verbatim}
{\small 
\begin{verbatim}
%% sent_Nw=1 sent_PP=23.22 sent_PPwp=0.00 sent_Nbo=0 sent_Noov=0 sent_OOV=0.00%
%% sent_Nw=8 sent_PP=7489.50 sent_PPwp=7356.27 sent_Nbo=7 sent_Noov=2 sent_OOV=25.00%
%% sent_Nw=9 sent_PP=1231.44 sent_PPwp=0.00 sent_Nbo=14 sent_Noov=0 sent_OOV=0.00%
%% sent_Nw=6 sent_PP=27759.10 sent_PPwp=25867.42 sent_Nbo=19 sent_Noov=1 sent_OOV=16.67%
.....
%% sent_Nw=5 sent_PP=378.38 sent_PPwp=0.00 sent_Nbo=39893 sent_Noov=0 sent_OOV=0.00%
%% sent_Nw=15 sent_PP=4300.44 sent_PPwp=2831.89 sent_Nbo=39907 sent_Noov=1 sent_OOV=6.67%
%% Nw=49984 PP=1019.69 PPwp=564.73 Nbo=39907 Noov=2503 OOV=5.01%
\end{verbatim}
}

\bigskip
\noindent
Finally, tracing information with the {\tt --eval }  option are shown by setting 
debug levels from 1 to 4 ({\tt --debug}):
\begin{enumerate}
\item reports the back-off level for each word
\item adds the log-prob 
\item adds the back-off weight
\item check if probabilities sum up to 1.
\end{enumerate}

\subsection{Probability Computations}
Word-by-word log-probabilities  can be computed as well from standard input with the command:
\begin{verbatim}
$> compile-lm train.lm --score=yes < test

> </s>  1 p= NULL
> <s> <unk>     1 p= NULL
> <s> <unk> of  1 p= -3.530047e+00 bo= 2
> <unk> of the  1 p= -1.250668e+00 bo= 1
> of the senate 1 p= -1.170901e+01 bo= 1
> the senate (  1 p= -5.457265e+00 bo= 2
> senate ( <unk>        1 p= -2.166440e+01 bo= 2
....
....
\end{verbatim}

\noindent
the command reports the currently observed n-gram, 
including {\tt\_unk\_} words, a dummy
constant frequency 1, the log-probability of the n-gram, and the 
number of back-offs performed by the LM.

\paragraph{Warning:} All cross-sentence $n$-grams are skipped. The 1-grams with the sentence start symbol are also skipped. In a $n$-grams all words before the sentence start symbol are removed. For $n$-grams, whose size is smaller than the LM order, probability is not computed, but a {\tt NULL} value is returned.

### 2017-06-08 tools/extras/irstlm/doc/LMQuantization.tex

A language model file in ARPA  format, created with the IRST LM toolkit or
with other tools, can be quantized and stored in a compact data structure, 
called language model table.  Quantization can be performed by the command:

\begin{verbatim}
$> quantize-lm  train.lm train.qlm
\end{verbatim}

\noindent
which  generates   the  quantized  version  {\tt train.qlm} that  encodes all probabilities and back-off 
weights in 8 bits. The  output is a  modified ARPA format, called qARPA. Notice that quantized
LMs reduce memory consumptions at the cost of some loss in performance. Moreover, probabilities
of quantized LMs are not supposed to be properly normalized.

### 2017-06-08 tools/extras/irstlm/doc/gettingStarted.tex

After a successful installation, you are ready to use {\IRSTLM}.

\noindent
In this Section, a basic 4-step procedure is given to estimate a LM and to compute its perplexity on a text.
Many changes to this procedure can be done in order to optimize effectiveness and efficiency according to your needs.

\noindent 
Please refer to Section~\ref{sec:commands} to learn more about each IRSTLM commands, and 
to Section~\ref{sec:functions} to get hints about IRSTLM functionalities.

\IMPORTANT{All programs assume that the environment variable {\bf IRSTLM} is correctly set to {\tt /path/to/install/doc}, and that that environment variable {\bf PATH} includes the command directory {\tt /path/to/install/bin}. see above}

\noindent
Data used in the following usage examples can be found in an archive you can download from the official website of {\IRSTLM}.
Most of them are very little, so the reported figures are not reliable.

\subsection{Preparation of Training Data}
In order to estimate a Language Model, you first need to prepare your training corpus. The corpus just consists of a text.
We assume that the text is already preprocessed according to the user needs; this means that lowercasing, uppercasing, tokenization, and any other text transformation has to be performed beforehand with other tools.

\noindent
You can only decide whether you are interested that {\IRSTLM} is aware of sentence boundaries, i.e. where a sentence starts and ends. Otherwise, it considers the corpus as a continuous stream of text, and does not identify sentence splits.

\noindent
The following script adds start and end symbols ({\tt <s>} and {\tt </s>}, respectively) to all lines in your training corpus.
\begin{verbatim}
$> cat train.txt | add-start-end.sh > train.txt.se
\end{verbatim}

\noindent
{\IRSTLM} does not compute probabilities for cross-sentence $n$-grams, i.e. $n$-grams including the pair {\tt </s>  <s>}.

\IMPORTANT{{\IRSTLM} assumes that each line corresponds to a sentence, regardless the presence of punctuation inside or at the end of the line.}
\IMPORTANT{Start and end symbols ({\tt <s>} and {\tt </s>}) should be considered reserved symbols, and used only as sentence boundaries.}

\subsection{Computation of $n$-gram statistics}
\noindent
You can now collect $n$-gram statistics for your training data (3-gram in this example) by running the command:

\begin{verbatim}
$> ngt -i=train.txt.se -n=3 -o=train.www -b=yes
\end{verbatim}

\noindent
The $n$-grams counts are saved in the binary file "train.www".

\subsection{Estimation of the LM}
\noindent
You can now estimate a $n$-gram LM (3-gram LM in this example) smoothed according to the Linear Witten Bell method by running the command:

\begin{verbatim}
$> tlm -tr=train.www -n=3 -lm=LinearWittenBell -obin=train.blm
\end{verbatim}
\noindent
The estimated LM is saved in the binary file "train.blm".

\subsection{Computation of the Perplexity}
\noindent
With the estimated LM, you can now compute the perplexity of any text contained in "test.txt" by running the commands below.

\noindent
To be compliant with the training data actually used to estimate the LM, start and end symbols are added to the text as well.

\begin{verbatim}
$> cat test.txt | add-start-end.sh > test.txt.se
$> compile-lm  train.blm  --eval=test.txt.se
\end{verbatim}

\noindent
which produces the output:
\begin{verbatim}
%% Nw=1009 PP=8547.90 PPwp=6870.51 Nbo=983 Noov=119 OOV=11.79%
\end{verbatim}

\noindent
The output shows the number of  words ({\tt Nw}), the LM perplexity ({\tt PP}), the portion of PP due to the out-of-vocabulary words ({\tt PPwp}), the amount of backoff calls({\tt Nbo}) required for computing PP, the amount of out-of-vocabulary words ({\tt Noov}), and the out-of-vocabulary rate ({\tt OOV}).

### 2017-06-08 tools/extras/irstlm/doc/giganticLM.tex

LM estimation starts with the collection of n-grams and their frequency counters. Then, 
smoothing parameters are estimated for each n-gram level; infrequent n-grams are
possibly pruned and, finally, a LM file is created containing n-grams with probabilities and 
back-off weights.  This procedure can be very demanding in terms of memory and
time if it applied on huge corpora.   We provide here a way to split LM training  into smaller and independent steps, that can be easily distributed among independent processes. The  
procedure relies on a training scripts that makes little use of computer RAM and implements 
the  Witten-Bell smoothing method in an exact way.

\noindent
Before starting, let us create a working directory under {\tt examples}, as many files will be created:

\begin{verbatim}
$> mkdir stat
\end{verbatim}

The script to generate the LM is:

\begin{verbatim}
$> build-lm.sh -i "gunzip -c train.gz" -n 3  -o train.ilm.gz -k 5
\end{verbatim}
where the available options are:

\begin{verbatim}
-i    Input training file e.g. 'gunzip -c train.gz'
-o    Output gzipped LM, e.g. lm.gz
-k    Number of splits (default 5)
-n    Order of language model (default 3)
-t    Directory for temporary files (default ./stat)
-p    Prune singleton n-grams (default false)
-s    Smoothing: witten-bell (default), kneser-ney, improved-kneser-ney 
-b    Include sentence boundary n-grams (optional)
-d    Define subdictionary for n-grams (optional)
-v    Verbose
\end{verbatim}

\noindent
The script splits the estimation procedure into 5 distinct jobs, that are explained in
the following section. There are other options that can be used. We recommend for instance to use pruning of singletons to get smaller LM files. 
Notice that {\tt build-lm.sh} produces a LM file {\tt train.ilm.gz} that is NOT in the final ARPA format, but in
an intermediate format called {\tt iARPA}, that is recognized by the {\tt compile-lm} 
command and by the Moses SMT decoder running with {\IRSTLM}. 
To convert the file into the standard ARPA format you can use the command:

\begin{verbatim}
$> compile-lm train.ilm.gz train.lm --text=yes
\end{verbatim}
this will create the proper ARPA file {\tt lm-final}.
To create a gzipped file you might also use:
\begin{verbatim}
$> compile-lm train.ilm.gz /dev/stdout --text=yes | gzip -c > train.lm.gz
\end{verbatim}

\noindent
In the following sections, we will discuss on LM file formats, on compiling LMs into a 
more compact and efficient binary format, and on querying LMs.

\subsection{Estimating a LM with a Partial Dictionary}

A sub-dictionary can be defined by just taking words occurring more than 5 times ({\tt -pf=5})
and at most the top frequent 5000 words ({\tt -pr=5000}):
\begin{verbatim}
$>dict -i="gunzip -c train.gz" -o=sdict -pr=5000 -pf=5 
\end{verbatim}

\noindent
The LM can be restricted to the defined sub-dictionary with the 
command {\tt build-lm.sh} by using the option {\tt -d}: 
\begin{verbatim}
$> build-lm.sh -i "gunzip -c train.gz" -n 3  -o  train.ilm.gz -k 5 -p -d sdict 
\end{verbatim}

\noindent
Notice that all words outside the sub-dictionary will be mapped into the {\tt <unk>}
class, the probability of which will be directly estimated from the corpus statistics.
A preferable alternative to this approach is to estimate a large LM and then to filter
it according to a list of words (see Filtering a LM).

### 2017-06-08 tools/extras/irstlm/doc/installation.tex

\IMPORTANT{The installation procedure has been tested using the {\tt bash} shell on the following operating systems: Mac OSx 10.6.8 (Snow Leopard), Ubuntu 14.04 LTS (trusty), Scientific Linux release 6.3 (carbon).}

\noindent In order to install {\IRSTLM} on your machine, please perform the following steps.

\subsection{Step 0: Preparation of the Configuration Scripts}
Run the following command to prepare up-to-date configuration scripts.
\begin{verbatim}
$> ./regenerate-makefiles.sh [--force]
\end{verbatim}

\WARNING{Run with the "--force" parameter if you want to recreate all links to the autotools.}

\subsection{Step 1: Configuration of the Compilation}
Run the following command to prepare up-to-date compilation scripts, and to optionally set the installation directory (parameter "{\tt -prefix}".
\begin{verbatim}
$> ./configure [--prefix=/path/to/install] [optional-parameters]
\end{verbatim}

You can set other optional parameters to modify the standard compilation behavior.
\begin{verbatim}
  --enable-doc|--disable-doc
    Enable or Disable (default) creation of documentation
  --enable-trace|--disable-trace
    Enable (default) or Disable trace info at run-time
  --enable-debugging|--disable-debugging
    Enable or Disable (default) debugging info ("-g -O2")
  --enable-profiling|--disable-profiling
    Enable or Disable (default) profiling info
  --enable-caching|--disable-caching
    Enable or Disable (default) internal caches
    to store probs and other info
  --enable-interpolatedsearch|--disable-interpolatedsearch
    Enable or Disable (default) interpolated search for n-grams
  --enable-optimization|--disable-optimization
    Enable or Disable (default) C++ optimization info ("-O3")
\end{verbatim}

\noindent
Run the following command to get more details on the compilation options.
\begin{verbatim}
$> configure --help
\end{verbatim}

\subsection{Step 2: Compilation}

\begin{verbatim}
$> make clean
$> make
\end{verbatim}

\subsection{Step 3: Installation}
\begin{verbatim}
$> make install
\end{verbatim}

\noindent
Libraries and commands are generated,  respectively, under the directories\newline {\tt /path/to/install/lib} and {\tt /path/to/install/bin}.

\noindent
If enabled and PdfLatex is installed, this user manual (in pdf) is generated under the directory\newline {\tt /path/to/install/doc}.

\noindent
Although caching is not enabled by default, it is highly recommended to activate through its compilation flag "{\tt --enable-caching}".
%See Section~\ref{sec:caching} to learn more.

\subsection{Step 4: Environment Settings}
Set the environment variable {\tt IRSTLM} to {\tt /path/to/install}.

\noindent
Include the command directory {\tt /path/to/install/bin} into your environment variable {\tt PATH}.
For instance, you can run the following commands

\begin{verbatim}
$> export IRSTLM=/path/to/install/
$> export PATH=${IRSTLM/bin:${PATH}
\end{verbatim}

\subsection{Step 5: Regression Tests}
If the installation procedure succeeds, you can also run the regression tests to double-check the integrity of the software.
Please go to Section~\ref{sec:regressionTests} to learn hot to run the regression tests.

\noindent Regression tests should be run also in the case of any change made in the source code.

### 2017-06-08 tools/extras/irstlm/doc/introduction.tex

This manual illustrates the functionalities of  the IRST Language  Modeling (LM)  toolkit ({\IRSTLM}). It  should  
put you quickly  in  the condition of:
\begin{itemize}
\item extracting the dictionary from a corpus
\item extracting n-gram statistics from it
\item estimating n-gram LMs using different smoothing criteria
\item saving a LM into several textual and binary file
\item adapting a LM on task-specific data
\item estimating and handling gigantic LMs
\item pruning a LM
\item reducing LM size through quantization
\item querying a LM through a command or script
\end{itemize}

\noindent
{\IRSTLM} features very efficient algorithms and data structures suitable to estimate, store, and access very  large LMs.

\noindent
{\IRSTLM} provides adaptation methods to effectively adapt generic LM to specific task when only little task-related data are available.

\noindent
{\IRSTLM} provides standalone programs for all its functionalities, as well as library for its exploitation in other softwares, like for instance speech recognizers, machine translation decoders, and POS taggers.

\noindent
{\IRSTLM} has been integrated into a popular open source SMT decoder  called {\tt Moses}\footnote{http://www.statmt.org/moses/}, and is compatible with LMs created with other tools, such as the SRILM Tooolkit\footnote{http://www.speech.sri.com/projects/srilm}.

\paragraph{Acknowledgments.}Users of this toolkit  might cite in their publications:
\begin{quote}
M. Federico,  N. Bertoldi,  M. Cettolo, {\em IRSTLM: an Open Source Toolkit for Handling Large Scale Language Models}, Proceedings of Interspeech, Brisbane, Australia, pp. 1618-1621, 2008.
\end{quote}

\noindent
References to introductory material on $n$-gram LMs are given in Appendix~\ref{sec:ReferenceMaterial}.

### 2017-06-08 tools/extras/irstlm/doc/referenceMaterial.tex

The following books contain basic introductions to statistical language modeling:
\begin{itemize}
\item {\em Spoken Dialogues with Computers}, by Renato DeMori, chapter 7.
\item {\em Speech  and Language Processing},  by Dan  Jurafsky and  Jim Martin, chapter 6.
\item {\em Foundations   of  Statistical   Natural  Language   Processing},  by C. Manning and H. Schuetze.
\item {\em Statistical Methods for Speech Recognition}, by Frederick Jelinek.
\item {\em Spoken Language Processing}, by Huang, Acero and Hon.
\end{itemize}

\noindent
The following papers describe the IRST LM toolkit:
\begin{itemize}

\item Efficient data structures to handle huge language models:
\begin{quote}
Marcello Federico and Mauro Cettolo, {\em Efficient Handling of N-gram Language Models for Statistical Machine Translation}, In Proc. of the Second Workshop on Statistical Machine Translation, pp. 88--95, ACL, Prague, Czech Republic, 2007.
\end{quote}

\item Language Model quantization:
\begin{quote}
Marcello Federico and Nicola Bertoldi, {\em How Many Bits Are Needed To Store Probabilities for Phrase-Based Translation?}, In Proc. of the Workshop on Statistical Machine Translation. pp. 94-101, NAACL, New York City, NY, 2006. 
\end{quote}

\item Language Model adaptation with mixtures:
\begin{quote}
Marcello Federico and Nicola Bertoldi, {\em Broadcast news LM adaptation over time}, Computer Speech and Language. 18(4): pp. 417-435, October, 2004.
\end{quote}
\item Language Model adaptation with MDI:
\begin{quote}
Marcello Federico, {\em Efficient LM Adaptation through MDI Estimation}. In Proc. of Eurospeech, Budapest, Hungary, 1999.
\end{quote}
\end{itemize}

### 2017-06-08 tools/extras/irstlm/doc/releaseNotes.tex

\IMPORTANT{If present, the index in parentheses refers to the revision number in IRSTLM repository (until 5.60.02) or SourceForge repository (from 5.60.03).}
\subsection{Version 3.2}
\begin{itemize}
\item Quantization of probabilities
\item Efficient run-time data structure for LM querying 
\item Dismissal of MT output format
\end{itemize}

\subsection{Version 4.2}
\begin{itemize}
\item Distinction between open source and internal Irstlm tools
\item More memory efficient versions of binarization and quantization commands
\item Memory mapping of run-time LM
\item Scripts and data structures for the estimation and handling of gigantic LMs 
\item Integration of {\IRSTLM} into Moses Decoder
\end{itemize}

\subsection{Version 5.00}
\begin{itemize}
\item Fixed bug in the documentation 
\item General script {\tt build-lm.sh} for the estimation of large LMs.
\item Management of iARPA file format.
\item Bug fixes
\item Estimation of LM over a partial dictionary.
\end{itemize}

\subsection{Version 5.04}
\begin{itemize}
\item Extended documentation with ShiftBeta smoothing. 
\item Smoothing parameter of ShiftBeta can be set manually.
\item Robust handling for smoothing parameters of ModifiedShiftBeta.
\item Fixed probability checks in TLM.
\item Parallel estimation of gigantic LM through SGE
\item Better management of sub dictionary with build-lm.sh   
\item Minor bug fixes
\end{itemize}

\subsection{Version 5.05}
\begin{itemize}
\item (Optional) computation of OOV penalty in terms of single OOV word instead of OOV class
\item Extended use of OOV penalty to the standard input LM scores of compile-lm. 
\item Minor bug fixes
\end{itemize}

\subsection{Version 5.10}
\begin{itemize}
\item Extended ngt to compute statistics for approximated Kneser-Ney smoothing
\item New implementation of approximated Kneser-Ney smoothing method
\item Minor bug fixes
\item More to be added here ....
\end{itemize}

\subsection{Version 5.20}
\begin{itemize}
\item Improved tracing of back-offs
\item Added command prune-lm  (thanks to Fabio Brugnara)
\item Extended lprob function to supply back-off weight/level information
\item Improved back-off handling of OOV words with quantized LM
\item Added more debug modalities to compile-lm
\item Fixed minor bugs in regression tests
\item Updated documentation
\end{itemize}

\subsection{Version 5.21}
\begin{itemize}
\item Addition of interpolate-lm 
\item Added LM filtering to compile-lm
\item Improved regression tests
\item Integration of interpolated LMs in Moses
\item Extended tests on compilers and platforms
\item Improved documentation with website
\end{itemize}

\subsection{Version 5.22}
\begin{itemize}
\item Use of AutoConf/AutoMake toolkit compilation and installation
\end{itemize}

\subsection{Version 5.30}
\begin{itemize}
\item Support for a safe management of LMs with a total amount of $n$-grams larger than 250 million
\item Use of a new parameter to specify a directory for temporary computation because the default ("/tmp") could be too small
\item Improved a safer method of concatenation of gzipped sub lms
\item Improved management of log files
\end{itemize}

\subsection{Version 5.40}
\begin{itemize}
\item Merging of internal-only tlm code into the public version
\item Updated documentation into the public version
\item Included documentation into the public version
\end{itemize}

\subsection{Version 5.50}

\begin{itemize}
\item {\bf 5.50.01}
\begin{itemize}
\item Binary saving directly with tlm
\item Speed improvement through 
\item Caching of probability and states of $n$-grams in the LM interface
\item Storing of $n$-grams in inverted order
\end{itemize}

\item {\bf 5.50.02}
\begin{itemize}
\item Optional creation of documentation
\item Improved documentation
\item Optional computation of the perplexity at sentence-level
\end{itemize}

\end{itemize}

\subsection{Version 5.60}
\begin{itemize}
\item {\bf 5.60.01}
\begin{itemize}
\item Handling of class/chunk LMs with both compile-lm and interpolate-lm
\item Improved pruning strategy to handle with sentence-start symbols
\item Improved documentation and examples
\end{itemize}
\item {\bf 5.60.02}
\begin{itemize}
\item Code cleanup 
\end{itemize}
\item {\bf 5.60.03 (r404)}
\begin{itemize}
\item Xcode project
\item import from IRSTLM repository (revision 4263)
\end{itemize}

\end{itemize}

\subsection{Version 5.70}
\begin{itemize}
\item {\bf 5.70.01 (r454)}
\begin{itemize}
\item Class-based LM
\item Added improved-kneser-ney smoothing for lm-build-qsub.sh
\item Enabled different singleton pruning policy for each submodel of mixture LM
\item Enabled the possibility to load an existing LM up to a specific level smaller than the actual LM order
\item Code tracing
\item Handling of error codes
\item Handling of long filenames and parameters
\item Improved parallel code compilation
\item Improved documentation and examples
\end{itemize}
\item {\bf 5.70.02 (r469)}
\begin{itemize}
\item Code optimization
\item Common interface for all LM types
\end{itemize}
\end{itemize}

\subsection{Version 5.80}
\begin{itemize}
\item {\bf 5.80.01 (r501)}
\begin{itemize}
\item Facility to {\em beautify} source code
\item Re-activation of filtering on a sub-dictionary
\item Code optimization related to LM dumping
\item Transformation of scripts into Bourne shell scripts 
\end{itemize}
\item {\bf 5.80.03 (r579)}
\begin{itemize}
\item Data selection tool
\item Handling of precision upper- and lower-bounds by means of constants.
\item Improved of Xcode project
\item Improved code compilation
\item Improved documentation
\item Improved handling of help
\item Facility to check whether IRSTLM is compile with or without caching
\end{itemize}
\item {\bf 5.80.05 (r642)}
\begin{itemize}
\item Introduction of {\em namespace irstlm}
\item Code optimization
\item Code compliant with OsX Maverick
\item Support for Redis output format to ngt
\item Support CRC16 algorithm
\item Improved plsa command and regression test
\item Improved handling of tracing
\item Improved handling of help
\item Improved handling of error messages
\end{itemize}
\item {\bf 5.80.06 (r647)}
\begin{itemize}
\item Improved code compilation
\end{itemize}
\item {\bf 5.80.07}
\begin{itemize}
\item Changes to LM smoothing types; removed Good-Turing, added Approximated Modified ShiftBeta, renaming
\item Added an additional pruning method, based on level-dependent pruning frequency
\item Improved code compilation
\item Improved documentation
\item Code cleanup
\item Added support for long names of parameters
\item Improved output format 
\end{itemize}
\item {\bf 5.80.08}
\begin{itemize}
\item Added functionality to score n-grams in isolation
\item Added level-based caches for storing prob, state, and statesize
\item Improved management of tracing assert
\item Improved management of tracing/assert/caching AutoConf compilation flags
\item Improved output format
\item Improved code compilation
\item Code cleanup
\end{itemize}
\end{itemize}

\COMMENT{
\subsection{Version 5.xx}
\begin{itemize}
\item {\bf 5.xx.01}
\begin{itemize}
\item 
\end{itemize}
\end{itemize}
}
